### 2014-06-04 docs/arxiv.tex
Moving experiments to here to generate final versions for paper

%\documentclass{nature}
\documentclass{article}
\usepackage[english,american]{babel}
\usepackage[backend=biber,style=nature]{biblatex}
\usepackage[disable]{todonotes}
\usepackage{grffile}
\usepackage{booktabs}
\usepackage{hyperref}
\usepackage{graphicx}
\usepackage{fullpage}
%\usepackage{algorithm}
%\usepackage{algorithmic}
\usepackage{amsmath}
%\usepackage{subcaption}
%\usepackage{multicol}
\usepackage{csquotes}
\usepackage{pdfpages}
\usepackage{placeins}

\renewcommand{\vec}[1]{\mathbf{#1}}

%\renewcommand{\includegraphics}[2][]{\fbox{}}

\addbibresource{connectomics motif.bib}
\addbibresource{connectomics motif-relational models.bib}

\DeclareGraphicsRule{.ai}{pdf}{.ai}{}

\title{Automatic discovery of circuit structure from neural connectomics}
\author{Eric Jonas$^1$, Konrad Kording$^{2, 3,4}$}

\begin{document}
\maketitle

\begin{small}
\begin{enumerate}
  \item Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology, Cambridge, Massachusetts
  \item Department of Physical Medicine and Rehabilitation, Northwestern University and Rehabilitation Institute of Chicago, Chicago, Illinois
  \item Department of Physiology, Northwestern University, Chicago, Illinois
  \item Department of Applied Mathematics, Northwestern University, Chicago, Illinois
\end{enumerate}
\end{small}

\begin{abstract}
  New techniques produce massive data about neural connectivity,
  necessitating new analysis methods to discover underlying
  structure. It has long been assumed that discovering the local
  patterns of microcircuitry is the key to understanding neural
  function. Here we developed a nonparametric Bayesian technique that
  identifies neuron types and circuitry patterns in connectomics
  data. We show that the approach recovers known neuron types in the
  retina, reveals interesting structure in the nervous system of
  c. elegans, and automatically discovers the structure of
  microprocessors. Our approach extracts structural meaning from
  connectomics, enabling new approaches of deriving anatomical
  insights from these emerging datasets.

\end{abstract}

\section*{Introduction}
Emerging techniques \autocite{Morgan2013,Zador2012} promise to
quantify the location of each neuron within a volume of interest,
along with its connections to all other neurons. Far exceeding the
capacity of neuroanatomists to trace small circuits, this leads to
enormous datasets that quantify aspects of the nervous system. The impact of
this kind of high-throughput transition has been seen before -- 
rise of sequencing techniques necessitated the
development of novel computational methods to understand genomic
structure, ushering in an era of bioinformatics as an independent
discipline \autocite{Koboldt2013}.

The brain consists of multiple kinds of neurons, each of which is
hypothesized to have a specific role in the overall
computation. Neuron types differ in many ways, e.g. chemical or
morphological, but they also differ in the way they connect to one
another. In fact, the idea of well defined local connectivity patterns
has been prominent in many areas, ranging from sensory (e.g. retina,
\autocite{Masland2001} via processing (e.g. neocortex \autocite{Mountcastle1997})
to movement (e.g. spinal cord) \autocite{Grillner2005}. Indeed, such
repeated patterns also exists in human-made computing circuits. It
remains an important challenge to develop algorithms to use anatomical
data, e.g. connectomics, to back out underlying connectivity motifs.

The discovery of structure is a crucial aspect of network
science. Early approaches focused on global graph properties, such as
the types of scaling present in the network \autocite
{WattsStrogatz1998} .  While this approach leads to an understanding
of global network, more recent work aims at identifying very small-scale
repeat patterns, or “motifs” in networks\autocite{Milo2002}

The discovery of structure in probabilistic graphs is a crucial aspect
in machine learning. Commonly used algorithms include
community-based-detection methods \autocite{Girvan2002}, and
stochastic block models \autocite{Nowicki2001}.  While these
approaches can incorporate the probabilistic nature of neural
connections \autocite{Hill2012} they do not generally consider the
known fact that the probability of connections between neurons depends
on distance. Combining such probabilistic models with a spatial
component to model distance promises to improve the detection of
microcircuitry from neural data.

When it comes to neuroscience and other computing systems, we expect
patterns of connectivity much more complex than traditional motifs,
exhibiting a strong spatial dependence arising from the complex
genetic, chemical, and activity-based neural development process.

Here we describe a Bayesian non-parametric model that can discover
circuit structure automatically from connectomics data: both the cell
types and their spatial patterns of interconnection. We apply it to
three computing systems: the mouse retina connectome
\autocite{Helmstaedter2013}, the c. elegans connectome
\autocite{White1986}, and a ``connectome'' of a classical
microprocessor \autocite{James2010}. Note we do not explicitly
consider cellular morphology, absolute position (cell layer), or any
chemical or genetic information. Yet in all cases, we discover
cell-type to cell-type connectivity that depends on spatial distance
and these connectivity functions depend on the two involved cell
types. Comparing the cell types discovered by the algorithms with
those obtained by anatomists based on orthogonal information reveals a
high degree of agreement. We present a scalable approach to infer
microcircuitry from connectivity data.

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=183mm]{overview.ai}}
  \caption{An overview of our method. a. Input data is
a connectivity matrix indicating whether cell $i$ (row)
synapses on cell $j$ (column). The location of each cell in space is known. 
b. Our algorithm discovers hidden cell types in this connectivity data
by assuming all cells of a type share a distance-dependent connectivity 
profile, c.), with cells of other types. This results in a clustering 
of the cells by those hidden types. b.) shows the cell connectivity
matrix with cells of the same type grouped together. d.) shows
the learned probability of connection between our different types
at various distances, illustrating short-range and long-range
connectivity patterns.}
\label{fig:overview}
\end{figure}

\section{Results}
We use a Bayesian structure discovery approach that is based on
prior knowledge about the way positions and cell types affect the
probability of connections between cells, a so-called generative
model. We term our model the ``infinite spatial-relational model'' (iSRM), 
in keeping with the existing nonparametric literature.  
The probability of connections between pairs of neurons
decreases with distance, but how high the baseline probability of
connection is and how quickly it decays with distance depends on the
two cell-types. Note that this simplistic prior information is all that 
we tell our model; no additional physiological information is included.

We combine annealing with Markov Chain Monte Carlo to
perform simultaneously joint posterior inference over the structure,
parameters, and hyperparameters of the model (see Methods). This
procedure automatically learns the number of cell types, which cells 
belong to which types, along with the 
the distance and type dependent connectivity probabilities (see
fig~\ref{fig:overview}). These results can then be compared with the
results of previous anatomical studies that have used other aspects of
neurons, such as chemical properties and cell-morphology to divide
neurons into types.

Before we can use a probabilistic model to analyze real data we first
need to assess if the model works properly. We thus started out by
simulating data for which we know the correct structure and comparing
the estimated structure based on the algorithm (see methods) with the
one we used for simulation. We find that the model does a good job of
recovering the structure we used to generate the data
(fig~\ref{fig:synthetic}a) Indeed, it does far better than a simpler
and well established model (the infinite stochastic block model, or
iSBM) that assumes that there is no distance dependence to the
connectivity. The model converges relatively quickly to an estimate of
the most probable values for the cell types, which is enabled by using
a combination of simulated annealing and parallelized Markov chain
Monte Carlo (see methods for details). It appears that the model
behaves the way it should and allows application to relatively large
datasets.

\begin{figure}
  \centering 
  \centerline{\includegraphics{synthetic.ai}}
  \caption{Correct recovery of true numbers of hidden classes
in synthetic data. a.) Other models over-cluster the number
of classes as they fail to take distance into account, whereas
our model finds close to the true number of classes. b.) ARI, a measure
of ``correct clustering'', for
different true class counts, between our model and a traditional
iSBM. c.) For synthetic data, 
the iSBM overclusters, finding a large number of spatially-localized types. 
In contrast, the infinite spatial-relational model recovers the true spatial
extent of the underlying types.}
\label{fig:synthetic}
\end{figure}

\subsubsection{Mouse Retina}

\begin{figure}
  \centering 
  \centerline{\includegraphics{mouseretina.ai}}
  \caption{Discovering cell classes in the mouse retina connectome. 
a.) Input connectivity data for 950 cells for which soma positions were known. b. clustered connectivity matrix. c.) connectivity diagram showing our clusters, as
well as the cell depth and known cell type of each cell. Our model agrees
with the cell types traditionally identified by anatomists, and identifies 
laminar-specific connectivity patterns. d.) Connectivity between our
clusters as a function of distance -- the cluster consisting primarily of
retinal ganglion cells (lower left-center) exhibits the expected near and
far connectivity. e. The spatial distribution of our cell types -- each cell type tesselates space.}
\label{fig:mouseretina}
\end{figure}

The retina of the mouse \autocite{Masland2001} is an example where we
should expect to have connectivity patterns that are well approximated
by our generative model. It is known that there are multiple classes
of cells that can be grouped into ganglion cells that transmit
information to the rest of the brain, bipolar cells that connect
between different cells, and amacrine cells that feed into the
ganglion cells \autocite{cite some review}. Recent research
\autocite{Helmstaedter2013} has produced a large dataset containing
both the types of cells from orthogonal approaches, and also the
connectivity matrix between all reconstructed cells. We want to know
to which level we can classify neurons into types based on this
connectivity only.

The algorithm took 8 hours to converge to a locally optimal solution,
dividing neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
wedge is a type). For each pair of neurons there is a specific
distance dependent connection probability (fig~\ref{fig:mouseretina}b,c,d),
which is well approximated by the model fit. Moreover, each type of
cell is rather isotropically distributed across space
(fig~\ref{fig:mouseretina}e) as should be expected for true cell types. The
algorithm is thus able to extract meaningful structure from the data.

Comparing the results of the algorithm to other information sources
allows evaluating the quality of the type determination. Our types
closely reflect the (anatomist-determined) segmentation of cells into
retinal ganglion, narrow amacrine, medium/wide amacrine, and bipolar
cells (fig~\ref{fig:mouseretina}c, outermost ring). We find that the
types we find tend to reflect the known laminar distribution in the
retina (fig~\ref{fig:mouseretina}c, middle ring), which is exciting as
the algorithm did not actually use this type of information.

The algorithm yields a separation of neurons into a smaller number of
types than the fully granular listing of 71 types found by the
original authors of the paper, although is still highly correlated
with those finer type distinctions (see
supp~\ref{supp:mouseretina}. It is our expectation that, with larger
datasets, even closer agreement would be found.

\todo{Do we assign any of the unknown cells correctly?  are types 72-79
“different” or simply unknown?}

We compared the results of our model to the traditional iSBM to see if
the addition of spatial information was necessary. We find that, as
expected from our synthetic experiments, the iSBM massively
over-segmented the data, finding spatially-localized “communities” of
cells which bore little relation to known types.  The average spatial
extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. However,
in the ground truth by anatomists the average distance was $34.3\pm7.6
\mu m$ and $35.3\pm 2.4 \mu m$, quite close to the $34.5\pm 1.0 \mu m$
discovered by our algorithm. Modeling the distance-dependence prevents
over-segmentation and thus allows more meaningful estimates.

% from experiments/mouseretina/spatial_var.txt

% Relational Model: mean = 14.586508 std=8.136334
% Spatial-Relational Model: mean = 34.540355 std=1.069483
% Truth (fine): mean = 34.348443 std=7.567564
% Truth (coarse): mean = 35.276636 std=2.372951

\subsection{C. elegans}

The roundworm Caenorhabditis elegans is a model system in
developmental neuroscience\autocite{White1986}, with the location and
connectivity of each of 302 neurons deterministically known. This
incredible cross-organism consistency led to early mapping of the
c. elegans chemical and electrical (gap junction) connectome. Unlike
the retina, only the motor neurons in c. elegans exhibit regular
distribution in space, being distributed along the body axis. Most
interneurons are concentrated in various ganglia that project
throughout the entire animal, and the sensory neurons are primarily
located in a small number of anterior ganglia.

\begin{figure}
  \centering 
  \centerline{\includegraphics[width=5in]{celegans.ai}}
  \caption{Discovering cell types in the \textit{c. elegans}
    connectome. a.) Initial cell connectivity matrix, red are
    (directional) chemical synapses, blue are electrical
    synapses. Point size is proportional to synapse count between
    those cells. b.) Clustered adjacency matrix showing identified
    clusters. c.) Inter-cluster connectivity, showing the position of
    each cell along the body axis as well as the ``true'' cell type --
    both coarse (motor, sensory, and interneuron) as well as fine
    (small text labels). }
  \label{fig:celegans}
\end{figure}

Using both the chemical and electrical connectivity (see methods), we
determined the underlying clusters explained by connectivity and
distance (fig~\ref{fig:celegans}a). A superficial inspection of the results shows
clustering into groups consisting roughly homogeneously of motor
neurons, sensory neurons, and interneurons. Closer examination reveals
agreement with the classifications originally outlined by Brenner in
1986.

We identify cell types that reflect the known motor/non-motor neuron
classes, even though this system lacks the strong repeat
microcircuitry our model was designed for.  Motor-neuron types AS, DA,
and VA, all exclusively postsynaptic, are clustered together, as are
motor-neuron types VD and DD. Traditional types VC, DB, and VB also
mostly share a cluster. Various head motor neurons, including types
SMD, RMD, and RIM are clustered together. Interneurons with known
anatomically-distinct connectivity patterns, such as RIA (2 cells) or
AVA (2 cells) are clustered into pure types. The algorithm even
correctly places the single-cell type DVB by itself. Note our
clustering does not perfectly reflect known divisions -- several
combinations of head and sensory neurons are combined, and a
difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are
grouped together. Our identified cell types thus reflect a
``coarsening'' of known types, based entirely on connectivity and
distance information, even when the organism exhibits substantially
less spatial regularity than the retina.

\subsection{Microprocessor}
\begin{figure}
  \centering 
  \centerline{\includegraphics{mos6502.ai}}
  \caption{Discovering transistor types in the MOS 6502
    microprocessor. a.) The clusters and connectivity versus distance
    for connections between Gate and C1, Gate and C2, and C1 and C2
    terminals on a transistor. Red and teal types have a terminal
    pulled down to ground and mostly function as inverters. Purple
    class are clocked, stateful transistors, orange control the ALU
    and yellow control the special data bus. b.) The spatial
    distribution of the transistors in each cluster. }
\label{fig:mos6502}
\end{figure}

To show the applicability of our method to other connectome-style
datasets, we obtained the spatial location and interconnectivity of
the transistors in a classic microprocessor, the MOS Technology 6502
(used in the Apple II) \autocite{James2010}. Computer architects use
common patterns of transistors when designing circuits, with each
transistor having a ``type'' in the circuit. We identified a region of
the processor with complex but known structure containing the primary
8-bit registers X, Y, and S (fig~\ref{fig:mos6502}).

Our algorithm identifies areas of spatial homogeneity that mirror the
known structure in the underlying architectural circuit, segmenting
transistor types recognizable to computer architects. Using the
original schematics, we see that one identified type contains the
``clocked'' transistors, which retain digital state. Two other types
contain transistors with pins C1 or C2 connected to ground, mostly
serving as inverters.  An additional identified type controls the
behavior of the three registers of interest (X, Y, and S) with respect
to the SB data bus, either allowing them to latch or drive data from
the bus. The repeat patterns of spatial connectivity are visible in
figure~\ref{fig:mos6502}a, showing the man-made horizontal and
vertical layout of the same types of transistors.

\section{Discussion}
We have presented a machine learning technique that allows cell types
and microcircuitry to be discovered using only distance and
connectivity data. We have shown its applicability to a newly
published connectomics dataset from the mouse retina, the classical
connectivity of the worm c. elegans, and the connectivity of a
historical man-made microprocessor. We have found that distance and
connectivity alone are sufficient to discover cell types and
microcircuitry that were known to exist in the systems based on decades
of previous research which had utilized extensive additional
information.

For probabilistic models like ours, no known solution exists to
exactly find the most probable parsing of the neurons into cell-types
and connectivity patterns . We employ a collection of Markov-chain
Monte carlo techniques (see Methods) but while different
initializations converge to similar ultimate values, we can never
realistically obtain the global optimum. There are a broad range of
techniques that may offer better approximations to the global optimum
\autocite{Wanga2012} and future work could adapt them to
the problem we addressed here.

Our inference becomes slower as the amount of data increases. Our
algorithm required 30 hours for 1000 neurons and could take months for
substantially longer datasets. Scaling this class of probabilistic
model is an active area of research, and recent results in both
variational methods \autocite{Hoffman2013} and spectral learning
  \autocite{Anandkumar2012} promise substantial speed increases.

  Larger datasets will allow algorithms to distinguish more distinct
  types \todo{Add supplemental data} and we expect closer agreement
  with existing anotomical knowledge as more data become available.
  Moreover, in general, for such problems precision increases with the
  size of the dataset and the cells that we have are not sufficient to
  statistically distinguish all the cell types known in anatomy (such
  as the $\sim 70$ in the retina). Still, using only connectivity and
  distance it is possible to meaningfully divide neurons into types.

Our small collection of hand-selected distance-dependent likelihood
functions are clearly non-exhaustive, and assume monotonicity
of connectivity probability -- for a given class, closer cells
are never less-likely to connect. This is known to be insufficient
for various neural systems. Future models could incorporate
a wider variety of likelihood functions, or even learn the global
functional form from the data.

There exist a range of previous approaches to the discovery of
neural microcircuitry \todo{Do we want a citation here?}. These
generally involve a great deal of manual labor and ad-hoc
determination of what constitutes a “type” of cell -- to this day
there are disagreements in the literature as to the “true” types in
the mammalian retina. Much as phylogenomics has changed our
understanding of animal ontologies, modern large scale data will allow
the efficient unbiased discovery of cell types and circuits. The sheer
amount of available data demands the introduction of algorithmic
approaches.

The development of automatic identification and quantification of cell
type may also provide a new computational phenotype for quantifying
the effect of disease, genetic interventions, and
developmentally-experienced neural activity. Our method can in
principle identify neuron-types across non-connected graphs,
e.g. across animals. For example, the types of neurons in one animal
can be associated with the types of neurons in another animal, in the
same way as this is already possible through molecular markers
\autocite{Brown2009}. This could be particularly important if cell
types appear that are due to properties of the stimuli and experience
as opposed to just the molecular properties of cells, such as color
and orientation selective types in primary visual cortex
\autocite{Sincich2005,Lennie2005}. This would allow comparative
quantitative anatomy across animals, and aid the search for the
ultimate causes of connectivity.

Ultimately, we will need to combine information across different
sources into a joint understanding of neural function. Distinct cell
types differ in morphology, connectivity, transcriptomics, relation to
behavior or stimuli and many other ways. Algorithms based on each type
information can then be combined to synthesize all the available
information from one experiment or even across experiments into a
joint model of brain function.

\subsection{Methods Summary}

We take as input a connectivity matrix $R$ defining the connections
between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
e_j)$ representing a (physical) distance between adjacent cells. See
the supplemental material for extension to multiple connectivity
matrices. We assume there exist an unknown number $K$ of latent
(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
observed connectivity between two cells $R(e_i, e_j)$ then depends
only on their latent type and their distance through a link function
$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
as a set of global hyper parameters $\theta$, such that the link
function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.

We then jointly infer the maximum a posteriori (MAP) estimate of the
class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
the global model hyperparameters $\theta$ :

\begin{equation}
  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
\end{equation}

Inference is performed in three steps via composible transition 
kernels -- one for structural, one for per-type parameters, and
one kernel for global parameters and hyperparameters. Details
of data preprocessing, inference parameters, and runtime can
be found in the Methods section.

\printbibliography

\begin{itemize}

 \item \textbf{Acknowledgments} We thank Josh Vogelstein for discussions and reading of the manuscript, Finale Doshi-Velez for early discussions on the model, and Erica Peterson, and Jonathan Glidden for extensive manuscript review. Funding for compute time was provided by Amazon Web Services ``AWS in Education'' grants. 
\item \textbf{Author Contributions} KK and EJ developed model. EJ derived inference, implemented code, tested, and ran experiments. KK and EJ wrote manuscript text and solicited feedback. 
 \item \textbf{Competing Interests} The authors declare that they have no
competing financial interests.
 \item \textbf{Correspondence} Correspondence and requests for materials
should be addressed to E.J.~(email: jonas@mit.edu).
\end{itemize}

\newpage
\section{Methods }

\subsection{Probabilistic Model}

Our model is a extension of the iSBM
\autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
inspired by attempts to extend these models with arbitrary
discriminative functions\autocite{Murphy2012}.

We take as input a connectivity matrix $R$ defining the connections
between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
e_j)$ representing a (physical) distance between adjacent cells. See
the supplemental material for extension to multiple connectivity
matrices. We assume there exist an unknown number $K$ of latent
(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
observed connectivity between two cells $R(e_i, e_j)$ then depends
only on their latent type and their distance through a link function
$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
as a set of global hyper parameters $\theta$, such that the link
function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.

We then jointly infer the maximum a posteriori (MAP) estimate of the
class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
the global model hyperparameters $\theta$ :

\begin{equation}
  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
\end{equation}

% Methods section 3000 words
% methods summary does not appear in the online text
% methods section can have no figures or tables [grrr]

We describe the spatial ``Logistic-distance Bernoulli''  function here,
and others in the supplemental material.

% fixme class component, NOT class
The ``logistic-distance Bernoulli'' spatial model assumes that, if cell
$e_i$ is of type $m$ and cell $e_j$ is of type $n$, then $\eta_{mn}
= (\mu_{mn}, \lambda_{mn})$, and the probability that two cells $e_i$
and $e_j$ are connected is given by
\begin{eqnarray}
p^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
p &= & p^* \cdot (p_{max} - p_{min}) + p_{min}
\end{eqnarray}
where $p_{max}$ and $p_{min}$ are global per-graph parameters.

We place an exponential priors on the latent parameters:
\begin{eqnarray}
 \mu_{mn} \sim \exp(\mu | \mu^{hp}) \\
\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})
\end{eqnarray}

using  $\lambda^{hp}$ and $\mu^{hp}$ as global per-graph hyperparameters.

We use a Dirichlet-process prior on class assignments, which allows
the number of classs to be determined automatically. In brief, for $N$
total cells, the probability of a cell belonging to a class is
proportional to the number of datapoints already in that class, $N_k$,
such that $p(c_i = k) \propto \frac{m_k}{N + \alpha}$ and the
probability of the cell belonging to a new class $k'$ is $p(c_i = k')
\propto \frac{\alpha}{N + \alpha}$. $\alpha$ is the global
concentration parameter -- larger values of $\alpha$ make the model
more likely to propose new classes. We grid the parameter $\alpha$ and
allow the best value to be learned from the data.

\subsection{Inference} 
We perform posterior inference via Markov-chain Monte Carlo (MCMC),
annealing on the global likelihood during the traditional burn-in
phase. MCMC transition kernels for different parts of the state space
can be chained together to construct a kernel whose ergodic
distribution is the target ergodic distribution over the entire state space.

Our first transition kernel (``structural'') performs gibbs sampling 
of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
The lack of conjugacy in our likelihood model makes an explicit 
evaluation of the conditional assignment probabilities impossible, 
motivating us to use an auxiliary variable method \autocite{Neal2000}
in which a collection of ephemeral classs are explicitly represented
for the duration of the Gibbs scan.

We then employ a transition kernel to update the per-component
parameter values $\eta_{mn}$. Conditioned on the assignment vector
$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
individual $\eta_{mn}$ are independent. We slice sample \autocite{Neal2003}
each component's parameters, choosing the slice width as a function
of the global hyperparameter range.

The global hyper-parameters, both $\alpha$ and $\theta$, are allowed
to take on a discrete set of possible values. As $\theta$ is often a
tuple of possible values, we explore the cartesian product of all
possible values. We then Gibbs sample (our final transition kernel),
which is always possible in a small, finite, discrete state space.

We chain these three kernels together, and then globally
anneal on the likelihood from a temperature of $T=64$ down to 
$T=1$ over 300 iterations unless otherwise indicated, and
then run the chain for another $100$ iterations.

To pick the MAP, we run up to $100$ simultaneous chains in parallel,
initialized from different random initial points in the state space,
and then pick the chain with the highest log likelihood. Early
experiments attempted to use the full ensemble of chains as an
estimate of the posterior but were abandoned due to the difficulty of
summarizing the posterior over such a complex state space, which is an
ongoing area of research in the probabilistic modeling community
\todo{cite more? missing}.

\todo{Validation: do we want to write extensively about posterior
  enumeration tests, etc. ? }

\subsection {Mouse Retina}
\label{supp:mouseretina}
Dense serial electron microscopy of a $114\mu m \times 80 \mu m $ area
in the mouse retina by \autocite{Helmstaedter2013} yielded a listing
of places where neurons come into contact. There were over 1000 cells
originally, and selected the $950$ for which the location of the soma
could be reconstructed from the provided cell plots (soma locations
were not provided by the study's authors in machine-readable
form). Ultimately this left a matrix between the total synapse-like
contact area between all pairs of 950 cells. Area was thresholded at
$0.5\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
matrix that served as input to our algorithm. We measured the distance
between cells using the reconstructed soma centers, and used the
Logistic-Distance spatial relation. Hyperprior griddings are shown in
supplemental section \ref{supp:hyperpriors}.

\subsection{C. elegans}

We obtained the connectome of c. elegans from data published
previously \autocite{Varshney2011}, and isolated the 279 nonpharyngeal
neurons, with a total of 6393 chemical synapses and 890 gap junctions
originally cleaned up in \autocite{Chen2006}. A cell's position was
its distance along the anterior-posterior axis normalized between zero
and one. We used both networks, the chemical network as a directed
graph and the electrical network as undirected graph. We use the
synapse counts with the logistic-distance poisson likelihood, scaling
the counts by 4.0 to compensate for the Poisson's overdispersion.

\subsection{Microprocessor}
We extracted the connection graph for the transistors in the MOS6502
\autocite{visual6502source}. Each transistor has three terminals (gate,
source, drain), but the methods of the original dataset were unable to
consistently resolve which of the C1 and C2 terminals were source and
drain, leading to ambiguity in our encoding. We identified a region
consisting of three registers X, Y, and S via visual
inspection and focused our efforts there. We created a total of six
connectivity matrices by examining possible terminal pairings. One
graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
$e_j$ are connected via pins $g$ and $c_1$.

% \begin{addendum}
%  \item Put acknowledgements here.
%  \item[Competing Interests] The authors declare that they have no
% competing financial interests.
%  \item[Correspondence] Correspondence and requests for materials
% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
% \end{addendum}
\newpage
\section{Supplemental}

\subsection{Other Likelihoods}
\label{supp:otherlikelihoods}

We reparameterized the Logistic-Distance Bernoulli likelihood to
better capture the microprocessor data structure. We are explicitly
setting the maximum probability $p$ of the logistic function on a
per-component basis, drawing from a global $p \sim \operatorname{Beta}(\alpha_{hp},
\beta_{hp})$. Then $\lambda$ is set for each component as a global
hyperparameter, $\lambda$.

The ``logstic-distance Poisson'' spatial model is used to explicitly mode the count of synapses, $c$, between two neurons. The probability of c synapses between two neurons is distributed $c \sim \textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a scaled logistic function (the logistic function has range $[0, 1]$. For each component $\eta_{mn}$ we learn both the threshold $\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$ and $n$ are likely to have on average $20$ synases if they are closer than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$  "

Thus the probability of $R(e_i, e_j) = c$ synapses between two cells $e_i$ and $e_j$ is given by:
\begin{eqnarray}
r^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda}}\\
r & = & r^* \cdot (r_{mn} - r_{min}) + r_{min} \\
R(e_i, e_j) \sim \textrm{Poisson}(c | r)
\end{eqnarray}

where $\lambda$ and $r_{min}$ are per-graph parameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $r_{mn} \sim \exp(r_{mn} | r_{scale}^{hp})$.

\subsection{Source code}

All source code and materials for running experiments can be
obtained from the project website, at \\

\href{http://spatialrelation.github.io }{http://spatialrelation.github.io}

\subsection{Extension to multiple graphs}
\label{supp:multigraph}
The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by extending the likelihood to include the product of the likelihoods of the individual graphs.

\begin{equation}
  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}, \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha) 
\end{equation}

\FloatBarrier

\subsection{Microprocessor Netlist annotation}
\label{supp:mos6502netlist}
This is a matrix consisting of the different clusters of electrical nets for 
the registers, with the colors consistent with figure~\ref{fig:mos6502}.

\includepdf[pages=-]{mos6502.clusters.pdf}

\FloatBarrier
\subsection{Model Comparison for Mouse Retina}

We compare the classs found by our model, the traditional infinite
stochastic block model, and a finite stochastic block model with the number
of classes set to the best-found number from our iSRM model ($k=12$).

We see from fig~\ref{fig:supp:irmspatialcomparemouse} that the spatial
variance of our classes closely matches the ground truth, while the
other two models find spatially-localized
classs. Figure~\ref{fig:supp:mousecompareari} shows the ARI of our
clustering, the clustering found by the iSBM,
as well as the result of parametrically varying the traditional stochastic
block model for various values of $k$.

The actual clusterings for iSRM, the iSBM, and $SBM_{k=12}$ are shown
in figures \ref{fig:supp:retinabbclusters},
\ref{fig:supp:retinaldclusters}, and
\ref{fig:supp:retinabbfixedclusters} respectively. Each block is a
cluster, with the true cell type indicated along the x-axis, and the
cell ID written to the left or right.

\begin{figure}[h!]
  \centering 
    \centerline{\includegraphics[width=5in]{mouseretina/spatial_var.pdf}}
  \caption{The distribution of the spatial extent of cell types found by each of
the above methods, and ground truth clusters}
\label{fig:supp:irmspatialcomparemouse}
\end{figure}

\begin{figure}[h!]
  \centering 
    \centerline{\includegraphics[width=5in]{mouseretina/ari_vs_cluster.pdf}}
  \caption{ARI vs cluster number for the iSRM, iSBM, and the SBM with varying K.}
\label{fig:supp:mousecompareari}
\end{figure}

\begin{figure}[h!]
  \centering 
    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
  \caption{The cell IDs and found types (grouped) vs known types (y-axis)  found by the infinite stochastic block model, dramatically overclustering relative to ground truth}
\label{fig:supp:retinabbclusters}
\end{figure}

\begin{figure}[h!]

  \centering 
    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
  \caption{The cell IDs and found types (grouped) vs known types (y-axis) by the infinite spatial-relational model.}
\label{fig:supp:retinaldclusters}
\end{figure}

\begin{figure}[h!]

  \centering 
    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_12-anneal_slow_fixed_400.truth_latent.pdf}}
  \caption{The cell IDs and found types (grouped) vs known types (y-axis) via the stochastic block model with $k=12$}
\label{fig:supp:retinabbfixedclusters}
\end{figure}

\newpage

\FloatBarrier
\subsection{Hyperprior grids}
\label{supp:hyperpriors}

For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80. $p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99]$.

For the c. elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0. We globally set $rate_{min}=0.01$.

For the microprocessor with the Logistic Distance fixed lambda Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced points between 10 and 500 and set $\lambda=\mu_{hp}/10$. $p_{min} \in \{0.001, 0.01, 0.02\}$  and both $p_\alpha$ and $p_\beta \in \{0.1, 1.0, 2.0\}$.

\section{Measuring clustering similarity}

The adjusted rand index (ARI) is a measure of the similarity of two
different clusterings \autocite{Hubert1985} -- two identical clusters
have an ARI of 1.0 while progressively more dissimilar clusters have
lower ARIs, becoming negative as the clustering gets anti-correlated.

\end{document}

### 2014-06-04 docs/nature-template.tex
Moving experiments to here to generate final versions for paper

%% Template for a preprint Letter or Article for submission
%% to the journal Nature.
%% Written by Peter Czoschke, 26 February 2004
%%

\documentclass{nature}

%% make sure you have the nature.cls and naturemag.bst files where
%% LaTeX can find them

\bibliographystyle{naturemag}

\title{Put title here (less than 90 characters).}

%% Notice placement of commas and superscripts and use of &
%% in the author list

\author{Aauthor$^{1,2}$, Bauthor$^2$ \& LastAuthor$^2$}

\begin{document}

\maketitle

\begin{affiliations}
 \item Put institutions in this environment and
 \item separate with \verb|\item| commands.
\end{affiliations}

\begin{abstract}
For Nature, the abstract is really an introductory paragraph set
in bold type.  This paragraph must be ``fully referenced'' and
less than 180 words for Letters.  This is the thing that is
supposed to be aimed at people from other disciplines and is
arguably the most important part to getting your paper past the
editors.  End this paragraph with a sentence like ``Here we
show...'' or something similar.
\end{abstract}

Then the body of the main text appears after the intro paragraph.
Figure environments can be left in place in the document.
\verb|\includegraphics| commands are ignored since Nature wants
the figures sent as separate files and the captions are
automatically moved to the end of the document (they are printed
out with the \verb|\end{document}| command. However, tables must
be manually moved to the end of the document, after the addendum.

\begin{figure}
\caption{Each figure legend should begin with a brief title for
the whole figure and continue with a short description of each
panel and the symbols used. For contributions with methods
sections, legends should not contain any details of methods, or
exceed 100 words (fewer than 500 words in total for the whole
paper). In contributions without methods sections, legends should
be fewer than 300 words (800 words or fewer in total for the whole
paper).}
\end{figure}

\section*{Another Section}

Sections can only be used in Articles.  Contributions should be
organized in the sequence: title, text, methods, references,
Supplementary Information line (if any), acknowledgements,
interest declaration, corresponding author line, tables, figure
legends.

Spelling must be British English (Oxford English Dictionary)

In addition, a cover letter needs to be written with the
following:
\begin{enumerate}
 \item A 100 word or less summary indicating on scientific grounds
why the paper should be considered for a wide-ranging journal like
\textsl{Nature} instead of a more narrowly focussed journal.
 \item A 100 word or less summary aimed at a non-scientific audience,
written at the level of a national newspaper.  It may be used for
\textsl{Nature}'s press release or other general publicity.
 \item The cover letter should state clearly what is included as the
submission, including number of figures, supporting manuscripts
and any Supplementary Information (specifying number of items and
format).
 \item The cover letter should also state the number of
words of text in the paper; the number of figures and parts of
figures (for example, 4 figures, comprising 16 separate panels in
total); a rough estimate of the desired final size of figures in
terms of number of pages; and a full current postal address,
telephone and fax numbers, and current e-mail address.
\end{enumerate}

See \textsl{Nature}'s website
(\texttt{http://www.nature.com/nature/submit/gta/index.html}) for
complete submission guidelines.

\begin{methods}
Put methods in here.  If you are going to subsection it, use
\verb|\subsection| commands.  Methods section should be less than
800 words and if it is less than 200 words, it can be incorporated
into the main text.

\subsection{Method subsection.}

Here is a description of a specific method used.  Note that the
subsection heading ends with a full stop (period) and that the
command is \verb|\subsection{}| not \verb|\subsection*{}|.

\end{methods}

%% Put the bibliography here, most people will use BiBTeX in
%% which case the environment below should be replaced with
%% the \bibliography{} command.

\begin{thebibliography}{1}
\bibitem{dummy} Articles are restricted to 50 references, Letters
to 30.
\bibitem{dummyb} No compound references -- only one source per
reference.
\end{thebibliography}

%% Here is the endmatter stuff: Supplementary Info, etc.
%% Use \item's to separate, default label is "Acknowledgements"

\begin{addendum}
 \item Put acknowledgements here.
 \item[Competing Interests] The authors declare that they have no
competing financial interests.
 \item[Correspondence] Correspondence and requests for materials
should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
\end{addendum}

%%
%% TABLES
%%
%% If there are any tables, put them here.
%%

\end{document}

### 2014-06-04 docs/paper.tex
Moving experiments to here to generate final versions for paper

--- 2014-06-04 docs/arxiv.tex
@@
-%\documentclass{nature}
+\documentclass{nature}
 \documentclass{article}
@@
 
-\title{Automatic discovery of circuit structure from neural connectomics}
+\todo{“Anatomist cell type” text}
+\todo{Consistent line thickness}
+\todo{on adjacency matrix remove text}
+
+\title{Automatic discovery of cell types and microcircuitry from neural connectomics}
 \author{Eric Jonas$^1$, Konrad Kording$^{2, 3,4}$}
@@
 \begin{abstract}
-  New techniques produce massive data about neural connectivity,
-  necessitating new analysis methods to discover underlying
-  structure. It has long been assumed that discovering the local
-  patterns of microcircuitry is the key to understanding neural
-  function. Here we developed a nonparametric Bayesian technique that
-  identifies neuron types and circuitry patterns in connectomics
-  data. We show that the approach recovers known neuron types in the
-  retina, reveals interesting structure in the nervous system of
-  c. elegans, and automatically discovers the structure of
-  microprocessors. Our approach extracts structural meaning from
-  connectomics, enabling new approaches of deriving anatomical
-  insights from these emerging datasets.
-
+  Connectomics produces massive amounts of data,
+  necessitating new analysis methods to discover the biological and
+  computational structure. It has long been assumed
+  that discovering neuron types and their relation to microcircuitry is crucial to
+  understanding neural function. Here we developed a nonparametric
+  Bayesian technique that identifies neuron types and microcircuitry
+  patterns in connectomics data.
+It combines the information traditionally used by biologists, including connectivity, cell body location and the spatial distribution of synapses.
+We show that the approach recovers known neuron types in the retina and enables predictions of connectivity, better than simpler algorithms.
+It also can reveal interesting structure in
+  the nervous system of c. elegans, and automatically discovers the
+  structure of a microprocessor. 
+ Our approach extracts structural meaning from connectomics, enabling new approaches of automatically deriving  anatomical insights from these emerging datasets.
 \end{abstract}
@@
 another. In fact, the idea of well defined local connectivity patterns
-has been prominent in many areas, ranging from sensory (e.g. retina,
-\autocite{Masland2001} via processing (e.g. neocortex \autocite{Mountcastle1997})
-to movement (e.g. spinal cord) \autocite{Grillner2005}. Indeed, such
-repeated patterns also exists in human-made computing circuits. It
-remains an important challenge to develop algorithms to use anatomical
-data, e.g. connectomics, to back out underlying connectivity motifs.
+(microcircuits) has been prominent in many areas, ranging from sensory
+(e.g. retina, \autocite{Masland2001} via processing (e.g. neocortex
+\autocite{Mountcastle1997}) to movement (e.g. spinal cord)
+\autocite{Grillner2005}. Indeed, such repeated patterns also exists in
+human-made computing circuits. It remains an important challenge to
+develop algorithms to use anatomical data, e.g. connectomics, to automatically back
+out underlying microcircuitry.
 
@@
 approaches can incorporate the probabilistic nature of neural
-connections \autocite{Hill2012} they do not generally consider the
-known fact that the probability of connections between neurons depends
-on distance. Combining such probabilistic models with a spatial
-component to model distance promises to improve the detection of
-microcircuitry from neural data.
+connections \autocite{Hill2012} they do not incorporate the additional
+richer structure present in connectomics data -- the location of cell
+bodies, the spatial distribution of synapses, and the distances
+between neurons. Of particular importance is that the probability of
+connections has a strong spatial component, a factor that is hard to
+reconcile with many other methods. A model attempting to fully capture
+the variation in the nervous system should take into account the broad
+set of available features.
 
@@
 exhibiting a strong spatial dependence arising from the complex
-genetic, chemical, and activity-based neural development process.
+genetic, chemical, and activity-based neural development processes.
 
 Here we describe a Bayesian non-parametric model that can discover
-circuit structure automatically from connectomics data: both the cell
-types and their spatial patterns of interconnection. We apply it to
-three computing systems: the mouse retina connectome
-\autocite{Helmstaedter2013}, the c. elegans connectome
-\autocite{White1986}, and a ``connectome'' of a classical
-microprocessor \autocite{James2010}. Note we do not explicitly
-consider cellular morphology, absolute position (cell layer), or any
-chemical or genetic information. Yet in all cases, we discover
-cell-type to cell-type connectivity that depends on spatial distance
-and these connectivity functions depend on the two involved cell
-types. Comparing the cell types discovered by the algorithms with
-those obtained by anatomists based on orthogonal information reveals a
+circuit structure automatically from connectomics data: the cell
+types, their spatial patterns of interconnection, and the locations of
+somata and synapses. We primarily focus on the recently-released mouse
+retina connectome \autocite{Helmstaedter2013}, but additionally
+examine the c. elegans connectome \autocite{White1986}, and then
+``connectome'' of a classical microprocessor
+\autocite{James2010}. Comparing the cell types discovered by the
+algorithms with those obtained manually by human anatomists reveals a
 high degree of agreement. We present a scalable approach to infer
-microcircuitry from connectivity data.
+microcircuitry from connectomics data.
 
@@
   \centering 
-  \centerline{\includegraphics[width=183mm]{overview.ai}}
-  \caption{An overview of our method. a. Input data is
-a connectivity matrix indicating whether cell $i$ (row)
-synapses on cell $j$ (column). The location of each cell in space is known. 
-b. Our algorithm discovers hidden cell types in this connectivity data
-by assuming all cells of a type share a distance-dependent connectivity 
-profile, c.), with cells of other types. This results in a clustering 
-of the cells by those hidden types. b.) shows the cell connectivity
-matrix with cells of the same type grouped together. d.) shows
-the learned probability of connection between our different types
-at various distances, illustrating short-range and long-range
-connectivity patterns.}
+  \centerline{\includegraphics[width=183mm]{fig1b.ai}}
+  \caption{An overview of our method. a. Input data is a connectivity
+    matrix indicating whether cell $i$ (row) synapses on cell $j$
+    (column), a distance matrix between cell soma, b. the depth of the
+    cell soma, and the c. depth profile of the cells synaptic
+    connectivity. d. Our algorithm discovers hidden cell types in this
+    connectivity data by assuming all cells of a type share a
+    distance-dependent connectivity profile, similar depth, and a
+    similar synpatic density profile, with cells of other types.  This
+    results in a clustering of the cells by those hidden
+    types. d. shows the cell connectivity matrix with cells of the
+    same type grouped together. e.) shows the learned probability of
+    connection between our different types at various distances -- in
+    this case, the cells are likely to connect when they are
+    close. f. shows a connectivity profile with virtually no
+    probability of connectivity. We can plot this these types g.) to
+    show the relationship and spatial connections, and how probable
+    various types are to connect to one another. h. The connectivity
+    between our discovered types changes as a function of distance. }
+
 \label{fig:overview}
@@
 \section{Results}
-We use a Bayesian structure discovery approach that is based on
-prior knowledge about the way positions and cell types affect the
-probability of connections between cells, a so-called generative
-model. We term our model the ``infinite spatial-relational model'' (iSRM), 
-in keeping with the existing nonparametric literature.  
-The probability of connections between pairs of neurons
-decreases with distance, but how high the baseline probability of
-connection is and how quickly it decays with distance depends on the
-two cell-types. Note that this simplistic prior information is all that 
-we tell our model; no additional physiological information is included.
-
-We combine annealing with Markov Chain Monte Carlo to
-perform simultaneously joint posterior inference over the structure,
-parameters, and hyperparameters of the model (see Methods). This
+We start with the generic assumption that cell types exist and affect
+soma depth, distribution of synapses, as well as a cell type and
+distance dependent connection probability. From this assumption we
+develop a Bayesian algorithm that estimates the underlying cell types.
+We start with a model for connectivity, the infinite stochastic block
+model (iSBM)\autocite{Kemp2006a,Xu2006}, which has been shown to
+meaningfully cluster connection graphs. We extend this approach by
+adding distance dependence to model salient aspects of
+microcircuitry. We additionally model cell body depth as a mixture of
+Gaussians and each cell’s synapse density profile as mixture of
+mixtures of Gaussians. For example, retinal ganglion cells may synapse
+on nearby, but not far away, amacrine cells, with ganglion cells being
+superficial and having a broad distribution of synapses.  This
 procedure automatically learns the number of cell types, which cells
-belong to which types, along with the 
-the distance and type dependent connectivity probabilities (see
-fig~\ref{fig:overview}). These results can then be compared with the
-results of previous anatomical studies that have used other aspects of
-neurons, such as chemical properties and cell-morphology to divide
-neurons into types.
+belong to which types, along with the the distance and type dependent
+connectivity probabilities (see fig~\ref{fig:overview}). We apply our
+algorithm to datasets from mouse retina, c-elegans and a historical
+microprocessor. Anatomists classify cells based on many features,
+giving us a meaningful baseline to compare against.
 
@@
 one we used for simulation. We find that the model does a good job of
-recovering the structure we used to generate the data
-(fig~\ref{fig:synthetic}a) Indeed, it does far better than a simpler
-and well established model (the infinite stochastic block model, or
-iSBM) that assumes that there is no distance dependence to the
-connectivity. The model converges relatively quickly to an estimate of
-the most probable values for the cell types, which is enabled by using
-a combination of simulated annealing and parallelized Markov chain
-Monte Carlo (see methods for details). It appears that the model
-behaves the way it should and allows application to relatively large
-datasets.
+recovering the correct number of cell types,
+(fig~\ref{fig:synthetic}a), the the cell identities
+(fig~\ref{fig:synthetic}b), and the spatial extent of each type
+(fig~\ref{fig:synthetic}c).  Indeed, it does far better than the
+existing infinite stochastic block model that assumes that there is no
+distance dependence to the connectivity. The model converges
+relatively quickly to an estimate of the most probable values for the
+cell types, which is enabled by using a combination of simulated
+annealing and parallelized Markov chain Monte Carlo (see methods for
+details). It appears that the model behaves the way it should and
+allows application to relatively large datasets.
+<XXX now add extra features>
 
@@
   \centerline{\includegraphics{synthetic.ai}}
-  \caption{Correct recovery of true numbers of hidden classes
-in synthetic data. a.) Other models over-cluster the number
-of classes as they fail to take distance into account, whereas
-our model finds close to the true number of classes. b.) ARI, a measure
-of ``correct clustering'', for
-different true class counts, between our model and a traditional
-iSBM. c.) For synthetic data, 
-the iSBM overclusters, finding a large number of spatially-localized types. 
-In contrast, the infinite spatial-relational model recovers the true spatial
-extent of the underlying types.}
+  \caption{Correct recovery of true numbers of hidden types in
+    synthetic data when incorporating spatial information. a.) The
+    iSBM (which only uses connectivity information) over-estimates the
+    number of classes as it fails to take distance into account,
+    whereas our modeling of the combination of distance and
+    connectivity finds close to the true number of classes. b.) ARI, a
+    measure of ``correct clustering'', for different true class
+    counts, between our model and a traditional iSBM. c.) For
+    synthetic data, type discovery based only on connectivity are
+    tiny spatially-localized groups -- in
+    contrast, our model recovers the true spatial extent of the
+    underlying types.  }
 \label{fig:synthetic}
@@
   \caption{Discovering cell classes in the mouse retina connectome.
-a.) Input connectivity data for 950 cells for which soma positions were known. b. clustered connectivity matrix. c.) connectivity diagram showing our clusters, as
-well as the cell depth and known cell type of each cell. Our model agrees
-with the cell types traditionally identified by anatomists, and identifies 
-laminar-specific connectivity patterns. d.) Connectivity between our
-clusters as a function of distance -- the cluster consisting primarily of
-retinal ganglion cells (lower left-center) exhibits the expected near and
-far connectivity. e. The spatial distribution of our cell types -- each cell type tesselates space.}
+    a.) Input connectivity data for 950 cells for which soma positions
+    were known. b. clustered connectivity matrix. c.) connectivity
+    diagram showing our clusters, as well as the cell depth and known
+    cell type of each cell. Our model shows moderate agreement with
+    the cell types traditionally identified by anatomist
+    laminar-specific connectivity patterns. \todo{Add more text here}
+    d.) Connectivity between our clusters as a function of distance --
+    the cluster consisting primarily of retinal ganglion cells (lower
+    left-center) exhibits the expected near and far connectivity. e.)
+    The spatial distribution of our cell types -- each cell type
+    tesselates space.}
 \label{fig:mouseretina}
@@
 between different cells, and amacrine cells that feed into the
-ganglion cells \autocite{cite some review}. Recent research
+ganglion cells. Recent research
 \autocite{Helmstaedter2013} has produced a large dataset containing
@@
 
-The algorithm took 8 hours to converge to a locally optimal solution,
+The algorithm took less than 2 hours to converge to a good solution,
 dividing neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
@@
 
-\todo{Do we assign any of the unknown cells correctly?  are types 72-79
-“different” or simply unknown?}
+<xxx discuss the various algorithms and plus and minus 
+   Other model classes
+   brielfly explain link prediction accuracy and ARI
+   show how it compares to human example	
+   the more humanlike the information it gets the more humanlike its classification but many techniques allow good link prediction
+>
 
@@
 cells which bore little relation to known types.  The average spatial
-extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. However,
-in the ground truth by anatomists the average distance was $34.3\pm7.6
-\mu m$ and $35.3\pm 2.4 \mu m$, quite close to the $34.5\pm 1.0 \mu m$
-discovered by our algorithm. Modeling the distance-dependence prevents
+extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. Using
+the anatomists' ground truth, the mean spatial extent of types was
+$34.3\pm7.6 \mu m$, quite close to the $34.5\pm 1.0 \mu m$ discovered
+by our algorithm. Modeling the distance-dependence prevents
 over-segmentation and thus allows more meaningful estimates.
@@
 developmental neuroscience\autocite{White1986}, with the location and
-connectivity of each of 302 neurons deterministically known. This
-incredible cross-organism consistency led to early mapping of the
-c. elegans chemical and electrical (gap junction) connectome. Unlike
-the retina, only the motor neurons in c. elegans exhibit regular
-distribution in space, being distributed along the body axis. Most
-interneurons are concentrated in various ganglia that project
-throughout the entire animal, and the sensory neurons are primarily
-located in a small number of anterior ganglia.
+connectivity of each of 302 neurons developmentally determined,
+leading to early measurement of the connectome. Unlike the retina,
+only the motor neurons in c. elegans exhibit regular distribution in
+space -- along the body axis. Most interneurons are
+concentrated in various ganglia that project throughout the entire
+animal, and the sensory neurons are primarily located in a small
+number of anterior ganglia.
 
@@
   \centering 
-  \centerline{\includegraphics[width=5in]{celegans.ai}}
-  \caption{Discovering cell types in the \textit{c. elegans}
-    connectome. a.) Initial cell connectivity matrix, red are
+  \centerline{\includegraphics[width=7in]{otherdata.ai}}
+  \caption{Discovering connectivity and type in both
+    \textit{c. elegans} and the MOS 6502 microprocessor.  a.) Initial
+    \textit{c. elegans} cell connectivity matrix, red are
     (directional) chemical synapses, blue are electrical
@@
     both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). }
-  \label{fig:celegans}
+    (small text labels). a.) The clusters and connectivity versus
+    distance for connections between Gate and C1, Gate and C2, and C1
+    and C2 terminals on a transistor. Red and teal types have a
+    terminal pulled down to ground and mostly function as
+    inverters. Purple class are clocked, stateful transistors, orange
+    control the ALU and yellow control the special data bus. b.) The
+    spatial distribution of the transistors in each cluster.}
+  \label{fig:otherdata}
 \end{figure}
@@
 microcircuitry our model was designed for.  Motor-neuron types AS, DA,
-and VA, all exclusively postsynaptic, are clustered together, as are
-motor-neuron types VD and DD. Traditional types VC, DB, and VB also
-mostly share a cluster. Various head motor neurons, including types
-SMD, RMD, and RIM are clustered together. Interneurons with known
-anatomically-distinct connectivity patterns, such as RIA (2 cells) or
-AVA (2 cells) are clustered into pure types. The algorithm even
-correctly places the single-cell type DVB by itself. Note our
-clustering does not perfectly reflect known divisions -- several
-combinations of head and sensory neurons are combined, and a
-difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are
-grouped together. Our identified cell types thus reflect a
-``coarsening'' of known types, based entirely on connectivity and
-distance information, even when the organism exhibits substantially
-less spatial regularity than the retina.
+and VA, all exclusively postsynaptic, are identified as a common type,
+as are motor-neuron types VD and DD. Traditional types VC, DB, and VB
+also mostly share a cluster. Various head motor neurons, including
+types SMD and RMD, are clustered together. Interneurons with known
+anatomically-distinct connectivity patterns, such as AVA (2 cells),
+are clustered into pure types.The algorithm even correctly places the
+single-cell types DVB and DVC by themselves.
+
+Note our clustering does not perfectly reflect known divisions --
+several combinations of head and sensory neurons are combined, and a
+difficult-to-explain group of mostly VB and DB motor neuron types,
+with VC split between various groups. Our identified cell types thus
+reflect a ``coarsening'' of known types, based entirely on
+connectivity and distance information, even when the organism exhibits
+substantially less spatial regularity than the retina.
 
 \subsection{Microprocessor}
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics{mos6502.ai}}
-  \caption{Discovering transistor types in the MOS 6502
-    microprocessor. a.) The clusters and connectivity versus distance
-    for connections between Gate and C1, Gate and C2, and C1 and C2
-    terminals on a transistor. Red and teal types have a terminal
-    pulled down to ground and mostly function as inverters. Purple
-    class are clocked, stateful transistors, orange control the ALU
-    and yellow control the special data bus. b.) The spatial
-    distribution of the transistors in each cluster. }
-\label{fig:mos6502}
-\end{figure}
 
@@
 
+TODO FIGURE 5: 
+We know we want the bar graph with ARI and AUC at the same time:
+Conn only
+conn and dist
+conn and dist and synapse profile and depth
+LPCM 
+Just cell depth 
+human dashed line (for link prediction)
+
 \section{Discussion}
 We have presented a machine learning technique that allows cell types
-and microcircuitry to be discovered using only distance and
-connectivity data. We have shown its applicability to a newly
-published connectomics dataset from the mouse retina, the classical
-connectivity of the worm c. elegans, and the connectivity of a
-historical man-made microprocessor. We have found that distance and
-connectivity alone are sufficient to discover cell types and
-microcircuitry that were known to exist in the systems based on decades
-of previous research which had utilized extensive additional
-information.
+and microcircuitry to be discovered from connectomics data.  We have
+shown its applicability to a newly published connectomics dataset from
+the mouse retina. When compared to existing methods, we show how the
+incorporation of all of this data yields results that combine both
+high link-prediction accuracy and high agreement with human
+anatomists.  I also reveals interesting structure when applied to the
+classical connectivity of the worm c. elegans, and the connectivity of
+a historical man-made microprocessor. We have found that combining the
+available data types allows us to discover cell types and
+microcircuitry that were known to exist in the systems based on
+decades of previous research and allows good prediction of
+connectivity.
 
-For probabilistic models like ours, no known solution exists to
+For our probabilistic models, no known solution exists to
 exactly find the most probable parsing of the neurons into cell-types
-and connectivity patterns . We employ a collection of Markov-chain
+and connectivity patterns. We employ a collection of Markov-chain
 Monte carlo techniques (see Methods) but while different
@@
 realistically obtain the global optimum. There are a broad range of
-techniques that may offer better approximations to the global optimum
-\autocite{Wanga2012} and future work could adapt them to
-the problem we addressed here.
+techniques that may offer good approximations to the global optimum
+and future work could adapt them to
+find more precise solutions to our problem.
 
-Our inference becomes slower as the amount of data increases. Our
-algorithm required 30 hours for 1000 neurons and could take months for
-substantially longer datasets. Scaling this class of probabilistic
-model is an active area of research, and recent results in both
-variational methods \autocite{Hoffman2013} and spectral learning
-  \autocite{Anandkumar2012} promise substantial speed increases.
+For our probabilistic model, inference becomes slower as the amount of
+data increases. Our algorithm required several hours for 1000
+neurons. Scaling this class of probabilistic model is an active area
+of research, and recent results in both variational methods
+\autocite{Hoffman2013} and spectral learning \autocite{Anandkumar2012}
+and future work could adapte them to find faster approximate solutions
+to our problem.
 
 Larger datasets will allow algorithms to distinguish more distinct
-  types \todo{Add supplemental data} and we expect closer agreement
-  with existing anotomical knowledge as more data become available.
-  Moreover, in general, for such problems precision increases with the
-  size of the dataset and the cells that we have are not sufficient to
-  statistically distinguish all the cell types known in anatomy (such
-  as the $\sim 70$ in the retina). Still, using only connectivity and
-  distance it is possible to meaningfully divide neurons into types.
+types and we expect closer agreement with existing anatomical
+knowledge as more data become available.  Moreover, in general, for
+such problems precision increases with the size of the dataset and the
+cells that we have are not sufficient to statistically distinguish all
+the cell types known in anatomy (such as the $\sim 70$ in the
+retina). Still, using only connectivity and distance it is possible to
+meaningfully divide neurons into types.
 
@@
 There exist a range of previous approaches to the discovery of
-neural microcircuitry \todo{Do we want a citation here?}. These
+neural microcircuitry \todo{ xxx as many cites as possible}. These
 generally involve a great deal of manual labor and ad-hoc
@@
 can be associated with the types of neurons in another animal, in the
-same way as this is already possible through molecular markers
+same way as this is already possible through molecular marker
 \autocite{Brown2009}. This could be particularly important if cell
@@
 
-Ultimately, we will need to combine information across different
-sources into a joint understanding of neural function. Distinct cell
+Our model combines connectivity, cellular, and synaptic properties, 
+and suggests the way towards combining even richer data. Distinct cell
 types differ in morphology, connectivity, transcriptomics, relation to
-behavior or stimuli and many other ways. Algorithms based on each type
-information can then be combined to synthesize all the available
+behavior or stimuli and many other ways. Algorithms combining this data and type type
+information may allow us to synthesize all the available
 information from one experiment or even across experiments into a
-joint model of brain function.
+joint model of brain structure and function.
+
+Eventually, algorithms will have to become a central tool for
+anatomists, as it will progressively become impossible for humans to
+parse the huge datasets. This transition may follow a similar
+transition to the one underway in molecular biology (with gene-finding
+algorithms) and evolutionary biology with (computational
+phylogenetics).  In fact, it is known that there are significant
+disagreements across human anatomists and computational approaches
+promise to make the debates more data driven.
 
@@
 
-We take as input a connectivity matrix $R$ defining the connections
-between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
-e_j)$ representing a (physical) distance between adjacent cells. See
+For the basic link-distance model, we take as input a connectivity
+matrix $R$ defining the connections between cell $e_i$ and $e_j$, as
+well as a distance function $d(e_i, e_j)$ representing a (physical)
+distance between adjacent cells. \todo{Add the cell soma caveat?}  See
 the supplemental material for extension to multiple connectivity
@@
 We then jointly infer the maximum a posteriori (MAP) estimate of the
-class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
-the global model hyperparameters $\theta$ :
+class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix
+$\eta_{mn}$, and the global model hyperparameters $\theta$ :
 
@@
 \end{equation}
+
+For the retina data, we then extend the model with the additional
+features indicated. Cell soma depth is modeled as a
+cell-type-dependent Gaussian with latent (unknown) per-type mean and
+variance. Similarly, each cell has some number $N_i$ of synapses, 
+each of which is drawn from a cell-type-specific density profile
+with up to three modes.
 
@@
 \section{Methods }
+XXX TODO : add n-way cross validation is not done because it would be a pain
+XXX TODO that said overfitting may be shared across all models
 
@@
 
+% TODO add in the thingie
 \begin{equation}
@@
 
+Where we model cell depth, we assume that 
+each cell type has a typical depth, and thus a Gaussian distribution
+of $s_i$. We assume $s_i \sim N(\mu^{(s)}_k,
+\sigma^{2(s)}_k)$, where the $(s)$ superscript indicates
+these model parameters are associated with the soma-depth 
+portion of our model.  We use a conjugate prior for $(\mu^{(s)}_k, \sigma^{2(s)}_k)$ with
+$\mu^{(s)}_k \sim N(\mu^{(s)}_{hp}, \sigma^{2(s)}_k/ \kappa^{(s)}_{hp})$ and $\sigma^{2(s)}_k \sim
+\chi^{-1}(\sigma^{2(s)}_{hp}, \nu^{(s)}_{hp}$. The use of conjugacy simplifies inference while allowing for each cell-type to have its own depth mean and distribution.
+
+Where we model synapse depth profile, we assume that each cell
+type has a characteristic depth distribution of synaptic contact points, and thus
+a mixture of Gaussians distribution over cell $i$’s $N_i$ contact points, $\vec{g^i}$.
+We do this by assuming the $g^i_j$
+are drawn from an $M=3$-component mixture of Gaussians. Thus associated with each cell type $k$ is a vector
+of $M$ Gaussian means $(\mu^g_{k,1}, \cdots, \mu^g_{k, M})$, and a
+mixture vector $\pi_k$.
+This representation can thus model depth distributions of contact points that have up to three modes, an assumption that is well matched in the bulk of anatomical studies of cell-type dependent connectivity.
+
 \subsection{Inference} 
@@
 To pick the MAP, we run up to $100$ simultaneous chains in parallel,
-initialized from different random initial points in the state space,
-and then pick the chain with the highest log likelihood. Early
-experiments attempted to use the full ensemble of chains as an
-estimate of the posterior but were abandoned due to the difficulty of
-summarizing the posterior over such a complex state space, which is an
-ongoing area of research in the probabilistic modeling community
-\todo{cite more? missing}.
+initialized from different random initial points in the state space. 
+For visualization we pick the chain with the highest log likelihood,
+but for all numerical comparisons (including link probability and ARI)
+we use this full collection of posterior samples to estimate
+the resulting statistics.
 
-\todo{Validation: do we want to write extensively about posterior
-  enumeration tests, etc. ? }
+\subsubsection{Mixing of our Markov chains} 
+Evaluating whether or not approximate inference methods,
+such as MCMC, produce samples which are valid approximations 
+of the posterior distribution is an ongoing area of research
+in the computational statistics community. Here we validate
+mixing in two ways:
+
+\being{itemize}
+\item \textbf{synthetic}: For synthetic datasets of sizes comparable to our
+real data size, do we recover known ground truth information
+after running our markov chains for the appropriate amount of time? 
+\todo{Add figure}
+\item \textbf{autocorrelation} For various summary statistics 
+of the latent space, after a certain amount of time 
+\todo{add figure}
+\end{itemize}
+
+But importantly, regardless of whether our model over- or under-estimates the exact posterior variance about the network, we find points in the latent variable space that are both predictive \textit{and} parsimonious, largely agreeing with the human anatomists and predicting existing connections.
 
@@
 \newpage
-\section{Supplemental}
+\section{Supplemental Material}
 
@@
 
-\href{http://spatialrelation.github.io }{http://spatialrelation.github.io}
+\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/netmotifs/}
+
+Please contact the author for pre-publication access.
 
@@
 \includepdf[pages=-]{mos6502.clusters.pdf}
+
+\newpage
+
+\FloatBarrier
+\subsection{Detailed learned connectivity for mouse retina}
+
+Here we show the recovered distance-dependent connectivty profile
+versus the empirical distance-dependent connectivty for the
+retina dataset. The Logistic-distance Poisson likelihood closely
+matches the empirical likelihood for most classes.
+
+\begin{figure}[h!]
+  \centering 
+    \centerline{\includegraphics[width=7in,page=2]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.0.latent.pdf}}
+  \caption{Learned connectivity vs distance between each of the discovered cell types in the mouse retina.
+The rows are the source type, and the columns are the destination type.
+Blue is the empirical plot of the probability of connection vs distance for those two types; red is the learned function}.
+
+\label{fig:supp:mousecompareari}
+\end{figure}
 

### 2014-06-04 docs/fig1.py

@@ -1,0 +1,301 @@
+import numpy as np
+from matplotlib import pylab
+import irm
+
+def dist(a, b):
+    return np.sqrt(np.sum((b-a)**2))
+
+np.random.seed(0)
+
+SIDE_N = 6
+
+# class_conn = {(0, 1) : ('d', 1.0, 0.7), 
+#               (1, 2) : ('d', 2.0, 0.8), 
+#               (3, 2) : ('p', 0.1), 
+#               (3, 0) : ('d', 1.7, 0.9), 
+#               (1, 3) : ('p', 0.1)}
+
+
+# nodes_with_class, connectivity = irm.data.generate.c_mixed_dist_block(SIDE_N, 
+#                                                                       class_conn, JITTER=0.1, rand_conn_prob=0.01)
+
+
+# class_conn = {(0, 1) : (1.0, 0.8, 0.1), 
+#               (1, 2) : (2.0, 0.4, 0.3), 
+#               (3, 2) : (0.7, 0.9, 0.2), 
+#               (3, 0) : (1.7, 0.7, 0.5), 
+#               (1, 3) : (0.5, 0.5, 0.2)}
+
+
+# nodes_with_class, connectivity = irm.data.generate.c_bump_dist_block(SIDE_N, 
+#                                                                       class_conn, JITTER=0.3, rand_conn_prob=0.01)
+
+
+
+class_conn = {(0, 1) : (0.7, 0.8), 
+              (1, 2) : (2.0, 0.6), 
+              #(3, 2) : (0.7, 0.9), 
+              (2, 1) : (0.7, 0.8), 
+              (0, 3) : (1.7, 0.7), 
+              (3, 4) : (2.4, 0.8), 
+              (4, 0) : (1.0, 0.3)}
+
+
+nodes_with_class, connectivity = irm.data.generate.c_class_neighbors(SIDE_N, 
+                                                                     class_conn, JITTER=0.3, default_param=0.01)
+
+
+
+CELL_N = len(connectivity)
+
+conn_and_dist = np.zeros((CELL_N, CELL_N), 
+                         dtype = [('link', np.uint8), 
+                                  ('distance', np.float32)])
+
+for ni, (ci, posi) in enumerate(nodes_with_class):
+    for nj, (cj, posj) in enumerate(nodes_with_class):
+        conn_and_dist[ni, nj]['link'] = connectivity[ni, nj]
[…]

### 2014-06-04 docs/svg2ai.py

@@ -1,0 +1,45 @@
+import subprocess
+import sys
+import time
+import os
+
+
+ILLUSTRATOR_BIN = r"/Applications/Adobe Illustrator CC/Adobe Illustrator.app/Contents/MacOS/Adobe Illustrator" 
+SCRIPT = "svg2ai.jsx" 
+
+fcount = len(sys.argv[1:])/2
+files = []
+for i in range(fcount):
+    files.append((sys.argv[2*i+1], sys.argv[2*i+2]))
+
+
+fid = file("filenames.txt", 'w')
+for f in files:
+    fid.write("%s %s\n" % f)
+fid.close()
+
+starttime = time.time() 
+
+sp = subprocess.Popen([ILLUSTRATOR_BIN, SCRIPT])
+print "sleeping on execution"
+
+time.sleep(10)
+# quitting -- all this because app.quit segfaults 
+waiting  = True
+while waiting:
+    waiting  = False
+
+    for f_in, f_out in files:
+        if os.path.exists(f_out):
+
+            ftime = os.path.getmtime(f_out)
+            if ftime < starttime:
+                waiting = True
+        else:
+            waiting = True
+
+    time.sleep(2)
+
+sp.terminate()
+
+

### 2014-06-04 predlinks.py

@@ -24,7 +24,7 @@ Hand anatomist
 
 
 
-def compute_prob_matrix(tgt_latent, tgt_data):
+def compute_prob_matrix(tgt_latent, tgt_data, model_name='LogisticDistance'):
     ss = tgt_latent['relations']['R1']['ss']
     ass = tgt_latent['domains']['d1']['assignment']
     hps = tgt_latent['relations']['R1']['hps']
@@ -38,16 +38,21 @@ def compute_prob_matrix(tgt_latent, tgt_data):
             c1 = ass[i]
             c2 = ass[j]
             c = ss[(c1, c2)]
-            dist = data_conn['distance'][i, j]
-            y = irm.util.logistic(dist, c['mu'], c['lambda']) 
-            y = y * (hps['p_max'] - hps['p_min']) + hps['p_min']
+            if model_name == "LogisticDistance":
+                dist = data_conn['distance'][i, j]
+                y = irm.util.logistic(dist, c['mu'], c['lambda']) 
+                y = y * (hps['p_max'] - hps['p_min']) + hps['p_min']
+            elif model_name == "BetaBernoulliNonConj":
+                y = c['p']
+            else:
+                raise NotImplementedError()
             pred[i, j] = y
     return pred
 
 THOLDS = [1, 2, 3]
 PRED_EVALS= np.logspace(-4, 0, 41) # np.linspace(0, 1.0, 41)
 
-@files("../preprocess/mouseretina/mouseretina.db", ['truth.%d.predlinks.pickle' % t for t in THOLDS])
+@files("../preprocess/mouseretina/mouseretina.db", ['truth.%d.ld.predlinks.pickle' % t for t in THOLDS])
 def create_truth(dbfile, outfiles):
     conn = sqlite3.connect(dbfile)
     for THOLD_i, outfile in zip(THOLDS, outfiles):
@@ -83,10 +88,12 @@ def create_truth_bb(dbfile, outfiles):
     for THOLD_i, outfile in zip(THOLDS, outfiles):
         cells, conn_mat, dist_mats = preprocess.create_data(conn, process.THOLDS[THOLD_i])
 
+        irm_latent, irm_data = irm.irmio.default_graph_init(conn_mat,
+                                                        'BetaBernoulliNonConj')
 
[…]

### 2014-06-04 predlinks_plot.py

@@ -15,6 +15,9 @@ clust_df = clust_df[clust_df['filename'].str.contains("retina.%d" % THOLD_IDX)]
 def f(x):
     x['scoreidx'] = np.array(np.argsort(np.argsort(x['score'])[::-1]))
     return x
+
+print clust_df.index.values
+
 clust_df = clust_df.groupby('filename').apply(f )
 
 clust_df['model'] = None
@@ -65,10 +68,28 @@ f = pylab.figure(figsize=(12, 8))
 # f.savefig('pl.comparison.allparam.ari.pdf')
 
 preddf = pickle.load(open("predlinks.pickle", 'r'))
-preddf['srm'] = preddf['filename'].str.contains('srm')
-preddf = preddf[preddf['filename'].str.contains('fixed')]
-preddf['distvar'] = preddf.apply(lambda x : x['filename'].split('.')[5], axis=1)
 
+#print preddf['filename']
+print preddf[preddf['filename'].str.contains("simple_bb")]['filename']
+
+
+
+preddf['srm'] = preddf['filename'].str.contains('srm')
+
+preddf = preddf[preddf['filename'].str.contains('fixed')]
+
+# get rid of the one-off value
+
+#preddf['model'] = preddf[preddf['filename'].str.contains('simple_bb') == False]
+def foo(x):
+    s = x['filename'].split('.')
+    if len(s) > 5:
+        return s[5]
+    else:
+        return ""
+
+preddf['distvar'] = preddf.apply(foo, axis=1)
+print 'here'
 preddf['tp'] = preddf['t_t'] / preddf['t_tot']
 preddf['fp'] = preddf['f_t'] / preddf['f_tot']
 preddf['frac_wrong'] = 1.0 - (preddf['t_t'] + preddf['f_f']) / (preddf['t_tot'] + preddf['f_tot'])
@@ -92,7 +113,7 @@ preddf['filename']=preddf['filename'].apply(lambda x : x[:-(len(".predlinks.pick
 aucsdf = preddf.groupby(['filename', 'chain_i']).apply(g_auc).reset_index()
 aucsdf['chain_i'] = aucsdf['chain_i'].apply(lambda x : int(x))
 
-aucsdf['distvar'] = aucsdf.apply(lambda x : x['filename'].split('.')[5], axis=1)
+aucsdf['distvar'] = aucsdf.apply(foo, axis=1)
[…]

### 2014-06-04 process.py

@@ -463,88 +463,48 @@ def get_results(exp_wait, exp_results):
                 open(exp_results, 'w'))
 
 
-# @transform(get_results, suffix(".samples"), 
-#            [(".%d.clusters.pdf" % d, ".%d.latent.pdf" % d )  for d in range(2)])
-# def plot_best_cluster_latent(exp_results, 
-#                      out_filenames):
-
-#     sample_d = pickle.load(open(exp_results))
-#     chains = sample_d['chains']
+@transform(get_results, suffix(".samples"), 
+           ".latent.pdf")
+def plot_best_cluster_latent(exp_results, 
+                     out_filename):
     
-#     exp = sample_d['exp']
-#     data_filename = exp['data_filename']
-#     data = pickle.load(open(data_filename))
-#     data_basename, _ = os.path.splitext(data_filename)
-#     meta = pickle.load(open(data_basename + ".meta"))
-
-#     meta_infile = meta['infile']
-#     print "meta_infile=", meta_infile
-
-#     d = pickle.load(open(meta_infile, 'r'))
-#     conn = d['dist_matrix']['link']
-#     cell_id_permutation = d['cell_id_permutation']
-
-#     dist_matrix = d['dist_matrix']
-#     orig_data = pickle.load(open(d['infile']))
-#     cell_types = d['types'][:len(conn)]
-
-#     type_metadata_df = pickle.load(open("type_metadata.pickle", 'r'))['type_metadata']
-#     type_color_map = {'gc' : 'r', 
-#                       'ac' : 'b', 
-#                       'bc' : 'g', 
-#                       'other' : 'k'}
-
-#     TYPE_N = np.max(cell_types) + 1
-
-#     type_colors = []
-#     for i in range(TYPE_N):
-#         if (i < 70):
-#             d = type_metadata_df.loc[i+1]['desig']
-#         else:
-#             d = "  "
-#         type_colors.append(type_color_map.get(d[:2], 'k'))
-
-#     print type_colors 
+    sample_d = pickle.load(open(exp_results))
+    chains = sample_d['chains']
     
+    exp = sample_d['exp']
+    data_filename = exp['data_filename']
+    data = pickle.load(open(data_filename))
+    data_basename, _ = os.path.splitext(data_filename)
+    meta = pickle.load(open(data_basename + ".meta"))
 
-#     chains = [c for c in chains if type(c['scores']) != int]
-#     CHAINN = len(chains)
[…]

### 2014-06-04 experiments/mouseretina/pipeline.py
final process code

@@ -1,3 +1,9 @@
+"""
+"""
+import sys
+# total hack
+sys.path.append("../../")
+
 from ruffus import *
 import cPickle as pickle
 import numpy as np
@@ -13,6 +19,7 @@ from sklearn import metrics
 import preprocess
 import sqlite3
 import connattribio 
+import multyvac
 
 import matplotlib.gridspec as gridspec
 import models
@@ -21,6 +28,8 @@ import irm
 import irm.data
 import util
 from irm import rand
+
+
 
 def dist(a, b):
     return np.sqrt(np.sum((b-a)**2))
@@ -32,11 +41,29 @@ def to_f32(x):
     return a
 
 
+VOLUME_NAME = "connattrib_mouseretina_paper"
+MULTYVAC_LAYER = "test11"
+DEFAULT_CORES = 8
+DEFAULT_RELATION = "ParRelation"
+WORKING_DIR = "data"
 
-BUCKET_BASE="connattrib/mouseretina"
+
+@files(None, "volume.%s.sentinel" % VOLUME_NAME)
+def create_volume(infile, outfile):
+    multyvac.volume.create(VOLUME_NAME, '/%s' % VOLUME_NAME)
+    vol = multyvac.volume.get(VOLUME_NAME) 
+    fname = "%s/dir.setinel" % WORKING_DIR
+    fid = file(fname, 'w')
+    fid.write("test")
+    fid.close()
+    vol.put_file(fname, fname)
+   
+    open(outfile, 'w').write("done\n")
+
+
 RETINA_DB = "../preprocess/mouseretina/mouseretina.db"
 
-WORKING_DIR = "data"
+
 
 def td(fname): # "to directory"
     return os.path.join(WORKING_DIR, fname)
@@ -48,7 +75,7 @@ EXPERIMENTS = [
 
 
 
-    ('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
+    #('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
     #('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
 ]
 
@@ -62,24 +89,32 @@ BB_ALPHAS = [1.0]
 BB_BETAS = [1.0]
 
 VAR_SCALES = [0.1, 1.0]
-COMP_KS = [1, 2, 3]
+COMP_KS = [3]
 
-for ti in [1, 2,  3]: # remember to add 2 back in ! 
-    for v in range(len(VAR_SCALES)):
[…]

### 2014-06-04 experiments/mouseretina/process.py
found old process with plotting code, whew

@@ -1,0 +1,1319 @@
+from ruffus import *
+import cPickle as pickle
+import numpy as np
+import copy
+import os, glob, sys
+import time
+from matplotlib import pylab
+import matplotlib
+import pandas
+import colorbrewer
+from nose.tools import assert_equal 
+from sklearn import metrics
+import preprocess
+import sqlite3
+import connattribio 
+
+import matplotlib.gridspec as gridspec
+import models
+
+import irm
+import irm.data
+import util
+from irm import rand
+
+def dist(a, b):
+    return np.sqrt(np.sum((b-a)**2))
+
+
+def to_f32(x):
+    a = np.array(x).astype(np.float32)
+    assert type(a) == np.ndarray
+    return a
+
+
+
+BUCKET_BASE="connattrib/mouseretina"
+RETINA_DB = "../preprocess/mouseretina/mouseretina.db"
+
+WORKING_DIR = "data"
+
+def td(fname): # "to directory"
+    return os.path.join(WORKING_DIR, fname)
+
+EXPERIMENTS = [
+    ('retina.0.ld.0.0.x', 'debug_2_100', 'debug_20'), 
+    ('retina.0.ld.0.0.xyz','debug_2_100', 'debug_20'), 
+    ('retina.0.ld.0.0.yz', 'debug_2_100', 'debug_20'), 
+
+
+
+    ('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
+    #('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
+]
+
+
+THOLDS = [0.01, 0.1, 0.5, 1.0]
+    
+MULAMBS = [1.0, 5.0, 10.0, 20.0, 50.0]
+PMAXS = [0.95, 0.9, 0.7]
+
+BB_ALPHAS = [1.0]
+BB_BETAS = [1.0]
+
+VAR_SCALES = [0.1, 1.0]
+COMP_KS = [1, 2, 3]
+
+for ti in [1, 2,  3]: # remember to add 2 back in ! 
+    for v in range(len(VAR_SCALES)):
+        for k in COMP_KS:
+            EXPERIMENTS.append(('retina.%d.clist.%d.%d' % (ti, v, k) , 
+                                'fixed_20_100', 'anneal_slow_400'))
+            pass
+
+for ti in range(len(THOLDS)):
+    for ml_i in range(len(MULAMBS)):
[…]

### 2014-06-04 experiments/mouseretina/pipeline.py
updates to pipeline

@@ -69,10 +69,7 @@ def td(fname): # "to directory"
     return os.path.join(WORKING_DIR, fname)
 
 EXPERIMENTS = [
-    ('retina.0.ld.0.0.x', 'debug_2_100', 'debug_20'), 
     ('retina.0.ld.0.0.xyz','debug_2_100', 'debug_20'), 
-    ('retina.0.ld.0.0.yz', 'debug_2_100', 'debug_20'), 
-
 
 
     #('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
@@ -105,23 +102,23 @@ COMP_KS = [3]
 #                 bs = 'retina.%d.ld.%d.%d.%s' % (ti, ml_i, pmax_i, vars)
 #                 EXPERIMENTS.append((bs, 'fixed_20_100', 'anneal_slow_400'))
 
-### 
-for ti in [1]: # (len(THOLDS)):
-    for ml_i in [2]:
-        for pmax_i in [0]:
-            for vars in ['xyz']: 
-                bs = 'retina.%d.ld.%d.%d.%s' % (ti, ml_i, pmax_i, vars)
-                EXPERIMENTS.append((bs, 'fixed_20_100', 'anneal_slow_400'))
+
+# for ti in [1]: # (len(THOLDS)):
+#     for ml_i in [2]:
+#         for pmax_i in [0]:
+#             for vars in ['xyz']: 
+#                 bs = 'retina.%d.ld.%d.%d.%s' % (ti, ml_i, pmax_i, vars)
+#                 EXPERIMENTS.append((bs, 'fixed_20_100', 'anneal_slow_400'))
 
 
-for ti in [1]:
-    for ml_i in [3] : # range(len(MULAMBS)):
-        for pmax_i in range(len(PMAXS)):
-            for vars in ['xyz']:
-                for var_scale in range(len(VAR_SCALES)):
-                    for comp_k in COMP_KS:
-                        bs = 'retina.%d.srm_clist_xsoma.%d.%d.%s.%d.%d' % (ti, ml_i, pmax_i, vars, var_scale, comp_k)
-                        EXPERIMENTS.append((bs, 'fixed_20_100', 'anneal_slow_400'))
+# for ti in [1]:
+#     for ml_i in [3] : # range(len(MULAMBS)):
+#         for pmax_i in range(len(PMAXS)):
+#             for vars in ['xyz']:
+#                 for var_scale in range(len(VAR_SCALES)):
[…]

### 2014-06-04 code/connattribio.py
moving all the code around

@@ -1,0 +1,55 @@
+import numpy as np
+import irm
+
+"""
+This is code that presents a mixture model interface
+for testing and other stuff
+
+
+"""
+MODEL_DTYPE = {'BetaBernoulli' : np.uint8, 
+               'NormalInverseChiSq' : np.float32, 
+               'GammaPoisson' : np.uint32, 
+               'MixtureModelDistribution' : irm.models.MixtureModelDistribution().data_dtype()}
+
+
+def create_mm(feature_desc):
+    """
+    feature_desc: {feature_name : {'data' : numpy data, 
+                                   'model' : model'}}
+
+    shared index domain will be 'd1'
+    dummy domain will be d_feature_name
+    relation will be r_feature_name
+    """
+
+    N = len(feature_desc[feature_desc.keys()[0]]['data'])
+    latent = {'domains' : {'d1' : {'hps' : {'alpha' : 1.0}, 
+                                   'assignment' : np.zeros(N, dtype=np.int32) }}, 
+              'relations' : {}}
+    data = {'domains' : {'d1': {'N' : N}}, 
+            'relations' : {}}
+    
+    for feature_name, desc in feature_desc.iteritems():
+        relation_name = "r_%s" % feature_name
+        dummy_domain_name = "d_%s" % feature_name
+
+        # model data sanity check
+        if desc['data'].dtype != MODEL_DTYPE[desc['model']] :
+            raise Exception("Dtype is incorrect for feature %s" % feature_name)
+
+        data['relations'][relation_name] = {'relation' : ('d1', dummy_domain_name), 
+                                            'model' : desc['model'], 
+                                            'data' : desc['data']}
+        latent['domains'][dummy_domain_name] ={'hps' : {'alpha' : 1.0}, 
+                                               'assignment' : [0]}
+        latent['relations'][relation_name] = {}
+        data['domains'][dummy_domain_name] = {'N' : 1}
+
+        if len(desc['data']) != N:
+            raise Exception("Unequal featur length: %s is not of length %d" % (feature_name, N))
+
+
+
+    return latent, data
+                                              

### 2014-06-04 code/features.py

@@ -1,0 +1,213 @@
+import pandas
+import numpy as np
+from ruffus import * 
+import sqlite3
+import cPickle as pickle
+import irm
+
+
+"""
+Create two sets of circos plots based on the features we extract
+
+Features: 
+Soma position in layer
+layer profile of synapse 
+Spatial extent of dendritic arbor [Y/Z]
+
+Fraction of conatcts at that depth ? Number of contacts? 
+
+"""
+
+MAX_CONTACT_AREA = 5.0
+
+BINS = np.linspace(65, 120, 40)
+
+@files("../preprocess/mouseretina/mouseretina.db", "features.pickle")
+def create_features(infile, outfile):
+    """ 
+    input is sqlite table
+    output is data frame with all the features
+
+    """
+
+    # soma depth
+    # dendritic arbor depth histogram area
+    # dendritic arbor depth histogram count
+
+    
+    
+    conn = sqlite3.connect(infile)
+
+
+    cells = pandas.io.sql.read_frame("select c.cell_id, c.type_id, s.x as soma_x, s.z as soma_z, s.y as soma_y, t.coarse as type_coarse from cells as c join somapositions as s on c.cell_id = s.cell_id join types as t on c.type_id = t.type_id", 
+                                     conn, index_col='cell_id')
+
+
+    contacts_df = pandas.io.sql.read_frame("select * from contacts where area < %f and area > %f" % (MAX_CONTACT_AREA, 1.0), 
+                                           conn, index_col='id')
+    
+    # contacts sanity check, make sure there is only ONE way of representing cell A contacts cell B
+    canon_set = set()
+    for from_id, to_id in zip(contacts_df['from_id'], contacts_df['to_id']):
+        if (from_id, to_id) in canon_set:
+            assert (to_id, from_id) not in canon_set
+        canon_set.add((from_id, to_id))
+
+    def f(group):
+        #row = group.irow(0)
+        gc = group.copy()
+        gc['cell_id'] = group['from_id']
+        g2 = group.copy()
+        g2['cell_id'] = group['to_id']
+
+        #new_df = group
+        #return DataFrame({'class': [row['class']] * row['count']})
+        return pandas.concat([gc, g2])
+    contacts_df_sym = contacts_df.groupby('from_id', group_keys=False).apply(f)
+
+
+    def feature_extract(group):
+        od = {}
+        od['contact_x_mean'] = group['x'].mean()
+        od['contact_x_mean_area_weight'] = np.average(group['x'], 
[…]

### 2014-06-04 code/gibbs.py

@@ -1,0 +1,133 @@
+import numpy as np
+from matplotlib import pylab
+import model
+
+# GIBBS SAMPLING
+def die_roll(v):
+    """
+    Take in a vector of probs and roll
+    """
+    x = np.cumsum(v)
+    r = np.random.rand()
+    return np.searchsorted(x, r)
+
+def scores_to_prob(x):
+    """
+    Take in a vector of scores
+    normalize, log-sumpadd, and return
+    
+    """
+    xn = x - np.max(x)
+    a = np.logaddexp.accumulate(xn)[-1]
+    xn = xn - a
+    return np.exp(xn)
+
+def sample_from_scores(scores):
+    return die_roll(scores_to_prob(scores))
+
+
+def gibbs_sample(domain_inf, rng, impotent=False):
+
+    T_N = domain_inf.entity_count()
+
+    if impotent:
+        print "gibbs_sample: IMPOTENT"
+
+    for entity_pos in np.random.permutation(T_N):
+        g = domain_inf.remove_entity_from_group(entity_pos)
+        if domain_inf.group_size(g) == 0:
+            temp_group = g
+        else:
+            temp_group = domain_inf.create_group(rng)
+
+
+        groups = domain_inf.get_groups()
+        scores = np.zeros(len(groups))
+        for gi, group_id in enumerate(groups):
+            scores[gi] = domain_inf.post_pred(group_id, entity_pos)
+        #print entity_pos, scores
+        sample_i = sample_from_scores(scores)
+        new_group = groups[sample_i]
+
+        if impotent:
+            new_group = g
+
+        domain_inf.add_entity_to_group(new_group, entity_pos)
+        if new_group != temp_group:
+            assert domain_inf.group_size(temp_group) == 0
+            domain_inf.delete_group(temp_group)
+
+def gibbs_sample_nonconj(domain_inf, M, rng, impotent=False):
+    """
+    Radford neal Algo 8 for non-conj models
+    
+    M is the number of ephemeral clusters
+    
+    We assume that every cluster in the model is currently used
+    
+    impotent: if true, we always assign the object back to its original
+    cluster. Useful for benchmarking
+    
+    """
+    T_N = domain_inf.entity_count()
+
+    if impotent:
+        print "gibbs_sample_nonconj IMPOTENT"
+
+    if T_N == 1:
+        return # nothing to do 
+
+    for entity_pos in range(T_N):
+        g = domain_inf.remove_entity_from_group(entity_pos)
[…]

### 2014-06-04 code/hopebackup.py

@@ -1,0 +1,1224 @@
+from ruffus import *
+import cPickle as pickle
+import numpy as np
+import copy
+import os, glob
+import time
+from matplotlib import pylab
+import matplotlib
+import pandas
+import colorbrewer
+from nose.tools import assert_equal 
+from sklearn import metrics
+import preprocess
+
+import matplotlib.gridspec as gridspec
+
+import irm
+import irm.data
+import util
+from irm import rand
+
+def dist(a, b):
+    return np.sqrt(np.sum((b-a)**2))
+
+
+import cloud
+
+BUCKET_BASE="connattrib/mouseretina"
+RETINA_DB = "../preprocess/mouseretina/mouseretina.db"
+
+WORKING_DIR = "data"
+
+def td(fname): # "to directory"
+    return os.path.join(WORKING_DIR, fname)
+
+EXPERIMENTS = [
+    ('retina.1.0.ld.0.0', 'fixed_20_100', 'anneal_slow_400'), 
+    ('retina.1.1.ld.0.0', 'fixed_20_100', 'anneal_slow_400'), 
+    ('retina.1.2.ld.0.0', 'fixed_20_100', 'anneal_slow_400'), 
+    ('retina.1.3.ld.0.0', 'fixed_20_100', 'anneal_slow_400'), 
+               
+]
+
+THOLDS = [0.01, 0.1, 0.5, 1.0]
+    
+MULAMBS = [1.0, 5.0, 10.0, 20.0, 50.0]
+PMAXS = [0.95, 0.9, 0.7]
+
+BB_ALPHAS = [1.0]
+BB_BETAS = [1.0]
+
+
+            
+INIT_CONFIGS = {'fixed_20_100' : {'N' : 20, 
+                                  'config' : {'type' : 'fixed', 
+                                              'group_num' : 100}}, 
+}
+
+
+slow_anneal = irm.runner.default_kernel_anneal()
+slow_anneal[0][1]['anneal_sched']['start_temp'] = 64.0
+slow_anneal[0][1]['anneal_sched']['iterations'] = 300
+
+def generate_ld_hypers():
+    space_vals =  irm.util.logspace(1.0, 80.0, 40)
+    p_mins = np.array([0.001, 0.005, 0.01])
+    p_maxs = np.array([0.99, 0.95, 0.90, 0.80])
+    res = []
+    for s in space_vals:
+        for p_min in p_mins:
+            for p_max in p_maxs:
[…]

### 2014-06-04 code/mixmodel.py

@@ -1,0 +1,162 @@
+import numpy as np
+from scipy.special import erf
+import irm
+
+NOT_ASSIGNED = -1
+
+class Feature(object):
+    def __init__(self, data, model):
+        self.components = {}
+        self.assignments = {}
+        self.mod = model
+        self.data = data
+        
+        self.hps = self.mod.create_hp()
+
+    def add_entity_to_group(self, group_id, entity_pos):
+        ss = self.components[group_id]
+        self.mod.add_data(ss, self.data[entity_pos])
+        self.assignments[group_id].add(entity_pos)
+
+    def remove_entity_from_group(self, group_id, entity_pos):
+        ss = self.components[group_id]
+        self.mod.remove_data(ss, self.data[entity_pos])
+        self.assignments[group_id].remove(entity_pos)
+
+    def create_group(self, group_id, rng):
+        self.components[group_id] = self.mod.create_ss(self.hps, rng)
+        self.assignments[group_id] = set()
+
+    def post_pred(self, group_id, entity_pos):
+        ss = self.components[group_id]
+        return self.mod.pred_prob(self.hps, ss, self.data[entity_pos])
+        
+    def delete_group(self, group_id):
+        del self.components[group_id]
+        del self.assignments[group_id]
+
+    def data_prob(self, group_id):
+        ss = self.components[group_id]
+        ds = [self.data[i] for i in self.assignments[group_id]]
+        return self.mod.data_prob(self.hps, ss, ds)
+
+
+class MixtureModel(object):
+    """
+    A single handle that we use to glue objects together
+    Also computes the CRP
+    """
+
+    def __init__(self, ENT_N, features):
+        """
+        features : {'relationname' : feature_obj}
+        where domainname is the name the relation knows this domain as
+
+
+        """
+        self.groups = set()
+        self.g_pos = 0
+        self.assignments = np.ones(ENT_N, dtype=np.int)
+        self.assignments[:] = NOT_ASSIGNED
+        self.temp = 1.0
+
+        self.features = features
+        self.alpha = 1.0
+
+    def entity_count(self):
+        return len(self.assignments)
+
+    def set_hps(self, hps):
+        self.alpha = hps['alpha']
+    
+    def get_hps(self):
+        return {'alpha' : self.alpha}
+
+    def get_groups(self):
+        return list(self.groups)
+
+    def group_count(self):
+        return len(self.groups)
+
+    def set_temp(self, t):
+        self.temp = t
+
[…]

### 2014-06-04 code/model.py

@@ -1,0 +1,458 @@
+import numpy as np
+from scipy.special import erf, gammaln
+import irm
+import copy
+from matplotlib import pylab
+from scipy.stats import chi2, t, norm
+
+
+"""
+Test mixture model of things-that-look-like-distributions
+"""
+
+
+def vec_gauss_prob(bins, mu, std):
+    """
+    Probability mass of N(mu, std^2) in each bin
+
+    """
+    
+    a = erf((bins[1:]-mu)/(std*np.sqrt(2.0))) - erf((bins[:-1]-mu)/(std*np.sqrt(2.0)))
+
+    return a/2.0
+    
+def compute_mm_probs(bins, params):
+    """
+    params are [(pi, mu, sigmasq)]
+    
+    returns len(bins)-1 bins
+    """
+    N = len(bins) - 1
+    K = len(params)
+    out = np.zeros(N, dtype=np.float32)
+    for pi, mu, sigmasq in params:
+        out += pi * vec_gauss_prob(bins, mu, np.sqrt(sigmasq))
+
+    return out
+
+def sample_from_mm(N, params):
+    """
+    Generate N from the params
+    params are [(pi, mu, sigmasq)]
+    """
+    
+    pi_vect = [p[0] for p in params]
+    clust_sizes = np.random.multinomial(N, pi_vect)
+    d = np.zeros(N)
+    pos = 0
+    for ci, c_size in enumerate(clust_sizes):
+        d[pos:pos+c_size] = np.random.normal(params[ci][1], 
+                                             np.sqrt(params[ci][2]), 
+                                             size=c_size)
+        pos += c_size
+    
+    return np.random.permutation(d)
+
+    
+
+fln_table = [0., 0., .69314, 1.79175, 3.17805, 4.78749,
+    6.57925, 8.52516, 10.60460, 12.80182]
+
+def factorialln(x):
+    if x < 10:
+        return fln_table[x]
+    else:
+        return gammaln(x + 1)
+
+def log_multinom_dens(x, p):
+    p_eps = 1e-10
+    l = np.dot(x, np.log(p + p_eps))
+    # normalizing factor
+
+    Z = factorialln(np.sum(x)) - np.sum(gammaln(x + 1))
+    return l + Z
+
+def log_dirichlet_dens(p, alphas):
+    score = 0.0
+    beta_alpha = 0.0
+    for i in range(len(p)):
+        score += (alphas[i] - 1.) * np.log(p[i])
[…]

### 2014-06-04 code/models.py

@@ -1,0 +1,121 @@
+import numpy as np
+import cPickle as pickle
+import irm
+import copy
+
+
+# this is code that goes from a dataset to a model (data, latent, meta)
+
+def create_conn_dist_lowlevel(conn_mat, dist_mats, dist_vars, 
+                              model_name= 'LogisticDistance'):
+
+    CELL_N = len(conn_mat)
+    conn_dist_matrix = np.zeros((CELL_N, CELL_N), 
+                           dtype=[('link', np.uint8), 
+                                  ('distance', np.float32)])
+    conn_dist_matrix['link'] = conn_mat
+    
+    for dist_var in dist_vars:
+        conn_dist_matrix['distance'] += dist_mats[dist_var]**2.0
+    conn_dist_matrix['distance'] = np.sqrt(conn_dist_matrix['distance'])
+
+    irm_latent, irm_data = irm.irmio.default_graph_init(conn_dist_matrix, model_name)
+
+    if model_name == "LogisticDistance":
+        HPS = {'mu_hp' : 10.0,
+               'lambda_hp' : 10.0,
+               'p_min' : 0.05, 
+               'p_max' : 0.90}
+    elif model_name == "LogisticDistanceFixedLambda":
+        HPS = {'lambda' : 20.0,
+               'mu_hp' : 20.0,
+               'p_min' : 0.01, 
+               'p_scale_alpha_hp' : 1.0, 
+               'p_scale_beta_hp' : 1.0}
+
+    else:
+        raise NotImplementedError()
+
+    irm_latent['relations']['R1']['hps'] = HPS
+
+    return irm_latent, irm_data
+
+def create_conn_dist(infile, dist_vars, model_name= 'LogisticDistance'):
+
+    d = pickle.load(open(infile, 'r'))
+    conn_mat = d['conn_mat']
+    dist_mats = d['dist_mats']
+    return create_conn_dist_lowlevel(conn_mat, dist_mats, 
+                                     dist_vars, model_name)
+
+
+def create_mixmodeldata(l, x_min, x_max):
+
+    N = len(l)
+    # convert into a real list of lists
+    contact_x_list = np.zeros(N, dtype=irm.models.MixtureModelDistribution().data_dtype())
+    
+    for xi, x in enumerate(l):
+        if type(x) == float:
+            x = []
+
+        y = np.array(x)
+        if len(y) > 1024:
+            y = np.random.permutation(y)[:1024]
+
[…]

### 2014-06-04 code/pipeline.py

@@ -1,0 +1,249 @@
+import numpy as np
+from ruffus import * 
+import cPickle as pickle
+import connattribio
+import irm
+import features
+
+def to_f32(x):
+    a = np.array(x).astype(np.float32)
+    assert type(a) == np.ndarray
+    return a
+
+def soma_x_hp_grid():
+    GRIDN = 10
+    mu = np.linspace(20, 120, GRIDN+1) 
+    sigmasq = irm.util.logspace(1.0, 10.0, GRIDN)
+    kappa = [0.1, 1.0]
+    nu = irm.util.logspace(10.0, 50.0, GRIDN) 
+    
+    hps = []
+    for m in mu:
+        for s in sigmasq:
+            for k in kappa:
+                for n in nu:
+                    hps.append({'mu' : m, 
+                                'kappa' : k, 
+                                'sigmasq' : s, 
+                                'nu' : n})
+    return hps
+
+@files('features.pickle', "assignment.pickle")
+def run_inference(infile, outfile):
+    data = pickle.load(open(infile, 'r'))
+    df = data['featuredf']
+    df_vals = df[np.isfinite(df['contact_x_mean'])][:100]
+
+    N = len(df_vals)
+
+    desc = {'soma_x' : {'data' : to_f32(df_vals['soma_x']), 
+                        'model' : 'NormalInverseChiSq'}, 
+            # 'contact_spatial_std' : {'data' : to_f32(df_vals['contact_spatial_std']),
+            #                          'model' : 'NormalInverseChiSq'}, 
+    }
+
+    for i, bi in enumerate(features.BINS[:-1]):
+        a = np.array([row['contact_area_hist'][i] for row_i, row in df_vals.iterrows()], 
+                     dtype=np.float32)
+        print a
+        desc['contact_x_hist_%d' % i] =  {'data' : a, 
+                                          'model' : 'NormalInverseChiSq'}
+
+    latent, data = connattribio.create_mm(desc)
+    latent['domains']['d1']['assignment'] = np.arange(N) % 40
+
+    latent, data = irm.data.synth.prior_generate(latent, data)
+    
+    
+    rng = irm.RNG()
+    irm_model = irm.irmio.create_model_from_data(data, rng=rng)
+    irm.irmio.set_model_latent(irm_model, latent, rng)
+    
+    kernel_config = irm.runner.default_kernel_anneal()
[…]

### 2014-06-04 code/predlinks.py

@@ -1,0 +1,305 @@
+import numpy as np
+import numpy as np
+import cPickle as pickle
+import sqlite3
+import preprocess
+import irm
+import models
+import glob
+from ruffus import * 
+import pandas
+import process
+
+"""
+Ground truth check: 
+
+1. Load ground truth assignments
+2. fit BB likelihood
+3. Predict each link 
+4. measure accuracy
+
+
+Hand anatomist 
+"""
+
+
+
+def compute_prob_matrix(tgt_latent, tgt_data, model_name='LogisticDistance'):
+    ss = tgt_latent['relations']['R1']['ss']
+    ass = tgt_latent['domains']['d1']['assignment']
+    hps = tgt_latent['relations']['R1']['hps']
+    data_conn = tgt_data['relations']['R1']['data']
+    
+    N = data_conn.shape[0]
+    pred = np.zeros((N, N))
+    
+    for i in range(N):
+        for j in range(N):
+            c1 = ass[i]
+            c2 = ass[j]
+            c = ss[(c1, c2)]
+            if model_name == "LogisticDistance":
+                dist = data_conn['distance'][i, j]
+                y = irm.util.logistic(dist, c['mu'], c['lambda']) 
+                y = y * (hps['p_max'] - hps['p_min']) + hps['p_min']
+            elif model_name == "BetaBernoulliNonConj":
+                y = c['p']
+            else:
+                raise NotImplementedError()
+            pred[i, j] = y
+    return pred
+
+THOLDS = [1, 2, 3]
+PRED_EVALS= np.logspace(-4, 0, 41) # np.linspace(0, 1.0, 41)
+
+@files("../preprocess/mouseretina/mouseretina.db", ['truth.%d.ld.predlinks.pickle' % t for t in THOLDS])
+def create_truth(dbfile, outfiles):
+    conn = sqlite3.connect(dbfile)
+    for THOLD_i, outfile in zip(THOLDS, outfiles):
+        cells, conn_mat, dist_mats = preprocess.create_data(conn, process.THOLDS[THOLD_i])
+
+
+        irm_latent, irm_data = models.create_conn_dist_lowlevel(conn_mat, dist_mats, 'xyz', model_name="LogisticDistance")
+
+        irm_latent['relations']['R1']['hps'] = {'lambda_hp': 50.0, 'mu_hp': 50.0, 'p_max': 0.9, 'p_min': 0.01}
+
+
[…]

### 2014-06-04 code/predlinks_plot.py

@@ -1,0 +1,130 @@
+import numpy as np
+import cPickle as pickle
+import pandas
+from matplotlib import pylab
+
+
+THOLD_IDX = 1
+
+
+clust_df = pickle.load(open("cluster_metrics.pickle"))['clust_df']
+clust_df = clust_df[ clust_df['filename'].str.contains("debug") == False]
+
+clust_df = clust_df[clust_df['filename'].str.contains("retina.%d" % THOLD_IDX)]
+
+def f(x):
+    x['scoreidx'] = np.array(np.argsort(np.argsort(x['score'])[::-1]))
+    return x
+
+print clust_df.index.values
+
+clust_df = clust_df.groupby('filename').apply(f )
+
+clust_df['model'] = None
+clust_df['model'][clust_df['filename'].str.contains('ld') & clust_df['filename'].str.contains('xyz') ] = 'srm.xyz'
+clust_df['model'][clust_df['filename'].str.contains('.xsoma.')] = 'xsoma'
+clust_df['model'][clust_df['filename'].str.contains('.clist.')] = 'clist'
+clust_df['model'][clust_df['filename'].str.contains('srm_clist_xsoma')]= 'srm_clist_xsoma'
+
+tgtdf = clust_df #[df['filename'].str.contains('retina.2')]
+
+measure = 'ari'
+#ndf = tgtdf[tgtdf['scoreidx']==0]
+b = tgtdf.sort(measure, ascending=False).groupby('model', as_index=False).first()
+
+filenames = b['filename']
+
+
+f = pylab.figure(figsize=(12, 8))
+
+                 
+# bins = np.linspace(0, 0.6, 20)
+# for fi, filename in enumerate(filenames):
+#     ax = f.add_subplot(2, 2, fi+1)
+#     subset = tgtdf[tgtdf['filename'] == filename]
+#     #print subset.head(1)
+#     label = subset.head(1).irow(0)['model']
+#     ax.hist(subset[measure], label=label, bins=bins, normed=True)
+#     ax.axvline(subset[measure].mean(), c='r', linewidth=4)
+#     ax.set_title(label)
+#     ax.set_xlim(bins[0], bins[-1])
+#     ax.grid()
+# f.savefig('pl.comparison.best.ari_coarse.pdf')
+
+
+# f = pylab.figure(figsize=(12, 8))
+
+                 
+# bins = np.linspace(0, 0.6, 40)
[…]

### 2014-06-04 code/preprocess.py

@@ -1,0 +1,120 @@
+import pandas
+import numpy as np
+from ruffus import * 
+import sqlite3
+import cPickle as pickle
+import sklearn.metrics
+
+MAX_CONTACT_AREA = 5.0
+
+# Take in the connectivity matrix and 
+# 1. randomize the ordering of the cells 
+# select the N cells we're going to use and take that data frame
+# return 3 distance matrices
+# return connectivity matrix with real areas to use as subsequent threshold
+
+
+# if we are going to say "above size X is a synapse" then we need to be consistent
+# with it.
+
+# create the python structures we want / need
+
+def create_adj_mat(con, 
+                   area_contact_min, 
+                   cell_data):
+    """
+
+    """
+    
+    df = pandas.io.sql.read_frame("select from_id, to_id, area, sum(area) as contact_area, count(area) as contact_count from contacts  where area < %f  and area > %f group by from_id, to_id" % (MAX_CONTACT_AREA, area_contact_min), con)
+    
+    CELL_N = len(cell_data)
+    id_to_pos = {id: pos for pos, id in enumerate(cell_data.index.values)}
+
+    area_mat = np.zeros((CELL_N, CELL_N), dtype=np.float32)
+
+    for c_i, c_row in df.iterrows():
+        i1 = id_to_pos.get(c_row['from_id'], -1)
+        i2 = id_to_pos.get(c_row['to_id'], -1)
+        if i1 >= 0 and i2 >= 0:
+            area_mat[i1, i2] = c_row['contact_area']
+
+ 
+    return area_mat
+            
+def create_contact_x_lists(con, area_thold_min):
+    """
+    for each cell collapse the list of synapse hists
+    """
+
+    contacts_df = pandas.io.sql.read_frame("select * from contacts where area < %f and area > %f" % (MAX_CONTACT_AREA, area_thold_min), 
+                                           con, index_col='id')
+    
+    # contacts sanity check, make sure there is only ONE way of representing cell A contacts cell B
+    canon_set = set()
+    for from_id, to_id in zip(contacts_df['from_id'], contacts_df['to_id']):
+        if (from_id, to_id) in canon_set:
+            assert (to_id, from_id) not in canon_set
+        canon_set.add((from_id, to_id))
+
+    def f(group):
+        #row = group.irow(0)
+        gc = group.copy()
+        gc['cell_id'] = group['from_id']
+        g2 = group.copy()
[…]

### 2014-06-04 code/process_debug.py

@@ -1,0 +1,1139 @@
+from ruffus import *
+import cPickle as pickle
+import numpy as np
+import copy
+import os, glob, sys
+import time
+from matplotlib import pylab
+import matplotlib
+import pandas
+import colorbrewer
+from nose.tools import assert_equal 
+from sklearn import metrics
+import preprocess
+import sqlite3
+import connattribio 
+
+import matplotlib.gridspec as gridspec
+import models
+
+import irm
+import irm.data
+import util
+from irm import rand
+
+def dist(a, b):
+    return np.sqrt(np.sum((b-a)**2))
+
+
+def to_f32(x):
+    a = np.array(x).astype(np.float32)
+    assert type(a) == np.ndarray
+    return a
+
+
+import cloud
+
+BUCKET_BASE="connattrib/mouseretina_debug"
+RETINA_DB = "../preprocess/mouseretina/mouseretina.db"
+
+WORKING_DIR = "data"
+
+def td(fname): # "to directory"
+    return os.path.join(WORKING_DIR, fname)
+
+EXPERIMENTS = [
+
+]
+
+
+THOLDS = [0.01, 0.1, 0.5, 1.0]
+    
+MULAMBS = [1.0, 5.0, 10.0, 20.0, 50.0]
+PMAXS = [0.95, 0.9, 0.7]
+
+BB_ALPHAS = [1.0]
+BB_BETAS = [1.0]
+
+VAR_SCALES = [0.1, 1.0]
+COMP_KS = [1, 2, 3]
+
+for ti in range(len(THOLDS)):
+    for v in range(len(VAR_SCALES)):
+        for k in COMP_KS:
+            EXPERIMENTS.append(('retina.%d.clistdebug.%d.%d' % (ti, v, k) , 
+                                'fixed_20_100', 'anneal_slow_400'))
+            pass
+
+            
+INIT_CONFIGS = {'fixed_20_100' : {'N' : 20, 
+                                  'config' : {'type' : 'fixed', 
+                                              'group_num' : 100}}, 
+                'debug_2_100' : {'N' : 2, 
+                                 'config' : {'type' : 'fixed', 
+                                             'group_num' : 100}}, 
+            }
+
+
+slow_anneal = irm.runner.default_kernel_anneal()
+slow_anneal[0][1]['anneal_sched']['start_temp'] = 64.0
+slow_anneal[0][1]['anneal_sched']['iterations'] = 300
+
[…]

### 2014-06-04 code/run_mm.py

@@ -1,0 +1,90 @@
+import numpy as np
+from matplotlib import pylab
+import model
+import mixmodel
+import irm
+import gibbs
+
+# range is always [0, 1]
+# synth_comps = [[(1.0, 0.1, 0.1)], 
+#                [(1.0, 0.9, 0.1)]]
+
+np.random.seed(0)
+
+
+synth_comps = [[(0.5, 0.1, 0.05),
+                (0.5, 0.9, 0.05)], 
+               [(0.2, 0.4, 0.05), 
+                (0.8, 0.6, 0.05)]]
+
+
+GROUP_N = len(synth_comps)
+BIN_N = 10
+BINS = np.linspace(0, 1.0, BIN_N + 1)
+BIN_WIDTH = BINS[1] - BINS[0]
+
+# now generate the fake data 
+DP_N = 100
+ENTITIES_PER_GROUP = 50
+ROW_N = ENTITIES_PER_GROUP * GROUP_N
+data = np.zeros((ROW_N, BIN_N), dtype=np.int32)
+
+# now generate the fake data:
+for ci, comp in enumerate(synth_comps):
+    p = model.compute_mm_probs(BINS, comp)
+    p = p / np.sum(p)
+
+    data[ci*ENTITIES_PER_GROUP:(ci+1)*ENTITIES_PER_GROUP] = np.random.multinomial(DP_N, p, size=ENTITIES_PER_GROUP)
+
+    
+# now let's do some fucking inference
+data = np.random.permutation(data)
+
+
+
+MODEL = model.BinnedDist(BIN_N, 2, 0.05)
+f = mixmodel.Feature(data, MODEL)
+
+mm = mixmodel.MixtureModel(ROW_N, {'f1' : f})
+
+rng = None
+# random init
+grp = {}
+for i, g in enumerate(np.random.permutation(np.arange(ROW_N) % 10)):
+    if g not in grp:
+        grp[g] = mm.create_group(rng)
+    mm.add_entity_to_group(grp[g], i)
+    
+print mm.score()
+
+for i in range(100):
+    gibbs.gibbs_sample_nonconj(mm, 20, rng)
+    for group_id, comp in f.components.iteritems():
+        di = list(f.assignments[group_id])
+        
+        new_ss = model.mh_comp(MODEL, f.hps, comp, f.data[di])
+        f.components[group_id] = new_ss
+    
[…]

### 2014-06-04 code/run_mm_cont.py

@@ -1,0 +1,297 @@
+import numpy as np
+from matplotlib import pylab
+import model
+import mixmodel
+import irm
+import gibbs
+import cPickle as pickle
+from sklearn import metrics
+
+from ruffus import * 
+
+# range is always [0, 1]
+# synth_comps = [[(1.0, 0.1, 0.01)], 
+#                [(1.0, 0.9, 0.01)]]
+
+np.random.seed(0)
+
+
+def create_data():
+    synth_comps = [[(0.5, 0.1, 0.01),
+                    (0.5, 0.9, 0.1)], 
+                   [(0.2, 0.4, 0.01), 
+                    (0.8, 0.6, 0.01)], 
+                   # [(0.8, 0.4, 0.01), 
+                   #  (0.2, 0.6, 0.01)], 
+                   # [(0.6, 0.4, 0.01), 
+                   #  (0.2, 0.6, 0.01),
+                   #  (0.2, 0.9, 0.001)],
+                   # [(0.25, 0.2, 0.001), 
+                   #  (0.25, 0.4, 0.001),
+                   #  (0.25, 0.6, 0.001),
+                   #  (0.25, 0.8, 0.001)], 
+               ]
+
+    def gen_synth_comps():
+        N = 20
+        out = []
+        for i in range(N):
+            outparam = []
+            K = np.random.poisson(3) + 1
+            weights = np.random.dirichlet(np.ones(K)*0.5)
+            for k in range(K):
+                outparam.append((weights[k], 
+                                 np.random.rand(), 
+                                 np.random.uniform(0.001, 0.05)))
+            out.append(outparam)
+        return out
+
+    synth_comps = gen_synth_comps()
+    #print synth_comps
+
+    GROUP_N = len(synth_comps)
+
+    # now generate the fake data 
+    ENTITIES_PER_GROUP = 20
+    ROW_N = ENTITIES_PER_GROUP * GROUP_N
+    data = []
+    true_source = []
+    BIN_N = 100
+    BINS = np.linspace(0, 1.0, BIN_N + 1)
+    # now generate the fake data:
+    for ci, comp in enumerate(synth_comps):
+        for ei in range(ENTITIES_PER_GROUP):
+            dp_n = np.random.poisson(250)
+
+            data.append(model.sample_from_mm(dp_n, comp))
+            true_source.append(ci)
+
[…]

### 2014-06-04 code/run_simple_bb.py

@@ -1,0 +1,72 @@
+"""
+The purpose of this is just to run the dang model on the beta-bernoulli
+data
+"""
+
+import numpy as np
+import cPickle as pickle
+import irm
+import copy
+import time
+
+
+
+THOLD_IDX = 1
+
+# load in the retina data
+rd = pickle.load(open("data/retina.%d.data.pickle" % THOLD_IDX, 'r'))
+
+conn_mat = rd['conn_mat']
+dist_mats = rd['dist_mats']
+
+irm_latent, irm_data = irm.irmio.default_graph_init(conn_mat, "BetaBernoulliNonConj")
+
+
+# set the hypers
+
+irm_latent['relations']['R1']['hps'] = {'alpha' : 1.0, 
+                                        'beta' : 1.0}
+
+basename = "data/retina.%d.simple_bb.data" % THOLD_IDX
+pickle.dump(irm_data, open(basename, 'w'))
+
+kernel_config = pickle.load(open('anneal_slow_400.config', 'r'))
+
+
+seed = 0
+
+chain_runner = irm.runner.Runner(irm_latent, irm_data,
+                                 kernel_config, seed)
+ITERS = 400
+
+latent_samp_freq = 50
+scores = []
+times = []
+latents = {}
+def logger(iter, model, logger2):
+    print "Iter", iter
+    scores.append(model.total_score())
+    times.append(time.time())
+    if iter % latent_samp_freq == 0:
+        latents[iter] = chain_runner.get_state(include_ss=False)
+chain_runner.run_iters(ITERS, logger)
+state = chain_runner.get_state()
+
+chains = []
+        
+chains.append({'scores' : scores, 
+               'state' : state, 
+               'times' : times, 
+               'latents' : latents})
+        
+outfilename = "%s-fixed_4-anneal_slow_400.samples" % basename
+
+pickle.dump({'chains' : chains, 
+             'exp' : None}, 
+            open(outfilename, 'w'))
+
+
+# run it for bb-nonconj
+
+
+# call it good

### 2014-06-04 code/test_io.py

@@ -1,0 +1,116 @@
+import numpy as np
+from nose.tools import * 
+from matplotlib import pylab
+
+import connattribio
+import irm
+
+def test_io():
+    N = 10
+    desc = {'f1' : {'data' : np.zeros(N, dtype=np.bool), 
+                    'model' : 'BetaBernoulli'}}
+    
+    latent, data = connattribio.create_mm(desc)
+
+    latent, data = irm.data.synth.prior_generate(latent, data)
+    print data
+    print latent
+
+    assert_equal(len(latent['domains']), 2)
+    assert_equal(len(latent['relations']), 1)
+    assert_equal(len(data['domains']), 2)
+    assert_equal(len(data['relations']), 1)
+    
+    
+    rng = irm.RNG()
+    irm_model = irm.irmio.create_model_from_data(data, rng=rng)
+    irm.irmio.set_model_latent(irm_model, latent, rng)
+
+def test_mixture():
+    N = 100
+
+    np.random.seed(0)
+
+    d = np.zeros(N, dtype=np.float32)
+    for i in range(N/2):
+        d[i] = np.random.normal(-4, 1)
+        d[i+N/2] = np.random.normal(4, 1)
+
+    d = np.random.permutation(d)
+
+    desc = {'f1' : {'data' : d, 
+                    'model' : 'NormalInverseChiSq'}}
+
+    latent, data = connattribio.create_mm(desc)
+    latent['domains']['d1']['assignment'] = np.arange(N) % 10
+
+
+    latent, data = irm.data.synth.prior_generate(latent, data)
+    
+    
+    rng = irm.RNG()
+    irm_model = irm.irmio.create_model_from_data(data, rng=rng)
+    irm.irmio.set_model_latent(irm_model, latent, rng)
+    
+    kernel_config = irm.runner.default_kernel_anneal()
+    
+    for i in range(200):
+        irm.runner.do_inference(irm_model, rng, kernel_config, i)
+
+        new_latent = irm.irmio.get_latent(irm_model)
+        a = new_latent['domains']['d1']['assignment']
+        
+        print irm.util.assign_to_counts(a)
+        print new_latent['relations']['r_f1']['hps']
+
+def test_mixture_bb():
+    ENTITY_PER_GROUP = 50
+    GROUPS = 4
+    N = ENTITY_PER_GROUP * GROUPS
+    DIM = 4 
+
+    np.random.seed(0)
+
+
[…]

### 2014-06-04 code/test_mixmodel.py

@@ -1,0 +1,36 @@
+import numpy as np
+from nose.tools import *
+from matplotlib import pylab
+import mixmodel
+import irm
+import gibbs
+
+def test_mm():
+    np.random.seed(0)
+
+    import distributions
+    N = 100
+
+    data = np.zeros(N, dtype=np.float32)
+    MODEL = distributions.conjugate.nich
+    data[::2] = -1
+    data[1::2] = 1
+    f = mixmodel.Feature(data, MODEL)
+    f.hps = MODEL.HP(0.0, 1.0, 1.0, 1.0)
+
+    mm = mixmodel.MixtureModel(N, {'f1' : f})
+    
+    rng = None
+    # random init
+    grp = {}
+    for i, g in enumerate(np.random.permutation(np.arange(N) % 10)):
+        if g not in grp:
+            grp[g] = mm.create_group(rng)
+        mm.add_entity_to_group(grp[g], i)
+    print mm.score()
+
+    for i in range(10):
+        gibbs.gibbs_sample(mm, rng)
+        print mm.score()
+
+    assert_equal(irm.util.count(mm.get_assignments()).values(), [50, 50])

### 2014-06-04 code/test_model.py

@@ -1,0 +1,119 @@
+import numpy as np
+from nose.tools import *
+from matplotlib import pylab
+import model
+import irm
+import mixmodel
+
+def test_bins():
+    bins = np.linspace(-20, 20, 100)
+
+    mu = 3.0
+    std = 2.0
+    p = model.vec_gauss_prob(bins, mu, std)
+
+    assert_almost_equal(np.sum(p), 1.0, 4)
+
+    samps = np.random.normal(mu, std, size=1000000)
+    h, _ = np.histogram(samps, bins)
+    h = h.astype(float) / np.sum(h)
+    assert irm.util.kl(h, p) < 0.001
+
+    
+    
+def test_compute_mm_probs():
+    bins = np.linspace(-14, 14, 100)
+
+    params = [(0.5, 3.0, 1.0),
+              (0.3, -2.0, 0.5), 
+              (0.2, 0, 3.0)]
+    N = 1000000
+    i = np.random.multinomial(N, [p[0] for p in params])
+    samps = np.zeros(N, dtype=np.float32)
+    pos_i = 0
+    for pos, count in enumerate(i):
+        samps[pos_i:pos_i+count] = np.random.normal(params[pos][1], 
+                                                    np.sqrt(params[pos][2]), 
+                                                    size=count)
+        pos_i += count
+    samps = np.random.permutation(samps)
+
+    h, _ = np.histogram(samps, bins)
+    h = h.astype(float) / np.sum(h)
+
+    p = model.compute_mm_probs(bins, params)
+    assert_almost_equal(np.sum(p), 1.0, 4)
+
+    assert irm.util.kl(h, p) < 0.001
+    
+
+
+def test_mm_mixmodel():
+    np.random.seed(0)
+
+    import distributions
+    N = 100
+
+    data = np.zeros(N, dtype=np.float32)
+    MODEL = model.NonConjGaussian(0.5)
+
+    data[::2] = np.random.normal(-8, 0.5, N/2)
+    data[1::2] = np.random.normal(8, 0.5, N/2)
+    f = mixmodel.Feature(data, MODEL)
+
+    mm = mixmodel.MixtureModel(N, {'f1' : f})
+    
+    rng = None
+    # random init
+    grp = {}
+    for i, g in enumerate(np.random.permutation(np.arange(N) % 10)):
+        if g not in grp:
+            grp[g] = mm.create_group(rng)
[…]

### 2014-06-04 code/testlogs.py

@@ -1,0 +1,32 @@
+import numpy as np
+from matplotlib import pylab
+import time
+
+import irm
+
+N = 1e8
+d = np.arange(1, N, dtype=np.float32)/N
+
+lc = irm.pyirm.LogCompare(d.tolist())
+
+
+t1 = time.time()
+lc.log()
+t2 = time.time()
+print "Log took", t2-t1, "secs"
+
+t1 = time.time()
+lc.logf()
+t2 = time.time()
+print "Logf took", t2-t1, "secs"
+
+
+t1 = time.time()
+lc.fastlog()
+t2 = time.time()
+print "fastlog took", t2-t1, "secs"
+
+t1 = time.time()
+lc.fasterlog()
+t2 = time.time()
+print "fasterlog took", t2-t1, "secs"

### 2014-06-04 experiments/others/others.py
other data

@@ -1,0 +1,78 @@
+import rpy2.robjects as robjects
+import rpy2.robjects.numpy2ri
+from ruffus import * 
+rpy2.robjects.numpy2ri.activate()
+from rpy2.robjects.packages import importr
+vblpcm = importr("VBLPCM")
+network = importr("network")
+
+
+import numpy as np
+from matplotlib import pylab
+import cPickle as pickle
+import pandas
+
+import irm
+
+robjects.r.source("vblpcmroc2.R")
+
+def vblpcm_cluster(adj_mat, G, d, edgecovs = None, STEPS=50, 
+                   maxiter=1000):
+
+    """
+    Variational Bayes LPCM clustering with G groups and d latent dimensions. 
+    no side attributes
+
+    returns assignment vector. 
+
+    hilariously, NOTHING apparently happens with unconnected vertices. 
+
+
+    """
+
+    assert adj_mat.dtype == np.float
+    print "creating network"
+    x_network = network.network(adj_mat)
+
+    if edgecovs == None:
+        print "starting"
+        v_start = vblpcm.vblpcmstart(x_network, G=G, d=d)
+    else:
+        assert edgecovs.shape == adj_mat.shape
+        assert edgecovs.dtype == np.float
+
+        v_start = vblpcm.vblpcmstart(x_network, 
+                                     G=G, d=d, edgecovs=edgecovs)
+    print "Fitting"
+    v_fit = vblpcm.vblpcmfit(v_start, STEPS=STEPS, maxiter=maxiter)
+    print v_fit
+
+    print "done"
+    r_v_lambda = v_fit.rx2('V_lambda')
+    
+    roc = robjects.r.vblpcmroc2(v_fit)
+    conv = v_fit.rx2('conv')[0]
+
+    # well this is fun 
+    robjects.r.assign("x_network", x_network)
+    robjects.r.assign("v_start", v_start)
+    robjects.r.assign("adj_mat", adj_mat)
+    robjects.r.assign("v_fit", v_fit)
+    robjects.r.assign("roc", roc)
+                    
+    robjects.r.save("x_network", "v_start", "adj_mat", "v_fit", "roc",
+                    file=("rdata.%d.%d.rda" % (G, d)))
+    
+    #v_lambda= np.array(r_v_lambda)
+    v_lambda = np.asarray(r_v_lambda)
+
+    v_lambda = np.squeeze(v_lambda)
+
+    assignments = []
+    for i in range(len(adj_mat)):
[…]

### 2014-06-04 experiments/others/rundata.py

@@ -1,0 +1,79 @@
+import numpy as np
+import cPickle as pickle
+import sqlite3
+import sys
+import others
+import pandas
+from ruffus import * 
+
+
+sys.path.append("../connattrib")
+
+import preprocess
+
+
+AREA_THOLD_MIN = 0.1
+dbname = "../preprocess/mouseretina/mouseretina.db"
+
+@files(None, "data/data.pickle")
+def create_data(infile, outfile):
+
+    conn = sqlite3.connect(dbname)
+    cells, conn_mat, dist_mats = preprocess.create_data(conn, AREA_THOLD_MIN)
+    dist_xyz = np.sqrt(dist_mats['x']**2 + dist_mats['y']**2 + dist_mats['z']**2)
+    dist_yz = np.sqrt(dist_mats['y']**2 + dist_mats['z']**2)
+
+    conn_mat = conn_mat.astype(np.float)
+
+    have_edges_i = (conn_mat.sum(axis=1) > 0)
+    conn_mat_have_edges = conn_mat[have_edges_i]
+    conn_mat_have_edges = conn_mat_have_edges[:, have_edges_i]
+
+    pickle.dump({'conn_mat' : conn_mat, 
+                 'cells' : cells, 'dist_mats' : dist_mats, 
+                 'dist_yz' : dist_yz, 
+                 'dist_xyz' : dist_xyz, 
+                 'have_edges_i' : have_edges_i, 
+                 'conn_mat_have_edges' : conn_mat_have_edges}, 
+                open(outfile, 'w'))
+
+def vblpcm_params():
+    for G in [3, 5, 15, 20, 30]:
+        for d in [2, 3, 4, 5]:
+            infile = 'data/data.pickle'
+            outfile = 'data/data.vblpcm.%d.%d.pickle' % (G, d)
+            yield infile, outfile, G, d
+
+@follows(create_data)
+@files(vblpcm_params)
+def vblpcm_cluster(infile, outfile, G, d):
+    data = pickle.load(open(infile, 'r'))
+    conn_mat_have_edges = data['conn_mat_have_edges']
+    
+    res = others.vblpcm_cluster(conn_mat_have_edges, G, d, 
+                                STEPS=1000, maxiter=100)
+
+    pickle.dump(res, 
+                open(outfile, 'w'))
+
+@merge(vblpcm_cluster, 'vblpcm.results.pickle')
+def vblpcm_cluster_merge(infiles, outfile):
+    res = []
+    for f in infiles:
[…]

### 2014-06-04 experiments/others/test.py

@@ -1,0 +1,107 @@
+import rpy2.robjects as robjects
+import rpy2.robjects.numpy2ri
+from ruffus import * 
+rpy2.robjects.numpy2ri.activate()
+from rpy2.robjects.packages import importr
+vblpcm = importr("VBLPCM")
+network = importr("network")
+
+
+import numpy as np
+from matplotlib import pylab
+import cPickle as pickle
+import pandas
+
+import irm
+import others
+
+
+@files(None, 'test.data.pickle')
+def create_data(_, outfile):
+
+    SIDE_N = 5
+    GROUP_N = 4
+    conn = {}
+    for i in range(GROUP_N):
+        conn[(i, i)] = (10.0, 0.1)
+        conn[(i, (i + 1) % GROUP_N)] = (1.0, 0.2)
+        conn[(i, (i + 2) % GROUP_N)] = (1.5, 0.7)
+
+
+    nodes, obs = irm.data.generate.c_class_neighbors(SIDE_N , conn)
+    obs = obs.astype(float)
+    dist_mat = np.zeros(obs.shape, dtype=np.float)
+    # compute the distance matrix
+    for n1_i, n1 in enumerate(nodes):
+        for n2_i, n2 in enumerate(nodes):
+            dist_mat[n1_i, n2_i] = np.sum((nodes[n1_i]['pos'] - nodes[n2_i]['pos'])**2)
+        
+    
+    
+    pickle.dump({'nodes' : nodes, 
+                 'obs' : obs, 
+                 'dist_mat' : dist_mat}, 
+                open(outfile, 'w'))
+    
+@files(create_data, 'test.data.png')
+def plot_data(infile, outfile):
+
+    d = pickle.load(open(infile, 'r'))
+    obs = d['obs']
+
+    f = pylab.figure()
+    ax = f.add_subplot(1, 1, 1)
+    ax.imshow(obs, interpolation='nearest', cmap=pylab.cm.Greys)
+    f.savefig(outfile)
+
+@files(create_data, 'test.output.pickle')
+def run(infile, outfile):
+    d = pickle.load(open(infile, 'r'))
+    obs = d['obs']
+    dist_mat = d['dist_mat']
+    results = []
+    # now try and generate the data from four gaussians
+    for g in range(4, 10):
+        for d in range (2, 5):
+            for dm_i, dist_mat in enumerate([None, 1./dist_mat, dist_mat]):
+                #perm = np.random.permutation(len(obs))
[…]

### 2014-06-04 experiments/others/test_r.py

@@ -1,0 +1,9 @@
+import numpy as np
+import rpy2.robjects as robjects
+import rpy2.robjects.numpy2ri
+
+a = robjects.r.list(10)
+robjects.r.assign("foo", a)
+b = robjects.r.list(10)
+robjects.r.assign("foo2", b)
+robjects.r.save("foo", "foo2", file="test.rda")

### 2014-06-04 connattribio.py
moved original code

@@ -1,55 +1,0 @@
-import numpy as np
-import irm
-
-"""
-This is code that presents a mixture model interface
-for testing and other stuff
-
-
-"""
-MODEL_DTYPE = {'BetaBernoulli' : np.uint8, 
-               'NormalInverseChiSq' : np.float32, 
-               'GammaPoisson' : np.uint32, 
-               'MixtureModelDistribution' : irm.models.MixtureModelDistribution().data_dtype()}
-
-
-def create_mm(feature_desc):
-    """
-    feature_desc: {feature_name : {'data' : numpy data, 
-                                   'model' : model'}}
-
-    shared index domain will be 'd1'
-    dummy domain will be d_feature_name
-    relation will be r_feature_name
-    """
-
-    N = len(feature_desc[feature_desc.keys()[0]]['data'])
-    latent = {'domains' : {'d1' : {'hps' : {'alpha' : 1.0}, 
-                                   'assignment' : np.zeros(N, dtype=np.int32) }}, 
-              'relations' : {}}
-    data = {'domains' : {'d1': {'N' : N}}, 
-            'relations' : {}}
-    
-    for feature_name, desc in feature_desc.iteritems():
-        relation_name = "r_%s" % feature_name
-        dummy_domain_name = "d_%s" % feature_name
-
-        # model data sanity check
-        if desc['data'].dtype != MODEL_DTYPE[desc['model']] :
-            raise Exception("Dtype is incorrect for feature %s" % feature_name)
-
-        data['relations'][relation_name] = {'relation' : ('d1', dummy_domain_name), 
-                                            'model' : desc['model'], 
-                                            'data' : desc['data']}
-        latent['domains'][dummy_domain_name] ={'hps' : {'alpha' : 1.0}, 
-                                               'assignment' : [0]}
-        latent['relations'][relation_name] = {}
-        data['domains'][dummy_domain_name] = {'N' : 1}
-
-        if len(desc['data']) != N:
-            raise Exception("Unequal featur length: %s is not of length %d" % (feature_name, N))
-
-
-
-    return latent, data
-                                              

### 2014-06-04 features.py

@@ -1,213 +1,0 @@
-import pandas
-import numpy as np
-from ruffus import * 
-import sqlite3
-import cPickle as pickle
-import irm
-
-
-"""
-Create two sets of circos plots based on the features we extract
-
-Features: 
-Soma position in layer
-layer profile of synapse 
-Spatial extent of dendritic arbor [Y/Z]
-
-Fraction of conatcts at that depth ? Number of contacts? 
-
-"""
-
-MAX_CONTACT_AREA = 5.0
-
-BINS = np.linspace(65, 120, 40)
-
-@files("../preprocess/mouseretina/mouseretina.db", "features.pickle")
-def create_features(infile, outfile):
-    """ 
-    input is sqlite table
-    output is data frame with all the features
-
-    """
-
-    # soma depth
-    # dendritic arbor depth histogram area
-    # dendritic arbor depth histogram count
-
-    
-    
-    conn = sqlite3.connect(infile)
-
-
-    cells = pandas.io.sql.read_frame("select c.cell_id, c.type_id, s.x as soma_x, s.z as soma_z, s.y as soma_y, t.coarse as type_coarse from cells as c join somapositions as s on c.cell_id = s.cell_id join types as t on c.type_id = t.type_id", 
-                                     conn, index_col='cell_id')
-
-
-    contacts_df = pandas.io.sql.read_frame("select * from contacts where area < %f and area > %f" % (MAX_CONTACT_AREA, 1.0), 
-                                           conn, index_col='id')
-    
-    # contacts sanity check, make sure there is only ONE way of representing cell A contacts cell B
-    canon_set = set()
-    for from_id, to_id in zip(contacts_df['from_id'], contacts_df['to_id']):
-        if (from_id, to_id) in canon_set:
-            assert (to_id, from_id) not in canon_set
-        canon_set.add((from_id, to_id))
-
-    def f(group):
-        #row = group.irow(0)
-        gc = group.copy()
-        gc['cell_id'] = group['from_id']
-        g2 = group.copy()
-        g2['cell_id'] = group['to_id']
-
-        #new_df = group
-        #return DataFrame({'class': [row['class']] * row['count']})
-        return pandas.concat([gc, g2])
-    contacts_df_sym = contacts_df.groupby('from_id', group_keys=False).apply(f)
-
-
-    def feature_extract(group):
-        od = {}
-        od['contact_x_mean'] = group['x'].mean()
-        od['contact_x_mean_area_weight'] = np.average(group['x'], 
[…]

### 2014-06-04 gibbs.py

@@ -1,133 +1,0 @@
-import numpy as np
-from matplotlib import pylab
-import model
-
-# GIBBS SAMPLING
-def die_roll(v):
-    """
-    Take in a vector of probs and roll
-    """
-    x = np.cumsum(v)
-    r = np.random.rand()
-    return np.searchsorted(x, r)
-
-def scores_to_prob(x):
-    """
-    Take in a vector of scores
-    normalize, log-sumpadd, and return
-    
-    """
-    xn = x - np.max(x)
-    a = np.logaddexp.accumulate(xn)[-1]
-    xn = xn - a
-    return np.exp(xn)
-
-def sample_from_scores(scores):
-    return die_roll(scores_to_prob(scores))
-
-
-def gibbs_sample(domain_inf, rng, impotent=False):
-
-    T_N = domain_inf.entity_count()
-
-    if impotent:
-        print "gibbs_sample: IMPOTENT"
-
-    for entity_pos in np.random.permutation(T_N):
-        g = domain_inf.remove_entity_from_group(entity_pos)
-        if domain_inf.group_size(g) == 0:
-            temp_group = g
-        else:
-            temp_group = domain_inf.create_group(rng)
-
-
-        groups = domain_inf.get_groups()
-        scores = np.zeros(len(groups))
-        for gi, group_id in enumerate(groups):
-            scores[gi] = domain_inf.post_pred(group_id, entity_pos)
-        #print entity_pos, scores
-        sample_i = sample_from_scores(scores)
-        new_group = groups[sample_i]
-
-        if impotent:
-            new_group = g
-
-        domain_inf.add_entity_to_group(new_group, entity_pos)
-        if new_group != temp_group:
-            assert domain_inf.group_size(temp_group) == 0
-            domain_inf.delete_group(temp_group)
-
-def gibbs_sample_nonconj(domain_inf, M, rng, impotent=False):
-    """
-    Radford neal Algo 8 for non-conj models
-    
-    M is the number of ephemeral clusters
-    
-    We assume that every cluster in the model is currently used
-    
-    impotent: if true, we always assign the object back to its original
-    cluster. Useful for benchmarking
-    
-    """
-    T_N = domain_inf.entity_count()
-
-    if impotent:
-        print "gibbs_sample_nonconj IMPOTENT"
-
-    if T_N == 1:
-        return # nothing to do 
-
-    for entity_pos in range(T_N):
-        g = domain_inf.remove_entity_from_group(entity_pos)
[…]

### 2014-06-04 mixmodel.py

@@ -1,162 +1,0 @@
-import numpy as np
-from scipy.special import erf
-import irm
-
-NOT_ASSIGNED = -1
-
-class Feature(object):
-    def __init__(self, data, model):
-        self.components = {}
-        self.assignments = {}
-        self.mod = model
-        self.data = data
-        
-        self.hps = self.mod.create_hp()
-
-    def add_entity_to_group(self, group_id, entity_pos):
-        ss = self.components[group_id]
-        self.mod.add_data(ss, self.data[entity_pos])
-        self.assignments[group_id].add(entity_pos)
-
-    def remove_entity_from_group(self, group_id, entity_pos):
-        ss = self.components[group_id]
-        self.mod.remove_data(ss, self.data[entity_pos])
-        self.assignments[group_id].remove(entity_pos)
-
-    def create_group(self, group_id, rng):
-        self.components[group_id] = self.mod.create_ss(self.hps, rng)
-        self.assignments[group_id] = set()
-
-    def post_pred(self, group_id, entity_pos):
-        ss = self.components[group_id]
-        return self.mod.pred_prob(self.hps, ss, self.data[entity_pos])
-        
-    def delete_group(self, group_id):
-        del self.components[group_id]
-        del self.assignments[group_id]
-
-    def data_prob(self, group_id):
-        ss = self.components[group_id]
-        ds = [self.data[i] for i in self.assignments[group_id]]
-        return self.mod.data_prob(self.hps, ss, ds)
-
-
-class MixtureModel(object):
-    """
-    A single handle that we use to glue objects together
-    Also computes the CRP
-    """
-
-    def __init__(self, ENT_N, features):
-        """
-        features : {'relationname' : feature_obj}
-        where domainname is the name the relation knows this domain as
-
-
-        """
-        self.groups = set()
-        self.g_pos = 0
-        self.assignments = np.ones(ENT_N, dtype=np.int)
-        self.assignments[:] = NOT_ASSIGNED
-        self.temp = 1.0
-
-        self.features = features
-        self.alpha = 1.0
-
-    def entity_count(self):
-        return len(self.assignments)
-
-    def set_hps(self, hps):
-        self.alpha = hps['alpha']
-    
-    def get_hps(self):
-        return {'alpha' : self.alpha}
-
-    def get_groups(self):
-        return list(self.groups)
-
-    def group_count(self):
-        return len(self.groups)
-
-    def set_temp(self, t):
-        self.temp = t
-
[…]

### 2014-06-04 model.py

@@ -1,458 +1,0 @@
-import numpy as np
-from scipy.special import erf, gammaln
-import irm
-import copy
-from matplotlib import pylab
-from scipy.stats import chi2, t, norm
-
-
-"""
-Test mixture model of things-that-look-like-distributions
-"""
-
-
-def vec_gauss_prob(bins, mu, std):
-    """
-    Probability mass of N(mu, std^2) in each bin
-
-    """
-    
-    a = erf((bins[1:]-mu)/(std*np.sqrt(2.0))) - erf((bins[:-1]-mu)/(std*np.sqrt(2.0)))
-
-    return a/2.0
-    
-def compute_mm_probs(bins, params):
-    """
-    params are [(pi, mu, sigmasq)]
-    
-    returns len(bins)-1 bins
-    """
-    N = len(bins) - 1
-    K = len(params)
-    out = np.zeros(N, dtype=np.float32)
-    for pi, mu, sigmasq in params:
-        out += pi * vec_gauss_prob(bins, mu, np.sqrt(sigmasq))
-
-    return out
-
-def sample_from_mm(N, params):
-    """
-    Generate N from the params
-    params are [(pi, mu, sigmasq)]
-    """
-    
-    pi_vect = [p[0] for p in params]
-    clust_sizes = np.random.multinomial(N, pi_vect)
-    d = np.zeros(N)
-    pos = 0
-    for ci, c_size in enumerate(clust_sizes):
-        d[pos:pos+c_size] = np.random.normal(params[ci][1], 
-                                             np.sqrt(params[ci][2]), 
-                                             size=c_size)
-        pos += c_size
-    
-    return np.random.permutation(d)
-
-    
-
-fln_table = [0., 0., .69314, 1.79175, 3.17805, 4.78749,
-    6.57925, 8.52516, 10.60460, 12.80182]
-
-def factorialln(x):
-    if x < 10:
-        return fln_table[x]
-    else:
-        return gammaln(x + 1)
-
-def log_multinom_dens(x, p):
-    p_eps = 1e-10
-    l = np.dot(x, np.log(p + p_eps))
-    # normalizing factor
-
-    Z = factorialln(np.sum(x)) - np.sum(gammaln(x + 1))
-    return l + Z
-
-def log_dirichlet_dens(p, alphas):
-    score = 0.0
-    beta_alpha = 0.0
-    for i in range(len(p)):
-        score += (alphas[i] - 1.) * np.log(p[i])
[…]

### 2014-06-04 models.py

@@ -1,121 +1,0 @@
-import numpy as np
-import cPickle as pickle
-import irm
-import copy
-
-
-# this is code that goes from a dataset to a model (data, latent, meta)
-
-def create_conn_dist_lowlevel(conn_mat, dist_mats, dist_vars, 
-                              model_name= 'LogisticDistance'):
-
-    CELL_N = len(conn_mat)
-    conn_dist_matrix = np.zeros((CELL_N, CELL_N), 
-                           dtype=[('link', np.uint8), 
-                                  ('distance', np.float32)])
-    conn_dist_matrix['link'] = conn_mat
-    
-    for dist_var in dist_vars:
-        conn_dist_matrix['distance'] += dist_mats[dist_var]**2.0
-    conn_dist_matrix['distance'] = np.sqrt(conn_dist_matrix['distance'])
-
-    irm_latent, irm_data = irm.irmio.default_graph_init(conn_dist_matrix, model_name)
-
-    if model_name == "LogisticDistance":
-        HPS = {'mu_hp' : 10.0,
-               'lambda_hp' : 10.0,
-               'p_min' : 0.05, 
-               'p_max' : 0.90}
-    elif model_name == "LogisticDistanceFixedLambda":
-        HPS = {'lambda' : 20.0,
-               'mu_hp' : 20.0,
-               'p_min' : 0.01, 
-               'p_scale_alpha_hp' : 1.0, 
-               'p_scale_beta_hp' : 1.0}
-
-    else:
-        raise NotImplementedError()
-
-    irm_latent['relations']['R1']['hps'] = HPS
-
-    return irm_latent, irm_data
-
-def create_conn_dist(infile, dist_vars, model_name= 'LogisticDistance'):
-
-    d = pickle.load(open(infile, 'r'))
-    conn_mat = d['conn_mat']
-    dist_mats = d['dist_mats']
-    return create_conn_dist_lowlevel(conn_mat, dist_mats, 
-                                     dist_vars, model_name)
-
-
-def create_mixmodeldata(l, x_min, x_max):
-
-    N = len(l)
-    # convert into a real list of lists
-    contact_x_list = np.zeros(N, dtype=irm.models.MixtureModelDistribution().data_dtype())
-    
-    for xi, x in enumerate(l):
-        if type(x) == float:
-            x = []
-
-        y = np.array(x)
-        if len(y) > 1024:
-            y = np.random.permutation(y)[:1024]
-
[…]

### 2014-06-04 pipeline.py

@@ -1,249 +1,0 @@
-import numpy as np
-from ruffus import * 
-import cPickle as pickle
-import connattribio
-import irm
-import features
-
-def to_f32(x):
-    a = np.array(x).astype(np.float32)
-    assert type(a) == np.ndarray
-    return a
-
-def soma_x_hp_grid():
-    GRIDN = 10
-    mu = np.linspace(20, 120, GRIDN+1) 
-    sigmasq = irm.util.logspace(1.0, 10.0, GRIDN)
-    kappa = [0.1, 1.0]
-    nu = irm.util.logspace(10.0, 50.0, GRIDN) 
-    
-    hps = []
-    for m in mu:
-        for s in sigmasq:
-            for k in kappa:
-                for n in nu:
-                    hps.append({'mu' : m, 
-                                'kappa' : k, 
-                                'sigmasq' : s, 
-                                'nu' : n})
-    return hps
-
-@files('features.pickle', "assignment.pickle")
-def run_inference(infile, outfile):
-    data = pickle.load(open(infile, 'r'))
-    df = data['featuredf']
-    df_vals = df[np.isfinite(df['contact_x_mean'])][:100]
-
-    N = len(df_vals)
-
-    desc = {'soma_x' : {'data' : to_f32(df_vals['soma_x']), 
-                        'model' : 'NormalInverseChiSq'}, 
-            # 'contact_spatial_std' : {'data' : to_f32(df_vals['contact_spatial_std']),
-            #                          'model' : 'NormalInverseChiSq'}, 
-    }
-
-    for i, bi in enumerate(features.BINS[:-1]):
-        a = np.array([row['contact_area_hist'][i] for row_i, row in df_vals.iterrows()], 
-                     dtype=np.float32)
-        print a
-        desc['contact_x_hist_%d' % i] =  {'data' : a, 
-                                          'model' : 'NormalInverseChiSq'}
-
-    latent, data = connattribio.create_mm(desc)
-    latent['domains']['d1']['assignment'] = np.arange(N) % 40
-
-    latent, data = irm.data.synth.prior_generate(latent, data)
-    
-    
-    rng = irm.RNG()
-    irm_model = irm.irmio.create_model_from_data(data, rng=rng)
-    irm.irmio.set_model_latent(irm_model, latent, rng)
-    
-    kernel_config = irm.runner.default_kernel_anneal()
[…]

### 2014-06-04 predlinks.py

@@ -1,305 +1,0 @@
-import numpy as np
-import numpy as np
-import cPickle as pickle
-import sqlite3
-import preprocess
-import irm
-import models
-import glob
-from ruffus import * 
-import pandas
-import process
-
-"""
-Ground truth check: 
-
-1. Load ground truth assignments
-2. fit BB likelihood
-3. Predict each link 
-4. measure accuracy
-
-
-Hand anatomist 
-"""
-
-
-
-def compute_prob_matrix(tgt_latent, tgt_data, model_name='LogisticDistance'):
-    ss = tgt_latent['relations']['R1']['ss']
-    ass = tgt_latent['domains']['d1']['assignment']
-    hps = tgt_latent['relations']['R1']['hps']
-    data_conn = tgt_data['relations']['R1']['data']
-    
-    N = data_conn.shape[0]
-    pred = np.zeros((N, N))
-    
-    for i in range(N):
-        for j in range(N):
-            c1 = ass[i]
-            c2 = ass[j]
-            c = ss[(c1, c2)]
-            if model_name == "LogisticDistance":
-                dist = data_conn['distance'][i, j]
-                y = irm.util.logistic(dist, c['mu'], c['lambda']) 
-                y = y * (hps['p_max'] - hps['p_min']) + hps['p_min']
-            elif model_name == "BetaBernoulliNonConj":
-                y = c['p']
-            else:
-                raise NotImplementedError()
-            pred[i, j] = y
-    return pred
-
-THOLDS = [1, 2, 3]
-PRED_EVALS= np.logspace(-4, 0, 41) # np.linspace(0, 1.0, 41)
-
-@files("../preprocess/mouseretina/mouseretina.db", ['truth.%d.ld.predlinks.pickle' % t for t in THOLDS])
-def create_truth(dbfile, outfiles):
-    conn = sqlite3.connect(dbfile)
-    for THOLD_i, outfile in zip(THOLDS, outfiles):
-        cells, conn_mat, dist_mats = preprocess.create_data(conn, process.THOLDS[THOLD_i])
-
-
-        irm_latent, irm_data = models.create_conn_dist_lowlevel(conn_mat, dist_mats, 'xyz', model_name="LogisticDistance")
-
-        irm_latent['relations']['R1']['hps'] = {'lambda_hp': 50.0, 'mu_hp': 50.0, 'p_max': 0.9, 'p_min': 0.01}
-
-
[…]

### 2014-06-04 predlinks_plot.py

@@ -1,130 +1,0 @@
-import numpy as np
-import cPickle as pickle
-import pandas
-from matplotlib import pylab
-
-
-THOLD_IDX = 1
-
-
-clust_df = pickle.load(open("cluster_metrics.pickle"))['clust_df']
-clust_df = clust_df[ clust_df['filename'].str.contains("debug") == False]
-
-clust_df = clust_df[clust_df['filename'].str.contains("retina.%d" % THOLD_IDX)]
-
-def f(x):
-    x['scoreidx'] = np.array(np.argsort(np.argsort(x['score'])[::-1]))
-    return x
-
-print clust_df.index.values
-
-clust_df = clust_df.groupby('filename').apply(f )
-
-clust_df['model'] = None
-clust_df['model'][clust_df['filename'].str.contains('ld') & clust_df['filename'].str.contains('xyz') ] = 'srm.xyz'
-clust_df['model'][clust_df['filename'].str.contains('.xsoma.')] = 'xsoma'
-clust_df['model'][clust_df['filename'].str.contains('.clist.')] = 'clist'
-clust_df['model'][clust_df['filename'].str.contains('srm_clist_xsoma')]= 'srm_clist_xsoma'
-
-tgtdf = clust_df #[df['filename'].str.contains('retina.2')]
-
-measure = 'ari'
-#ndf = tgtdf[tgtdf['scoreidx']==0]
-b = tgtdf.sort(measure, ascending=False).groupby('model', as_index=False).first()
-
-filenames = b['filename']
-
-
-f = pylab.figure(figsize=(12, 8))
-
-                 
-# bins = np.linspace(0, 0.6, 20)
-# for fi, filename in enumerate(filenames):
-#     ax = f.add_subplot(2, 2, fi+1)
-#     subset = tgtdf[tgtdf['filename'] == filename]
-#     #print subset.head(1)
-#     label = subset.head(1).irow(0)['model']
-#     ax.hist(subset[measure], label=label, bins=bins, normed=True)
-#     ax.axvline(subset[measure].mean(), c='r', linewidth=4)
-#     ax.set_title(label)
-#     ax.set_xlim(bins[0], bins[-1])
-#     ax.grid()
-# f.savefig('pl.comparison.best.ari_coarse.pdf')
-
-
-# f = pylab.figure(figsize=(12, 8))
-
-                 
-# bins = np.linspace(0, 0.6, 40)
[…]

### 2014-06-04 preprocess.py

@@ -1,120 +1,0 @@
-import pandas
-import numpy as np
-from ruffus import * 
-import sqlite3
-import cPickle as pickle
-import sklearn.metrics
-
-MAX_CONTACT_AREA = 5.0
-
-# Take in the connectivity matrix and 
-# 1. randomize the ordering of the cells 
-# select the N cells we're going to use and take that data frame
-# return 3 distance matrices
-# return connectivity matrix with real areas to use as subsequent threshold
-
-
-# if we are going to say "above size X is a synapse" then we need to be consistent
-# with it.
-
-# create the python structures we want / need
-
-def create_adj_mat(con, 
-                   area_contact_min, 
-                   cell_data):
-    """
-
-    """
-    
-    df = pandas.io.sql.read_frame("select from_id, to_id, area, sum(area) as contact_area, count(area) as contact_count from contacts  where area < %f  and area > %f group by from_id, to_id" % (MAX_CONTACT_AREA, area_contact_min), con)
-    
-    CELL_N = len(cell_data)
-    id_to_pos = {id: pos for pos, id in enumerate(cell_data.index.values)}
-
-    area_mat = np.zeros((CELL_N, CELL_N), dtype=np.float32)
-
-    for c_i, c_row in df.iterrows():
-        i1 = id_to_pos.get(c_row['from_id'], -1)
-        i2 = id_to_pos.get(c_row['to_id'], -1)
-        if i1 >= 0 and i2 >= 0:
-            area_mat[i1, i2] = c_row['contact_area']
-
- 
-    return area_mat
-            
-def create_contact_x_lists(con, area_thold_min):
-    """
-    for each cell collapse the list of synapse hists
-    """
-
-    contacts_df = pandas.io.sql.read_frame("select * from contacts where area < %f and area > %f" % (MAX_CONTACT_AREA, area_thold_min), 
-                                           con, index_col='id')
-    
-    # contacts sanity check, make sure there is only ONE way of representing cell A contacts cell B
-    canon_set = set()
-    for from_id, to_id in zip(contacts_df['from_id'], contacts_df['to_id']):
-        if (from_id, to_id) in canon_set:
-            assert (to_id, from_id) not in canon_set
-        canon_set.add((from_id, to_id))
-
-    def f(group):
-        #row = group.irow(0)
-        gc = group.copy()
-        gc['cell_id'] = group['from_id']
-        g2 = group.copy()
[…]

### 2014-06-04 process_debug.py

@@ -1,1139 +1,0 @@
-from ruffus import *
-import cPickle as pickle
-import numpy as np
-import copy
-import os, glob, sys
-import time
-from matplotlib import pylab
-import matplotlib
-import pandas
-import colorbrewer
-from nose.tools import assert_equal 
-from sklearn import metrics
-import preprocess
-import sqlite3
-import connattribio 
-
-import matplotlib.gridspec as gridspec
-import models
-
-import irm
-import irm.data
-import util
-from irm import rand
-
-def dist(a, b):
-    return np.sqrt(np.sum((b-a)**2))
-
-
-def to_f32(x):
-    a = np.array(x).astype(np.float32)
-    assert type(a) == np.ndarray
-    return a
-
-
-import cloud
-
-BUCKET_BASE="connattrib/mouseretina_debug"
-RETINA_DB = "../preprocess/mouseretina/mouseretina.db"
-
-WORKING_DIR = "data"
-
-def td(fname): # "to directory"
-    return os.path.join(WORKING_DIR, fname)
-
-EXPERIMENTS = [
-
-]
-
-
-THOLDS = [0.01, 0.1, 0.5, 1.0]
-    
-MULAMBS = [1.0, 5.0, 10.0, 20.0, 50.0]
-PMAXS = [0.95, 0.9, 0.7]
-
-BB_ALPHAS = [1.0]
-BB_BETAS = [1.0]
-
-VAR_SCALES = [0.1, 1.0]
-COMP_KS = [1, 2, 3]
-
-for ti in range(len(THOLDS)):
-    for v in range(len(VAR_SCALES)):
-        for k in COMP_KS:
-            EXPERIMENTS.append(('retina.%d.clistdebug.%d.%d' % (ti, v, k) , 
-                                'fixed_20_100', 'anneal_slow_400'))
-            pass
-
-            
-INIT_CONFIGS = {'fixed_20_100' : {'N' : 20, 
-                                  'config' : {'type' : 'fixed', 
-                                              'group_num' : 100}}, 
-                'debug_2_100' : {'N' : 2, 
-                                 'config' : {'type' : 'fixed', 
-                                             'group_num' : 100}}, 
-            }
-
-
-slow_anneal = irm.runner.default_kernel_anneal()
-slow_anneal[0][1]['anneal_sched']['start_temp'] = 64.0
-slow_anneal[0][1]['anneal_sched']['iterations'] = 300
-
[…]

### 2014-06-04 run_mm.py

@@ -1,90 +1,0 @@
-import numpy as np
-from matplotlib import pylab
-import model
-import mixmodel
-import irm
-import gibbs
-
-# range is always [0, 1]
-# synth_comps = [[(1.0, 0.1, 0.1)], 
-#                [(1.0, 0.9, 0.1)]]
-
-np.random.seed(0)
-
-
-synth_comps = [[(0.5, 0.1, 0.05),
-                (0.5, 0.9, 0.05)], 
-               [(0.2, 0.4, 0.05), 
-                (0.8, 0.6, 0.05)]]
-
-
-GROUP_N = len(synth_comps)
-BIN_N = 10
-BINS = np.linspace(0, 1.0, BIN_N + 1)
-BIN_WIDTH = BINS[1] - BINS[0]
-
-# now generate the fake data 
-DP_N = 100
-ENTITIES_PER_GROUP = 50
-ROW_N = ENTITIES_PER_GROUP * GROUP_N
-data = np.zeros((ROW_N, BIN_N), dtype=np.int32)
-
-# now generate the fake data:
-for ci, comp in enumerate(synth_comps):
-    p = model.compute_mm_probs(BINS, comp)
-    p = p / np.sum(p)
-
-    data[ci*ENTITIES_PER_GROUP:(ci+1)*ENTITIES_PER_GROUP] = np.random.multinomial(DP_N, p, size=ENTITIES_PER_GROUP)
-
-    
-# now let's do some fucking inference
-data = np.random.permutation(data)
-
-
-
-MODEL = model.BinnedDist(BIN_N, 2, 0.05)
-f = mixmodel.Feature(data, MODEL)
-
-mm = mixmodel.MixtureModel(ROW_N, {'f1' : f})
-
-rng = None
-# random init
-grp = {}
-for i, g in enumerate(np.random.permutation(np.arange(ROW_N) % 10)):
-    if g not in grp:
-        grp[g] = mm.create_group(rng)
-    mm.add_entity_to_group(grp[g], i)
-    
-print mm.score()
-
-for i in range(100):
-    gibbs.gibbs_sample_nonconj(mm, 20, rng)
-    for group_id, comp in f.components.iteritems():
-        di = list(f.assignments[group_id])
-        
-        new_ss = model.mh_comp(MODEL, f.hps, comp, f.data[di])
-        f.components[group_id] = new_ss
-    
[…]

### 2014-06-04 run_mm_cont.py

@@ -1,297 +1,0 @@
-import numpy as np
-from matplotlib import pylab
-import model
-import mixmodel
-import irm
-import gibbs
-import cPickle as pickle
-from sklearn import metrics
-
-from ruffus import * 
-
-# range is always [0, 1]
-# synth_comps = [[(1.0, 0.1, 0.01)], 
-#                [(1.0, 0.9, 0.01)]]
-
-np.random.seed(0)
-
-
-def create_data():
-    synth_comps = [[(0.5, 0.1, 0.01),
-                    (0.5, 0.9, 0.1)], 
-                   [(0.2, 0.4, 0.01), 
-                    (0.8, 0.6, 0.01)], 
-                   # [(0.8, 0.4, 0.01), 
-                   #  (0.2, 0.6, 0.01)], 
-                   # [(0.6, 0.4, 0.01), 
-                   #  (0.2, 0.6, 0.01),
-                   #  (0.2, 0.9, 0.001)],
-                   # [(0.25, 0.2, 0.001), 
-                   #  (0.25, 0.4, 0.001),
-                   #  (0.25, 0.6, 0.001),
-                   #  (0.25, 0.8, 0.001)], 
-               ]
-
-    def gen_synth_comps():
-        N = 20
-        out = []
-        for i in range(N):
-            outparam = []
-            K = np.random.poisson(3) + 1
-            weights = np.random.dirichlet(np.ones(K)*0.5)
-            for k in range(K):
-                outparam.append((weights[k], 
-                                 np.random.rand(), 
-                                 np.random.uniform(0.001, 0.05)))
-            out.append(outparam)
-        return out
-
-    synth_comps = gen_synth_comps()
-    #print synth_comps
-
-    GROUP_N = len(synth_comps)
-
-    # now generate the fake data 
-    ENTITIES_PER_GROUP = 20
-    ROW_N = ENTITIES_PER_GROUP * GROUP_N
-    data = []
-    true_source = []
-    BIN_N = 100
-    BINS = np.linspace(0, 1.0, BIN_N + 1)
-    # now generate the fake data:
-    for ci, comp in enumerate(synth_comps):
-        for ei in range(ENTITIES_PER_GROUP):
-            dp_n = np.random.poisson(250)
-
-            data.append(model.sample_from_mm(dp_n, comp))
-            true_source.append(ci)
-
[…]

### 2014-06-04 test_io.py

@@ -1,116 +1,0 @@
-import numpy as np
-from nose.tools import * 
-from matplotlib import pylab
-
-import connattribio
-import irm
-
-def test_io():
-    N = 10
-    desc = {'f1' : {'data' : np.zeros(N, dtype=np.bool), 
-                    'model' : 'BetaBernoulli'}}
-    
-    latent, data = connattribio.create_mm(desc)
-
-    latent, data = irm.data.synth.prior_generate(latent, data)
-    print data
-    print latent
-
-    assert_equal(len(latent['domains']), 2)
-    assert_equal(len(latent['relations']), 1)
-    assert_equal(len(data['domains']), 2)
-    assert_equal(len(data['relations']), 1)
-    
-    
-    rng = irm.RNG()
-    irm_model = irm.irmio.create_model_from_data(data, rng=rng)
-    irm.irmio.set_model_latent(irm_model, latent, rng)
-
-def test_mixture():
-    N = 100
-
-    np.random.seed(0)
-
-    d = np.zeros(N, dtype=np.float32)
-    for i in range(N/2):
-        d[i] = np.random.normal(-4, 1)
-        d[i+N/2] = np.random.normal(4, 1)
-
-    d = np.random.permutation(d)
-
-    desc = {'f1' : {'data' : d, 
-                    'model' : 'NormalInverseChiSq'}}
-
-    latent, data = connattribio.create_mm(desc)
-    latent['domains']['d1']['assignment'] = np.arange(N) % 10
-
-
-    latent, data = irm.data.synth.prior_generate(latent, data)
-    
-    
-    rng = irm.RNG()
-    irm_model = irm.irmio.create_model_from_data(data, rng=rng)
-    irm.irmio.set_model_latent(irm_model, latent, rng)
-    
-    kernel_config = irm.runner.default_kernel_anneal()
-    
-    for i in range(200):
-        irm.runner.do_inference(irm_model, rng, kernel_config, i)
-
-        new_latent = irm.irmio.get_latent(irm_model)
-        a = new_latent['domains']['d1']['assignment']
-        
-        print irm.util.assign_to_counts(a)
-        print new_latent['relations']['r_f1']['hps']
-
-def test_mixture_bb():
-    ENTITY_PER_GROUP = 50
-    GROUPS = 4
-    N = ENTITY_PER_GROUP * GROUPS
-    DIM = 4 
-
-    np.random.seed(0)
-
-
[…]

### 2014-06-04 test_mixmodel.py

@@ -1,36 +1,0 @@
-import numpy as np
-from nose.tools import *
-from matplotlib import pylab
-import mixmodel
-import irm
-import gibbs
-
-def test_mm():
-    np.random.seed(0)
-
-    import distributions
-    N = 100
-
-    data = np.zeros(N, dtype=np.float32)
-    MODEL = distributions.conjugate.nich
-    data[::2] = -1
-    data[1::2] = 1
-    f = mixmodel.Feature(data, MODEL)
-    f.hps = MODEL.HP(0.0, 1.0, 1.0, 1.0)
-
-    mm = mixmodel.MixtureModel(N, {'f1' : f})
-    
-    rng = None
-    # random init
-    grp = {}
-    for i, g in enumerate(np.random.permutation(np.arange(N) % 10)):
-        if g not in grp:
-            grp[g] = mm.create_group(rng)
-        mm.add_entity_to_group(grp[g], i)
-    print mm.score()
-
-    for i in range(10):
-        gibbs.gibbs_sample(mm, rng)
-        print mm.score()
-
-    assert_equal(irm.util.count(mm.get_assignments()).values(), [50, 50])

### 2014-06-04 test_model.py

@@ -1,119 +1,0 @@
-import numpy as np
-from nose.tools import *
-from matplotlib import pylab
-import model
-import irm
-import mixmodel
-
-def test_bins():
-    bins = np.linspace(-20, 20, 100)
-
-    mu = 3.0
-    std = 2.0
-    p = model.vec_gauss_prob(bins, mu, std)
-
-    assert_almost_equal(np.sum(p), 1.0, 4)
-
-    samps = np.random.normal(mu, std, size=1000000)
-    h, _ = np.histogram(samps, bins)
-    h = h.astype(float) / np.sum(h)
-    assert irm.util.kl(h, p) < 0.001
-
-    
-    
-def test_compute_mm_probs():
-    bins = np.linspace(-14, 14, 100)
-
-    params = [(0.5, 3.0, 1.0),
-              (0.3, -2.0, 0.5), 
-              (0.2, 0, 3.0)]
-    N = 1000000
-    i = np.random.multinomial(N, [p[0] for p in params])
-    samps = np.zeros(N, dtype=np.float32)
-    pos_i = 0
-    for pos, count in enumerate(i):
-        samps[pos_i:pos_i+count] = np.random.normal(params[pos][1], 
-                                                    np.sqrt(params[pos][2]), 
-                                                    size=count)
-        pos_i += count
-    samps = np.random.permutation(samps)
-
-    h, _ = np.histogram(samps, bins)
-    h = h.astype(float) / np.sum(h)
-
-    p = model.compute_mm_probs(bins, params)
-    assert_almost_equal(np.sum(p), 1.0, 4)
-
-    assert irm.util.kl(h, p) < 0.001
-    
-
-
-def test_mm_mixmodel():
-    np.random.seed(0)
-
-    import distributions
-    N = 100
-
-    data = np.zeros(N, dtype=np.float32)
-    MODEL = model.NonConjGaussian(0.5)
-
-    data[::2] = np.random.normal(-8, 0.5, N/2)
-    data[1::2] = np.random.normal(8, 0.5, N/2)
-    f = mixmodel.Feature(data, MODEL)
-
-    mm = mixmodel.MixtureModel(N, {'f1' : f})
-    
-    rng = None
-    # random init
-    grp = {}
-    for i, g in enumerate(np.random.permutation(np.arange(N) % 10)):
-        if g not in grp:
-            grp[g] = mm.create_group(rng)
[…]

### 2014-06-04 util.py

@@ -1,35 +1,0 @@
-import numpy as np
-import cPickle as pickle
-from matplotlib import pylab
-from ruffus import * 
-from preprocess import * 
-import pandas
-import sqlite3
-
-# plot spatial distribution of each cell type
-# plot area vs cell body distance
-
-
-def create_adj_mat(con, area_thold, cell_data):
-    """
-    returns (upper-triangular) contact area matrix, cell_ids in order
-
-    """
-    
-    df = pandas.io.sql.read_frame("select from_id, to_id, area, sum(area) as contact_area, count(area) as contact_count from contacts  where area < %f group by from_id, to_id" % area_thold, 
-                                  con)
-    
-    CELL_N = len(cell_data)
-    id_to_pos = {id: pos for pos, id in enumerate(cell_data.index.values)}
-
-    area_mat = np.zeros((CELL_N, CELL_N), dtype=np.float32)
-
-    for c_i, c_row in df.iterrows():
-        i1 = id_to_pos.get(c_row['from_id'], -1)
-        i2 = id_to_pos.get(c_row['to_id'], -1)
-        if i1 >= 0 and i2 >= 0:
-            area_mat[i1, i2] = c_row['contact_area']
-
- 
-    return area_mat, cell_data.index.values
-            

### 2014-06-15 experiments/mouseretina/pipeline.py
updates

@@ -72,7 +72,8 @@ EXPERIMENTS = [
     ('retina.0.ld.0.0.xyz','debug_2_100', 'debug_20'), 
 
 
-    #('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
+    ('retina.xsoma' , 'fixed_20_100', 'anneal_slow_1000'), 
+    ('retina.1.bb' , 'fixed_20_100', 'anneal_slow_1000'), 
     #('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
 ]
 
@@ -88,12 +89,12 @@ BB_BETAS = [1.0]
 VAR_SCALES = [0.01, 0.1, 1.0]
 COMP_KS = [2, 3]
 
-# for ti in [1, 2,  3]: # remember to add 2 back in ! 
-#     for v in range(len(VAR_SCALES)):
-#         for k in COMP_KS:
-#             EXPERIMENTS.append(('retina.%d.clist.%d.%d' % (ti, v, k) , 
-#                                 'fixed_20_100', 'anneal_slow_400'))
-#             pass
+for ti in [1]: # remember to add 2 back in ! 
+    for v in range(len(VAR_SCALES)):
+        for k in COMP_KS:
+            EXPERIMENTS.append(('retina.%d.clist.%d.%d' % (ti, v, k) , 
+                                'fixed_20_100', 'anneal_slow_1000'))
+
 
 # for ti in range(len(THOLDS)):
 #     for ml_i in range(len(MULAMBS)):
@@ -247,6 +248,38 @@ def create_latents_srm(infile,
 
 
 
+
+def create_latents_bb_params():
+    for a in create_tholds():
+        inf = a[1][0]
+        outf_base = inf[:-len('.data.pickle')]
+        outf = "%s.bb" % (outf_base)
+        yield inf, [outf + '.data', 
+                    outf + '.latent', outf + '.meta']
+
+
+@follows(data_create_thold)
+@files(create_latents_bb_params)
+def create_latents_bb(infile, 
+                       (data_filename, latent_filename, meta_filename)):
+
+
+    d = pickle.load(open(infile, 'r'))
+    conn_mat = d['conn_mat']
+    model_name = "BetaBernoulliNonConj"
+    irm_latent, irm_data = irm.irmio.default_graph_init(conn_mat, model_name)
+
+    HPS = {'alpha' : BB_ALPHAS[0], 
[…]

### 2014-06-15 experiments/mouseretina/predlinks.py

@@ -1,4 +1,7 @@
-import numpy as np
+
+import sys
+# total hack, I should really know better
+sys.path.append("../../code")
 import numpy as np
 import cPickle as pickle
 import sqlite3
@@ -52,7 +55,7 @@ def compute_prob_matrix(tgt_latent, tgt_data, model_name='LogisticDistance'):
 THOLDS = [1, 2, 3]
 PRED_EVALS= np.logspace(-4, 0, 41) # np.linspace(0, 1.0, 41)
 
-@files("../preprocess/mouseretina/mouseretina.db", ['truth.%d.ld.predlinks.pickle' % t for t in THOLDS])
+@files("../../../preprocess/mouseretina/mouseretina.db", ['truth.%d.ld.predlinks.pickle' % t for t in THOLDS])
 def create_truth(dbfile, outfiles):
     conn = sqlite3.connect(dbfile)
     for THOLD_i, outfile in zip(THOLDS, outfiles):
@@ -82,7 +85,7 @@ def create_truth(dbfile, outfiles):
                     'thold_i' : THOLD_i}, 
                     open(outfile, 'w'))
 
-@files("../preprocess/mouseretina/mouseretina.db", ['truth.%d.bb.predlinks.pickle' % t for t in THOLDS])
+@files("../../../preprocess/mouseretina/mouseretina.db", ['truth.%d.bb.predlinks.pickle' % t for t in THOLDS])
 def create_truth_bb(dbfile, outfiles):
     conn = sqlite3.connect(dbfile)
     for THOLD_i, outfile in zip(THOLDS, outfiles):
@@ -115,9 +118,9 @@ def create_truth_bb(dbfile, outfiles):
                     'thold_i' : THOLD_i}, 
                     open(outfile, 'w'))
 
-@transform(["data/retina.1.*.ld*.samples", 
-            "data/retina.1.*.srm_*.samples", 
-            "data/retina.1.simple_bb.data*.samples", 
+@transform(["data/retina.1.ld*.samples", 
+            "data/retina.1.srm_*.samples", 
+            "data/retina.1.bb.data*.samples", 
         ], 
            suffix(".samples"), ".predlinks.pickle")
 def compute_pred_mat(sample_name, outfile):
@@ -173,7 +176,7 @@ def compute_pred_mat(sample_name, outfile):
 
 @transform(["data/retina.1.ld*.samples", 
             "data/retina.1.srm_*.samples", 
[…]

### 2014-06-15 experiments/mouseretina/predlinks_plot.py

@@ -1,0 +1,127 @@
+import numpy as np
+import cPickle as pickle
+import pandas
+from matplotlib import pylab
+
+
+THOLD_IDX = 1
+
+
+clust_df = pickle.load(open("cluster_metrics.pickle"))['clust_df']
+clust_df = clust_df[ clust_df['filename'].str.contains("debug") == False]
+
+clust_df = clust_df[clust_df['filename'].str.contains("retina.%d" % THOLD_IDX)]
+
+def f(x):
+    x['scoreidx'] = np.array(np.argsort(np.argsort(x['score'])[::-1]))
+    return x
+
+
+clust_df = clust_df.groupby('filename').apply(f )
+
+clust_df['model'] = None
+clust_df['model'][clust_df['filename'].str.contains('ld') & clust_df['filename'].str.contains('xyz') ] = 'srm.xyz'
+clust_df['model'][clust_df['filename'].str.contains('.xsoma.')] = 'xsoma'
+clust_df['model'][clust_df['filename'].str.contains('.clist.')] = 'clist'
+clust_df['model'][clust_df['filename'].str.contains('srm_clist_xsoma')]= 'srm_clist_xsoma'
+
+tgtdf = clust_df #[df['filename'].str.contains('retina.2')]
+
+measure = 'ari'
+#ndf = tgtdf[tgtdf['scoreidx']==0]
+b = tgtdf.sort(measure, ascending=False).groupby('model', as_index=False).first()
+
+filenames = b['filename']
+
+
+f = pylab.figure(figsize=(12, 8))
+
+                 
+# bins = np.linspace(0, 0.6, 20)
+# for fi, filename in enumerate(filenames):
+#     ax = f.add_subplot(2, 2, fi+1)
+#     subset = tgtdf[tgtdf['filename'] == filename]
+#     #print subset.head(1)
+#     label = subset.head(1).irow(0)['model']
+#     ax.hist(subset[measure], label=label, bins=bins, normed=True)
+#     ax.axvline(subset[measure].mean(), c='r', linewidth=4)
+#     ax.set_title(label)
+#     ax.set_xlim(bins[0], bins[-1])
+#     ax.grid()
+# f.savefig('pl.comparison.best.ari_coarse.pdf')
+
+
+# f = pylab.figure(figsize=(12, 8))
+
+                 
+# bins = np.linspace(0, 0.6, 40)
[…]

### 2014-06-17 docs/paper.tex
build process works
updates to all figs

@@
 \documentclass{nature}
-\documentclass{article}
 \usepackage[english,american]{babel}
@@
 
-\todo{“Anatomist cell type” text}
 \todo{Consistent line thickness}
@@
 \begin{enumerate}
-  \item Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology, Cambridge, Massachusetts
+  \item Electrical Engineering and Computer Science, University of California, Berkeley
   \item Department of Physical Medicine and Rehabilitation, Northwestern University and Rehabilitation Institute of Chicago, Chicago, Illinois
@@
 >
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics{comparison.ai}}
+  \caption{Comparison of the value of incorporating additional metadata 
+for the mouse retina connectome.}
+\label{fig:mouseretina_compare}
+\end{figure}
+
+\todo{spatial extent comparison}
 
@@
 
-\being{itemize}
+\begin{itemize}
 \item \textbf{synthetic}: For synthetic datasets of sizes comparable to our

### 2014-06-17 experiments/mouseretina/pipeline.py

@@ -601,6 +601,99 @@ def plot_best_cluster_latent(exp_results,
         # dont do the clist ones
         file(out_filename, 'w').write("test")
 
+@transform(get_results, suffix(".samples"), 
+           ".z.pdf")
+def plot_z_matrix(exp_results, 
+                  out_filename):
+    # debug
+    # fid = open(out_filename, 'w')
+    # fid.write("test")
+    # fid.close()
+    # return
+
+    sample_d = pickle.load(open(exp_results))
+    chains = sample_d['chains']
+    
+    exp = sample_d['exp']
+    data_filename = exp['data_filename']
+    data = pickle.load(open(data_filename))
+    data_basename, _ = os.path.splitext(data_filename)
+    meta = pickle.load(open(data_basename + ".meta"))
+
+    meta_infile = meta['infile']
+
+    d = pickle.load(open(meta_infile, 'r'))
+    conn = d['conn_mat']
+    cells = d['cells']
+
+    cell_types = cells['type_id']
+
+    chains = [c for c in chains if type(c['scores']) != int]
+    CHAINN = len(chains)
+
+    # compute the z matrix 
+    z_mat = np.zeros((len(cells), len(cells)))
+    for ci, c in enumerate(chains):
+        print "z for chain", ci
+        sample_latent = c['state']
+        cell_assignment = np.array(sample_latent['domains']['d1']['assignment'])
+        ca = irm.util.canonicalize_assignment(cell_assignment)
+        for u in np.unique(ca):
+            ent_i = np.argwhere(ca == u).flatten()
+            for ci in ent_i:
+                for cj in ent_i:
+                    z_mat[ci, cj] += 1
+
+    import scipy.cluster
+    l = scipy.cluster.hierarchy.linkage(z_mat, method='ward')
+    
+    
+    ca = np.array(scipy.cluster.hierarchy.leaves_list(l))
+    z_mat = z_mat[ca]
+    z_mat = z_mat[:, ca]
+
+
+    import matplotlib.gridspec as grd
+    f = pylab.figure()
+    gs = grd.GridSpec(1, 2, width_ratios=[1,6], wspace=0.02)
+    ax = f.add_subplot(gs[0, 1])
+    ax.imshow(z_mat, cmap=pylab.cm.Greys, interpolation='nearest')
+    ax.set_xticks([])
+    ax.set_yticks([])
[…]

### 2014-06-17 docs/paper.tex
updates to paper text

@@
-\documentclass{nature}
+%\documentclass{nature}
+\documentclass{article}
 \usepackage[english,american]{babel}
@@
 \begin{abstract}
-  Connectomics produces massive amounts of data,
-  necessitating new analysis methods to discover the biological and
-  computational structure. It has long been assumed
-  that discovering neuron types and their relation to microcircuitry is crucial to
-  understanding neural function. Here we developed a nonparametric
-  Bayesian technique that identifies neuron types and microcircuitry
-  patterns in connectomics data.
-It combines the information traditionally used by biologists, including connectivity, cell body location and the spatial distribution of synapses.
-We show that the approach recovers known neuron types in the retina and enables predictions of connectivity, better than simpler algorithms.
-It also can reveal interesting structure in
-  the nervous system of c. elegans, and automatically discovers the
-  structure of a microprocessor. 
- Our approach extracts structural meaning from connectomics, enabling new approaches of automatically deriving  anatomical insights from these emerging datasets.
+  Connectomics produces massive amounts of data, necessitating new
+  analysis methods to discover the biological and computational
+  structure. It has long been assumed that discovering neuron types
+  and their relation to microcircuitry is crucial to understanding
+  neural function. Here we developed a nonparametric Bayesian
+  technique that identifies neuron types and microcircuitry patterns
+  in connectomics data. It combines the information traditionally used
+  by biologists, including connectivity, cell body location and the
+  spatial distribution of synapses.  We show that the approach
+  recovers known neuron types in the retina and enables predictions of
+  connectivity, better than simpler algorithms. It also can reveal
+  interesting structure in the nervous system of c. elegans, and
+  automatically discovers the structure of a microprocessor.  Our
+  approach extracts structural meaning from connectomics, enabling new
+  approaches of automatically deriving anatomical insights from these
+  emerging datasets.
 \end{abstract}
@@
   \centerline{\includegraphics{comparison.ai}}
-  \caption{Comparison of the value of incorporating additional metadata 
-for the mouse retina connectome.}
+  \caption{a.
+Cell-cell assignment matrix reflecting the posterior distribution of clusterings found by our model.
+Large contiguous black areas are cells that were considered the same type in the majority of samples.
+b.
+The cells and their true types sorted to agree with the coassignment matrix.
+c.
+A comparison of the predictive accuracy (Area under the curve) for hand-labeled anatomical data, versus inclusion of additional sources of information.
+d.
+) The spatial extent (in depth and area) of the types identified by humans and our various algorithmic approaches.}
+
 \label{fig:mouseretina_compare}
@@
 of the latent space, after a certain amount of time
-\todo{add figure}
+
 \end{itemize}
@@
 
+\todo{Concerns about parameter sensitivity}
+\todo{How did we pick the best one we often show?}
+
 \end{document}

### 2014-06-17 docs/paper.tex
mixing text

@@
 
-\todo{Consistent line thickness}
-\todo{on adjacency matrix remove text}
-
 \title{Automatic discovery of cell types and microcircuitry from neural connectomics}
@@
   \item Electrical Engineering and Computer Science, University of California, Berkeley
-  \item Department of Physical Medicine and Rehabilitation, Northwestern University and Rehabilitation Institute of Chicago, Chicago, Illinois
+  \item Department of Physical Medicine and Rehabilitation,
+    Northwestern University and Rehabilitation Institute of Chicago,
+    Chicago, Illinois
   \item Department of Physiology, Northwestern University, Chicago, Illinois
@@
 \begin{abstract}
-  Connectomics produces massive amounts of data, necessitating new
-  analysis methods to discover the biological and computational
-  structure. It has long been assumed that discovering neuron types
-  and their relation to microcircuitry is crucial to understanding
-  neural function. Here we developed a nonparametric Bayesian
-  technique that identifies neuron types and microcircuitry patterns
-  in connectomics data. It combines the information traditionally used
-  by biologists, including connectivity, cell body location and the
-  spatial distribution of synapses.  We show that the approach
-  recovers known neuron types in the retina and enables predictions of
-  connectivity, better than simpler algorithms. It also can reveal
-  interesting structure in the nervous system of c. elegans, and
-  automatically discovers the structure of a microprocessor.  Our
-  approach extracts structural meaning from connectomics, enabling new
-  approaches of automatically deriving anatomical insights from these
-  emerging datasets.
+  Neural connectomics has begun producing massive amounts of data,
+  necessitating new analysis methods to discover the biological and
+  computational structure. It has long been assumed that discovering
+  neuron types and their relation to microcircuitry is crucial to
+  understanding neural function. Here we developed a nonparametric
+  Bayesian technique that identifies neuron types and microcircuitry
+  patterns in connectomics data. It combines the information
+  traditionally used by biologists, including connectivity, cell body
+  location and the spatial distribution of synapses.  We show that the
+  approach recovers known neuron types in the retina and enables
+  predictions of connectivity, better than simpler algorithms. It also
+  can reveal interesting structure in the nervous system of
+  c. elegans, and automatically discovers the structure of a
+  microprocessor.  Our approach extracts structural meaning from
+  connectomics, enabling new approaches of automatically deriving
+  anatomical insights from these emerging datasets.
 \end{abstract}
@@
 \section*{Introduction}
-Emerging techniques \autocite{Morgan2013,Zador2012} promise to
-quantify the location of each neuron within a volume of interest,
-along with its connections to all other neurons. Far exceeding the
-capacity of neuroanatomists to trace small circuits, this leads to
-enormous datasets that quantify aspects of the nervous system. The impact of
-this kind of high-throughput transition has been seen before -- 
-rise of sequencing techniques necessitated the
-development of novel computational methods to understand genomic
-structure, ushering in an era of bioinformatics as an independent
-discipline \autocite{Koboldt2013}.
+Emerging connectomics techniques \autocite{Morgan2013,Zador2012}
+promise to quantify the location and connectivity of each neuron
+within a tissue volume. These massive datasets will far exceed the
+capacity of neuroanatomists to manually trace small circuits, thus
+necessitating computational, quantitative, and automatic methods for
+understanding neural structure.  The impact of this kind of
+high-throughput transition has been seen before -- rise of sequencing
+techniques necessitated the development of novel computational methods
+to understand genomic structure, ushering in an era of bioinformatics
+as an independent discipline \autocite{Koboldt2013}.
 
@@
 morphological, but they also differ in the way they connect to one
-another. In fact, the idea of well defined local connectivity patterns
-(microcircuits) has been prominent in many areas, ranging from sensory
-(e.g. retina, \autocite{Masland2001} via processing (e.g. neocortex
-\autocite{Mountcastle1997}) to movement (e.g. spinal cord)
-\autocite{Grillner2005}. Indeed, such repeated patterns also exists in
-human-made computing circuits. It remains an important challenge to
-develop algorithms to use anatomical data, e.g. connectomics, to automatically back
-out underlying microcircuitry.
+another. In fact, the idea of well defined, type-dependent local
+connectivity patterns (microcircuits) has been prominent in many
+areas, ranging from sensory (e.g. retina, \autocite{Masland2001} via
+processing (e.g. neocortex \autocite{Mountcastle1997}) to movement
+(e.g. spinal cord) \autocite{Grillner2005}. These sorts of repeated
+computing patterns are a common feature of computing systems, even
+arising in human-made computing circuits. It remains an important
+challenge to develop algorithms to use anatomical data,
+e.g. connectomics, to automatically back out underlying
+microcircuitry.
 
@@
 the types of scaling present in the network \autocite
-{WattsStrogatz1998} .  While this approach leads to an understanding
-of global network, more recent work aims at identifying very small-scale
-repeat patterns, or “motifs” in networks\autocite{Milo2002}
+{WattsStrogatz1998}.  While this approach leads to an understanding
+of the global network, more recent work aims at identifying very small-scale
+repeat patterns, or “motifs” in networks\autocite{Milo2002}. \todo{motifs vs type}.
 
@@
 
-Here we describe a Bayesian non-parametric model that can discover
-circuit structure automatically from connectomics data: the cell
-types, their spatial patterns of interconnection, and the locations of
-somata and synapses. We primarily focus on the recently-released mouse
-retina connectome \autocite{Helmstaedter2013}, but additionally
-examine the c. elegans connectome \autocite{White1986}, and then
-``connectome'' of a classical microprocessor
-\autocite{James2010}. Comparing the cell types discovered by the
-algorithms with those obtained manually by human anatomists reveals a
-high degree of agreement. We present a scalable approach to infer
-microcircuitry from connectomics data.
+To address these challenges, here we describe a Bayesian
+nonparametric model that can discover circuit structure automatically
+from connectomics data: the cell types, their spatial patterns of
+interconnection, and the locations of somata and synapses. We show
+that by incorporating this additional information, our model both
+accurately predicts the connection as well as agrees
+with human neuroanatomists as to the identification of cell types.
+
+We primarily focus on the recently-released mouse retina connectome
+\autocite{Helmstaedter2013}, but additionally examine the c. elegans
+connectome \autocite{White1986}, and then ``connectome'' of a
+classical microprocessor \autocite{James2010}. Comparing the cell
+types discovered by the algorithms with those obtained manually by
+human anatomists reveals a high degree of agreement. We present a
+scalable approach to infer microcircuitry from connectomics data.
 
@@
 \section{Results}
-We start with the generic assumption that cell types exist and affect
+
+We start with the generic assumption that cell types exist -- a cell
+is a member of a single, unobserved type -- and these types affect
 soma depth, distribution of synapses, as well as a cell type and
@@
 model (iSBM)\autocite{Kemp2006a,Xu2006}, which has been shown to
-meaningfully cluster connection graphs. We extend this approach by
-adding distance dependence to model salient aspects of
-microcircuitry. We additionally model cell body depth as a mixture of
-Gaussians and each cell’s synapse density profile as mixture of
-mixtures of Gaussians. For example, retinal ganglion cells may synapse
-on nearby, but not far away, amacrine cells, with ganglion cells being
-superficial and having a broad distribution of synapses.  This
-procedure automatically learns the number of cell types, which cells
-belong to which types, along with the the distance and type dependent
-connectivity probabilities (see fig~\ref{fig:overview}). We apply our
-algorithm to datasets from mouse retina, c-elegans and a historical
-microprocessor. Anatomists classify cells based on many features,
-giving us a meaningful baseline to compare against.
+meaningfully cluster connection graphs while learning the number of
+hidden groups, or types. We extend this approach by adding distance
+dependence to model salient aspects of microcircuitry. We additionally
+model cell body depth as a mixture of Gaussians and each cell’s
+synapse density profile as mixture of mixtures of Gaussians.
+
+For example, retinal ganglion cells may synapse on nearby, but not far
+away, amacrine cells, with ganglion cells being superficial and having
+a broad distribution of synapses.  This procedure automatically learns
+the number of cell types, which cells belong to which types, along
+with the distance and type dependent connectivity probabilities (see
+fig~\ref{fig:overview}). We apply our algorithm to datasets from mouse
+retina, c-elegans and a historical microprocessor. Anatomists classify
+cells based on many features, giving us a meaningful baseline to
+compare against.
 
 Before we can use a probabilistic model to analyze real data we first
-need to assess if the model works properly. We thus started out by
-simulating data for which we know the correct structure and comparing
-the estimated structure based on the algorithm (see methods) with the
-one we used for simulation. We find that the model does a good job of
-recovering the correct number of cell types,
-(fig~\ref{fig:synthetic}a), the the cell identities
-(fig~\ref{fig:synthetic}b), and the spatial extent of each type
-(fig~\ref{fig:synthetic}c).  Indeed, it does far better than the
-existing infinite stochastic block model that assumes that there is no
-distance dependence to the connectivity. The model converges
+need to assess if the model works properly. We thus simulate data for
+which we know the correct structure and comparing the estimated
+structure based on the algorithm (see methods) with the one we used
+for simulation. We find that the model does a good job of recovering
+the correct number of cell types, (fig~\ref{fig:synthetic}a), the the
+cell identities (fig~\ref{fig:synthetic}b), and the spatial extent of
+each type (fig~\ref{fig:synthetic}c).  Indeed, it does far better than
+the existing infinite stochastic block model that assumes that there
+is no distance dependence to the connectivity. The model converges
 relatively quickly to an estimate of the most probable values for the
@@
 allows application to relatively large datasets.
-<XXX now add extra features>
 
@@
     synthetic data when incorporating spatial information. a.) The
-    iSBM (which only uses connectivity information) over-estimates the
-    number of classes as it fails to take distance into account,
-    whereas our modeling of the combination of distance and
-    connectivity finds close to the true number of classes. b.) ARI, a
-    measure of ``correct clustering'', for different true class
-    counts, between our model and a traditional iSBM. c.) For
-    synthetic data, type discovery based only on connectivity are
-    tiny spatially-localized groups -- in
+    infinite stochastic block model (which only uses connectivity
+    information) over-estimates the number of classes as it fails to
+    take distance into account, whereas our modeling of the
+    combination of distance and connectivity finds close to the true
+    number of classes. b.) ARI, a measure of ``correct clustering'',
+    for different true class counts, between our model and a
+    traditional iSBM. c.) For synthetic data, type discovery based
+    only on connectivity are tiny spatially-localized groups -- in
     contrast, our model recovers the true spatial extent of the
@@
     were known. b. clustered connectivity matrix. c.) connectivity
-    diagram showing our clusters, as well as the cell depth and known
-    cell type of each cell. Our model shows moderate agreement with
-    the cell types traditionally identified by anatomist
-    laminar-specific connectivity patterns. \todo{Add more text here}
-    d.) Connectivity between our clusters as a function of distance --
-    the cluster consisting primarily of retinal ganglion cells (lower
-    left-center) exhibits the expected near and far connectivity. e.)
-    The spatial distribution of our cell types -- each cell type
-    tesselates space.}
+    diagram showing our clusters, as well as the cell depth and
+    anatomist-labeled cell type of each cell. Our model shows moderate
+    agreement with the cell types traditionally identified by
+    anatomist laminar-specific connectivity patterns. 
+    d.) Connectivity between our clusters as a function
+    of distance -- the cluster consisting primarily of retinal
+    ganglion cells (lower left-center) exhibits the expected near and
+    far connectivity. e.)  The spatial distribution of our cell types
+    -- each cell type tesselates space.}
 \label{fig:mouseretina}
@@
 
-The retina of the mouse \autocite{Masland2001} is an example where we
-should expect to have connectivity patterns that are well approximated
-by our generative model. It is known that there are multiple classes
-of cells that can be grouped into ganglion cells that transmit
+The mouse retina \autocite{Masland2001} is a neural circuit which we
+expect to have connectivity patterns that are well approximated by our
+generative model. It is known that there are multiple classes of cells
+that can be broadly grouped into: ganglion cells that transmit
 information to the rest of the brain, bipolar cells that connect
 between different cells, and amacrine cells that feed into the
-ganglion cells. Recent research
-\autocite{Helmstaedter2013} has produced a large dataset containing
-both the types of cells from orthogonal approaches, and also the
-connectivity matrix between all reconstructed cells. We want to know
-to which level we can classify neurons into types based on this
-connectivity only.
+ganglion cells. Recent research \autocite{Helmstaedter2013} has
+produced a large dataset containing both the types of cells from
+orthogonal approaches, and also the connectivity matrix between all
+reconstructed cells.
 
-The algorithm took less than 2 hours to converge to a good solution,
-dividing neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
+The algorithm took less than 2 hours to perform inference, dividing
+neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
 wedge is a type). For each pair of neurons there is a specific
-distance dependent connection probability (fig~\ref{fig:mouseretina}b,c,d),
-which is well approximated by the model fit. Moreover, each type of
-cell is rather isotropically distributed across space
-(fig~\ref{fig:mouseretina}e) as should be expected for true cell types. The
-algorithm is thus able to extract meaningful structure from the data.
+distance dependent connection probability
+(fig~\ref{fig:mouseretina}b,c,d), which is well approximated by the
+model fit. Moreover, each type of cell is rather isotropically
+distributed across space (fig~\ref{fig:mouseretina}e) as should be
+expected for true cell types.
 
@@
 types we find tend to reflect the known laminar distribution in the
-retina (fig~\ref{fig:mouseretina}c, middle ring), which is exciting as
-the algorithm did not actually use this type of information.
+retina (fig~\ref{fig:mouseretina}c, middle ring).
 
@@
 
-<xxx discuss the various algorithms and plus and minus 
-   Other model classes
-   brielfly explain link prediction accuracy and ARI
-   show how it compares to human example	
-   the more humanlike the information it gets the more humanlike its classification but many techniques allow good link prediction
->
+Our fully Bayesian model produces a distribution over probable
+clusterings.  Figure~\ref{fig:mouseretina_compare} shows this
+posterior distribution as a cell-cell coassignment matrix, sorted to
+find maximum block structure. Each large, dark block represents a
+collection of cells believed with strong probability to be of the same
+type. When we plot (fig~\ref{fig:mouseretina_compare}b) the
+anatomist-derived cell types along the left, we can see that each
+block consists of a roughly-homogeneous collection of types.
+
+We can thus evaluate our model along three sets of parameters: how
+closely does our clustering agree with neuroanatomists' knowledge? 
+Given two cells, how accurately can our model predict the link
+between them? And how closely does the spatial extent (within a layer)
+of our identified types agree with the known neuroanatomists.
+
+We find that given the dataset many techniques allow for good
+link-predictive accuracy but poor agreement with neuroanatomists. In
+fact, when we fit a version of our model fixing the clusterings given
+by neuroanatomists \todo{Add something here}. As we incorporate
+additional information beyond connectivity into our model (soma
+distance, soma depth, synapse density profile) we more closely
+approximate human clustering.
 
@@
   \centerline{\includegraphics{comparison.ai}}
-  \caption{a.
-Cell-cell assignment matrix reflecting the posterior distribution of clusterings found by our model.
-Large contiguous black areas are cells that were considered the same type in the majority of samples.
-b.
-The cells and their true types sorted to agree with the coassignment matrix.
-c.
-A comparison of the predictive accuracy (Area under the curve) for hand-labeled anatomical data, versus inclusion of additional sources of information.
-d.
-) The spatial extent (in depth and area) of the types identified by humans and our various algorithmic approaches.}
+  \caption{a. Cell-cell assignment matrix reflecting the posterior
+    distribution of clusterings found by our model. Large contiguous
+    black areas are cells that were considered the same type in the
+    majority of samples. b. The cells and their true types sorted to
+    agree with the coassignment matrix. c. A comparison of the
+    predictive accuracy (Area under the curve) for hand-labeled
+    anatomical data, versus inclusion of additional sources of
+    information. d. ) The spatial extent (in depth and area) of the
+    types identified by humans and our various algorithmic
+    approaches.}
 
@@
 \end{figure}
-
-\todo{spatial extent comparison}
-
-We compared the results of our model to the traditional iSBM to see if
-the addition of spatial information was necessary. We find that, as
-expected from our synthetic experiments, the iSBM massively
-over-segmented the data, finding spatially-localized “communities” of
-cells which bore little relation to known types.  The average spatial
-extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. Using
-the anatomists' ground truth, the mean spatial extent of types was
-$34.3\pm7.6 \mu m$, quite close to the $34.5\pm 1.0 \mu m$ discovered
-by our algorithm. Modeling the distance-dependence prevents
-over-segmentation and thus allows more meaningful estimates.
 
@@
 
-TODO FIGURE 5: 
-We know we want the bar graph with ARI and AUC at the same time:
-Conn only
-conn and dist
-conn and dist and synapse profile and depth
-LPCM 
-Just cell depth 
-human dashed line (for link prediction)
-
 \section{Discussion}
@@
 types differ in morphology, connectivity, transcriptomics, relation to
-behavior or stimuli and many other ways. Algorithms combining this data and type type
-information may allow us to synthesize all the available
-information from one experiment or even across experiments into a
-joint model of brain structure and function.
+behavior or stimuli and many other ways. Algorithms combining this
+data and type type information may allow us to synthesize all the
+available information from one experiment or even across experiments
+into a joint model of brain structure and function.
 
@@
  \item \textbf{Correspondence} Correspondence and requests for materials
-should be addressed to E.J.~(email: jonas@mit.edu).
+should be addressed to E.J.~(email: jonas@eecs.berkeley.edu).
 \end{itemize}
@@
 \section{Methods }
-XXX TODO : add n-way cross validation is not done because it would be a pain
-XXX TODO that said overfitting may be shared across all models
 
@@
 
-% TODO add in the thingie
 \begin{equation}
@@
 
-We chain these three kernels together, and then globally
-anneal on the likelihood from a temperature of $T=64$ down to 
-$T=1$ over 300 iterations unless otherwise indicated, and
-then run the chain for another $100$ iterations.
+We chain these three kernels together, and then globally anneal on the
+likelihood from a temperature of $T=64$ down to $T=1$ over $900$
+iterations unless otherwise indicated, and then run the chain for
+another $100$ iterations. We then generate at least $20$ samples, each
+taken from the end of a single Markov chain initialized from different
+random initial points in the state space. For visualization we pick the
+chain with the highest log likelihood, but for all numerical
+comparisons (including link probability and ARI) we use this full
+collection of samples from the posterior distribution to estimate the
+resulting statistics.
 
-To pick the MAP, we run up to $100$ simultaneous chains in parallel,
-initialized from different random initial points in the state space. 
-For visualization we pick the chain with the highest log likelihood,
-but for all numerical comparisons (including link probability and ARI)
-we use this full collection of posterior samples to estimate
-the resulting statistics.
+\subsection{Link Prediction}
+As a proxy for link-prediction accuracy we compute the probability of a link
+between two cells using each model, trained fully on the data. While this method
+is potentially prone to overfitting, the overfitting will be shared across models
+and in fact will preferentially bias in favor of competing models
+which over-cluster the data. We use a full collection of posterior samples when
+computing the link probability, and then compute the area under the ROC curve for
+each.
+
+\subsection{Model Comparison}
+We compare our model with a standard network clustering model, the
+latent-position clutsering model. This model assumes each cell belongs
+to one of K clusters, and each cluster is associated with a
+$d-$dimensional Gaussian distribution. The probability of a link is
+then a function of the distance between the data points in this
+continuous-space. We use \autocite{VBLPCM} a variational
+implementation provided in R, parametrically varying the number of
+latent dimensions and the number of requested groups.  While this
+model provides reasonable link predictive accuracy, the clusterings
+dramatically disagree with those from human anatomists.
+
+\subsection{Parameters}
+
+Hierarchical generative models can be sensitive to hyperparameter
+settings, thus for most hyperparamters we perform inference. In cases
+where we cannot we run separate collections of markov chains at
+separate settings and show the results across all pooled
+parameters. For the case of the mouse retina data, we consisder
+maximum link probability $p_{max} \in \{0.95, 0.9, 0.7\}$, variance
+scales for the synapse density profile of $\sigma^2 \in \{0.01, 0.1,
+1.0\}$ (of normalized depth), and $K \in \{2, 3\}$ possible synapse
+density profile mixture components. For the connectivity-distance-only
+model we actually perform inference over both $p_{max}$ and $p_{min}$.
 
 \subsubsection{Mixing of our Markov chains} 
-Evaluating whether or not approximate inference methods,
-such as MCMC, produce samples which are valid approximations 
-of the posterior distribution is an ongoing area of research
-in the computational statistics community. Here we validate
-mixing in two ways:
+Evaluating whether or not approximate inference methods, such as MCMC,
+produce samples which are valid approximations of the posterior
+distribution is an ongoing area of research in the computational
+statistics community. We use a rough proxy here -- synthetic
+likelihood evaluation.  For synthetic datasets of sizes comparable to
+our real data size, do we recover known ground truth information after
+running our markov chains for the appropriate amount of time?
 
-\begin{itemize}
-\item \textbf{synthetic}: For synthetic datasets of sizes comparable to our
-real data size, do we recover known ground truth information
-after running our markov chains for the appropriate amount of time? 
-\todo{Add figure}
-\item \textbf{autocorrelation} For various summary statistics 
-of the latent space, after a certain amount of time
+Figures \ref{fig:mixing:ari} and \ref{fig:mixing:scorevstime} shows
+the ARI to ground truth and the total log score as a function of
+runtime.  We see dramatic changes in log score initially as we vary
+the temperature, stabilizing as runtime progresses, for each
+chain. Then we see the characteristic jumps between nearby modes
+towards the end of the run, in both log score and ARI.  Importantly,
+regardless of whether our model over- or under-estimates the exact
+posterior variance about the network, we find points in the latent
+variable space that are both predictive \textit{and} parsimonious,
+largely agreeing with the human anatomists and predicting existing
+connections.
 
-\end{itemize}
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=60mm]{mixing.ari.pdf}}
+  \caption{}
+\label{fig:mixing:ari}
+\end{figure}
 
-But importantly, regardless of whether our model over- or under-estimates the exact posterior variance about the network, we find points in the latent variable space that are both predictive \textit{and} parsimonious, largely agreeing with the human anatomists and predicting existing connections.
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=130mm]{mixing.scorevstime.pdf}}
+  \caption{}
+\label{fig:mixing:scorevstime}
+\end{figure}
 
@@
 contact area between all pairs of 950 cells. Area was thresholded at
-$0.5\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
+$0.1\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
 matrix that served as input to our algorithm. We measured the distance
@@
 
-The ``logstic-distance Poisson'' spatial model is used to explicitly mode the count of synapses, $c$, between two neurons.
-The probability of c synapses between two neurons is distributed $c \sim \textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a scaled logistic function (the logistic function has range $[0, 1]$.
-For each component $\eta_{mn}$ we learn both the threshold $\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$ and $n$ are likely to have on average $20$ synases if they are closer than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$  "
+The ``logstic-distance Poisson'' spatial model is used to explicitly
+mode the count of synapses, $c$, between two neurons. The probability
+of c synapses between two neurons is distributed $c \sim
+\textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a
+scaled logistic function (the logistic function has range $[0,
+1]$. For each component $\eta_{mn}$ we learn both the threshold
+$\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$
+and $n$ are likely to have on average $20$ synases if they are closer
+than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$ "
 
@@
 
-\subsection{Source code}
+\subsection{Source code and data}
 
@@
 \href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/netmotifs/}
+
+and the content of this paper along with scripts to run
+experiments and generate all figures can be found at
+
+\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/connect-disco-paper/}
+
+All preprocessed data has been made publicaly available as well.
 
@@
 \FloatBarrier
-
-\subsection{Microprocessor Netlist annotation}
-\label{supp:mos6502netlist}
-This is a matrix consisting of the different clusters of electrical nets for 
-the registers, with the colors consistent with figure~\ref{fig:mos6502}.
-
-\includepdf[pages=-]{mos6502.clusters.pdf}
-
-\newpage
-
-\FloatBarrier
-\subsection{Detailed learned connectivity for mouse retina}
-
-Here we show the recovered distance-dependent connectivty profile
-versus the empirical distance-dependent connectivty for the
-retina dataset. The Logistic-distance Poisson likelihood closely
-matches the empirical likelihood for most classes.
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=7in,page=2]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.0.latent.pdf}}
-  \caption{Learned connectivity vs distance between each of the discovered cell types in the mouse retina.
-The rows are the source type, and the columns are the destination type.
-Blue is the empirical plot of the probability of connection vs distance for those two types; red is the learned function}.
-
-\label{fig:supp:mousecompareari}
-\end{figure}
-
-\FloatBarrier
-\subsection{Model Comparison for Mouse Retina}
-
-We compare the classs found by our model, the traditional infinite
-stochastic block model, and a finite stochastic block model with the number
-of classes set to the best-found number from our iSRM model ($k=12$).
-
-We see from fig~\ref{fig:supp:irmspatialcomparemouse} that the spatial
-variance of our classes closely matches the ground truth, while the
-other two models find spatially-localized
-classs. Figure~\ref{fig:supp:mousecompareari} shows the ARI of our
-clustering, the clustering found by the iSBM,
-as well as the result of parametrically varying the traditional stochastic
-block model for various values of $k$.
-
-The actual clusterings for iSRM, the iSBM, and $SBM_{k=12}$ are shown
-in figures \ref{fig:supp:retinabbclusters},
-\ref{fig:supp:retinaldclusters}, and
-\ref{fig:supp:retinabbfixedclusters} respectively. Each block is a
-cluster, with the true cell type indicated along the x-axis, and the
-cell ID written to the left or right.
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=5in]{mouseretina/spatial_var.pdf}}
-  \caption{The distribution of the spatial extent of cell types found by each of
-the above methods, and ground truth clusters}
-\label{fig:supp:irmspatialcomparemouse}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=5in]{mouseretina/ari_vs_cluster.pdf}}
-  \caption{ARI vs cluster number for the iSRM, iSBM, and the SBM with varying K.}
-\label{fig:supp:mousecompareari}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis)  found by the infinite stochastic block model, dramatically overclustering relative to ground truth}
-\label{fig:supp:retinabbclusters}
-\end{figure}
-
-\begin{figure}[h!]
-
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis) by the infinite spatial-relational model.}
-\label{fig:supp:retinaldclusters}
-\end{figure}
-
-\begin{figure}[h!]
-
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_12-anneal_slow_fixed_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis) via the stochastic block model with $k=12$}
-\label{fig:supp:retinabbfixedclusters}
-\end{figure}
-
-\newpage
-
-\FloatBarrier
-\subsection{Hyperprior grids}
+\subsection{Hyperprior grids and hyperprior inference}
 \label{supp:hyperpriors}
 
-For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80.
-$p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99]$.
+For the mouse retina Logistic-Distance Bernoulli model, we gridded
+$\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0
+and 80.
 
-For the c.
-elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0.
-We globally set $rate_{min}=0.01$.
+For the c. elegans data with the Logistic Distance poisson model, we
+gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points
+bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20
+$\log_{10}$-spaced points between 2.0 and 20.0. We globally set
+$rate_{min}=0.01$.
 
-For the microprocessor with the Logistic Distance fixed lambda Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced points between 10 and 500 and set $\lambda=\mu_{hp}/10$.
-$p_{min} \in \{0.001, 0.01, 0.02\}$  and both $p_\alpha$ and $p_\beta \in \{0.1, 1.0, 2.0\}$.
+For the microprocessor with the Logistic Distance fixed lambda
+Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced
+points between 10 and 500 and set $\lambda=\mu_{hp}/10$. $p_{min} \in
+\{0.001, 0.01, 0.02\}$ and both $p_\alpha$ and $p_\beta \in \{0.1,
+1.0, 2.0\}$.
 
@@
 
-\todo{Concerns about parameter sensitivity}
-\todo{How did we pick the best one we often show?}
-
 \end{document}

### 2014-06-17 experiments/mouseretina/pipeline.py

@@ -210,7 +210,24 @@ def data_create_thold(dbname,
                  'dist_mats' : dist_mats}, 
                 open(retina_outfile, 'w'))
 
+@transform(data_create_thold, suffix(".pickle"), ".png")
+def plot_data_raw((infile, ), outfile):
+    x = pickle.load(open(infile, 'r'))
+    cells = x['cells']
+    print cells
+    f = pylab.figure()
+    ax  = f.add_subplot(1, 1, 1)
+    ai = np.argsort(cells['y'])[::-1]
+    cm = x['conn_mat']
+    cm = cm[ai]
+    cm = cm[:, ai]
 
+    ax.imshow(cm, interpolation='nearest', cmap=pylab.cm.Greys)
+    ax.set_xticks([])
+    ax.set_yticks([])
+    f.tight_layout()
+    f.savefig(outfile, dpi=200)
+    
 
 
 def create_latents_srm_params():
@@ -695,7 +712,7 @@ def plot_z_matrix(exp_results,
     f.savefig(out_filename)
 
 @transform(get_results, suffix(".samples"), [".hypers.pdf"])
-def plot_hypers(exp_results, (plot_hypers_filename,)):
+def plot_hypers_vs_iter(exp_results, (plot_hypers_filename,)):
 
     sample_d = pickle.load(open(exp_results))
     chains = sample_d['chains']
@@ -710,6 +727,37 @@ def plot_hypers(exp_results, (plot_hypers_filename,)):
     chains = [c for c in chains if type(c['scores']) != int]
     if "srm" or "ld" in plot_hypers_filename:
         irm.experiments.plot_chains_hypers(f, chains, data)
+
+    f.savefig(plot_hypers_filename)
+
+@transform(get_results, suffix(".samples"), [".hypers_individual.pdf"])
+def plot_hypers(exp_results, (plot_hypers_filename,)):
+
+    sample_d = pickle.load(open(exp_results))
+    chains = sample_d['chains']
+    
+    exp = sample_d['exp']
+    data_filename = exp['data_filename']
+    data = pickle.load(open(data_filename))
+
+    f = pylab.figure(figsize= (12, 8))
+
+    
+    chains = [c for c in chains if type(c['scores']) != int]
+    if "srm" in plot_hypers_filename:
+        # this is sort of a gross hack
+        states = [c['state'] for c in chains]
+        mu_hps = [s['relations']['R1']['hps']['mu_hp'] for s in states]
[…]

### 2014-06-17 docs/paper.tex
text updates etc

@@
 \usepackage[backend=biber,style=nature]{biblatex}
-\usepackage[disable]{todonotes}
+\usepackage{todonotes}
 \usepackage{grffile}
@@
 of the global network, more recent work aims at identifying very small-scale
-repeat patterns, or “motifs” in networks\autocite{Milo2002}. \todo{motifs vs type}.
+repeat patterns, or “motifs” in networks\autocite{Milo2002}. These motifs
+are defined not between different node types, but rather represent repeated
+patterns of topology.
 
@@
 
-We can thus evaluate our model along three sets of parameters: how
+We evaluate our model along three sets of parameters: how
 closely does our clustering agree with neuroanatomists' knowledge? 

### 2014-06-19 docs/paper.tex
updates to main text

@@
 \documentclass{article}
-\usepackage[english,american]{babel}
+\usepackage[american]{babel}
 \usepackage[backend=biber,style=nature]{biblatex}
@@
 
-We evaluate our model along three sets of parameters: how
-closely does our clustering agree with neuroanatomists' knowledge? 
-Given two cells, how accurately can our model predict the link
-between them? And how closely does the spatial extent (within a layer)
-of our identified types agree with the known neuroanatomists.
+We evaluate our model along three sets of parameters
+(Fig.\ref{fig:mouseretina_compare}): how closely does our clustering agree
+with neuroanatomists' knowledge?  Given two cells, how accurately can
+our model predict the link between them? And how closely does the
+spatial extent (within a layer) of our identified types agree with the
+known neuroanatomists.
 
-We find that given the dataset many techniques allow for good
-link-predictive accuracy but poor agreement with neuroanatomists. In
-fact, when we fit a version of our model fixing the clusterings given
-by neuroanatomists \todo{Add something here}. As we incorporate
-additional information beyond connectivity into our model (soma
-distance, soma depth, synapse density profile) we more closely
-approximate human clustering.
+As a first measure we compare link prediction accuracy across the
+methods (Fig.\ref{fig:mouseretina_compare} B AUC, red).  We find that
+given the dataset many techniques allow for good link-predictive
+accuracy. All the methods allow decent link prediction with an AUC in
+the .9 range. However, our algorithm clearly outperforms the simple
+statistical models that only use connectivity.
+
+As a second measure we compare link prediction accuracy across the
+methods (Fig.\ref{fig:mouseretina_compare} B ARI, blue). We find that our
+algorithm far outperforms the controls. We also find that when it is
+based on more of the same information used by anatomists use then it
+gets better at agreeing with these anatomists. In particular, using
+connectivity, distance, synapse distribution and soma depth leads to
+the highest ARI. When using the available information the algorithm
+produces a good fit to human anatomist judgments.
+
+Finally we look at the spatial extent of the discovered types both
+within a layer and between layers (Fig.\ref{fig:mouseretina_compare}
+C). We see that, in the absence of distance information, mere
+connectivity information results in types which only span a small
+region of space -- essentially “local cliques”. Incorporation of
+distance information results in types which span the entire extent of
+the layer. The depth variance of all models continues to be
+substantially larger than that predicted by human anatomists -- future
+directions of work include attempting to more strongly encode this
+prior belief of laminarity.
 
@@
     both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). a.) The clusters and connectivity versus
+    (small text labels). d.) The clusters and connectivity versus
     distance for connections between Gate and C1, Gate and C2, and C1
@@
     inverters. Purple class are clocked, stateful transistors, orange
-    control the ALU and yellow control the special data bus. b.) The
+    control the ALU and yellow control the special data bus. e.) The
     spatial distribution of the transistors in each cluster.}
@@
 determined the underlying clusters explained by connectivity and
-distance (fig~\ref{fig:celegans}a). A superficial inspection of the results shows
+distance (fig~\ref{fig:otherdata}a). A superficial inspection of the results shows
 clustering into groups consisting roughly homogeneously of motor
@@
 the processor with complex but known structure containing the primary
-8-bit registers X, Y, and S (fig~\ref{fig:mos6502}).
+8-bit registers X, Y, and S (fig~\ref{fig:otherdata}).
 
@@
 the bus. The repeat patterns of spatial connectivity are visible in
-figure~\ref{fig:mos6502}a, showing the man-made horizontal and
+figure~\ref{fig:otherdata}a, showing the man-made horizontal and
 vertical layout of the same types of transistors.
@@
 
-There exist a range of previous approaches to the discovery of
-neural microcircuitry \todo{ xxx as many cites as possible}. These
-generally involve a great deal of manual labor and ad-hoc
-determination of what constitutes a “type” of cell -- to this day
-there are disagreements in the literature as to the “true” types in
-the mammalian retina. Much as phylogenomics has changed our
+There exist a range of previous approaches to the discovery of neural
+microcircuitry\autocite{Mountcastle1957, Douglas1991, Bartho2004,
+  Freund1998}.  These generally involve a great deal of manual labor
+and ad-hoc determination of what constitutes a “type” of cell -- to
+this day there are disagreements in the literature as to the “true”
+types in the mammalian retina. Much as phylogenomics has changed our
 understanding of animal ontologies, modern large scale data will allow
@@
 well as a distance function $d(e_i, e_j)$ representing a (physical)
-distance between adjacent cells. \todo{Add the cell soma caveat?}  See
+distance between adjacent cells.  See
 the supplemental material for extension to multiple connectivity

### 2014-06-23 docs/arxiv.tex
incorporated yarden's first set of suggestions

@@
 \documentclass{article}
-\usepackage[english,american]{babel}
+\usepackage[american]{babel}
 \usepackage[backend=biber,style=nature]{biblatex}
-\usepackage[disable]{todonotes}
+\usepackage{todonotes}
 \usepackage{grffile}
@@
 
-\title{Automatic discovery of circuit structure from neural connectomics}
+\title{Automatic discovery of cell types and microcircuitry from neural connectomics}
 \author{Eric Jonas$^1$, Konrad Kording$^{2, 3,4}$}
@@
 \begin{enumerate}
-  \item Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology, Cambridge, Massachusetts
-  \item Department of Physical Medicine and Rehabilitation, Northwestern University and Rehabilitation Institute of Chicago, Chicago, Illinois
+  \item Electrical Engineering and Computer Science, University of California, Berkeley
+  \item Department of Physical Medicine and Rehabilitation,
+    Northwestern University and Rehabilitation Institute of Chicago,
+    Chicago, Illinois
   \item Department of Physiology, Northwestern University, Chicago, Illinois
@@
 \begin{abstract}
-  New techniques produce massive data about neural connectivity,
-  necessitating new analysis methods to discover underlying
-  structure. It has long been assumed that discovering the local
-  patterns of microcircuitry is the key to understanding neural
-  function. Here we developed a nonparametric Bayesian technique that
-  identifies neuron types and circuitry patterns in connectomics
-  data. We show that the approach recovers known neuron types in the
-  retina, reveals interesting structure in the nervous system of
-  c. elegans, and automatically discovers the structure of
-  microprocessors. Our approach extracts structural meaning from
-  connectomics, enabling new approaches of deriving anatomical
-  insights from these emerging datasets.
-
+  Neural connectomics has begun producing massive amounts of data,
+  necessitating new analysis methods to discover the biological and
+  computational structure. It has long been assumed that discovering
+  neuron types and their relation to microcircuitry is crucial to
+  understanding neural function. Here we developed a nonparametric
+  Bayesian technique that identifies neuron types and microcircuitry
+  patterns in connectomics data. It combines the information
+  traditionally used by biologists, including connectivity, cell body
+  location and the spatial distribution of synapses.  We show that the
+  approach recovers known neuron types in the retina and enables
+  predictions of connectivity, better than simpler algorithms. It also
+  can reveal interesting structure in the nervous system of
+  c. elegans, and automatically discovers the structure of a
+  microprocessor.  Our approach extracts structural meaning from
+  connectomics, enabling new approaches of automatically deriving
+  anatomical insights from these emerging datasets.
 \end{abstract}
@@
 \section*{Introduction}
-Emerging techniques \autocite{Morgan2013,Zador2012} promise to
-quantify the location of each neuron within a volume of interest,
-along with its connections to all other neurons. Far exceeding the
-capacity of neuroanatomists to trace small circuits, this leads to
-enormous datasets that quantify aspects of the nervous system. The impact of
-this kind of high-throughput transition has been seen before -- 
-rise of sequencing techniques necessitated the
-development of novel computational methods to understand genomic
-structure, ushering in an era of bioinformatics as an independent
-discipline \autocite{Koboldt2013}.
+Emerging connectomics techniques \autocite{Morgan2013,Zador2012}
+promise to quantify the location and connectivity of each neuron
+within a tissue volume. These massive datasets will far exceed the
+capacity of neuroanatomists to manually trace small circuits, thus
+necessitating computational, quantitative, and automatic methods for
+understanding neural circuit structure.  The impact of this kind of
+high-throughput transition has been seen before -- rise of sequencing
+techniques necessitated the development of novel computational methods
+to understand genomic structure, ushering in  bioinformatics
+as an independent discipline \autocite{Koboldt2013}.
 
@@
 morphological, but they also differ in the way they connect to one
-another. In fact, the idea of well defined local connectivity patterns
-has been prominent in many areas, ranging from sensory (e.g. retina,
-\autocite{Masland2001} via processing (e.g. neocortex \autocite{Mountcastle1997})
-to movement (e.g. spinal cord) \autocite{Grillner2005}. Indeed, such
-repeated patterns also exists in human-made computing circuits. It
-remains an important challenge to develop algorithms to use anatomical
-data, e.g. connectomics, to back out underlying connectivity motifs.
+another. In fact, the idea of well defined, type-dependent local
+connectivity patterns (microcircuits) has been prominent in many
+areas, from sensory (e.g. retina, \autocite{Masland2001} to processing
+(e.g. neocortex \autocite{Mountcastle1997}) to movement (e.g. spinal
+cord) \autocite{Grillner2005}. These sorts of repeated computing
+patterns are a common feature of computing systems, even arising in
+human-made computing circuits. It remains an important challenge to
+develop algorithms to use anatomical data, e.g. connectomics, to
+automatically back out underlying microcircuitry.
 
@@
 the types of scaling present in the network \autocite
-{WattsStrogatz1998} .  While this approach leads to an understanding
-of global network, more recent work aims at identifying very small-scale
-repeat patterns, or “motifs” in networks\autocite{Milo2002}
+{WattsStrogatz1998}.  While this approach leads to an understanding
+of the global network, more recent work aims at identifying very small-scale
+repeat patterns, or “motifs” in networks\autocite{Milo2002}. These motifs
+are defined not between different node types, but rather represent repeated
+patterns of topology.
 
@@
 exhibiting a strong spatial dependence arising from the complex
-genetic, chemical, and activity-based neural development process.
+genetic, chemical, and activity-based neural development processes.
 
-Here we describe a Bayesian non-parametric model that can discover
-circuit structure automatically from connectomics data: both the cell
-types and their spatial patterns of interconnection. We apply it to
-three computing systems: the mouse retina connectome
-\autocite{Helmstaedter2013}, the c. elegans connectome
-\autocite{White1986}, and a ``connectome'' of a classical
-microprocessor \autocite{James2010}. Note we do not explicitly
-consider cellular morphology, absolute position (cell layer), or any
-chemical or genetic information. Yet in all cases, we discover
-cell-type to cell-type connectivity that depends on spatial distance
-and these connectivity functions depend on the two involved cell
-types. Comparing the cell types discovered by the algorithms with
-those obtained by anatomists based on orthogonal information reveals a
-high degree of agreement. We present a scalable approach to infer
-microcircuitry from connectivity data.
+To address these challenges, here we describe a Bayesian
+nonparametric model that can discover circuit structure automatically
+from connectomics data: the cell types, their spatial patterns of
+interconnection, and the locations of somata and synapses. We show
+that by incorporating this additional information, our model both
+accurately predicts the connection as well as agrees
+with human neuroanatomists as to the identification of cell types.
+
+We primarily focus on the recently-released mouse retina connectome
+\autocite{Helmstaedter2013}, but additionally examine the c. elegans
+connectome \autocite{White1986}, and then ``connectome'' of a
+classical microprocessor \autocite{James2010}. Comparing the cell
+types discovered by the algorithms with those obtained manually by
+human anatomists reveals a high degree of agreement. We present a
+scalable approach to infer microcircuitry from connectomics data.
 
@@
 \section{Results}
-We use a Bayesian structure discovery approach that is based on
-prior knowledge about the way positions and cell types affect the
-probability of connections between cells, a so-called generative
-model. We term our model the ``infinite spatial-relational model'' (iSRM), 
-in keeping with the existing nonparametric literature.  
-The probability of connections between pairs of neurons
-decreases with distance, but how high the baseline probability of
-connection is and how quickly it decays with distance depends on the
-two cell-types. Note that this simplistic prior information is all that 
-we tell our model; no additional physiological information is included.
 
-We combine annealing with Markov Chain Monte Carlo to
-perform simultaneously joint posterior inference over the structure,
-parameters, and hyperparameters of the model (see Methods). This
-procedure automatically learns the number of cell types, which cells 
-belong to which types, along with the 
-the distance and type dependent connectivity probabilities (see
-fig~\ref{fig:overview}). These results can then be compared with the
-results of previous anatomical studies that have used other aspects of
-neurons, such as chemical properties and cell-morphology to divide
-neurons into types.
+We start with the generic assumption that cell types exist -- a cell
+is a member of a single, unobserved type -- and these types affect
+soma depth, distribution of synapses, as well as a cell type and
+distance dependent connection probability. From this assumption we
+develop a generative Bayesian model that estimates the underlying cell
+types and how they connect.  We start with a model for connectivity,
+the infinite stochastic block model (iSBM)\autocite{Kemp2006a,Xu2006},
+which has been shown to meaningfully cluster connection graphs while
+learning the number of hidden groups, or types. We extend this
+approach by adding distance dependence to model salient aspects of
+microcircuitry. We additionally model cell body depth as a mixture of
+Gaussians and each cell’s synapse density profile as mixture of
+mixtures of Gaussians.
+
+For example, retinal ganglion cells may synapse on nearby, but not far
+away, amacrine cells, with ganglion cells being superficial and having
+a broad distribution of synapses.  This procedure automatically learns
+the number of cell types, which cells belong to which types, along
+with the distance and type dependent connectivity probabilities (see
+fig~\ref{fig:overview}). We apply our algorithm to datasets from mouse
+retina, c-elegans and a historical microprocessor. Anatomists classify
+cells based on many features, giving us a meaningful baseline to
+compare against.
 
 Before we can use a probabilistic model to analyze real data we first
-need to assess if the model works properly. We thus started out by
-simulating data for which we know the correct structure and comparing
-the estimated structure based on the algorithm (see methods) with the
-one we used for simulation. We find that the model does a good job of
-recovering the structure we used to generate the data
-(fig~\ref{fig:synthetic}a) Indeed, it does far better than a simpler
-and well established model (the infinite stochastic block model, or
-iSBM) that assumes that there is no distance dependence to the
-connectivity. The model converges relatively quickly to an estimate of
-the most probable values for the cell types, which is enabled by using
-a combination of simulated annealing and parallelized Markov chain
-Monte Carlo (see methods for details). It appears that the model
-behaves the way it should and allows application to relatively large
-datasets.
+need to assess if the model works properly. We thus simulate data for
+which we know the correct structure and comparing the estimated
+structure based on the algorithm (see methods) with the one we used
+for simulation. We find that the model does a good job of recovering
+the correct number of cell types, (fig~\ref{fig:synthetic}a), the the
+cell identities (fig~\ref{fig:synthetic}b), and the spatial extent of
+each type (fig~\ref{fig:synthetic}c).  Indeed, it does far better than
+the existing infinite stochastic block model that assumes that there
+is no distance dependence to the connectivity. The model converges
+relatively quickly to an estimate of the most probable values for the
+cell types, which is enabled by using a combination of simulated
+annealing and parallelized Markov chain Monte Carlo (see methods for
+details). It appears that the model behaves the way it should and
+allows application to relatively large datasets.
 
@@
   \centerline{\includegraphics{synthetic.ai}}
-  \caption{Correct recovery of true numbers of hidden classes
-in synthetic data. a.) Other models over-cluster the number
-of classes as they fail to take distance into account, whereas
-our model finds close to the true number of classes. b.) ARI, a measure
-of ``correct clustering'', for
-different true class counts, between our model and a traditional
-iSBM. c.) For synthetic data, 
-the iSBM overclusters, finding a large number of spatially-localized types. 
-In contrast, the infinite spatial-relational model recovers the true spatial
-extent of the underlying types.}
+  \caption{Correct recovery of true numbers of hidden types in
+    synthetic data when incorporating spatial information. a.) The
+    infinite stochastic block model (which only uses connectivity
+    information) over-estimates the number of classes as it fails to
+    take distance into account, whereas our modeling of the
+    combination of distance and connectivity finds close to the true
+    number of classes. b.) ARI, a measure of ``correct clustering'',
+    for different true class counts, between our model and a
+    traditional iSBM. c.) For synthetic data, type discovery based
+    only on connectivity are tiny spatially-localized groups -- in
+    contrast, our model recovers the true spatial extent of the
+    underlying types.  }
 \label{fig:synthetic}
@@
   \caption{Discovering cell classes in the mouse retina connectome.
-a.) Input connectivity data for 950 cells for which soma positions were known. b. clustered connectivity matrix. c.) connectivity diagram showing our clusters, as
-well as the cell depth and known cell type of each cell. Our model agrees
-with the cell types traditionally identified by anatomists, and identifies 
-laminar-specific connectivity patterns. d.) Connectivity between our
-clusters as a function of distance -- the cluster consisting primarily of
-retinal ganglion cells (lower left-center) exhibits the expected near and
-far connectivity. e. The spatial distribution of our cell types -- each cell type tesselates space.}
+    a.) Input connectivity data for 950 cells for which soma positions
+    were known. b. clustered connectivity matrix. c.) connectivity
+    diagram showing our clusters, as well as the cell depth and
+    anatomist-labeled cell type of each cell. Our model shows moderate
+    agreement with the cell types traditionally identified by
+    anatomist laminar-specific connectivity patterns. 
+    d.) Connectivity between our clusters as a function
+    of distance -- the cluster consisting primarily of retinal
+    ganglion cells (lower left-center) exhibits the expected near and
+    far connectivity. e.)  The spatial distribution of our cell types
+    -- each cell type tesselates space.}
 \label{fig:mouseretina}
@@
 
-The retina of the mouse \autocite{Masland2001} is an example where we
-should expect to have connectivity patterns that are well approximated
-by our generative model. It is known that there are multiple classes
-of cells that can be grouped into ganglion cells that transmit
+The mouse retina \autocite{Masland2001} is a neural circuit which we
+expect to have connectivity patterns that are well approximated by our
+generative model. It is known that there are multiple classes of cells
+that can be broadly grouped into: ganglion cells that transmit
 information to the rest of the brain, bipolar cells that connect
 between different cells, and amacrine cells that feed into the
-ganglion cells \autocite{cite some review}. Recent research
-\autocite{Helmstaedter2013} has produced a large dataset containing
-both the types of cells from orthogonal approaches, and also the
-connectivity matrix between all reconstructed cells. We want to know
-to which level we can classify neurons into types based on this
-connectivity only.
+ganglion cells. Recent research \autocite{Helmstaedter2013} has
+produced a large dataset containing both the types of cells from
+orthogonal approaches, and also the connectivity matrix between all
+reconstructed cells.
 
-The algorithm took 8 hours to converge to a locally optimal solution,
-dividing neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
+The algorithm took less than 2 hours to perform inference, dividing
+neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
 wedge is a type). For each pair of neurons there is a specific
-distance dependent connection probability (fig~\ref{fig:mouseretina}b,c,d),
-which is well approximated by the model fit. Moreover, each type of
-cell is rather isotropically distributed across space
-(fig~\ref{fig:mouseretina}e) as should be expected for true cell types. The
-algorithm is thus able to extract meaningful structure from the data.
+distance dependent connection probability
+(fig~\ref{fig:mouseretina}b,c,d), which is well approximated by the
+model fit. Moreover, each type of cell is rather isotropically
+distributed across space (fig~\ref{fig:mouseretina}e) as should be
+expected for true cell types.
 
@@
 
-\todo{Do we assign any of the unknown cells correctly?  are types 72-79
-“different” or simply unknown?}
+Our fully Bayesian model produces a distribution over probable
+clusterings.  Figure~\ref{fig:mouseretina_compare} shows this
+posterior distribution as a cell-cell coassignment matrix, sorted to
+find maximum block structure. Each large, dark block represents a
+collection of cells believed with strong probability to be of the same
+type. When we plot (fig~\ref{fig:mouseretina_compare}b) the
+anatomist-derived cell types along the left, we can see that each
+block consists of a roughly-homogeneous collection of types.
 
-We compared the results of our model to the traditional iSBM to see if
-the addition of spatial information was necessary. We find that, as
-expected from our synthetic experiments, the iSBM massively
-over-segmented the data, finding spatially-localized “communities” of
-cells which bore little relation to known types.  The average spatial
-extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. However,
-in the ground truth by anatomists the average distance was $34.3\pm7.6
-\mu m$ and $35.3\pm 2.4 \mu m$, quite close to the $34.5\pm 1.0 \mu m$
-discovered by our algorithm. Modeling the distance-dependence prevents
-over-segmentation and thus allows more meaningful estimates.
+We evaluate our model along three sets of parameters
+(Fig.\ref{fig:mouseretina_compare}): how closely does our clustering agree
+with neuroanatomists' knowledge?  Given two cells, how accurately can
+our model predict the link between them? And how closely does the
+spatial extent (within a layer) of our identified types agree with the
+known neuroanatomists.
+
+As a first measure we compare link prediction accuracy across the
+methods (Fig.\ref{fig:mouseretina_compare} B AUC, red).  We find that
+given the dataset many techniques allow for good link-predictive
+accuracy. All the methods allow decent link prediction with an AUC in
+the .9 range. However, our algorithm clearly outperforms the simple
+statistical models that only use connectivity.
+
+As a second measure we compare link prediction accuracy across the
+methods (Fig.\ref{fig:mouseretina_compare} B ARI, blue). We find that our
+algorithm far outperforms the controls. We also find that when it is
+based on more of the same information used by anatomists use then it
+gets better at agreeing with these anatomists. In particular, using
+connectivity, distance, synapse distribution and soma depth leads to
+the highest ARI. When using the available information the algorithm
+produces a good fit to human anatomist judgments.
+
+Finally we look at the spatial extent of the discovered types both
+within a layer and between layers (Fig.\ref{fig:mouseretina_compare}
+C). We see that, in the absence of distance information, mere
+connectivity information results in types which only span a small
+region of space -- essentially “local cliques”. Incorporation of
+distance information results in types which span the entire extent of
+the layer. The depth variance of all models continues to be
+substantially larger than that predicted by human anatomists -- future
+directions of work include attempting to more strongly encode this
+prior belief of laminarity.
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics{comparison.ai}}
+  \caption{a. Cell-cell assignment matrix reflecting the posterior
+    distribution of clusterings found by our model. Large contiguous
+    black areas are cells that were considered the same type in the
+    majority of samples. b. The cells and their true types sorted to
+    agree with the coassignment matrix. c. A comparison of the
+    predictive accuracy (Area under the curve) for hand-labeled
+    anatomical data, versus inclusion of additional sources of
+    information. d. ) The spatial extent (in depth and area) of the
+    types identified by humans and our various algorithmic
+    approaches.}
+
+\label{fig:mouseretina_compare}
+\end{figure}
 
@@
     both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). }
-  \label{fig:celegans}
+    (small text labels). d.) The clusters and connectivity versus
+    distance for connections between Gate and C1, Gate and C2, and C1
+    and C2 terminals on a transistor. Red and teal types have a
+    terminal pulled down to ground and mostly function as
+    inverters. Purple class are clocked, stateful transistors, orange
+    control the ALU and yellow control the special data bus. e.) The
+    spatial distribution of the transistors in each cluster.}
+  \label{fig:otherdata}
 \end{figure}
@@
 We have presented a machine learning technique that allows cell types
-and microcircuitry to be discovered using only distance and
-connectivity data. We have shown its applicability to a newly
-published connectomics dataset from the mouse retina, the classical
-connectivity of the worm c. elegans, and the connectivity of a
-historical man-made microprocessor. We have found that distance and
-connectivity alone are sufficient to discover cell types and
-microcircuitry that were known to exist in the systems based on decades
-of previous research which had utilized extensive additional
-information.
+and microcircuitry to be discovered from connectomics data.  We have
+shown its applicability to a newly published connectomics dataset from
+the mouse retina. When compared to existing methods, we show how the
+incorporation of all of this data yields results that combine both
+high link-prediction accuracy and high agreement with human
+anatomists.  I also reveals interesting structure when applied to the
+classical connectivity of the worm c. elegans, and the connectivity of
+a historical man-made microprocessor. We have found that combining the
+available data types allows us to discover cell types and
+microcircuitry that were known to exist in the systems based on
+decades of previous research and allows good prediction of
+connectivity.
 
-For probabilistic models like ours, no known solution exists to
+For our probabilistic models, no known solution exists to
 exactly find the most probable parsing of the neurons into cell-types
-and connectivity patterns . We employ a collection of Markov-chain
+and connectivity patterns. We employ a collection of Markov-chain
 Monte carlo techniques (see Methods) but while different
@@
 
-There exist a range of previous approaches to the discovery of
-neural microcircuitry \todo{Do we want a citation here?}. These
-generally involve a great deal of manual labor and ad-hoc
-determination of what constitutes a “type” of cell -- to this day
-there are disagreements in the literature as to the “true” types in
-the mammalian retina. Much as phylogenomics has changed our
+There exist a range of previous approaches to the discovery of neural
+microcircuitry\autocite{Mountcastle1957, Douglas1991, Bartho2004,
+  Freund1998}.  These generally involve a great deal of manual labor
+and ad-hoc determination of what constitutes a “type” of cell -- to
+this day there are disagreements in the literature as to the “true”
+types in the mammalian retina. Much as phylogenomics has changed our
 understanding of animal ontologies, modern large scale data will allow
@@
 
-Ultimately, we will need to combine information across different
-sources into a joint understanding of neural function. Distinct cell
+Our model combines connectivity, cellular, and synaptic properties,
+and suggests the way towards combining even richer data. Distinct cell
 types differ in morphology, connectivity, transcriptomics, relation to
-behavior or stimuli and many other ways. Algorithms based on each type
-information can then be combined to synthesize all the available
-information from one experiment or even across experiments into a
-joint model of brain function.
+behavior or stimuli and many other ways. Algorithms combining this
+data and type type information may allow us to synthesize all the
+available information from one experiment or even across experiments
+into a joint model of brain structure and function.
+
+Eventually, algorithms will have to become a central tool for
+anatomists, as it will progressively become impossible for humans to
+parse the huge datasets. This transition may follow a similar
+transition to the one underway in molecular biology (with gene-finding
+algorithms) and evolutionary biology with (computational
+phylogenetics).  In fact, it is known that there are significant
+disagreements across human anatomists and computational approaches
+promise to make the debates more data driven.
 
@@
 
-We take as input a connectivity matrix $R$ defining the connections
-between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
-e_j)$ representing a (physical) distance between adjacent cells. See
+For the basic link-distance model, we take as input a connectivity
+matrix $R$ defining the connections between cell $e_i$ and $e_j$, as
+well as a distance function $d(e_i, e_j)$ representing a (physical)
+distance between adjacent cells.  See
 the supplemental material for extension to multiple connectivity
@@
 
-We chain these three kernels together, and then globally
-anneal on the likelihood from a temperature of $T=64$ down to 
-$T=1$ over 300 iterations unless otherwise indicated, and
-then run the chain for another $100$ iterations.
+We chain these three kernels together, and then globally anneal on the
+likelihood from a temperature of $T=64$ down to $T=1$ over $900$
+iterations unless otherwise indicated, and then run the chain for
+another $100$ iterations. We then generate at least $20$ samples, each
+taken from the end of a single Markov chain initialized from different
+random initial points in the state space. For visualization we pick the
+chain with the highest log likelihood, but for all numerical
+comparisons (including link probability and ARI) we use this full
+collection of samples from the posterior distribution to estimate the
+resulting statistics.
 
-To pick the MAP, we run up to $100$ simultaneous chains in parallel,
-initialized from different random initial points in the state space,
-and then pick the chain with the highest log likelihood. Early
-experiments attempted to use the full ensemble of chains as an
-estimate of the posterior but were abandoned due to the difficulty of
-summarizing the posterior over such a complex state space, which is an
-ongoing area of research in the probabilistic modeling community
-\todo{cite more? missing}.
+\subsection{Link Prediction}
+As a proxy for link-prediction accuracy we compute the probability of a link
+between two cells using each model, trained fully on the data. While this method
+is potentially prone to overfitting, the overfitting will be shared across models
+and in fact will preferentially bias in favor of competing models
+which over-cluster the data. We use a full collection of posterior samples when
+computing the link probability, and then compute the area under the ROC curve for
+each.
 
-\todo{Validation: do we want to write extensively about posterior
-  enumeration tests, etc. ? }
+\subsection{Model Comparison}
+We compare our model with a standard network clustering model, the
+latent-position clutsering model. This model assumes each cell belongs
+to one of K clusters, and each cluster is associated with a
+$d-$dimensional Gaussian distribution. The probability of a link is
+then a function of the distance between the data points in this
+continuous-space. We use \autocite{VBLPCM} a variational
+implementation provided in R, parametrically varying the number of
+latent dimensions and the number of requested groups.  While this
+model provides reasonable link predictive accuracy, the clusterings
+dramatically disagree with those from human anatomists.
+
+\subsection{Parameters}
+
+Hierarchical generative models can be sensitive to hyperparameter
+settings, thus for most hyperparamters we perform inference. In cases
+where we cannot we run separate collections of markov chains at
+separate settings and show the results across all pooled
+parameters. For the case of the mouse retina data, we consisder
+maximum link probability $p_{max} \in \{0.95, 0.9, 0.7\}$, variance
+scales for the synapse density profile of $\sigma^2 \in \{0.01, 0.1,
+1.0\}$ (of normalized depth), and $K \in \{2, 3\}$ possible synapse
+density profile mixture components. For the connectivity-distance-only
+model we actually perform inference over both $p_{max}$ and $p_{min}$.
+
+\subsubsection{Mixing of our Markov chains} 
+Evaluating whether or not approximate inference methods, such as MCMC,
+produce samples which are valid approximations of the posterior
+distribution is an ongoing area of research in the computational
+statistics community. We use a rough proxy here -- synthetic
+likelihood evaluation.  For synthetic datasets of sizes comparable to
+our real data size, do we recover known ground truth information after
+running our markov chains for the appropriate amount of time?
+
+Figures \ref{fig:mixing:ari} and \ref{fig:mixing:scorevstime} shows
+the ARI to ground truth and the total log score as a function of
+runtime.  We see dramatic changes in log score initially as we vary
+the temperature, stabilizing as runtime progresses, for each
+chain. Then we see the characteristic jumps between nearby modes
+towards the end of the run, in both log score and ARI.  Importantly,
+regardless of whether our model over- or under-estimates the exact
+posterior variance about the network, we find points in the latent
+variable space that are both predictive \textit{and} parsimonious,
+largely agreeing with the human anatomists and predicting existing
+connections.
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=60mm]{mixing.ari.pdf}}
+  \caption{Adjusted rand index for synthetic data as a function of run iteration. }
+\label{fig:mixing:ari}
+\end{figure}
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=130mm]{mixing.scorevstime.pdf}}
+  \caption{Total model score (log score) vs iteration}
+\label{fig:mixing:scorevstime}
+\end{figure}
 
@@
 
-\href{http://spatialrelation.github.io }{http://spatialrelation.github.io}
+\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/netmotifs/}
+
+and the content of this paper along with scripts to run
+experiments and generate all figures can be found at
+
+\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/connect-disco-paper/}
+
+All preprocessed data has been made publicaly available as well.
+
+Please contact the author for pre-publication access.
 
@@
 \FloatBarrier
-
-\subsection{Microprocessor Netlist annotation}
-\label{supp:mos6502netlist}
-This is a matrix consisting of the different clusters of electrical nets for 
-the registers, with the colors consistent with figure~\ref{fig:mos6502}.
-
-\includepdf[pages=-]{mos6502.clusters.pdf}
-
-\FloatBarrier
-\subsection{Model Comparison for Mouse Retina}
-
-We compare the classs found by our model, the traditional infinite
-stochastic block model, and a finite stochastic block model with the number
-of classes set to the best-found number from our iSRM model ($k=12$).
-
-We see from fig~\ref{fig:supp:irmspatialcomparemouse} that the spatial
-variance of our classes closely matches the ground truth, while the
-other two models find spatially-localized
-classs. Figure~\ref{fig:supp:mousecompareari} shows the ARI of our
-clustering, the clustering found by the iSBM,
-as well as the result of parametrically varying the traditional stochastic
-block model for various values of $k$.
-
-The actual clusterings for iSRM, the iSBM, and $SBM_{k=12}$ are shown
-in figures \ref{fig:supp:retinabbclusters},
-\ref{fig:supp:retinaldclusters}, and
-\ref{fig:supp:retinabbfixedclusters} respectively. Each block is a
-cluster, with the true cell type indicated along the x-axis, and the
-cell ID written to the left or right.
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=5in]{mouseretina/spatial_var.pdf}}
-  \caption{The distribution of the spatial extent of cell types found by each of
-the above methods, and ground truth clusters}
-\label{fig:supp:irmspatialcomparemouse}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=5in]{mouseretina/ari_vs_cluster.pdf}}
-  \caption{ARI vs cluster number for the iSRM, iSBM, and the SBM with varying K.}
-\label{fig:supp:mousecompareari}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis)  found by the infinite stochastic block model, dramatically overclustering relative to ground truth}
-\label{fig:supp:retinabbclusters}
-\end{figure}
-
-\begin{figure}[h!]
-
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis) by the infinite spatial-relational model.}
-\label{fig:supp:retinaldclusters}
-\end{figure}
-
-\begin{figure}[h!]
-
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_12-anneal_slow_fixed_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis) via the stochastic block model with $k=12$}
-\label{fig:supp:retinabbfixedclusters}
-\end{figure}
-
-\newpage
-
-\FloatBarrier
-\subsection{Hyperprior grids}
+\subsection{Hyperprior grids and hyperprior inference}
 \label{supp:hyperpriors}
 
-For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80.
-$p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99]$.
+For the mouse retina Logistic-Distance Bernoulli model, we gridded
+$\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0
+and 80.
 
-For the c.
-elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0.
-We globally set $rate_{min}=0.01$.
+For the c. elegans data with the Logistic Distance poisson model, we
+gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points
+bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20
+$\log_{10}$-spaced points between 2.0 and 20.0. We globally set
+$rate_{min}=0.01$.
 
-For the microprocessor with the Logistic Distance fixed lambda Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced points between 10 and 500 and set $\lambda=\mu_{hp}/10$.
-$p_{min} \in \{0.001, 0.01, 0.02\}$  and both $p_\alpha$ and $p_\beta \in \{0.1, 1.0, 2.0\}$.
+For the microprocessor with the Logistic Distance fixed lambda
+Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced
+points between 10 and 500 and set $\lambda=\mu_{hp}/10$. $p_{min} \in
+\{0.001, 0.01, 0.02\}$ and both $p_\alpha$ and $p_\beta \in \{0.1,
+1.0, 2.0\}$.
 

### 2014-06-23 docs/paper.tex
incorporated yarden's first set of suggestions

@@
   traditionally used by biologists, including connectivity, cell body
-  location and the spatial distribution of synapses.  We show that the
+  location and the spatial distribution of synapses, in a
+  principled and probabilstically-coherent manner. We show that the
   approach recovers known neuron types in the retina and enables
@@
   can reveal interesting structure in the nervous system of
-  c. elegans, and automatically discovers the structure of a
+  C. elegans, and automatically discovers the structure of a
   microprocessor.  Our approach extracts structural meaning from
@@
 necessitating computational, quantitative, and automatic methods for
-understanding neural structure.  The impact of this kind of
+understanding neural circuit structure.  The impact of this kind of
 high-throughput transition has been seen before -- rise of sequencing
 techniques necessitated the development of novel computational methods
-to understand genomic structure, ushering in an era of bioinformatics
+to understand genomic structure, ushering in  bioinformatics
 as an independent discipline \autocite{Koboldt2013}.
@@
 connectivity patterns (microcircuits) has been prominent in many
-areas, ranging from sensory (e.g. retina, \autocite{Masland2001} via
-processing (e.g. neocortex \autocite{Mountcastle1997}) to movement
-(e.g. spinal cord) \autocite{Grillner2005}. These sorts of repeated
-computing patterns are a common feature of computing systems, even
-arising in human-made computing circuits. It remains an important
-challenge to develop algorithms to use anatomical data,
-e.g. connectomics, to automatically back out underlying
-microcircuitry.
+areas, from sensory (e.g. retina, \autocite{Masland2001} to processing
+(e.g. neocortex \autocite{Mountcastle1997}) to movement (e.g. spinal
+cord) \autocite{Grillner2005}. These sorts of repeated computing
+patterns are a common feature of computing systems, even arising in
+human-made computing circuits. It remains an important challenge to
+develop algorithms to use anatomical data, e.g. connectomics, to
+automatically back out underlying microcircuitry.
 
@@
 
-The discovery of structure in probabilistic graphs is a crucial aspect
-in machine learning. Commonly used algorithms include
+The discovery of structure in probabilistic graphs is a well-known
+problem in machine learning. Commonly used algorithms include
 community-based-detection methods \autocite{Girvan2002}, and
@@
 We primarily focus on the recently-released mouse retina connectome
-\autocite{Helmstaedter2013}, but additionally examine the c. elegans
+\autocite{Helmstaedter2013}, but additionally examine the C. elegans
 connectome \autocite{White1986}, and then ``connectome'' of a
@@
 types discovered by the algorithms with those obtained manually by
-human anatomists reveals a high degree of agreement. We present a
-scalable approach to infer microcircuitry from connectomics data.
+human anatomists reveals a high degree of agreement. We thus present a
+scalable probabilistic approach to infer microcircuitry from
+connectomics data available today and in the future.
 
@@
     types. d. shows the cell connectivity matrix with cells of the
-    same type grouped together. e.) shows the learned probability of
+    same type grouped together. e. shows the learned probability of
     connection between our different types at various distances -- in
     this case, the cells are likely to connect when they are
-    close. f. shows a connectivity profile with virtually no
-    probability of connectivity. We can plot this these types g.) to
-    show the relationship and spatial connections, and how probable
-    various types are to connect to one another. h. The connectivity
-    between our discovered types changes as a function of distance. }
+    close. f. shows the probability of connection between two cell types
+that very rarely connect -- there's a background ``base'' connection rate
+to account for errors in data, but the probability is very low. 
+We can plot this these types g.) to show the relationship and spatial
+connections, and how probable various types are to connect to one
+another. h. The connectivity between our discovered types changes as a
+function of distance. }
 
@@
 
-We start with the generic assumption that cell types exist -- a cell
-is a member of a single, unobserved type -- and these types affect
-soma depth, distribution of synapses, as well as a cell type and
-distance dependent connection probability. From this assumption we
-develop a Bayesian algorithm that estimates the underlying cell types.
+We build a structured probabilistic model which begins with the
+generic notion of a cell being a member of a single, unobserved type
+-- and these types affect soma depth, distribution of synapses, as
+well as a cell type and distance dependent connection probability. For
+example, retinal ganglion cells may synapse on nearby, but not far
+away, amacrine cells, with ganglion cells being superficial and having
+a broad distribution of synapses.
+
+From these assumptions (priors) develop a generative Bayesian model
+that estimates the underlying cell types and how they connect. We take
+as input (fig ~\ref{fig:overview}a) the connectivity matrix of cells,
+a matrix of the distance between cells, and the per-cell soma depth
+(fig ~\ref{fig:overview}b) and the depth profile of the cell's
+synapses (fig ~\ref{fig:overview}c). We perform joint probabilistic
+inference to automatically learn the number of cell types, which cells
+belong to which type, their type-specific connectivity, and how
+connections between types vary with distance. We also simultaneously
+learn the soma depth associated with each type and the typical
+synaptic density profile (fig ~\ref{fig:overview}d.).
+
+We apply our algorithm to datasets from mouse
+retina, C. elegans and a historical microprocessor. Anatomists
+classify cells based on many features, giving us a meaningful baseline
+to compare against.
+
 We start with a model for connectivity, the infinite stochastic block
@@
 hidden groups, or types. We extend this approach by adding distance
-dependence to model salient aspects of microcircuitry. We additionally
-model cell body depth as a mixture of Gaussians and each cell’s
-synapse density profile as mixture of mixtures of Gaussians.
+dependence to model salient aspects of microcircuitry via logistic and
+exponential distance-link functions.  We additionally model cell body
+depth unimodally and synapse density profile multimodially (see
+Methods for mathematical details).
 
-For example, retinal ganglion cells may synapse on nearby, but not far
-away, amacrine cells, with ganglion cells being superficial and having
-a broad distribution of synapses.  This procedure automatically learns
-the number of cell types, which cells belong to which types, along
-with the distance and type dependent connectivity probabilities (see
-fig~\ref{fig:overview}). We apply our algorithm to datasets from mouse
-retina, c-elegans and a historical microprocessor. Anatomists classify
-cells based on many features, giving us a meaningful baseline to
-compare against.
-
-Before we can use a probabilistic model to analyze real data we first
-need to assess if the model works properly. We thus simulate data for
-which we know the correct structure and comparing the estimated
-structure based on the algorithm (see methods) with the one we used
-for simulation. We find that the model does a good job of recovering
-the correct number of cell types, (fig~\ref{fig:synthetic}a), the the
-cell identities (fig~\ref{fig:synthetic}b), and the spatial extent of
-each type (fig~\ref{fig:synthetic}c).  Indeed, it does far better than
-the existing infinite stochastic block model that assumes that there
-is no distance dependence to the connectivity. The model converges
+To validate our model, we performed a series of simulations to test if
+the model can accurately recover the true underlying network structure
+and cell type identity.  We thus simulate data for which we know the
+correct structure and comparing the estimated structure based on the
+algorithm (see methods) with the one we used for simulation. We find
+that the model does a good job of recovering the correct number of
+cell types, (fig~\ref{fig:synthetic}a), the the cell identities
+(fig~\ref{fig:synthetic}b), and the spatial extent of each type
+(fig~\ref{fig:synthetic}c).  Indeed, it does far better than the
+existing infinite stochastic block model that assumes that there is no
+distance dependence to the connectivity. The model converges
 relatively quickly to an estimate of the most probable values for the
@@
 annealing and parallelized Markov chain Monte Carlo (see methods for
-details). It appears that the model behaves the way it should and
-allows application to relatively large datasets.
+details). Thus we can apply our model to simulated datasets with structure
+and scale similar to that of our biological datasets and recover the known 
+correct structure.
 
@@
 leading to early measurement of the connectome. Unlike the retina,
-only the motor neurons in c. elegans exhibit regular distribution in
+only the motor neurons in C. elegans exhibit regular distribution in
 space -- along the body axis. Most interneurons are
@@
   \caption{Discovering connectivity and type in both
-    \textit{c. elegans} and the MOS 6502 microprocessor.  a.) Initial
-    \textit{c. elegans} cell connectivity matrix, red are
+    \textit{C. elegans} and the MOS 6502 microprocessor.  a.) Initial
+    \textit{C. elegans} cell connectivity matrix, red are
     (directional) chemical synapses, blue are electrical
@@
   \centerline{\includegraphics[width=60mm]{mixing.ari.pdf}}
-  \caption{}
+  \caption{Adjusted rand index for synthetic data as a function of run iteration. }
 \label{fig:mixing:ari}
@@
   \centerline{\includegraphics[width=130mm]{mixing.scorevstime.pdf}}
-  \caption{}
+  \caption{Total model score (log score) vs iteration}
 \label{fig:mixing:scorevstime}

### 2014-06-24 experiments/mouseretina/pipeline.py
marginal attempt to plot more stuff

@@ -630,8 +630,8 @@ def plot_z_matrix(exp_results,
 
     sample_d = pickle.load(open(exp_results))
     chains = sample_d['chains']
-    
     exp = sample_d['exp']
+    print "exp=", exp
     data_filename = exp['data_filename']
     data = pickle.load(open(data_filename))
     data_basename, _ = os.path.splitext(data_filename)
@@ -651,7 +651,6 @@ def plot_z_matrix(exp_results,
     # compute the z matrix 
     z_mat = np.zeros((len(cells), len(cells)))
     for ci, c in enumerate(chains):
-        print "z for chain", ci
         sample_latent = c['state']
         cell_assignment = np.array(sample_latent['domains']['d1']['assignment'])
         ca = irm.util.canonicalize_assignment(cell_assignment)
@@ -671,16 +670,19 @@ def plot_z_matrix(exp_results,
 
 
     import matplotlib.gridspec as grd
-    f = pylab.figure()
-    gs = grd.GridSpec(1, 2, width_ratios=[1,6], wspace=0.02)
-    ax = f.add_subplot(gs[0, 1])
-    ax.imshow(z_mat, cmap=pylab.cm.Greys, interpolation='nearest')
+    f = pylab.figure(figsize=(10, 6))
+    gs = grd.GridSpec(4, 3, width_ratios=[2, 2,10 ], wspace=0.02)
+    ax = f.add_subplot(gs[:, 2])
+    im = ax.imshow(z_mat/CHAINN, cmap=pylab.cm.Greys, interpolation='nearest')
     ax.set_xticks([])
     ax.set_yticks([])
     ax.set_title("Cell coassignment probability") 
     ax.set_xlabel("cells")
-    ax = f.add_subplot(gs[0, 0])
+    cbar_ax = f.add_axes([0.35, 0.2, 0.1, 0.02])
+    f.colorbar(im, cax=cbar_ax, orientation='horizontal', ticks=[0.0, 1.0])
 
+
+    typeax = f.add_subplot(gs[:, 1])
     cells['coarse'][cells['type_id'] == 58] = 'bc'
     cmap = {None : 4, 
             'other': 4, 
@@ -691,24 +693,65 @@ def plot_z_matrix(exp_results,
     color_idx = [cmap[o] for o in cells['coarse']]        
     import brewer2mpl
     cmap = brewer2mpl.get_map('Set1', 'qualitative', 5).mpl_colormap
-    ax.scatter(cell_types, np.argsort(ca), edgecolor='none', 
[…]

### 2014-06-25 experiments/mouseretina/pipeline.py
updates to plots for connect

@@ -670,19 +670,21 @@ def plot_z_matrix(exp_results,
 
 
     import matplotlib.gridspec as grd
-    f = pylab.figure(figsize=(10, 6))
-    gs = grd.GridSpec(4, 3, width_ratios=[2, 2,10 ], wspace=0.02)
-    ax = f.add_subplot(gs[:, 2])
+    f = pylab.figure(figsize=(8, 6))
+    gs = grd.GridSpec(1, 2, width_ratios=[2,10 ], wspace=0.02)
+    ax = f.add_subplot(gs[:, 1])
     im = ax.imshow(z_mat/CHAINN, cmap=pylab.cm.Greys, interpolation='nearest')
     ax.set_xticks([])
     ax.set_yticks([])
     ax.set_title("Cell coassignment probability") 
     ax.set_xlabel("cells")
-    cbar_ax = f.add_axes([0.35, 0.2, 0.1, 0.02])
+
+
+    cbar_ax = f.add_axes([0.30, 0.15, 0.1, 0.02])
     f.colorbar(im, cax=cbar_ax, orientation='horizontal', ticks=[0.0, 1.0])
 
 
-    typeax = f.add_subplot(gs[:, 1])
+    typeax = f.add_subplot(gs[:, 0])
     cells['coarse'][cells['type_id'] == 58] = 'bc'
     cmap = {None : 4, 
             'other': 4, 
@@ -698,12 +700,16 @@ def plot_z_matrix(exp_results,
                c=color_idx, 
                alpha=0.8,
                cmap = cmap)
-    typeax.set_xticks([12, 24, 57, 72])
+    p = np.array([0, 12, 24, 57, 72, 80])
+    typeax.set_xticks(p)
+    typeax.set_xticks(p[1:] - np.diff(p)/2. , minor=True)
     typeax.set_yticks([])
+    typeax.set_xticklabels(['gc', 'bc', 'nac', 'mwac', 'other'], minor=True,
+                           fontsize=8, rotation=90)
     typeax.set_xticklabels([])
-    typeax.set_title("Type ID")
-    typeax.set_ylabel("cells")
 
+    typeax.set_title("anatomist type (0-72)", fontsize=8)
+    typeax.set_ylabel("cells")
     typeax.grid()
 
     # create colors
@@ -712,46 +718,45 @@ def plot_z_matrix(exp_results,
     typeax.set_ylim(950, 0)
 
 
[…]

### 2014-06-25 docs/paper.tex
updates to all

@@
   location and the spatial distribution of synapses, in a
-  principled and probabilstically-coherent manner. We show that the
+  principled and probabilistically-coherent manner. We show that the
   approach recovers known neuron types in the retina and enables
@@
   \centerline{\includegraphics[width=183mm]{fig1b.ai}}
-  \caption{An overview of our method. a. Input data is a connectivity
-    matrix indicating whether cell $i$ (row) synapses on cell $j$
-    (column), a distance matrix between cell soma, b. the depth of the
-    cell soma, and the c. depth profile of the cells synaptic
-    connectivity. d. Our algorithm discovers hidden cell types in this
+  \caption{Deriving circuitry and types from connectomics data a.
+As input we take the connectivity betwen cells (b), the distance between them (c), the depth of the cell bodies (d), and the depth profile of the syanpses (e).
+f.
+Our algorithm discovers hidden cell types in this
     connectivity data by assuming all cells of a type share a
@@
     results in a clustering of the cells by those hidden
-    types. d. shows the cell connectivity matrix with cells of the
-    same type grouped together. e. shows the learned probability of
+    types. f.) shows the cell connectivity matrix with cells of the
+    same type grouped together. g.) shows the learned probability of
     connection between our different types at various distances -- in
     this case, the cells are likely to connect when they are
-    close. f. shows the probability of connection between two cell types
+    close. h. shows the probability of connection between two cell types
 that very rarely connect -- there's a background ``base'' connection rate
 to account for errors in data, but the probability is very low. 
-We can plot this these types g.) to show the relationship and spatial
+We can plot this these types (i) to show the relationship and spatial
 connections, and how probable various types are to connect to one
-another. h. The connectivity between our discovered types changes as a
+another. j.) The connectivity between our discovered types changes as a
 function of distance. }
@@
 
-From these assumptions (priors) develop a generative Bayesian model
+From these assumptions (priors) we develop a generative Bayesian model
 that estimates the underlying cell types and how they connect. We take
-as input (fig ~\ref{fig:overview}a) the connectivity matrix of cells,
-a matrix of the distance between cells, and the per-cell soma depth
-(fig ~\ref{fig:overview}b) and the depth profile of the cell's
-synapses (fig ~\ref{fig:overview}c). We perform joint probabilistic
+as input (fig ~\ref{fig:overview}a) the connectivity matrix of cells (fig ~\ref{fig:overview}b) ,
+a matrix of the distance between cells (fig ~\ref{fig:overview}c), the per-cell soma depth
+(fig ~\ref{fig:overview}d) and the depth profile of the cell's
+synapses (fig ~\ref{fig:overview}e). We perform joint probabilistic
 inference to automatically learn the number of cell types, which cells
@@
 learn the soma depth associated with each type and the typical
-synaptic density profile (fig ~\ref{fig:overview}d.).
+synaptic density profile (fig ~\ref{fig:overview}f-h.).
 
@@
 (fig~\ref{fig:synthetic}b), and the spatial extent of each type
-(fig~\ref{fig:synthetic}c).  Indeed, it does far better than the
-existing infinite stochastic block model that assumes that there is no
-distance dependence to the connectivity. The model converges
+(fig~\ref{fig:synthetic}c).
+For comparison, existing infinite stochastic block model assumes cell type alone matters, and thus finds small neighborhoods of connected nodes (instead of global connectivity patterns).
+The model converges
 relatively quickly to an estimate of the most probable values for the
@@
 
-\subsubsection{Mouse Retina}
+\subsubsection{Learning types and circuitry in the retina}
 
@@
   \centerline{\includegraphics{comparison.ai}}
-  \caption{a. Cell-cell assignment matrix reflecting the posterior
-    distribution of clusterings found by our model. Large contiguous
-    black areas are cells that were considered the same type in the
-    majority of samples. b. The cells and their true types sorted to
-    agree with the coassignment matrix. c. A comparison of the
-    predictive accuracy (Area under the curve) for hand-labeled
+  \caption{\textbf{Visualizing type inference uncertainty.} Our fully Bayesian model gives a confidence estimate (posterior probability) that any two given cells are of the same type.
+In b.) we visualize that cell-cell coassignment matrix, showing the probability that cell i is of the same type as cell j on a range from 0.0 to 1.0.
+The block structure shows subsets of cells which are believed to all belong to the same type.
+For comparison, a.) shows the anatomist-defined type for each cell, grouped broadly into the coarse types identified in the previous figure.
+\textbf{Link vs cluster accuracy} c.
+A comparison of the
+    predictive accuracy (area under the curve) for hand-labeled
     anatomical data, versus inclusion of additional sources of
-    information. d. ) The spatial extent (in depth and area) of the
+    information, as well as the clustering accuracy.
+Note that our model sacrifies very little predictive accuracy for additional clustering accuracy.
+By comparison, conventional methods fail at one or both.
+d.
+) The spatial extent (in depth and area) of the
     types identified by humans and our various algorithmic
@@
 
-\subsection{C. elegans}
+\subsection{Recovering spatial connectivity in multiple graphs simultaneously}
 
-The roundworm Caenorhabditis elegans is a model system in
+Having shown our model to work on the repeating tessellated, laminar structure of the mammilian retina, we then apply our model to a structurally very different connectome -- the whole body of a small roundworm:
+\textit{Caenorhabditis elegans} is a model system in
 developmental neuroscience\autocite{White1986}, with the location and
@@
 number of anterior ganglia.
+C.
+elegans also differs from the retina in that the measured connectome is actually two separate graphs -- one of directed chemical synapses and another of undirected electrical synapses.
 
@@
   \centering 
-  \centerline{\includegraphics[width=7in]{otherdata.ai}}
-  \caption{Discovering connectivity and type in both
-    \textit{C. elegans} and the MOS 6502 microprocessor.  a.) Initial
+  \centerline{\includegraphics[width=4.5in]{celegans.ai}}
+  \caption{Discovering connectivity and type in \textit{C. elegans}. 
+ a.) Initial
     \textit{C. elegans} cell connectivity matrix, red are
@@
     both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). d.) The clusters and connectivity versus
-    distance for connections between Gate and C1, Gate and C2, and C1
-    and C2 terminals on a transistor. Red and teal types have a
-    terminal pulled down to ground and mostly function as
-    inverters. Purple class are clocked, stateful transistors, orange
-    control the ALU and yellow control the special data bus. e.) The
-    spatial distribution of the transistors in each cluster.}
-  \label{fig:otherdata}
+    (small text labels). }
+  \label{fig:celegans}
 \end{figure}
@@
 determined the underlying clusters explained by connectivity and
-distance (fig~\ref{fig:otherdata}a). A superficial inspection of the results shows
+distance (fig~\ref{fig:celegans}a). A superficial inspection of the results shows
 clustering into groups consisting roughly homogeneously of motor
 neurons, sensory neurons, and interneurons. Closer examination reveals
-agreement with the classifications originally outlined by Brenner in
+agreement with the classifications originally outlined by White in
 1986.
@@
 
-\subsection{Microprocessor}
+\subsection{Types and connectivity in artificial structures}
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=4in]{mos6502.ai}}
+  \caption{Discovering connectivity and type the 6502 microprocessor.
+    a.) is the micrograph of the original microprocessor, with the
+    region containing the registers under study highlighted. b.) Our
+    graph consists of the interconnections of MOS field-effect
+    transistors with three terminals, Gate, C1, and C2. Reconstruction
+    technique did not permit resolution of C1 and C2 into source and
+    drain. c.) The spatial distribution of the transistors in each
+    cluster show a clear pattern d.) The clusters and connectivity
+    versus distance for connections between Gate and C1, Gate and C2,
+    and C1 and C2 terminals on a transistor. Red and teal types have a
+    terminal pulled down to ground and mostly function as
+    inverters. Purple class are clocked, stateful transistors, orange
+    control the ALU and yellow control the special data bus (SDB).}
+  \label{fig:mos6502}
+\end{figure}
 
@@
 the processor with complex but known structure containing the primary
-8-bit registers X, Y, and S (fig~\ref{fig:otherdata}).
+8-bit registers X, Y, and S (fig~\ref{fig:mos6502}).
 
@@
 the bus. The repeat patterns of spatial connectivity are visible in
-figure~\ref{fig:otherdata}a, showing the man-made horizontal and
+figure~\ref{fig:mos6502}c, showing the man-made horizontal and
 vertical layout of the same types of transistors.
@@
 and microcircuitry to be discovered from connectomics data.  We have
-shown its applicability to a newly published connectomics dataset from
-the mouse retina. When compared to existing methods, we show how the
+shown its applicability to regularly structured laminar neural circuits like the retina, 
+as well as a less structured whole neuronal organism (\textit{C. elegans}) and an artificial processor. When compared to existing methods, we show how the
 incorporation of all of this data yields results that combine both
 high link-prediction accuracy and high agreement with human
-anatomists.  I also reveals interesting structure when applied to the
-classical connectivity of the worm c. elegans, and the connectivity of
-a historical man-made microprocessor. We have found that combining the
+anatomists. We have found that combining the
 available data types allows us to discover cell types and
@@
 
+Our work shows how rich probabilistic models can contribute to computational neuroanatomy. 
 Eventually, algorithms will have to become a central tool for
@@
 parse the huge datasets. This transition may follow a similar
-transition to the one underway in molecular biology (with gene-finding
+transition to that of molecular biology (with gene-finding
 algorithms) and evolutionary biology with (computational
-phylogenetics).  In fact, it is known that there are significant
-disagreements across human anatomists and computational approaches
-promise to make the debates more data driven.
+phylogenetics). Ultimately, computational approaches may help resolve the significant
+disagreements across human anatomists.
 
@@
 
-Inference is performed in three steps via composible transition 
+Inference is performed in three steps via composable transition 
 kernels -- one for structural, one for per-type parameters, and
@@
 
- \item \textbf{Acknowledgments} We thank Josh Vogelstein for discussions and reading of the manuscript, Finale Doshi-Velez for early discussions on the model, and Erica Peterson, and Jonathan Glidden for extensive manuscript review.
+ \item \textbf{Acknowledgments} We thank Josh Vogelstein for discussions and reading of the manuscript, Finale Doshi-Velez for early discussions on the model, and Erica Peterson, and Jonathan Glidden, and Yarden Katz for extensive manuscript review.
 Funding for compute time was provided by Amazon Web Services ``AWS in Education'' grants.
@@
 chain with the highest log likelihood, but for all numerical
-comparisons (including link probability and ARI) we use this full
+comparisons (including link probability and cluster accuracy) we use this full
 collection of samples from the posterior distribution to estimate the
@@
 We compare our model with a standard network clustering model, the
-latent-position clutsering model. This model assumes each cell belongs
+latent-position clustering model. This model assumes each cell belongs
 to one of K clusters, and each cluster is associated with a
@@
 then a function of the distance between the data points in this
-continuous-space. We use \autocite{VBLPCM} a variational
+continuous-space. We use \autocite{Salter-Townshend2013} a variational
 implementation provided in R, parametrically varying the number of
@@
 Figures \ref{fig:mixing:ari} and \ref{fig:mixing:scorevstime} shows
-the ARI to ground truth and the total log score as a function of
+the cluster accuracy (ARI) to ground truth and the total log score as a function of
 runtime.  We see dramatic changes in log score initially as we vary

### 2014-06-26 docs/paper.tex
text we committed

@@
 
-\section{Results}
+\section*{Results}
 
@@
 
-\subsubsection{Learning types and circuitry in the retina}
+\subsubsection*{Learning types and circuitry in the retina}
 
@@
 
-\subsection{Recovering spatial connectivity in multiple graphs simultaneously}
+\subsection*{Recovering spatial connectivity in multiple graphs simultaneously}
 
@@
 
-\subsection{Types and connectivity in artificial structures}
+\subsection*{Types and connectivity in artificial structures}
 \begin{figure}
@@
 
-\section{Discussion}
+\section*{Discussion}
 We have presented a machine learning technique that allows cell types
@@
 
-\subsection{Methods Summary}
+\subsection*{Methods Summary}
 
@@
 \newpage
-\section{Methods }
+\section*{Methods }
 
-\subsection{Probabilistic Model}
+\subsection*{Probabilistic Model}
 
@@
 
-\subsection{Inference} 
+\subsection*{Inference} 
 We perform posterior inference via Markov-chain Monte Carlo (MCMC),
@@
 
-\subsection{Link Prediction}
+\subsection*{Link Prediction}
 As a proxy for link-prediction accuracy we compute the probability of a link
@@
 
-\subsection{Model Comparison}
+\subsection*{Model Comparison}
 We compare our model with a standard network clustering model, the
@@
 
-\subsection{Parameters}
+\subsection*{Parameters}
 
@@
 
-\subsubsection{Mixing of our Markov chains} 
+\subsubsection*{Mixing of our Markov chains} 
 Evaluating whether or not approximate inference methods, such as MCMC,
@@
 
-\subsection {Mouse Retina}
+\subsection* {Mouse Retina}
 \label{supp:mouseretina}
@@
 
-\subsection{C. elegans}
+\subsection*{C. elegans}
 
@@
 
-\subsection{Microprocessor}
+\subsection*{Microprocessor}
 We extracted the connection graph for the transistors in the MOS6502
@@
 \newpage
-\section{Supplemental Material}
+\section*{Supplemental Material}
 
-\subsection{Other Likelihoods}
+\subsection*{Other Likelihoods}
 \label{supp:otherlikelihoods}
@@
 
-\subsection{Source code and data}
+\subsection*{Source code and data}
 
@@
 
-\subsection{Extension to multiple graphs}
+\subsection*{Extension to multiple graphs}
 \label{supp:multigraph}
@@
 \FloatBarrier
-\subsection{Hyperprior grids and hyperprior inference}
+\subsection*{Hyperprior grids and hyperprior inference}
 \label{supp:hyperpriors}
@@
 
-\section{Measuring clustering similarity}
+\section*{Measuring clustering similarity}
 

### 2014-12-10 docs/paper.tex
arxiv sub
elife sub
lots of updates need to fix

@@
-%\documentclass{nature}
 \documentclass{article}
@@
 \usepackage[backend=biber,style=nature]{biblatex}
-\usepackage{todonotes}
+\usepackage{fixme}
 \usepackage{grffile}
@@
 
-\section*{Results}
+\section*{Methods}
 
@@
 that estimates the underlying cell types and how they connect. We take
-as input (fig ~\ref{fig:overview}a) the connectivity matrix of cells (fig ~\ref{fig:overview}b) ,
-a matrix of the distance between cells (fig ~\ref{fig:overview}c), the per-cell soma depth
-(fig ~\ref{fig:overview}d) and the depth profile of the cell's
-synapses (fig ~\ref{fig:overview}e). We perform joint probabilistic
-inference to automatically learn the number of cell types, which cells
-belong to which type, their type-specific connectivity, and how
-connections between types vary with distance. We also simultaneously
-learn the soma depth associated with each type and the typical
-synaptic density profile (fig ~\ref{fig:overview}f-h.).
+as input (fig ~\ref{fig:overview}a) the connectivity matrix of cells
+(fig ~\ref{fig:overview}b) , a matrix of the distance between cells
+(fig ~\ref{fig:overview}c), the per-cell soma depth (fig
+~\ref{fig:overview}d) and the depth profile of the cell's synapses
+(fig ~\ref{fig:overview}e). We perform joint probabilistic inference
+to automatically learn the number of cell types, which cells belong to
+which type, their type-specific connectivity, and how connections
+between types vary with distance. We also simultaneously learn the
+soma depth associated with each type and the typical synaptic density
+profile (fig ~\ref{fig:overview}f-h.).
 
-We apply our algorithm to datasets from mouse
-retina, C. elegans and a historical microprocessor. Anatomists
-classify cells based on many features, giving us a meaningful baseline
-to compare against.
+As an illustrative example, consider a network with only three cell
+types, labeled A, B, and C. Assume the network has three types of
+cells, they are uniformly distributed in space, and that the
+probability of connection between any two cells, $c_i$ and $c_j$,
+depends only on their type and their distance, according to a logistic
+(sigmoidal) function. Let A connect only to nearby B and C cells, but B
+connect to any C regardless of distance.
+
+For the basic link-distance model, we take as input a connectivity
+matrix $R$ defining the connections between cell $e_i$ and $e_j$, as
+well as a distance function $d(e_i, e_j)$ representing a (physical)
+distance between adjacent cells.  See
+the supplemental material for extension to multiple connectivity
+matrices. We assume there exist an unknown number $K$ of latent
+(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
+cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
+of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
+observed connectivity between two cells $R(e_i, e_j)$ then depends
+only on their latent type and their distance through a link function
+$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
+latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
+as a set of global hyper parameters $\theta$, such that the link
+function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
+
+We then jointly infer the maximum a posteriori (MAP) estimate of the
+class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix
+$\eta_{mn}$, and the global model hyperparameters $\theta$ :
+
+\begin{equation}
+  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
+\end{equation}
+
+For the retina data, we then extend the model with the additional
+features indicated. Cell soma depth is modeled as a
+cell-type-dependent Gaussian with latent (unknown) per-type mean and
+variance. Similarly, each cell has some number $N_i$ of synapses, 
+each of which is drawn from a cell-type-specific density profile
+with up to three modes.
+
+Inference is performed in three steps via composable transition 
+kernels -- one for structural, one for per-type parameters, and
+one kernel for global parameters and hyperparameters. Details
+of data preprocessing, inference parameters, and runtime can
+be found in the Supplementary Methods section.
 
@@
 
+\fxwarn{Do we talk about cross-validation early on? What about posterior uncertainty?}
+
+\subsection{Metrics}
+We focus on two questions we can ask from the learned model: which
+cells are of the same type? and given any two cells, what is the
+probability of their connection?
+
+To assess the accuracy of a clustering compared to that determined by
+neuroanatomists, we employ the adjustded Rand index (ARI). ARI
+measures the similarity of two clusterings. While there are no perfect
+clustering comparison metrics (see supplemental for a discussion of
+ARI's shortcomings) all suffer from the need to balance
+``overclustering'' (calling a single type two different types) with
+``merging'' (calling two distinct types a single type).
+
+To test whether our model accurately captures the true structure of
+the data, we learn the model on the data with a subset of the links
+missing and then compute our predictive accuracy. We perform 10-way
+cross validation on a given dataset, learn the resulting model, and
+use that model to predict the missing synapses. Each potential
+link between cells is assigned a probability, and we compute
+the area under the curve (AUC) for the resulting ROC curve. 
+\fxwarn{Add additional ARI} 
+\section*{Results}
+
+\subsection{Ground Truth Validation}
 To validate our model, we performed a series of simulations to test if
@@
 (fig~\ref{fig:synthetic}b), and the spatial extent of each type
-(fig~\ref{fig:synthetic}c).
-For comparison, existing infinite stochastic block model assumes cell type alone matters, and thus finds small neighborhoods of connected nodes (instead of global connectivity patterns).
-The model converges
-relatively quickly to an estimate of the most probable values for the
-cell types, which is enabled by using a combination of simulated
-annealing and parallelized Markov chain Monte Carlo (see methods for
-details). Thus we can apply our model to simulated datasets with structure
-and scale similar to that of our biological datasets and recover the known 
-correct structure.
+(fig~\ref{fig:synthetic}c).  For comparison, existing infinite
+stochastic block model assumes cell type alone matters, and thus finds
+small neighborhoods of connected nodes (instead of global connectivity
+patterns). The model converges relatively quickly to an estimate of
+the most probable values for the cell types, which is enabled by using
+a combination of simulated annealing and parallelized Markov chain
+Monte Carlo (see methods for details). Thus we can apply our model to
+simulated datasets with structure and scale similar to that of our
+biological datasets and recover the known correct structure.
+
+\subsection{Model Mismatch}
+Our type-and-distance assumptions for our connectivty model are
+simplifying, and reflect a general consensus about the organization of
+some (but not all) types of neural tissue. To show where our model
+performs poorly, we attempt connectivity inference on five sets of
+synthetic data. Our goal is to show that our model does not find
+``structure'' when there is none. \fxwarning{Fix this phrasing -- we want to say 'if your model thinks there are types, there probably are!}.
+
+\begin{itemize}
+\item Data with type-specific connectivity and distance dependence
+\item Data with type-specific connectivity and no distance dependence
+\item Data generated according to a latent-position clustering model,
+  where each cell belongs to a type which is part of a cluster in a
+  low-dimensional space \fxwarning{Cite this paper}
+\item cell type is combinatorial -- there are three possible
+  properties a cell can have, and each subset of properties influences
+  connectivity to neighboring cells. 
+\item cell type is a continuum -- each cell's type continuously varies
+  from zero to one, and the probability of a connection between cells
+  depends on this type.
+\end{itemize}
+
+We generate ten synthetic datasets based on each of these assumptions,
+and perform 10-way cross validation to predict link accuracy as well
+as ARI to the known (``true'') type, type spatial extent, and type
+number.
+
+\fxwarn{Add results}
+
+\subsection{Sensitivity to Edge Effects}
+
+Initial connectomic efforts reconstruct only a small section of tissue
+-- as a result, many connections to cells outside that tissue volume
+will be lost. We performed simulations to check if 1. edge effects
+could destroy spatial structure and if edge effects could introduce
+artifical, spurious spatial structure.
+
+\fxwarn{Add ``intrudoce spatial structure where there is none}
 
@@
 of data preprocessing, inference parameters, and runtime can
-be found in the Methods section.
+be found in the Supplementary Methods section.
 
@@
 \newpage
-\section*{Methods }
+\section*{Supplementary Methods }
 

### 2014-12-10 experiments/mouseretina/sparkpipelinecv.py

@@ -81,7 +81,7 @@ EXPERIMENTS = [
 THOLDS = [0.01, 0.1, 0.5, 1.0]
     
 MULAMBS = [1.0, 5.0, 10.0, 20.0, 50.0]
-PMAXS = [0.95] # , 0.9, 0.7]
+PMAXS = [0.95, 0.9] # , 0.9, 0.7]
 
 BB_ALPHAS = [1.0]
 BB_BETAS = [1.0]
@@ -119,8 +119,11 @@ for ti in [1]:
                 for var_scale in range(len(VAR_SCALES)):
                     for comp_k in COMP_KS:
                         bs = 'retina.%d.srm_clist_xsoma.%d.%d.%s.%d.%d' % (ti, ml_i, pmax_i, vars, var_scale, comp_k)
-                        EXPERIMENTS.append((bs, 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'))
+                        #EXPERIMENTS.append((bs, 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'))
                 
+# EXTRA ONE FROM MAKEFILE
+bs = 'retina.%d.srm_clist_xsoma.%d.%d.%s.%d.%d' % (1, 3, 1, 'xyz', 1,2)
+EXPERIMENTS.append((bs, 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'))
 
             
 INIT_CONFIGS = {'fixed_20_100' : {'N' : 20, 
@@ -790,6 +793,7 @@ def samples_organize(infile, outfile):
 
 @follows(get_samples)
 @follows(get_cvdata)
+@follows(samples_organize)
 @subdivide(get_cvdata, formatter(".+/(?P<base>.*).cvdata.pickle"), 
            "{path[0]}/{base[0]}.samples.organized/*/cv.data",
            # Output parameter: Glob matches any number of output file names

### 2014-12-10 experiments/synthdifferent/sparkanalysis.py

@@ -1,16 +1,17 @@
 """
 """
 import sys
-# total hack, I should really know better
-sys.path.append("../../code")
 import matplotlib
 matplotlib.use('Agg')
+# total hack, I should really know better
+sys.path.append("../../code")
 
 from ruffus import *
 import cPickle as pickle
 import numpy as np
 import copy
-import os, glob, sys
+import os, glob, sys, shutil
+from glob import glob
 import time
 from matplotlib import pylab
 import matplotlib
@@ -21,21 +22,20 @@ from sklearn import metrics
 import preprocess
 import sqlite3
 import connattribio 
-import multyvac
 
 import matplotlib.gridspec as gridspec
 import models
+import multyvac
 
 import irm
 import irm.data
 import util
 from irm import rand
-
 import predutil
 import boto
+
 import cvpipelineutil as cv
 import sparkutil
-
 
 
 def dist(a, b):
@@ -47,81 +47,38 @@ def to_f32(x):
     assert type(a) == np.ndarray
     return a
 
-DEFAULT_CORES = 8
-DEFAULT_RELATION = "ParRelation"
-WORKING_DIR = "sparkdatacv"
+
+WORKING_DIR = "sparkdata"
 S3_BUCKET = "jonas-testbucket2"
-S3_PATH= "netmotifs/paper/experiments/mouseretina"
-
-
-RETINA_DB = "../../../preprocess/mouseretina/mouseretina.db"
-
-CV_CONFIGS = {'cv_nfold_10' : {'N' : 10,
-                               'type' : 'nfold'},
-              'cv_nfold_2' : {'N' : 2,
-                              'type' : 'nfold'}}
+S3_PATH= "netmotifs/paper/experiments/synthdifferent"
 
 
 
 
+    
 def td(fname): # "to directory"
     return os.path.join(WORKING_DIR, fname)
 
+def get_dataset(data_name):
+    return glob(td("%s.data" %  data_name))
+
 EXPERIMENTS = [
-    ('retina.0.ld.0.0.xyz', 'cv_nfold_2', 'debug_2_100', 'debug_2'), 
+    ('srm', 'cv_nfold_2', 'debug_2_100', 'debug_20'), 
+    ('srm', 'cv_nfold_2', 'debug_2_100', 'debug_10'), 
[…]

### 2014-12-10 experiments/synthdifferent/sparkpipeline.py

@@ -77,10 +77,12 @@ def get_dataset(data_name):
 EXPERIMENTS = [
     ('srm', 'cv_nfold_2', 'debug_2_100', 'debug_20'), 
     ('srm', 'cv_nfold_2', 'debug_2_100', 'debug_10'), 
-    ('srm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
-    ('sbmnodist', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
-    ('lpcm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
-    ('mm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
+
+
+    #('srm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
+    #('sbmnodist', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
+    #('lpcm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
+    #('mm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
     # ('lpcm', 'cx_4_5', 'debug_2_100', 'debug_200'), 
     # ('mm', 'cx_4_5', 'debug_2_100', 'debug_200'), 
 
@@ -146,9 +148,15 @@ KERNEL_CONFIGS = {
 
 # 
 
-@files(None, td("srm.sourcedata"))
-def create_data_srm(infile, outfile):
-    np.random.seed(0)
+
+def srm_params():
+    for seed in range(10):
+        filename = "srm.%02d.sourcedata" % seed 
+    yield None, td(filename), seed
+
+@files(srm_params)
+def create_data_srm(infile, outfile, seed):
+    np.random.seed(seed)
     conn_config = {}
     CLASS_N = 4
     SIDE_N = 10
@@ -183,9 +191,14 @@ def create_data_srm(infile, outfile):
                 open(outfile, 'w'))
 
 
-@files(None, td("sbmnodist.sourcedata"))
-def create_data_sbm_nodist(infile, outfile):
-    np.random.seed(0)
+def sbmnodist_params():
+    for seed in range(10):
+        filename = "sbmnodist.%02d.sourcedata" % seed 
+    yield None, td(filename), seed
+
+@files(sbmnodist_params)
[…]

### 2014-12-10 experiments/synthdifferent/sparkreadtest.py

@@ -1,0 +1,7 @@
+from pyspark import SparkContext
+
+sc = SparkContext()
+tmpFile = "sparkdata/srm.data.samples"
+
+r = sc.pickleFile(tmpFile, 5).collect()
+

### 2014-12-10 fabfile.py

@@ -1,0 +1,127 @@
+from fabric.api import local, env, run, put, cd, task
+from fabric.contrib import project
+
+
+from sparkutil import sparkfab as s
+
+s.config['cluster_name'] = 'netmotifs' # ALWAYS SET THIS
+env.roledefs = s.create_roles()
+
+@task
+def deploy_files():
+    put('netmotifs.egg', '/data/netmotifs.egg')
+
+@task
+def add_packages():
+    env['forward_agent']= True
+    run("conda update --all --yes")
+    run("pip install ruffus")
+    run("pip install --upgrade boto")
+    run("pip install colorbrewer")
+    run("pip install multyvac")
+    run("pip install fabric")
+    run("pip install --upgrade git+ssh://git@github.com/ericmjonas/sparkutils.git")
+
+@task
+def deploy():
+    """
+    Uses rsync. Note directory trailing-slash behavior is anti-intuitive
+    """
+    local('git ls-tree --full-tree --name-only -r HEAD > .git-files-list')
+    
+    project.rsync_project("/data/connect-disco-paper/", local_dir="./",
+                          exclude=['*.pickle', 'experiments/synthdifferent/sparkdata'],
+                          extra_opts='--files-from=.git-files-list')
+    
+    env['forward_agent']= True
+    with cd("/data/connect-disco-paper"):
+        run('~/spark-ec2/copy-dir .')
+
+    project.rsync_project("/data/preprocess/", local_dir="../preprocess/")
+
+""""
+What's the deal with the multivac error
+org.apache.spark.api.python.PythonException: Traceback (most recent call last):
+  File "/root/spark/python/pyspark/worker.py", line 75, in main
+    command = pickleSer._read_with_length(infile)
+  File "/root/spark/python/pyspark/serializers.py", line 150, in _read_with_length
+    return self.loads(obj)
+  File "/root/spark/python/pyspark/cloudpickle.py", line 811, in subimport
+    __import__(name)
+  File "./netmotifs.egg/irm/__init__.py", line 10, in <module>
+    import experiments
+  File "./netmotifs.egg/irm/experiments.py", line 11, in <module>
+  File "/opt/anaconda/lib/python2.7/site-packages/multyvac/__init__.py", line 27, in <module>
+    _multyvac = Multyvac()
+  File "/opt/anaconda/lib/python2.7/site-packages/multyvac/multyvac.py", line 81, in __init__
[…]

### 2014-12-10 experiments/synthdifferent/cvpipelineutil.py
updates

@@ -1,0 +1,1 @@
+../../code/cvpipelineutil.py

### 2014-12-10 experiments/synthdifferent/sparkpipeline.py

@@ -75,9 +75,7 @@ def get_dataset(data_name):
     return glob(td("%s.data" %  data_name))
 
 EXPERIMENTS = [
-    ('srm', 'cv_nfold_2', 'debug_2_100', 'debug_20'), 
-    ('srm', 'cv_nfold_2', 'debug_2_100', 'debug_10'), 
-
+    ('srm.00', 'cv_nfold_2', 'debug_2_100', 'debug_2'), 
 
     #('srm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
     #('sbmnodist', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
@@ -137,8 +135,8 @@ KERNEL_CONFIGS = {
     'debug_20' : {'ITERS' : 2, 
                   'kernels': irm.runner.default_kernel_anneal(1.0, 20)
               },
-    'debug_10' : {'ITERS' : 2, 
-                  'kernels': irm.runner.default_kernel_anneal(1.0, 20)
+    'debug_2' : {'ITERS' : 2, 
+                  'kernels': irm.runner.default_kernel_anneal(1.0, 2)
               },
     'debug_200' : {'ITERS' : 200, 
                   'kernels': irm.runner.default_kernel_anneal(1.0, 160)
@@ -151,7 +149,7 @@ KERNEL_CONFIGS = {
 
 def srm_params():
     for seed in range(10):
-        filename = "srm.%02d.sourcedata" % seed 
+        filename = "srm_%02d.sourcedata" % seed 
     yield None, td(filename), seed
 
 @files(srm_params)
@@ -193,7 +191,7 @@ def create_data_srm(infile, outfile, seed):
 
 def sbmnodist_params():
     for seed in range(10):
-        filename = "sbmnodist.%02d.sourcedata" % seed 
+        filename = "sbmnodist_%02d.sourcedata" % seed 
     yield None, td(filename), seed
 
 @files(sbmnodist_params)
@@ -236,7 +234,7 @@ def create_data_sbm_nodist(infile, outfile, seed):
 
 def lpcm_params():
     for seed in range(10):
-        filename = "lpcm.%02d.sourcedata" % seed 
+        filename = "lpcm_%02d.sourcedata" % seed 
     yield None, td(filename), seed
 
 @files(lpcm_params)
@@ -306,7 +304,7 @@ def create_data_latent_position(infile, outfile, seed):
 
 def mm_params():
[…]

### 2014-12-10 fabfile.py

@@ -21,6 +21,8 @@ def add_packages():
     run("pip install multyvac")
     run("pip install fabric")
     run("pip install --upgrade git+ssh://git@github.com/ericmjonas/sparkutils.git")
+    run("yum -y install tmux")
+    run("yum -y update")
 
 @task
 def deploy():

### 2014-12-10 docs/paper.tex
updates to experiment runners

@@
   can reveal interesting structure in the nervous system of
-  C. elegans, and automatically discovers the structure of a
-  microprocessor.  Our approach extracts structural meaning from
+  C. elegans.
+  Our approach extracts structural meaning from
   connectomics, enabling new approaches of automatically deriving
@@
 \autocite{Helmstaedter2013}, but additionally examine the C. elegans
-connectome \autocite{White1986}, and then ``connectome'' of a
-classical microprocessor \autocite{James2010}. Comparing the cell
+connectome \autocite{White1986}. 
+Comparing the cell
 types discovered by the algorithms with those obtained manually by
@@
 
-\subsection*{Types and connectivity in artificial structures}
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics[width=4in]{mos6502.ai}}
-  \caption{Discovering connectivity and type the 6502 microprocessor.
-    a.) is the micrograph of the original microprocessor, with the
-    region containing the registers under study highlighted. b.) Our
-    graph consists of the interconnections of MOS field-effect
-    transistors with three terminals, Gate, C1, and C2. Reconstruction
-    technique did not permit resolution of C1 and C2 into source and
-    drain. c.) The spatial distribution of the transistors in each
-    cluster show a clear pattern d.) The clusters and connectivity
-    versus distance for connections between Gate and C1, Gate and C2,
-    and C1 and C2 terminals on a transistor. Red and teal types have a
-    terminal pulled down to ground and mostly function as
-    inverters. Purple class are clocked, stateful transistors, orange
-    control the ALU and yellow control the special data bus (SDB).}
-  \label{fig:mos6502}
-\end{figure}
-
-To show the applicability of our method to other connectome-style
-datasets, we obtained the spatial location and interconnectivity of
-the transistors in a classic microprocessor, the MOS Technology 6502
-(used in the Apple II) \autocite{James2010}. Computer architects use
-common patterns of transistors when designing circuits, with each
-transistor having a ``type'' in the circuit. We identified a region of
-the processor with complex but known structure containing the primary
-8-bit registers X, Y, and S (fig~\ref{fig:mos6502}).
-
-Our algorithm identifies areas of spatial homogeneity that mirror the
-known structure in the underlying architectural circuit, segmenting
-transistor types recognizable to computer architects. Using the
-original schematics, we see that one identified type contains the
-``clocked'' transistors, which retain digital state. Two other types
-contain transistors with pins C1 or C2 connected to ground, mostly
-serving as inverters.  An additional identified type controls the
-behavior of the three registers of interest (X, Y, and S) with respect
-to the SB data bus, either allowing them to latch or drive data from
-the bus. The repeat patterns of spatial connectivity are visible in
-figure~\ref{fig:mos6502}c, showing the man-made horizontal and
-vertical layout of the same types of transistors.
-
 \section*{Discussion}

### 2014-12-10 docs/processor.tex
updates to experiment runners

--- 2014-06-04 docs/arxiv.tex
@@
-%\documentclass{nature}
-\documentclass{article}
-\usepackage[english,american]{babel}
-\usepackage[backend=biber,style=nature]{biblatex}
-\usepackage[disable]{todonotes}
-\usepackage{grffile}
-\usepackage{booktabs}
-\usepackage{hyperref}
-\usepackage{graphicx}
-\usepackage{fullpage}
-%\usepackage{algorithm}
-%\usepackage{algorithmic}
-\usepackage{amsmath}
-%\usepackage{subcaption}
-%\usepackage{multicol}
-\usepackage{csquotes}
-\usepackage{pdfpages}
-\usepackage{placeins}
-
-\renewcommand{\vec}[1]{\mathbf{#1}}
-
-%\renewcommand{\includegraphics}[2][]{\fbox{}}
-
-\addbibresource{connectomics motif.bib}
-\addbibresource{connectomics motif-relational models.bib}
-
-\DeclareGraphicsRule{.ai}{pdf}{.ai}{}
-
-\title{Automatic discovery of circuit structure from neural connectomics}
-\author{Eric Jonas$^1$, Konrad Kording$^{2, 3,4}$}
-
-\begin{document}
-\maketitle
-
-\begin{small}
-\begin{enumerate}
-  \item Department of Brain and Cognitive Sciences, Massachusetts Institute of Technology, Cambridge, Massachusetts
-  \item Department of Physical Medicine and Rehabilitation, Northwestern University and Rehabilitation Institute of Chicago, Chicago, Illinois
-  \item Department of Physiology, Northwestern University, Chicago, Illinois
-  \item Department of Applied Mathematics, Northwestern University, Chicago, Illinois
-\end{enumerate}
-\end{small}
-
-\begin{abstract}
-  New techniques produce massive data about neural connectivity,
-  necessitating new analysis methods to discover underlying
-  structure. It has long been assumed that discovering the local
-  patterns of microcircuitry is the key to understanding neural
-  function. Here we developed a nonparametric Bayesian technique that
-  identifies neuron types and circuitry patterns in connectomics
-  data. We show that the approach recovers known neuron types in the
-  retina, reveals interesting structure in the nervous system of
-  c. elegans, and automatically discovers the structure of
-  microprocessors. Our approach extracts structural meaning from
-  connectomics, enabling new approaches of deriving anatomical
-  insights from these emerging datasets.
-
-\end{abstract}
-
-\section*{Introduction}
-Emerging techniques \autocite{Morgan2013,Zador2012} promise to
-quantify the location of each neuron within a volume of interest,
-along with its connections to all other neurons. Far exceeding the
-capacity of neuroanatomists to trace small circuits, this leads to
-enormous datasets that quantify aspects of the nervous system. The impact of
-this kind of high-throughput transition has been seen before -- 
-rise of sequencing techniques necessitated the
-development of novel computational methods to understand genomic
-structure, ushering in an era of bioinformatics as an independent
-discipline \autocite{Koboldt2013}.
-
-The brain consists of multiple kinds of neurons, each of which is
-hypothesized to have a specific role in the overall
-computation. Neuron types differ in many ways, e.g. chemical or
-morphological, but they also differ in the way they connect to one
-another. In fact, the idea of well defined local connectivity patterns
-has been prominent in many areas, ranging from sensory (e.g. retina,
-\autocite{Masland2001} via processing (e.g. neocortex \autocite{Mountcastle1997})
-to movement (e.g. spinal cord) \autocite{Grillner2005}. Indeed, such
-repeated patterns also exists in human-made computing circuits. It
-remains an important challenge to develop algorithms to use anatomical
-data, e.g. connectomics, to back out underlying connectivity motifs.
-
-The discovery of structure is a crucial aspect of network
-science. Early approaches focused on global graph properties, such as
-the types of scaling present in the network \autocite
-{WattsStrogatz1998} .  While this approach leads to an understanding
-of global network, more recent work aims at identifying very small-scale
-repeat patterns, or “motifs” in networks\autocite{Milo2002}
-
-The discovery of structure in probabilistic graphs is a crucial aspect
-in machine learning. Commonly used algorithms include
-community-based-detection methods \autocite{Girvan2002}, and
-stochastic block models \autocite{Nowicki2001}.  While these
-approaches can incorporate the probabilistic nature of neural
-connections \autocite{Hill2012} they do not generally consider the
-known fact that the probability of connections between neurons depends
-on distance. Combining such probabilistic models with a spatial
-component to model distance promises to improve the detection of
-microcircuitry from neural data.
-
-When it comes to neuroscience and other computing systems, we expect
-patterns of connectivity much more complex than traditional motifs,
-exhibiting a strong spatial dependence arising from the complex
-genetic, chemical, and activity-based neural development process.
-
-Here we describe a Bayesian non-parametric model that can discover
-circuit structure automatically from connectomics data: both the cell
-types and their spatial patterns of interconnection. We apply it to
-three computing systems: the mouse retina connectome
-\autocite{Helmstaedter2013}, the c. elegans connectome
-\autocite{White1986}, and a ``connectome'' of a classical
-microprocessor \autocite{James2010}. Note we do not explicitly
-consider cellular morphology, absolute position (cell layer), or any
-chemical or genetic information. Yet in all cases, we discover
-cell-type to cell-type connectivity that depends on spatial distance
-and these connectivity functions depend on the two involved cell
-types. Comparing the cell types discovered by the algorithms with
-those obtained by anatomists based on orthogonal information reveals a
-high degree of agreement. We present a scalable approach to infer
-microcircuitry from connectivity data.
-
+\subsection*{Types and connectivity in artificial structures}
 \begin{figure}
   \centering 
-  \centerline{\includegraphics[width=183mm]{overview.ai}}
-  \caption{An overview of our method. a. Input data is
-a connectivity matrix indicating whether cell $i$ (row)
-synapses on cell $j$ (column). The location of each cell in space is known. 
-b. Our algorithm discovers hidden cell types in this connectivity data
-by assuming all cells of a type share a distance-dependent connectivity 
-profile, c.), with cells of other types. This results in a clustering 
-of the cells by those hidden types. b.) shows the cell connectivity
-matrix with cells of the same type grouped together. d.) shows
-the learned probability of connection between our different types
-at various distances, illustrating short-range and long-range
-connectivity patterns.}
-\label{fig:overview}
-\end{figure}
-
-\section{Results}
-We use a Bayesian structure discovery approach that is based on
-prior knowledge about the way positions and cell types affect the
-probability of connections between cells, a so-called generative
-model. We term our model the ``infinite spatial-relational model'' (iSRM), 
-in keeping with the existing nonparametric literature.  
-The probability of connections between pairs of neurons
-decreases with distance, but how high the baseline probability of
-connection is and how quickly it decays with distance depends on the
-two cell-types. Note that this simplistic prior information is all that 
-we tell our model; no additional physiological information is included.
-
-We combine annealing with Markov Chain Monte Carlo to
-perform simultaneously joint posterior inference over the structure,
-parameters, and hyperparameters of the model (see Methods). This
-procedure automatically learns the number of cell types, which cells 
-belong to which types, along with the 
-the distance and type dependent connectivity probabilities (see
-fig~\ref{fig:overview}). These results can then be compared with the
-results of previous anatomical studies that have used other aspects of
-neurons, such as chemical properties and cell-morphology to divide
-neurons into types.
-
-Before we can use a probabilistic model to analyze real data we first
-need to assess if the model works properly. We thus started out by
-simulating data for which we know the correct structure and comparing
-the estimated structure based on the algorithm (see methods) with the
-one we used for simulation. We find that the model does a good job of
-recovering the structure we used to generate the data
-(fig~\ref{fig:synthetic}a) Indeed, it does far better than a simpler
-and well established model (the infinite stochastic block model, or
-iSBM) that assumes that there is no distance dependence to the
-connectivity. The model converges relatively quickly to an estimate of
-the most probable values for the cell types, which is enabled by using
-a combination of simulated annealing and parallelized Markov chain
-Monte Carlo (see methods for details). It appears that the model
-behaves the way it should and allows application to relatively large
-datasets.
-
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics{synthetic.ai}}
-  \caption{Correct recovery of true numbers of hidden classes
-in synthetic data. a.) Other models over-cluster the number
-of classes as they fail to take distance into account, whereas
-our model finds close to the true number of classes. b.) ARI, a measure
-of ``correct clustering'', for
-different true class counts, between our model and a traditional
-iSBM. c.) For synthetic data, 
-the iSBM overclusters, finding a large number of spatially-localized types. 
-In contrast, the infinite spatial-relational model recovers the true spatial
-extent of the underlying types.}
-\label{fig:synthetic}
-\end{figure}
-
-\subsubsection{Mouse Retina}
-
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics{mouseretina.ai}}
-  \caption{Discovering cell classes in the mouse retina connectome. 
-a.) Input connectivity data for 950 cells for which soma positions were known. b. clustered connectivity matrix. c.) connectivity diagram showing our clusters, as
-well as the cell depth and known cell type of each cell. Our model agrees
-with the cell types traditionally identified by anatomists, and identifies 
-laminar-specific connectivity patterns. d.) Connectivity between our
-clusters as a function of distance -- the cluster consisting primarily of
-retinal ganglion cells (lower left-center) exhibits the expected near and
-far connectivity. e. The spatial distribution of our cell types -- each cell type tesselates space.}
-\label{fig:mouseretina}
-\end{figure}
-
-The retina of the mouse \autocite{Masland2001} is an example where we
-should expect to have connectivity patterns that are well approximated
-by our generative model. It is known that there are multiple classes
-of cells that can be grouped into ganglion cells that transmit
-information to the rest of the brain, bipolar cells that connect
-between different cells, and amacrine cells that feed into the
-ganglion cells \autocite{cite some review}. Recent research
-\autocite{Helmstaedter2013} has produced a large dataset containing
-both the types of cells from orthogonal approaches, and also the
-connectivity matrix between all reconstructed cells. We want to know
-to which level we can classify neurons into types based on this
-connectivity only.
-
-The algorithm took 8 hours to converge to a locally optimal solution,
-dividing neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
-wedge is a type). For each pair of neurons there is a specific
-distance dependent connection probability (fig~\ref{fig:mouseretina}b,c,d),
-which is well approximated by the model fit. Moreover, each type of
-cell is rather isotropically distributed across space
-(fig~\ref{fig:mouseretina}e) as should be expected for true cell types. The
-algorithm is thus able to extract meaningful structure from the data.
-
-Comparing the results of the algorithm to other information sources
-allows evaluating the quality of the type determination. Our types
-closely reflect the (anatomist-determined) segmentation of cells into
-retinal ganglion, narrow amacrine, medium/wide amacrine, and bipolar
-cells (fig~\ref{fig:mouseretina}c, outermost ring). We find that the
-types we find tend to reflect the known laminar distribution in the
-retina (fig~\ref{fig:mouseretina}c, middle ring), which is exciting as
-the algorithm did not actually use this type of information.
-
-The algorithm yields a separation of neurons into a smaller number of
-types than the fully granular listing of 71 types found by the
-original authors of the paper, although is still highly correlated
-with those finer type distinctions (see
-supp~\ref{supp:mouseretina}. It is our expectation that, with larger
-datasets, even closer agreement would be found.
-
-\todo{Do we assign any of the unknown cells correctly?  are types 72-79
-“different” or simply unknown?}
-
-We compared the results of our model to the traditional iSBM to see if
-the addition of spatial information was necessary. We find that, as
-expected from our synthetic experiments, the iSBM massively
-over-segmented the data, finding spatially-localized “communities” of
-cells which bore little relation to known types.  The average spatial
-extent of the types found by the iSBM was $14.6\pm8.1 \mu m$. However,
-in the ground truth by anatomists the average distance was $34.3\pm7.6
-\mu m$ and $35.3\pm 2.4 \mu m$, quite close to the $34.5\pm 1.0 \mu m$
-discovered by our algorithm. Modeling the distance-dependence prevents
-over-segmentation and thus allows more meaningful estimates.
-
-% from experiments/mouseretina/spatial_var.txt
-
-% Relational Model: mean = 14.586508 std=8.136334
-% Spatial-Relational Model: mean = 34.540355 std=1.069483
-% Truth (fine): mean = 34.348443 std=7.567564
-% Truth (coarse): mean = 35.276636 std=2.372951
-
-\subsection{C. elegans}
-
-The roundworm Caenorhabditis elegans is a model system in
-developmental neuroscience\autocite{White1986}, with the location and
-connectivity of each of 302 neurons deterministically known. This
-incredible cross-organism consistency led to early mapping of the
-c. elegans chemical and electrical (gap junction) connectome. Unlike
-the retina, only the motor neurons in c. elegans exhibit regular
-distribution in space, being distributed along the body axis. Most
-interneurons are concentrated in various ganglia that project
-throughout the entire animal, and the sensory neurons are primarily
-located in a small number of anterior ganglia.
-
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics[width=5in]{celegans.ai}}
-  \caption{Discovering cell types in the \textit{c. elegans}
-    connectome. a.) Initial cell connectivity matrix, red are
-    (directional) chemical synapses, blue are electrical
-    synapses. Point size is proportional to synapse count between
-    those cells. b.) Clustered adjacency matrix showing identified
-    clusters. c.) Inter-cluster connectivity, showing the position of
-    each cell along the body axis as well as the ``true'' cell type --
-    both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). }
-  \label{fig:celegans}
-\end{figure}
-
-Using both the chemical and electrical connectivity (see methods), we
-determined the underlying clusters explained by connectivity and
-distance (fig~\ref{fig:celegans}a). A superficial inspection of the results shows
-clustering into groups consisting roughly homogeneously of motor
-neurons, sensory neurons, and interneurons. Closer examination reveals
-agreement with the classifications originally outlined by Brenner in
-1986.
-
-We identify cell types that reflect the known motor/non-motor neuron
-classes, even though this system lacks the strong repeat
-microcircuitry our model was designed for.  Motor-neuron types AS, DA,
-and VA, all exclusively postsynaptic, are clustered together, as are
-motor-neuron types VD and DD. Traditional types VC, DB, and VB also
-mostly share a cluster. Various head motor neurons, including types
-SMD, RMD, and RIM are clustered together. Interneurons with known
-anatomically-distinct connectivity patterns, such as RIA (2 cells) or
-AVA (2 cells) are clustered into pure types. The algorithm even
-correctly places the single-cell type DVB by itself. Note our
-clustering does not perfectly reflect known divisions -- several
-combinations of head and sensory neurons are combined, and a
-difficult-to-explain group of 2 VD, 2 VB, a VA, and a DD neuron are
-grouped together. Our identified cell types thus reflect a
-``coarsening'' of known types, based entirely on connectivity and
-distance information, even when the organism exhibits substantially
-less spatial regularity than the retina.
-
-\subsection{Microprocessor}
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics{mos6502.ai}}
-  \caption{Discovering transistor types in the MOS 6502
-    microprocessor. a.) The clusters and connectivity versus distance
-    for connections between Gate and C1, Gate and C2, and C1 and C2
-    terminals on a transistor. Red and teal types have a terminal
-    pulled down to ground and mostly function as inverters. Purple
-    class are clocked, stateful transistors, orange control the ALU
-    and yellow control the special data bus. b.) The spatial
-    distribution of the transistors in each cluster. }
+  \centerline{\includegraphics[width=4in]{mos6502.ai}}
+  \caption{Discovering connectivity and type the 6502 microprocessor.
+    a.) is the micrograph of the original microprocessor, with the
+    region containing the registers under study highlighted. b.) Our
+    graph consists of the interconnections of MOS field-effect
+    transistors with three terminals, Gate, C1, and C2. Reconstruction
+    technique did not permit resolution of C1 and C2 into source and
+    drain. c.) The spatial distribution of the transistors in each
+    cluster show a clear pattern d.) The clusters and connectivity
+    versus distance for connections between Gate and C1, Gate and C2,
+    and C1 and C2 terminals on a transistor. Red and teal types have a
+    terminal pulled down to ground and mostly function as
+    inverters. Purple class are clocked, stateful transistors, orange
+    control the ALU and yellow control the special data bus (SDB).}
   \label{fig:mos6502}
@@
 the bus. The repeat patterns of spatial connectivity are visible in
-figure~\ref{fig:mos6502}a, showing the man-made horizontal and
+figure~\ref{fig:mos6502}c, showing the man-made horizontal and
 vertical layout of the same types of transistors.
-
-\section{Discussion}
-We have presented a machine learning technique that allows cell types
-and microcircuitry to be discovered using only distance and
-connectivity data. We have shown its applicability to a newly
-published connectomics dataset from the mouse retina, the classical
-connectivity of the worm c. elegans, and the connectivity of a
-historical man-made microprocessor. We have found that distance and
-connectivity alone are sufficient to discover cell types and
-microcircuitry that were known to exist in the systems based on decades
-of previous research which had utilized extensive additional
-information.
-
-For probabilistic models like ours, no known solution exists to
-exactly find the most probable parsing of the neurons into cell-types
-and connectivity patterns . We employ a collection of Markov-chain
-Monte carlo techniques (see Methods) but while different
-initializations converge to similar ultimate values, we can never
-realistically obtain the global optimum. There are a broad range of
-techniques that may offer better approximations to the global optimum
-\autocite{Wanga2012} and future work could adapt them to
-the problem we addressed here.
-
-Our inference becomes slower as the amount of data increases. Our
-algorithm required 30 hours for 1000 neurons and could take months for
-substantially longer datasets. Scaling this class of probabilistic
-model is an active area of research, and recent results in both
-variational methods \autocite{Hoffman2013} and spectral learning
-  \autocite{Anandkumar2012} promise substantial speed increases.
-
-  Larger datasets will allow algorithms to distinguish more distinct
-  types \todo{Add supplemental data} and we expect closer agreement
-  with existing anotomical knowledge as more data become available.
-  Moreover, in general, for such problems precision increases with the
-  size of the dataset and the cells that we have are not sufficient to
-  statistically distinguish all the cell types known in anatomy (such
-  as the $\sim 70$ in the retina). Still, using only connectivity and
-  distance it is possible to meaningfully divide neurons into types.
-
-Our small collection of hand-selected distance-dependent likelihood
-functions are clearly non-exhaustive, and assume monotonicity
-of connectivity probability -- for a given class, closer cells
-are never less-likely to connect. This is known to be insufficient
-for various neural systems. Future models could incorporate
-a wider variety of likelihood functions, or even learn the global
-functional form from the data.
-
-There exist a range of previous approaches to the discovery of
-neural microcircuitry \todo{Do we want a citation here?}. These
-generally involve a great deal of manual labor and ad-hoc
-determination of what constitutes a “type” of cell -- to this day
-there are disagreements in the literature as to the “true” types in
-the mammalian retina. Much as phylogenomics has changed our
-understanding of animal ontologies, modern large scale data will allow
-the efficient unbiased discovery of cell types and circuits. The sheer
-amount of available data demands the introduction of algorithmic
-approaches.
-
-The development of automatic identification and quantification of cell
-type may also provide a new computational phenotype for quantifying
-the effect of disease, genetic interventions, and
-developmentally-experienced neural activity. Our method can in
-principle identify neuron-types across non-connected graphs,
-e.g. across animals. For example, the types of neurons in one animal
-can be associated with the types of neurons in another animal, in the
-same way as this is already possible through molecular markers
-\autocite{Brown2009}. This could be particularly important if cell
-types appear that are due to properties of the stimuli and experience
-as opposed to just the molecular properties of cells, such as color
-and orientation selective types in primary visual cortex
-\autocite{Sincich2005,Lennie2005}. This would allow comparative
-quantitative anatomy across animals, and aid the search for the
-ultimate causes of connectivity.
-
-Ultimately, we will need to combine information across different
-sources into a joint understanding of neural function. Distinct cell
-types differ in morphology, connectivity, transcriptomics, relation to
-behavior or stimuli and many other ways. Algorithms based on each type
-information can then be combined to synthesize all the available
-information from one experiment or even across experiments into a
-joint model of brain function.
-
-\subsection{Methods Summary}
-
-We take as input a connectivity matrix $R$ defining the connections
-between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
-e_j)$ representing a (physical) distance between adjacent cells. See
-the supplemental material for extension to multiple connectivity
-matrices. We assume there exist an unknown number $K$ of latent
-(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
-cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
-of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
-observed connectivity between two cells $R(e_i, e_j)$ then depends
-only on their latent type and their distance through a link function
-$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
-latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
-as a set of global hyper parameters $\theta$, such that the link
-function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
-
-We then jointly infer the maximum a posteriori (MAP) estimate of the
-class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
-the global model hyperparameters $\theta$ :
-
-\begin{equation}
-  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
-\end{equation}
-
-Inference is performed in three steps via composible transition 
-kernels -- one for structural, one for per-type parameters, and
-one kernel for global parameters and hyperparameters. Details
-of data preprocessing, inference parameters, and runtime can
-be found in the Methods section.
-
-\printbibliography
-
-\begin{itemize}
-
- \item \textbf{Acknowledgments} We thank Josh Vogelstein for discussions and reading of the manuscript, Finale Doshi-Velez for early discussions on the model, and Erica Peterson, and Jonathan Glidden for extensive manuscript review.
-Funding for compute time was provided by Amazon Web Services ``AWS in Education'' grants.
-\item \textbf{Author Contributions} KK and EJ developed model. EJ derived inference, implemented code, tested, and ran experiments. KK and EJ wrote manuscript text and solicited feedback. 
- \item \textbf{Competing Interests} The authors declare that they have no
-competing financial interests.
- \item \textbf{Correspondence} Correspondence and requests for materials
-should be addressed to E.J.~(email: jonas@mit.edu).
-\end{itemize}
-
-\newpage
-\section{Methods }
-
-\subsection{Probabilistic Model}
-
-Our model is a extension of the iSBM
-\autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
-inspired by attempts to extend these models with arbitrary
-discriminative functions\autocite{Murphy2012}.
-
-We take as input a connectivity matrix $R$ defining the connections
-between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
-e_j)$ representing a (physical) distance between adjacent cells. See
-the supplemental material for extension to multiple connectivity
-matrices. We assume there exist an unknown number $K$ of latent
-(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
-cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
-of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
-observed connectivity between two cells $R(e_i, e_j)$ then depends
-only on their latent type and their distance through a link function
-$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
-latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
-as a set of global hyper parameters $\theta$, such that the link
-function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
-
-We then jointly infer the maximum a posteriori (MAP) estimate of the
-class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
-the global model hyperparameters $\theta$ :
-
-\begin{equation}
-  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
-\end{equation}
-
-% Methods section 3000 words
-% methods summary does not appear in the online text
-% methods section can have no figures or tables [grrr]
-
-We describe the spatial ``Logistic-distance Bernoulli''  function here,
-and others in the supplemental material.
-
-% fixme class component, NOT class
-The ``logistic-distance Bernoulli'' spatial model assumes that, if cell
-$e_i$ is of type $m$ and cell $e_j$ is of type $n$, then $\eta_{mn}
-= (\mu_{mn}, \lambda_{mn})$, and the probability that two cells $e_i$
-and $e_j$ are connected is given by
-\begin{eqnarray}
-p^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
-p &= & p^* \cdot (p_{max} - p_{min}) + p_{min}
-\end{eqnarray}
-where $p_{max}$ and $p_{min}$ are global per-graph parameters.
-
-We place an exponential priors on the latent parameters:
-\begin{eqnarray}
- \mu_{mn} \sim \exp(\mu | \mu^{hp}) \\
-\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})
-\end{eqnarray}
-
-using  $\lambda^{hp}$ and $\mu^{hp}$ as global per-graph hyperparameters.
-
-We use a Dirichlet-process prior on class assignments, which allows
-the number of classs to be determined automatically. In brief, for $N$
-total cells, the probability of a cell belonging to a class is
-proportional to the number of datapoints already in that class, $N_k$,
-such that $p(c_i = k) \propto \frac{m_k}{N + \alpha}$ and the
-probability of the cell belonging to a new class $k'$ is $p(c_i = k')
-\propto \frac{\alpha}{N + \alpha}$. $\alpha$ is the global
-concentration parameter -- larger values of $\alpha$ make the model
-more likely to propose new classes. We grid the parameter $\alpha$ and
-allow the best value to be learned from the data.
-
-\subsection{Inference} 
-We perform posterior inference via Markov-chain Monte Carlo (MCMC),
-annealing on the global likelihood during the traditional burn-in
-phase. MCMC transition kernels for different parts of the state space
-can be chained together to construct a kernel whose ergodic
-distribution is the target ergodic distribution over the entire state space.
-
-Our first transition kernel (``structural'') performs gibbs sampling 
-of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
-The lack of conjugacy in our likelihood model makes an explicit 
-evaluation of the conditional assignment probabilities impossible, 
-motivating us to use an auxiliary variable method \autocite{Neal2000}
-in which a collection of ephemeral classs are explicitly represented
-for the duration of the Gibbs scan.
-
-We then employ a transition kernel to update the per-component
-parameter values $\eta_{mn}$. Conditioned on the assignment vector
-$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
-individual $\eta_{mn}$ are independent. We slice sample \autocite{Neal2003}
-each component's parameters, choosing the slice width as a function
-of the global hyperparameter range.
-
-The global hyper-parameters, both $\alpha$ and $\theta$, are allowed
-to take on a discrete set of possible values. As $\theta$ is often a
-tuple of possible values, we explore the cartesian product of all
-possible values. We then Gibbs sample (our final transition kernel),
-which is always possible in a small, finite, discrete state space.
-
-We chain these three kernels together, and then globally
-anneal on the likelihood from a temperature of $T=64$ down to 
-$T=1$ over 300 iterations unless otherwise indicated, and
-then run the chain for another $100$ iterations.
-
-To pick the MAP, we run up to $100$ simultaneous chains in parallel,
-initialized from different random initial points in the state space,
-and then pick the chain with the highest log likelihood. Early
-experiments attempted to use the full ensemble of chains as an
-estimate of the posterior but were abandoned due to the difficulty of
-summarizing the posterior over such a complex state space, which is an
-ongoing area of research in the probabilistic modeling community
-\todo{cite more? missing}.
-
-\todo{Validation: do we want to write extensively about posterior
-  enumeration tests, etc. ? }
-
-\subsection {Mouse Retina}
-\label{supp:mouseretina}
-Dense serial electron microscopy of a $114\mu m \times 80 \mu m $ area
-in the mouse retina by \autocite{Helmstaedter2013} yielded a listing
-of places where neurons come into contact. There were over 1000 cells
-originally, and selected the $950$ for which the location of the soma
-could be reconstructed from the provided cell plots (soma locations
-were not provided by the study's authors in machine-readable
-form). Ultimately this left a matrix between the total synapse-like
-contact area between all pairs of 950 cells. Area was thresholded at
-$0.5\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
-matrix that served as input to our algorithm. We measured the distance
-between cells using the reconstructed soma centers, and used the
-Logistic-Distance spatial relation. Hyperprior griddings are shown in
-supplemental section \ref{supp:hyperpriors}.
-
-\subsection{C. elegans}
-
-We obtained the connectome of c. elegans from data published
-previously \autocite{Varshney2011}, and isolated the 279 nonpharyngeal
-neurons, with a total of 6393 chemical synapses and 890 gap junctions
-originally cleaned up in \autocite{Chen2006}. A cell's position was
-its distance along the anterior-posterior axis normalized between zero
-and one. We used both networks, the chemical network as a directed
-graph and the electrical network as undirected graph. We use the
-synapse counts with the logistic-distance poisson likelihood, scaling
-the counts by 4.0 to compensate for the Poisson's overdispersion.
-
-\subsection{Microprocessor}
-We extracted the connection graph for the transistors in the MOS6502
-\autocite{visual6502source}. Each transistor has three terminals (gate,
-source, drain), but the methods of the original dataset were unable to
-consistently resolve which of the C1 and C2 terminals were source and
-drain, leading to ambiguity in our encoding. We identified a region
-consisting of three registers X, Y, and S via visual
-inspection and focused our efforts there. We created a total of six
-connectivity matrices by examining possible terminal pairings. One
-graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
-$e_j$ are connected via pins $g$ and $c_1$.
-
-% \begin{addendum}
-%  \item Put acknowledgements here.
-%  \item[Competing Interests] The authors declare that they have no
-% competing financial interests.
-%  \item[Correspondence] Correspondence and requests for materials
-% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
-% \end{addendum}
-\newpage
-\section{Supplemental}
-
-\subsection{Other Likelihoods}
-\label{supp:otherlikelihoods}
-
-We reparameterized the Logistic-Distance Bernoulli likelihood to
-better capture the microprocessor data structure. We are explicitly
-setting the maximum probability $p$ of the logistic function on a
-per-component basis, drawing from a global $p \sim \operatorname{Beta}(\alpha_{hp},
-\beta_{hp})$. Then $\lambda$ is set for each component as a global
-hyperparameter, $\lambda$.
-
-The ``logstic-distance Poisson'' spatial model is used to explicitly mode the count of synapses, $c$, between two neurons.
-The probability of c synapses between two neurons is distributed $c \sim \textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a scaled logistic function (the logistic function has range $[0, 1]$.
-For each component $\eta_{mn}$ we learn both the threshold $\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$ and $n$ are likely to have on average $20$ synases if they are closer than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$  "
-
-Thus the probability of $R(e_i, e_j) = c$ synapses between two cells $e_i$ and $e_j$ is given by:
-\begin{eqnarray}
-r^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda}}\\
-r & = & r^* \cdot (r_{mn} - r_{min}) + r_{min} \\
-R(e_i, e_j) \sim \textrm{Poisson}(c | r)
-\end{eqnarray}
-
-where $\lambda$ and $r_{min}$ are per-graph parameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $r_{mn} \sim \exp(r_{mn} | r_{scale}^{hp})$.
-
-\subsection{Source code}
-
-All source code and materials for running experiments can be
-obtained from the project website, at \\
-
-\href{http://spatialrelation.github.io }{http://spatialrelation.github.io}
-
-\subsection{Extension to multiple graphs}
-\label{supp:multigraph}
-The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by extending the likelihood to include the product of the likelihoods of the individual graphs.
-
-\begin{equation}
-  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}, \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha)
-\end{equation}
-
-\FloatBarrier
-
-\subsection{Microprocessor Netlist annotation}
-\label{supp:mos6502netlist}
-This is a matrix consisting of the different clusters of electrical nets for 
-the registers, with the colors consistent with figure~\ref{fig:mos6502}.
-
-\includepdf[pages=-]{mos6502.clusters.pdf}
-
-\FloatBarrier
-\subsection{Model Comparison for Mouse Retina}
-
-We compare the classs found by our model, the traditional infinite
-stochastic block model, and a finite stochastic block model with the number
-of classes set to the best-found number from our iSRM model ($k=12$).
-
-We see from fig~\ref{fig:supp:irmspatialcomparemouse} that the spatial
-variance of our classes closely matches the ground truth, while the
-other two models find spatially-localized
-classs. Figure~\ref{fig:supp:mousecompareari} shows the ARI of our
-clustering, the clustering found by the iSBM,
-as well as the result of parametrically varying the traditional stochastic
-block model for various values of $k$.
-
-The actual clusterings for iSRM, the iSBM, and $SBM_{k=12}$ are shown
-in figures \ref{fig:supp:retinabbclusters},
-\ref{fig:supp:retinaldclusters}, and
-\ref{fig:supp:retinabbfixedclusters} respectively. Each block is a
-cluster, with the true cell type indicated along the x-axis, and the
-cell ID written to the left or right.
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=5in]{mouseretina/spatial_var.pdf}}
-  \caption{The distribution of the spatial extent of cell types found by each of
-the above methods, and ground truth clusters}
-\label{fig:supp:irmspatialcomparemouse}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=5in]{mouseretina/ari_vs_cluster.pdf}}
-  \caption{ARI vs cluster number for the iSRM, iSBM, and the SBM with varying K.}
-\label{fig:supp:mousecompareari}
-\end{figure}
-
-\begin{figure}[h!]
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis)  found by the infinite stochastic block model, dramatically overclustering relative to ground truth}
-\label{fig:supp:retinabbclusters}
-\end{figure}
-
-\begin{figure}[h!]
-
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.ld.0.0.data-fixed_20_100-anneal_slow_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis) by the infinite spatial-relational model.}
-\label{fig:supp:retinaldclusters}
-\end{figure}
-
-\begin{figure}[h!]
-
-  \centering 
-    \centerline{\includegraphics[width=6.0in]{mouseretina/retina.1.2.bb.0.0.data-fixed_20_12-anneal_slow_fixed_400.truth_latent.pdf}}
-  \caption{The cell IDs and found types (grouped) vs known types (y-axis) via the stochastic block model with $k=12$}
-\label{fig:supp:retinabbfixedclusters}
-\end{figure}
-
-\newpage
-
-\FloatBarrier
-\subsection{Hyperprior grids}
-\label{supp:hyperpriors}
-
-For the mouse retina Logistic-Distance Bernoulli model, we gridded $\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0 and 80.
-$p_{min} \in [0.001, 0.005, 0.01]$ and $p_{max} \in [0.80, 0.90, 0.95, 0.99]$.
-
-For the c.
-elegans data with the Logistic Distance poisson model, we gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20 $\log_{10}$-spaced points between 2.0 and 20.0.
-We globally set $rate_{min}=0.01$.
-
-For the microprocessor with the Logistic Distance fixed lambda Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced points between 10 and 500 and set $\lambda=\mu_{hp}/10$.
-$p_{min} \in \{0.001, 0.01, 0.02\}$  and both $p_\alpha$ and $p_\beta \in \{0.1, 1.0, 2.0\}$.
-
-\section{Measuring clustering similarity}
-
-The adjusted rand index (ARI) is a measure of the similarity of two
-different clusterings \autocite{Hubert1985} -- two identical clusters
-have an ARI of 1.0 while progressively more dissimilar clusters have
-lower ARIs, becoming negative as the clustering gets anti-correlated.
-
-\end{document}

### 2014-12-10 code/cvpipelineutil.py

@@ -229,7 +229,7 @@ def run_exp_pure(data, init, kernel_config_name, seed, kc):
 
 def experiment_generator(EXPERIMENTS, CV_CONFIGS, INIT_CONFIGS, get_dataset, td):
     for data_name, cv_config_name, init_config_name, kernel_config_name in EXPERIMENTS:
-        data_filename = get_dataset(data_name)[0]
+        data_filename = get_dataset(data_name) # [0]
 
         df = "%s-%s-%s-%s" % (data_name, cv_config_name, init_config_name, kernel_config_name)
         

### 2014-12-10 experiments/synthdifferent/sparkpipeline.py

@@ -72,10 +72,11 @@ def td(fname): # "to directory"
     return os.path.join(WORKING_DIR, fname)
 
 def get_dataset(data_name):
-    return glob(td("%s.data" %  data_name))
+    # why were we globbing here? I'm not sure
+    return td("%s.data" %  data_name)
 
 EXPERIMENTS = [
-    ('srm.00', 'cv_nfold_2', 'debug_2_100', 'debug_2'), 
+    ('srm_00', 'cv_nfold_2', 'debug_2_100', 'debug_2'), 
 
     #('srm', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
     #('sbmnodist', 'cv_nfold_10', 'fixed_20_100', 'anneal_slow_1000'), 
@@ -87,6 +88,17 @@ EXPERIMENTS = [
 
     #('retina.xsoma' , 'fixed_20_100', 'anneal_slow_400'), 
 ]
+
+# Addditional experiments
+for seed in range(10):
+    for model in ['srm', 'sbmnodist', 'lpcm', 'mm']:
+        
+        EXPERIMENTS.append(('%s_%02d' %(model, seed),
+                            'cv_nfold_10',
+                            'fixed_20_100',
+                            'anneal_slow_1000'))
+        
+
 
 PRED_EVALS= np.logspace(-4, 0, 41) # np.linspace(0, 1.0, 41)
 
@@ -150,7 +162,7 @@ KERNEL_CONFIGS = {
 def srm_params():
     for seed in range(10):
         filename = "srm_%02d.sourcedata" % seed 
-    yield None, td(filename), seed
+        yield None, td(filename), seed
 
 @files(srm_params)
 def create_data_srm(infile, outfile, seed):
@@ -192,7 +204,7 @@ def create_data_srm(infile, outfile, seed):
 def sbmnodist_params():
     for seed in range(10):
         filename = "sbmnodist_%02d.sourcedata" % seed 
-    yield None, td(filename), seed
+        yield None, td(filename), seed
 
 @files(sbmnodist_params)
 def create_data_sbm_nodist(infile, outfile, seed):
@@ -235,7 +247,7 @@ def create_data_sbm_nodist(infile, outfile, seed):
 def lpcm_params():
     for seed in range(10):
         filename = "lpcm_%02d.sourcedata" % seed 
-    yield None, td(filename), seed
+        yield None, td(filename), seed
 
 @files(lpcm_params)
[…]

### 2014-12-10 fabfile.py

@@ -33,7 +33,7 @@ def deploy():
     
     project.rsync_project("/data/connect-disco-paper/", local_dir="./",
                           exclude=['*.pickle', 'experiments/synthdifferent/sparkdata'],
-                          extra_opts='--files-from=.git-files-list')
+                          extra_opts='--files-from=.git-files-list -l')
     
     env['forward_agent']= True
     with cd("/data/connect-disco-paper"):

### 2014-12-14 docs/paper.tex
text updates

@@
 
-\fxwarn{Do we talk about cross-validation early on? What about posterior uncertainty?}
+\fxwarning{Do we talk about cross-validation early on? What about
+  posterior uncertainty?}
 
@@
 the area under the curve (AUC) for the resulting ROC curve. 
-\fxwarn{Add additional ARI} 
+\fxwarning{Add additional ARI} 
 \section*{Results}
@@
 
-\fxwarn{Add results}
+\fxwarning{Add results}
 
@@
 
-\fxwarn{Add ``intrudoce spatial structure where there is none}
+\fxwarning{Add ``intrudoce spatial structure where there is none}
 
@@
 and microcircuitry to be discovered from connectomics data.  We have
-shown its applicability to regularly structured laminar neural circuits like the retina, 
-as well as a less structured whole neuronal organism (\textit{C. elegans}) and an artificial processor. When compared to existing methods, we show how the
-incorporation of all of this data yields results that combine both
-high link-prediction accuracy and high agreement with human
-anatomists. We have found that combining the
-available data types allows us to discover cell types and
-microcircuitry that were known to exist in the systems based on
+shown its applicability to regularly structured laminar neural
+circuits like the retina, as well as a less structured whole neuronal
+organism (\textit{C. elegans}) and an artificial processor. When
+compared to existing methods, we show how the incorporation of all of
+this data yields results that combine both high link-prediction
+accuracy and high agreement with human anatomists. We have found that
+combining the available data types allows us to discover cell types
+and microcircuitry that were known to exist in the systems based on
 decades of previous research and allows good prediction of
@@
 
-\subsection*{Microprocessor}
-We extracted the connection graph for the transistors in the MOS6502
-\autocite{visual6502source}. Each transistor has three terminals (gate,
-source, drain), but the methods of the original dataset were unable to
-consistently resolve which of the C1 and C2 terminals were source and
-drain, leading to ambiguity in our encoding. We identified a region
-consisting of three registers X, Y, and S via visual
-inspection and focused our efforts there. We created a total of six
-connectivity matrices by examining possible terminal pairings. One
-graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
-$e_j$ are connected via pins $g$ and $c_1$.
-
 % \begin{addendum}

### 2014-12-14 docs/processor.tex
text updates

@@
 vertical layout of the same types of transistors.
+
+\subsection*{Microprocessor}
+We extracted the connection graph for the transistors in the MOS6502
+\autocite{visual6502source}. Each transistor has three terminals (gate,
+source, drain), but the methods of the original dataset were unable to
+consistently resolve which of the C1 and C2 terminals were source and
+drain, leading to ambiguity in our encoding. We identified a region
+consisting of three registers X, Y, and S via visual
+inspection and focused our efforts there. We created a total of six
+connectivity matrices by examining possible terminal pairings. One
+graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
+$e_j$ are connected via pins $g$ and $c_1$.

### 2015-01-03 docs/paper.tex
update docs post konrad writing

@@
   can reveal interesting structure in the nervous system of
-  C. elegans.
+  C. elegans and a historic microprocessor.
   Our approach extracts structural meaning from
@@
 as an independent discipline \autocite{Koboldt2013}.
+\fxwarning{cite background article, cite the paper from Yuste compare supervised unsupervised on morphology, cite seungs new article}
 
@@
 When it comes to neuroscience and other computing systems, we expect
-patterns of connectivity much more complex than traditional motifs,
+patterns of connectivity much more complex than traditional motifs \fxwarning{Add citation of Science “motifs” paper},
 exhibiting a strong spatial dependence arising from the complex
@@
   \caption{Deriving circuitry and types from connectomics data a.
-As input we take the connectivity betwen cells (b), the distance between them (c), the depth of the cell bodies (d), and the depth profile of the syanpses (e).
+As input we take the connectivity between cells (b), the distance between them (c), the depth of the cell bodies (d), and the depth profile of the synapses (e).
 f.
@@
     distance-dependent connectivity profile, similar depth, and a
-    similar synpatic density profile, with cells of other types.  This
+    similar synaptic density profile, with cells of other types.  This
     results in a clustering of the cells by those hidden
@@
 exponential distance-link functions.  We additionally model cell body
-depth unimodally and synapse density profile multimodially (see
+depth unimodally and synapse density profile multimodally (see
 Methods for mathematical details).
@@
 \subsection{Metrics}
-We focus on two questions we can ask from the learned model: which
-cells are of the same type? and given any two cells, what is the
-probability of their connection?
+To evaluate the quality of the model fit, we need to use information that quantifies aspects of the data for which we have ground truth information.
+We focus on two aspects of performance.
+First, if the model works well, then the probability that a pair of neurons is of the same type should be high if the neurons actually are of the same type.
+Second, the model should assign a high probability of connection between two cells if they have a connection in the underlying data.
+We term these two factors clustering accuracy and link prediction accuracy.
 
 To assess the accuracy of a clustering compared to that determined by
-neuroanatomists, we employ the adjustded Rand index (ARI). ARI
-measures the similarity of two clusterings. While there are no perfect
-clustering comparison metrics (see supplemental for a discussion of
-ARI's shortcomings) all suffer from the need to balance
+neuroanatomists, we employ the adjusted Rand index (ARI).
+ARI is an established metric which measures the similarity of two clusterings -- two identical clusterings will have an ARI of 1.0.
+There are no perfect clustering comparison metrics as all suffer from the need to balance
 ``overclustering'' (calling a single type two different types) with
-``merging'' (calling two distinct types a single type).
+``merging'' (calling two distinct types a single type).  See supplemental for a discussion of
+ARI's shortcomings.
+Despite its shortcomings we use the ARI as a simple measure of clustering accuracy as it quantitatively captures the difference between the types discovered and the types believed by neuroanatomists.
 
-To test whether our model accurately captures the true structure of
-the data, we learn the model on the data with a subset of the links
-missing and then compute our predictive accuracy. We perform 10-way
+To assess the accuracy of the model for connections, we use link prediction
+accuracy. If our model accurately captures the true structure of
+the data, it should be good at predicting if a link exists.
+We thus train the model on the data with a subset of the links
+marked as unobserved and thus compute our predictive accuracy. We perform 10-way
 cross validation on a given dataset, learn the resulting model, and
@@
 the area under the curve (AUC) for the resulting ROC curve.
+An AUC of 1.0 means that we perfectly predict the presence and absence of the missing synapses.
+We use link prediction accuracy to quantify how good the model is at discovering the underlying connectivity.
 \fxwarning{Add additional ARI}
+
 \section*{Results}
+We will first establish that our algorithm works properly and try to understand its properties using simulated data.
+Subsequently we will analyze in detail a dataset on the retina.
+Lastly we will briefly discuss the analysis of data from the worm \textit{c.elegans} and from an ancient microprocessor.
 
@@
 and cell type identity.  We thus simulate data for which we know the
-correct structure and comparing the estimated structure based on the
+correct structure and compare the estimated structure based on the
 algorithm (see methods) with the one we used for simulation. We find
@@
 (fig~\ref{fig:synthetic}b), and the spatial extent of each type
-(fig~\ref{fig:synthetic}c).  For comparison, existing infinite
-stochastic block model assumes cell type alone matters, and thus finds
+(fig~\ref{fig:synthetic}c).  For comparison, the infinite
+stochastic block model (fig~\ref{fig:synthetic}a, b, c, black line) assumes that only cell type matters, and thus finds
 small neighborhoods of connected nodes (instead of global connectivity
-patterns). The model converges relatively quickly to an estimate of
+patterns). Our model converges relatively quickly (see supplemental) to an estimate of
 the most probable values for the cell types, which is enabled by using
 a combination of simulated annealing and parallelized Markov chain
-Monte Carlo (see methods for details). Thus we can apply our model to
-simulated datasets with structure and scale similar to that of our
-biological datasets and recover the known correct structure.
+Monte Carlo (see methods for details). Thus our model at least is promising to be applied to biological datasets.
 
 \subsection{Model Mismatch}
-Our type-and-distance assumptions for our connectivty model are
+To understand how our model would perform on real datasets that do not perfectly adhere to its generative model we next want to analyze its performance in such cases. To show where our model
+performs poorly, we attempt connectivity inference on five sets of
+synthetic data. Our type-and-distance assumptions for our connectivity model are
 simplifying, and reflect a general consensus about the organization of
-some (but not all) types of neural tissue. To show where our model
-performs poorly, we attempt connectivity inference on five sets of
-synthetic data. Our goal is to show that our model does not find
-``structure'' when there is none. \fxwarning{Fix this phrasing -- we want to say 'if your model thinks there are types, there probably are!}.
+some (but not all) types of neural tissue.
+We assume that type is discrete (as opposed to the lpcm).
+We assume that each neuron only has one type (as opposed to the mixed membership model).
+We assume that connectivity is a function of distance and type (as opposed to the classic stochastic block model).
+Our goal is to show that our model does not find
+``structure'' when there is none.
 
+We generate ten synthetic datasets based on assumptions of other models and analyze our model’s performance on these datasets.
+More specifically we perform 10-way cross validation to predict link accuracy as well as ARI to the known (``true'') type, type spatial extent, and type number.
+We find that x.
+We find that y.
+We find that z.
+Our results thus strongly suggest even when the data is not generated according to our own generative model that our results are meaningful.
+
+\fxwarning{Add results} 
+\fxwarning{add spatial extent of types}
+
+\fxwarn{add to figure comments}
 \begin{itemize}
@@
 
-We generate ten synthetic datasets based on each of these assumptions,
-and perform 10-way cross validation to predict link accuracy as well
-as ARI to the known (``true'') type, type spatial extent, and type
-number.
-
-\fxwarning{Add results}
-
 \subsection{Sensitivity to Edge Effects}
 
-Initial connectomic efforts reconstruct only a small section of tissue
--- as a result, many connections to cells outside that tissue volume
-will be lost. We performed simulations to check if 1. edge effects
+Connectomic efforts so far have reconstructed only small sections of tissue at a time. Consequently, many connections to cells outside that tissue volume
+will be lost.
+We are concerned that this selective elimination of connectivity along the boundary might give the appearance of distance-dependent connectivity when there is none.
+We thus performed simulations to check if edge effects
 could destroy spatial structure and if edge effects could introduce
-artifical, spurious spatial structure.
+artificial, spurious spatial structure.
+We measure the degree to which the model infers distance dependent connectivity for two synthetic datasets -- one with distance-dependence and one without.
+We find that, if there is no distance-dependence, edge effects do not artificially introduce distance dependence.
+However, if the section we are examining is too small, our model can miss the distance dependence.
+Thus with respect to distance-dependent connectivity inference, our model errs on the side of caution.
+But we also find that for spatial extend that is similar to the currently available datasets the effects of this are quite limited.
 
-\fxwarning{Add ``intrudoce spatial structure where there is none}
+\fxwarning{Add in a line on this figure to show where the existing data is by looking at old-school tracer reconstruction in the retina. CITE }
 
@@
     information, as well as the clustering accuracy.
-Note that our model sacrifies very little predictive accuracy for additional clustering accuracy.
+Note that our model sacrifices very little predictive accuracy for additional clustering accuracy.
 By comparison, conventional methods fail at one or both.

### 2015-01-03 experiments/celegans/sparkpipelinecv.py

@@ -5,7 +5,7 @@ SPARK_DRIVER_MEMORY=4g ~/projects/sparkutil/src/spark-1.1.0-bin-cdh4/bin/spark-s
 
 To run remotely:
 
-SPARK_DRIVER_MEMORY=32g ~/spark/bin/spark-subm --conf spark.exutorEnv.PYTHONPATH=`pwd` --conf spark.executor.memory=6g --conf spark.task.cpus=4  --conf spark.kryoserializer.buffer.mb=512 --conf spark.akka.frameSize=1000 --py-files=/data/netmotifs.egg,../../code/cvpipelineutil.py sparkpipelinecv.py
+SPARK_DRIVER_MEMORY=32g ~/spark/bin/spark-submit --conf spark.exutorEnv.PYTHONPATH=`pwd` --conf spark.executor.memory=6g --conf spark.task.cpus=4  --conf spark.kryoserializer.buffer.mb=512 --conf spark.akka.frameSize=1000 --py-files=/data/netmotifs.egg,../../code/cvpipelineutil.py sparkpipelinecv.py
 
 
 """

### 2015-01-03 experiments/synthdifferent/sparkanalysis.py

@@ -141,7 +141,8 @@ def extract_metadata(fname):
     meta['kernel_config'] = fields[3]
 
     return meta
-@merge(td("*.samples.organized/assign.pickle"),
+
+@merge(td("*.samples.organized/aggstats.pickle"),
          "aggregate.stats.pickle")
 def aggregate_stats(infiles, outfile):
     """
@@ -153,8 +154,10 @@ def aggregate_stats(infiles, outfile):
     res = []
     for infile in infiles:
         d = pickle.load(open(infile, 'r'))
-
+        print "The file is", infile
         assigndf = d['df']
+        print assigndf.dtypes
+        print assigndf['heldout_link_predprob']
 
         m = extract_metadata(infile)
         if len(m) == 0:
@@ -167,16 +170,16 @@ def aggregate_stats(infiles, outfile):
         # compute the statistics
         assigndf['ari'] = assigndf.apply(lambda x : metrics.adjusted_rand_score(x['true_assign'], irm.util.canonicalize_assignment(x['assign'])), axis=1)
 
-        assigndf['homogeneity'] = assigndf.apply(lambda x : metrics.homogeneity(x['true_assign'], irm.util.canonicalize_assignment(x['assign'])), axis=1)
+        assigndf['homogeneity'] = assigndf.apply(lambda x : metrics.homogeneity_score(x['true_assign'], irm.util.canonicalize_assignment(x['assign'])), axis=1)
 
-        assigndf['completeness'] = assigndf.apply(lambda x : metrics.completeness(x['true_assign'], irm.util.canonicalize_assignment(x['assign'])), axis=1)
+        assigndf['completeness'] = assigndf.apply(lambda x : metrics.completeness_score(x['true_assign'], irm.util.canonicalize_assignment(x['assign'])), axis=1)
 
 
 
         assigndf['type_n_true'] = assigndf.apply(lambda x : len(np.unique(x['true_assign'])), axis=1)
         assigndf['type_n_learned'] = assigndf.apply(lambda x : len(np.unique(x['assign'])), axis=1)
-        assigndf['auc'] = assigndf.apply(lambda x: metrics.roc_auc_score(x['heldout_link_truth'], x['heldout_link_predprob']))
-        assigndf['auc'] = assigndf.apply(lambda x: metrics.f1_score(x['heldout_link_truth'], x['heldout_link_predprob']))
[…]

### 2015-01-03 experiments/synthdifferent/sparkpipeline.py

@@ -72,7 +72,7 @@ S3_PATH= "netmotifs/paper/experiments/synthdifferent"
 
 JITTER = 0.4
 
-DATASET_N = 1
+DATASET_N = 10
 
     
 def td(fname): # "to directory"
@@ -271,7 +271,7 @@ def create_data_latent_position(infile, outfile, seed):
 
         nodes_with_meta[i]['latentpos'][:] = latentpos
 
-    beta = 2.0
+    beta = 1.0
 
     # now the connectivity
     connectivity = np.zeros((NODE_N, NODE_N), dtype=np.uint8)
@@ -484,7 +484,7 @@ def spark_run_experiments(data_filename, (out_samples, out_cv_data, out_inits),
     sc.stop()
 
     
-
+@jobs_limit(1)
 @transform(spark_run_experiments, suffix('.samples'), '.samples.pickle')
 def get_samples((exp_samples, exp_cvdata, exp_inits), out_filename):
     sample_metadata = pickle.load(open(exp_samples, 'r'))
@@ -496,7 +496,7 @@ def get_samples((exp_samples, exp_cvdata, exp_inits), out_filename):
     
     sc.stop()
 
-    
+@jobs_limit(1) 
 @transform(spark_run_experiments, suffix('.samples'), '.cvdata.pickle')
 def get_cvdata((exp_samples, exp_cvdata, exp_inits), out_filename):
     cvdata_metadata = pickle.load(open(exp_cvdata, 'r'))
@@ -748,6 +748,6 @@ if __name__ == "__main__":
         #plot_predlinks_roc,
         #plot_ari, 
         # plot_circos_latent,
-    ], multiprocess=6)
+    ])
     
     

### 2015-02-13 docs/paper.tex
updates to figure plots, etc

@@
   \centerline{\includegraphics[width=183mm]{fig1b.ai}}
-  \caption{Deriving circuitry and types from connectomics data a.
-As input we take the connectivity between cells (b), the distance between them (c), the depth of the cell bodies (d), and the depth profile of the synapses (e).
-f.
-Our algorithm discovers hidden cell types in this
-    connectivity data by assuming all cells of a type share a
-    distance-dependent connectivity profile, similar depth, and a
-    similar synaptic density profile, with cells of other types.  This
-    results in a clustering of the cells by those hidden
+  \caption{Deriving circuitry and types from connectomics data a. As
+    input we take the connectivity between cells (b), the distance
+    between them (c), the depth of the cell bodies (d), and the depth
+    profile of the synapses (e). f. Our algorithm discovers hidden
+    cell types in this connectivity data by assuming all cells of a
+    type share a distance-dependent connectivity profile, similar
+    depth, and a similar synaptic density profile, with cells of other
+    types.  This results in a clustering of the cells by those hidden
     types. f.) shows the cell connectivity matrix with cells of the
@@
     this case, the cells are likely to connect when they are
-    close. h. shows the probability of connection between two cell types
-that very rarely connect -- there's a background ``base'' connection rate
-to account for errors in data, but the probability is very low. 
-We can plot this these types (i) to show the relationship and spatial
-connections, and how probable various types are to connect to one
-another. j.) The connectivity between our discovered types changes as a
-function of distance. }
-
+    close. h. shows the probability of connection between two cell
+    types that very rarely connect -- there's a background ``base''
+    connection rate to account for errors in data, but the probability
+    is very low.  We can plot this these types (i) to show the
+    relationship and spatial connections, and how probable various
+    types are to connect to one another. j.) The connectivity between
+    our discovered types changes as a function of distance. }
 \label{fig:overview}
@@
 
-\subsection{Ground Truth Validation}
+\subsection{Validation with simulated data where ground truth is known}
 To validate our model, we performed a series of simulations to test if
@@
 (fig~\ref{fig:synthetic}b), and the spatial extent of each type
-(fig~\ref{fig:synthetic}c).  For comparison, the infinite
-stochastic block model (fig~\ref{fig:synthetic}a, b, c, black line) assumes that only cell type matters, and thus finds
+(fig~\ref{fig:synthetic}c).  For comparison, we show the results from instead using the infinite
+stochastic block model (fig~\ref{fig:synthetic}a, b, c, black line) which assumes that only cell type matters, and thus finds
 small neighborhoods of connected nodes (instead of global connectivity
-patterns). Our model converges relatively quickly (see supplemental) to an estimate of
+patterns).
+This contrast shows that while the regular block model can not correctly deal with distance dependent connectivity, our model can.
+Our model converges relatively quickly (see supplemental) to an estimate of
 the most probable values for the cell types, which is enabled by using
@@
 \subsection{Model Mismatch}
-To understand how our model would perform on real datasets that do not perfectly adhere to its generative model we next want to analyze its performance in such cases. To show where our model
-performs poorly, we attempt connectivity inference on five sets of
-synthetic data. Our type-and-distance assumptions for our connectivity model are
-simplifying, and reflect a general consensus about the organization of
-some (but not all) types of neural tissue.
-We assume that type is discrete (as opposed to the lpcm).
-We assume that each neuron only has one type (as opposed to the mixed membership model).
-We assume that connectivity is a function of distance and type (as opposed to the classic stochastic block model).
-Our goal is to show that our model does not find
-``structure'' when there is none.
-
-We generate ten synthetic datasets based on assumptions of other models and analyze our model’s performance on these datasets.
-More specifically we perform 10-way cross validation to predict link accuracy as well as ARI to the known (``true'') type, type spatial extent, and type number.
-We find that x.
-We find that y.
-We find that z.
-Our results thus strongly suggest even when the data is not generated according to our own generative model that our results are meaningful.
-
-\fxwarning{Add results} 
-\fxwarning{add spatial extent of types}
-
-\fxwarn{add to figure comments}
-\begin{itemize}
-\item Data with type-specific connectivity and distance dependence
-\item Data with type-specific connectivity and no distance dependence
-\item Data generated according to a latent-position clustering model,
-  where each cell belongs to a type which is part of a cluster in a
-  low-dimensional space \fxwarning{Cite this paper}
-\item cell type is combinatorial -- there are three possible
-  properties a cell can have, and each subset of properties influences
-  connectivity to neighboring cells. 
-\item cell type is a continuum -- each cell's type continuously varies
-  from zero to one, and the probability of a connection between cells
-  depends on this type.
-\end{itemize}
+
+\fxwarning{Add xlabel, ylabel, make line width of axes a bit bigger, larger number font, 
+all label fonts should be the same size }
+
+To understand how our model would perform on real datasets that do not
+perfectly adhere to its generative model we next want to analyze its
+performance in such cases. To understand the properties of our model, we
+attempt connectivity inference on four sets of synthetic data. This helps us understand what our model would do if the data does not obey our assumptions.
+
+We thus generate ten sets of synthetic data from each of four existing models. 
+The distance-dependent stochastic block model assumes type depends on distance, 
+the traditional stochastic block model has no notion of distance, the mixed membership block model assumes type is combinatorial, and the latent position cluster model assumes that type is clustered-but-continuous.
+
+If the data is sampled from our model, inference according to our model, unsurprisingly, is good by all measures.
+It correctly estimates the number of cell types; it is good at predicting connectivity (high AUC); it agrees with human classification (Rand index); it discovers all types, and leads to homogenous estimates (Fig.
+\ref{fig:othermodel} first row).
+If the data comes from a block model without distance dependence, we see that it still does pretty well on all meaningful measures (Fig.
+\ref{fig:othermodel} second row).
+It basically estimates correctly that there is no distance dependence.
+For the mixed membership model (Fig.
+\ref{fig:othermodel} third row), the model grossly overestimates the number of types, by basically allocating a type for each combination of memberships.
+Otherwise it still performs relatively well.
+Lastly, for the latent position clustering model (Fig.
+\ref{fig:othermodel} fourth row), the model does poorly.
+If type is continuous instead of discrete then our model is basically trying to cover a continuous set with a discrete scenario leading to rather poor performance.
+However, as we do expect cell types to be real in biology we might expect our model to do well with real data.
+
+\begin{figure}
+  \centering
+
+  \includegraphics[width=6.0in]{synthdifferent.ai}
+  \caption{Model inferences when true generating model differs from our distance-block-model prior.
+Horizontal columns show results with synthetic data generated according to the latent position cluster model, the mixed membership block model, the non-distance dependent stochastic block model, and the distance dependent stochastic block model.
+In all cases histograms represent posterior distribution over indicated metric.
+a.) The number of types found by the model, vertical dash line indicates ``true'' type number (not applicable to mixed membership model).
+b.) The area under the ROC curve, indicating link prediction accuracy.
+c, d, e.) Clustering metrics
+quantifying degree of type agreement with known ground truth. }
+
+  \label{fig:othermodel}
+\end{figure}
 
@@
 
-Connectomic efforts so far have reconstructed only small sections of tissue at a time. Consequently, many connections to cells outside that tissue volume
-will be lost.
-We are concerned that this selective elimination of connectivity along the boundary might give the appearance of distance-dependent connectivity when there is none.
-We thus performed simulations to check if edge effects
-could destroy spatial structure and if edge effects could introduce
-artificial, spurious spatial structure.
-We measure the degree to which the model infers distance dependent connectivity for two synthetic datasets -- one with distance-dependence and one without.
-We find that, if there is no distance-dependence, edge effects do not artificially introduce distance dependence.
-However, if the section we are examining is too small, our model can miss the distance dependence.
-Thus with respect to distance-dependent connectivity inference, our model errs on the side of caution.
-But we also find that for spatial extend that is similar to the currently available datasets the effects of this are quite limited.
+\begin{figure}
+  \centering 
+  % generated from experiments/edgeeffects/edgeeffects.ipynb
+  \centerline{\includegraphics[width=3.5in]{edgeeffects.pdf}}
+  \caption{edge effects}
+\label{fig:synthetic}
+\end{figure}
+
+\fxwarning{cut down on the number of ticks, this will be a small figure, homogenize fonts}
+
+Connectomic efforts so far have reconstructed only small sections of
+tissue at a time. Consequently, many connections to cells outside that
+tissue volume will be lost. We are concerned that this selective
+elimination of connectivity along the boundary might give the
+appearance of distance-dependent connectivity when there is none. We
+thus performed simulations to check if edge effects could destroy
+spatial structure and if edge effects could introduce artificial,
+spurious spatial structure. We measure the degree to which the model
+infers distance dependent connectivity for two synthetic datasets --
+one with distance-dependence and one without. We find that, if there
+is no distance-dependence, edge effects do not artificially introduce
+distance dependence. However, if the section we are examining is too
+small, our model can miss the distance dependence. Thus with respect
+to distance-dependent connectivity inference, our model errs on the
+side of caution. But we also find that for spatial extend that is
+similar to the currently available datasets the effects of this are
+quite limited.
 
@@
 
-Having shown our model to work on the repeating tessellated, laminar structure of the mammilian retina, we then apply our model to a structurally very different connectome -- the whole body of a small roundworm:
+Having shown our model to work on the repeating tessellated, laminar structure of the mammalian retina, we then apply our model to a structurally very different connectome -- the whole body of a small roundworm:
 \textit{Caenorhabditis elegans} is a model system in
@@
 number of anterior ganglia.
-C.
-elegans also differs from the retina in that the measured connectome is actually two separate graphs -- one of directed chemical synapses and another of undirected electrical synapses.
+\textit{C.
+elegans} also differs from the retina in that the measured connectome is actually two separate graphs -- one of directed chemical synapses and another of undirected electrical synapses.
+As this is a very different connectome, it allows an interesting generalization test: how well will our model work on such a distinct data set.
+
+\fxwarning{Fix linewidth too small, align d.e.f. callouts, change multiple fonts. Boxes same size. Need y labels (number of samples in bin) }
 
@@
   \centering 
-  \centerline{\includegraphics[width=4.5in]{celegans.ai}}
-  \caption{Discovering connectivity and type in \textit{C. elegans}. 
- a.) Initial
-    \textit{C. elegans} cell connectivity matrix, red are
-    (directional) chemical synapses, blue are electrical
-    synapses. Point size is proportional to synapse count between
-    those cells. b.) Clustered adjacency matrix showing identified
-    clusters. c.) Inter-cluster connectivity, showing the position of
-    each cell along the body axis as well as the ``true'' cell type --
-    both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). }
+  \centerline{\includegraphics[width=4.7in]{celegans2.ai}}
+  \caption{Discovering connectivity and type in
+    \textit{C. elegans}. Discovered types a.) The cell/cell
+    coassignment matrix identifies the posterior probability of any
+    two cells in \textit{c. elegans} being of the same type. The large
+    visible clusters roughly correspond to a more-coarse version of
+    types identified by neuroanatomists seen in b.) In particular,
+    robust subpopulations are found corresponding to divisions of
+    moter, interneuron, and sensory neurons. c.) The location of the
+    cell soma, revealing how spatial localizatinion correlates with
+    type. d.) The ROC curves for held-out link probability for both
+    the electrical synapses (gap junctions) and chemical synapses in
+    c. elegans. e.) The posterior distribution of the area under the
+    ROC curve for the curves in d. f.) Measurements of the
+    agreement of our identified cell types compared to neuroanatomists. 
+    See methods for an explanation of metrics. }
   \label{fig:celegans}
@@
 Using both the chemical and electrical connectivity (see methods), we
-determined the underlying clusters explained by connectivity and
+determined the underlying cell types explained by connectivity and
 distance (fig~\ref{fig:celegans}a). A superficial inspection of the results shows
@@
 1986.
-
-We identify cell types that reflect the known motor/non-motor neuron
-classes, even though this system lacks the strong repeat
-microcircuitry our model was designed for.  Motor-neuron types AS, DA,
-and VA, all exclusively postsynaptic, are identified as a common type,
-as are motor-neuron types VD and DD. Traditional types VC, DB, and VB
-also mostly share a cluster. Various head motor neurons, including
-types SMD and RMD, are clustered together. Interneurons with known
-anatomically-distinct connectivity patterns, such as AVA (2 cells),
-are clustered into pure types.The algorithm even correctly places the
-single-cell types DVB and DVC by themselves.
 
@@
 
+\fxwarning{Add chip back in}
+
 \section*{Discussion}
@@
 circuits like the retina, as well as a less structured whole neuronal
-organism (\textit{C. elegans}) and an artificial processor. When
+organism (\textit{C. elegans}) and a classical processor. When
 compared to existing methods, we show how the incorporation of all of
@@
 \autocite{Hoffman2013} and spectral learning \autocite{Anandkumar2012}
-and future work could adapte them to find faster approximate solutions
+and future work could adapt them to find faster approximate solutions
 to our problem.

### 2015-02-13 experiments/celegans/sparkpipelineanalyze2.py

@@ -1,0 +1,150 @@
+"""
+"""
+import sys
+import matplotlib
+matplotlib.use('Agg')
+# total hack, I should really know better
+sys.path.append("../../code")
+
+from ruffus import *
+import cPickle as pickle
+import numpy as np
+import copy
+import os, glob, sys, shutil
+from glob import glob
+import time
+from matplotlib import pylab
+import matplotlib
+import pandas
+import colorbrewer
+from sklearn import metrics
+import models
+import multyvac
+
+import irm
+import irm.data
+import util
+from irm import rand
+import predutil
+import boto
+
+import cvpipelineutil as cv
+import sparkutil
+
+
+def dist(a, b):
+    return np.sqrt(np.sum((b-a)**2))
+
+
+def to_f32(x):
+    a = np.array(x).astype(np.float32)
+    assert type(a) == np.ndarray
+    return a
+
+
+WORKING_DIR = "sparkdatacv"
+S3_BUCKET = "jonas-testbucket2"
+
+
+
+
+
+    
+def td(fname): # "to directory"
+    return os.path.join(WORKING_DIR, fname)
+
+def get_dataset(data_name):
+    return glob(td("%s.data" %  data_name))
+
+
+def extract_metadata(fname):
+    # extract metadata from the filename
+    dataset_name = os.path.basename(os.path.dirname(fname))[:-len('.samples.organized')]
+    fields = dataset_name.split('-')
+
+    # skip the stupid non-replicated ones
+    data_name, rel_status, source_model, rep_id = fields[0].split(".")
+    meta = {}
+    meta['source_model'] = source_model
+    meta['rep_id'] = rep_id
+
+    meta['cv_type'] = fields[1]
+    meta['init_config'] = fields[2]
+    meta['kernel_config'] = fields[3]
+
+    return meta
+
+@merge(td("*fixed_20_200*.samples.organized/aggstats.pickle"),
+         "aggregate.stats.pickle")
+def aggregate_stats(infiles, outfile):
+    """
+    Combine all the aggstats into a single file
+    
+    Compute summary statistics
+    """
+
+    res = []
+    for infile in infiles:
+        d = pickle.load(open(infile, 'r'))
+        print "The file is", infile
+        assigndf = d['df']
+        meta = d['meta']
+        neurons = meta['neurons']
+
+
+        m = extract_metadata(infile)
+        if len(m) == 0:
+            # skip the stupid non-replicated ones
+            continue 
+
+        for k, v in m.iteritems():
+            assigndf[k] = v
+        
+
[…]

### 2015-02-13 experiments/celegans/sparkpipelinecv.py

@@ -1275,6 +1275,112 @@ def cv_collate_predlinks_assign(cv_dirs, (predlinks_outfile,
                  open(assign_outfile, 'w'))
 
                   
+@follows(samples_organize)
+@follows(cvdata_organize)
+@collate("sparkdatacv/*.samples.organized/*",
+         regex(r"(sparkdatacv/.+.samples.organized)/(\d+)"),
+         #[td(r"\1-\2\4.predlinks"), td(r"\1-\2\4.assign")])
+         
+         r'\1/aggstats.pickle')
+def cv_collate_aggstats(cv_dirs,  assign_outfile):
+                                         
+    """
+    aggregate across samples and cross-validations
+    """
+    for d in cv_dirs:
+        print "cv dir", d
+
+    
+    base = os.path.dirname(cv_dirs[0])[:-len('.samples.organized')]
+    s = base.split('-')
+
+    input_basename = s[0]
+    input_data = pickle.load(open(input_basename + ".data"))
+    input_latent = pickle.load(open(input_basename + ".latent"))
+
+    input_meta = pickle.load(open(input_basename + ".meta"))['infile']
+    meta = pickle.load(open(input_meta, 'r'))
+
+
+    assignments_outputs = []
+
+    for relation_name in input_data['relations']:
+        print "computing predlinks for relation", relation_name
+        data_conn = input_data['relations'][relation_name]['data']
+        model_name= input_data['relations'][relation_name]['model']
+
+
+        print "META IS", meta.keys()
+        print meta['neurons'].dtypes
+
+        N = len(data_conn)
+
+        true_assign = meta['neurons']['cell_class'] # input_latent['domains']['d1']['assignment']
+
+
+        print "for", input_basename, "there are", len(np.unique(true_assign)), "classes"
+        if model_name == "BetaBernoulliNonConj":
+            truth_mat = data_conn
+        elif model_name == "LogisticDistance":
+            truth_mat = data_conn['link']
+        elif model_name == "LogisticDistancePoisson":
+            # For count data we just binarize with > 0 
+
+            truth_mat = data_conn['link'] > 0 
+
+
+
+        for cv_dir in cv_dirs:
+            cv_data = pickle.load(open(os.path.join(cv_dir, 'cv.data'), 'r'))
+            # get the cv idx for later use
+            cv_idx = int(os.path.basename(cv_dir))
+
+
+
+
+            # FIGURE OUT WHICH ENTRIES WERE MISSING
[…]

### 2015-02-13 experiments/mouseretina/sparkpipelineanalyze.py

@@ -242,9 +242,149 @@ def plot_ari(infile, outfile):
     ax.set_xlim(0, 1)
     f.savefig(outfile)
 
+@transform(td("retina.*.samples.organized/assign.pickle"),
+           suffix("assign.pickle"), "z.pdf")
+
+def plot_z_matrix(exp_results, 
+                  out_filename):
+    assigndf = pickle.load(open(infile, 'r'))['df']
+
+
+    sample_d = pickle.load(open(exp_results))
+    chains = sample_d['chains']
+    exp = sample_d['exp']
+    print "exp=", exp
+    data_filename = exp['data_filename']
+    data = pickle.load(open(data_filename))
+    data_basename, _ = os.path.splitext(data_filename)
+    meta = pickle.load(open(data_basename + ".meta"))
+
+    meta_infile = meta['infile']
+
+    d = pickle.load(open(meta_infile, 'r'))
+    conn = d['conn_mat']
+    cells = d['cells']
+
+    cell_types = cells['type_id']
+
+    chains = [c for c in chains if type(c['scores']) != int]
+    CHAINN = len(chains)
+
+    # compute the z matrix 
+    z_mat = np.zeros((len(cells), len(cells)))
+    for ci, c in enumerate(chains):
+        sample_latent = c['state']
+        cell_assignment = np.array(sample_latent['domains']['d1']['assignment'])
+        ca = irm.util.canonicalize_assignment(cell_assignment)
+        for u in np.unique(ca):
+            ent_i = np.argwhere(ca == u).flatten()
+            for ci in ent_i:
+                for cj in ent_i:
+                    z_mat[ci, cj] += 1
+
+    import scipy.cluster
+    l = scipy.cluster.hierarchy.linkage(z_mat, method='ward')
+    
+    
+    ca = np.array(scipy.cluster.hierarchy.leaves_list(l))
+    z_mat = z_mat[ca]
+    z_mat = z_mat[:, ca]
+
+
+    import matplotlib.gridspec as grd
+    f = pylab.figure(figsize=(8, 6))
+    gs = grd.GridSpec(1, 2, width_ratios=[2,10 ], wspace=0.02)
+    ax = f.add_subplot(gs[:, 1])
+    im = ax.imshow(z_mat/CHAINN, cmap=pylab.cm.Greys, interpolation='nearest')
+    ax.set_xticks([])
+    ax.set_yticks([])
[…]

### 2015-02-13 experiments/mouseretina/sparkpipelineanalyze2.py

@@ -1,0 +1,149 @@
+"""
+"""
+import sys
+import matplotlib
+matplotlib.use('Agg')
+# total hack, I should really know better
+sys.path.append("../../code")
+
+from ruffus import *
+import cPickle as pickle
+import numpy as np
+import copy
+import os, glob, sys, shutil
+from glob import glob
+import time
+from matplotlib import pylab
+import matplotlib
+import pandas
+import colorbrewer
+from sklearn import metrics
+import models
+import multyvac
+
+import irm
+import irm.data
+import util
+from irm import rand
+import predutil
+import boto
+
+import cvpipelineutil as cv
+import sparkutil
+
+
+def dist(a, b):
+    return np.sqrt(np.sum((b-a)**2))
+
+
+def to_f32(x):
+    a = np.array(x).astype(np.float32)
+    assert type(a) == np.ndarray
+    return a
+
+
+WORKING_DIR = "sparkdatacv"
+S3_BUCKET = "jonas-testbucket2"
+
+
+
+
+
+    
+def td(fname): # "to directory"
+    return os.path.join(WORKING_DIR, fname)
+
+def get_dataset(data_name):
+    return glob(td("%s.data" %  data_name))
+
+
+def extract_metadata(fname):
+    # extract metadata from the filename
+    dataset_name = os.path.basename(os.path.dirname(fname))[:-len('.samples.organized')]
+    fields = dataset_name.split('-')
+    print "The fields are", fields
+    # skip the stupid non-replicated ones
+    data_name, rel_status, source_model = fields[0].split(".")[:3]
+    meta = {}
+    meta['source_model'] = source_model
+    #meta['rep_id'] = rep_id
+
+    meta['cv_type'] = fields[1]
+    meta['init_config'] = fields[2]
+    meta['kernel_config'] = fields[3]
+
+    return meta
+
+@merge(td("*fixed_20_100*.samples.organized/aggstats.pickle"),
+         "aggregate.stats.pickle")
+def aggregate_stats(infiles, outfile):
+    """
+    Combine all the aggstats into a single file
+    
+    Compute summary statistics
+    """
+
+    res = []
+    for infile in infiles:
+        d = pickle.load(open(infile, 'r'))
+        print "The file is", infile
+        assigndf = d['df']
+        meta = d['meta']
+        cells = meta['cells']
+        print cells.dtypes
+
+        m = extract_metadata(infile)
+        if len(m) == 0:
+            # skip the stupid non-replicated ones
+            continue 
+
+        for k, v in m.iteritems():
[…]

### 2015-02-13 experiments/mouseretina/sparkpipelinecv.py

@@ -1927,6 +1927,120 @@ def cv_collate_predlinks_assign(cv_dirs, (predlinks_outfile,
 # #     fid = open(outfile_rpt, 'w')
 # #     fid.write(clust_df.to_html())
 # #     fid.close()
+
+@follows(samples_organize)
+@follows(cvdata_organize)
+@collate("sparkdatacv/*.samples.organized/*",
+         regex(r"(sparkdatacv/.+.samples.organized)/(\d+)"),
+         #[td(r"\1-\2\4.predlinks"), td(r"\1-\2\4.assign")])
+         
+         r'\1/aggstats.pickle')
+def cv_collate_aggstats(cv_dirs,  assign_outfile):
+                                         
+    """
+    aggregate across samples and cross-validations
+    """
+    for d in cv_dirs:
+        print "cv dir", d
+    USE_FIXED_RELATIONS = True
+    FIXED_RELATIONS = ['R1']
+
+    
+    base = os.path.dirname(cv_dirs[0])[:-len('.samples.organized')]
+    s = base.split('-')
+
+    input_basename = s[0]
+    input_data = pickle.load(open(input_basename + ".data"))
+    input_latent = pickle.load(open(input_basename + ".latent"))
+
+    input_meta = pickle.load(open(input_basename + ".meta"))['infile']
+    meta = pickle.load(open(input_meta, 'r'))
+
+
+    assignments_outputs = []
+    
+    if USE_FIXED_RELATIONS:
+        relation_set = FIXED_RELATIONS
+    else:
+        relation_set = input_data['relations']
+    # FIXME different
+    for relation_name in relation_set:
+        print "computing predlinks for relation", relation_name
+        data_conn = input_data['relations'][relation_name]['data']
+        model_name= input_data['relations'][relation_name]['model']
+
+
+        print "META IS", meta.keys()
+        print meta['cells'].dtypes
+
+        N = len(data_conn)
+
+        true_assign = meta['cells']['type_id'] # input_latent['domains']['d1']['assignment']
+
+
+        print "for", input_basename, "there are", len(np.unique(true_assign)), "classes"
+        if model_name == "BetaBernoulliNonConj":
+            truth_mat = data_conn
+        elif model_name == "LogisticDistance":
+            truth_mat = data_conn['link']
+        elif model_name == "LogisticDistancePoisson":
+            # For count data we just binarize with > 0 
+
+            truth_mat = data_conn['link'] > 0 
+
[…]

### 2015-02-13 fabfile.py

@@ -111,14 +111,14 @@ def get_data():
 
     DATASETS = [('experiments/mouseretina/sparkdatacv',
                  'experiments/mouseretina/',
-                 ['predlinks.pickle', 'assign.pickle']),
+                 ['predlinks.pickle', 'assign.pickle', 'aggstats.pickle']),
                 
                  ('experiments/synthdifferent/sparkdata',
                   'experiments/synthdifferent/',
                   ['aggstats.pickle']),
                  ('experiments/celegans/sparkdatacv',
                   'experiments/celegans/',
-                  ['predlinks.pickle', 'assign.pickle'])]
+                  ['predlinks.pickle', 'assign.pickle', 'aggstats.pickle'])]
     
    
     for remote_dir, local_dir, globs in DATASETS:

### 2015-02-21 code/linkplots.py
figure text and plotting updates

@@ -1,0 +1,111 @@
+import numpy as np
+from matplotlib.collections import PatchCollection, LineCollection
+from matplotlib.patches import Circle
+import util
+
+
+# group_colors_opaque = group_colors.copy()
+# group_colors_opaque[:, -1] = 1.0
+# fig = pylab.figure(figsize=(20, 6))
+# ax = fig.add_subplot(1, 1, 1, aspect='equal') 
+# X_SCALE = 6.0
+# CIRCLE_SIZE = 0.4
+
+# def color_func(x):
+#     return np.array([0.0, 0.0, 0.0, x**3])
+# YOFFSET = mats.shape[1]
+# for i in range(len(mats)):
+#     points = []
+#     colors = []
+#     widths = []
+    
+#     patches = []
+#     m = mats[i]
+#     for ci in range(m.shape[0]):
+#         for cj in range(m.shape[1]):
+#             if ci <= cj   and ci != cj:
+#                 p = m[ci, cj]
+#                 if p >= 0.5:
+#                     points.append([[i*X_SCALE, YOFFSET - ci], [(i+1)*X_SCALE, YOFFSET - cj]])
+#                     colors.append(color_func(p))
+#                     #print p
+#                     widths.append((1+p))
+
+    
+#     lc = LineCollection(points, colors=colors, linewidths=np.array(widths), zorder=1)
+#     ax.add_collection(lc)
+
+# # draw circles
+# for i in range(len(mats) + 1):     
+#     patches = [Circle((i*X_SCALE,YOFFSET - ci), CIRCLE_SIZE) for ci in range(mats.shape[1])]
+#     p = PatchCollection(patches, color = group_colors_opaque, edgecolor='k')
+#     ax.add_collection(p)
+
+# ax.set_yticks([])
+# ax.set_xticks(np.arange(mats.shape[0] + 1) * X_SCALE)
+# ax.set_xticklabels([""] + ["%2.0f um"% x for x in  dists], fontsize=20)
+# ax.set_xlim(-1, len(mats)*X_SCALE + 4)
+# ax.set_ylim(0, YOFFSET + 1)
+# ax.set_xlabel("cell distance", fontsize=20)
+# ax.set_ylabel("discovered cell type", fontsize=20)
+
+# color_eval_points = np.linspace(0, 1, 120)
+# util.draw_manual_color_bar(ax, color_eval_points, color_func, len(mats)*X_SCALE + 1, 4, 1, 4)
+
+
+# fig.savefig("tgt.bestsample.connections.pdf")
+
[…]

### 2015-02-21 code/util.py

@@ -6,6 +6,9 @@ from preprocess import *
 import pandas
 import sqlite3
 
+from matplotlib.collections import PatchCollection
+from matplotlib.patches import Circle
+
 # plot spatial distribution of each cell type
 # plot area vs cell body distance
 
@@ -33,3 +36,36 @@ def create_adj_mat(con, area_thold, cell_data):
  
     return area_mat, cell_data.index.values
             
+
+
+
+def draw_manual_color_bar(ax, eval_points, color_func, xoffset, yoffset, 
+                          xwidth, yscale, 
+                          border=True):
+    import matplotlib
+    from matplotlib.collections import PatchCollection
+
+    patches = []
+    colors = []
+
+    for ystart, yend in zip(eval_points[:-1], eval_points[1:]):
+        rect = matplotlib.patches.Rectangle((xoffset, ystart*yscale + yoffset), xwidth, yscale*(yend-ystart), 
+                                                edgecolor="none")
+        patches.append(rect)
+        colors.append(color_func((yend-ystart)/2.0 + ystart))
+    
+    
+    
+    p = PatchCollection(patches, edgecolor='none', color=colors)
+    
+    ax.add_collection(p)
+    
+    if border:
+        border_patch = matplotlib.patches.Rectangle((xoffset, yoffset), xwidth, 
+                                                    yscale * (np.max(eval_points) - np.min(eval_points)))
+        p = PatchCollection([border_patch], edgecolor='k', facecolor='none')
+        ax.add_collection(p)
+    
+
+
+

### 2015-02-21 docs/fig1.py

@@ -1,6 +1,12 @@
 import numpy as np
 from matplotlib import pylab
 import irm
+from matplotlib.collections import PatchCollection, LineCollection
+from matplotlib.patches import Circle    
+import sys
+sys.path.append("../code")
+import linkplots
+import util
 
 def dist(a, b):
     return np.sqrt(np.sum((b-a)**2))
@@ -160,6 +166,11 @@ f.savefig("source.f1.somadepth.raw.pdf")
 
 CLASS_N = len(np.unique(c_class))
 
+from brewer2mpl import qualitative
+bmap = qualitative.Dark1[CLASS_N]
+print bmap
+group_colors = bmap.mpl_colors # (np.linspace(0, 1, CLASS_N))
+
 ai = np.argsort(c_class).flatten()
 c_sorted = connectivity[ai, :]
 c_sorted = c_sorted[:, ai]
@@ -176,6 +187,36 @@ for d in di:
 ax.set_xticks([])
 ax.set_yticks([])
 f2.savefig("source.f1.sorted.pdf")
+
+f2 = pylab.figure()
+
+ax = f2.add_subplot(1, 1, 1)
+mat = c_sorted
+ax.imshow(mat, cmap=pylab.cm.Blues, interpolation='nearest')
+borders =  di # np.argwhere(np.diff(ca[cell_order])).flatten()
+
+ax.set_xticks([])
+ax.set_yticks([])
+
+#group_colors = np.random.permutation(group_colors)
+prev = 0
+
+BORDER_WIDTH = 2
+N = mat.shape[0]
+for b, fc in zip(borders.tolist()  + [mat.shape[0]], group_colors):
+    
+    ax.plot([prev, b], [-BORDER_WIDTH, -BORDER_WIDTH], c=fc, linewidth=10, solid_capstyle="butt")
+    
+    ax.plot([-BORDER_WIDTH, -BORDER_WIDTH], [prev, b], c=fc, linewidth=10, solid_capstyle="butt")
+    ax.plot([prev, b], [N+BORDER_WIDTH, N+BORDER_WIDTH], c=fc, linewidth=10, solid_capstyle="butt")
+    ax.plot([N+BORDER_WIDTH, N+BORDER_WIDTH], [prev, b], c=fc, linewidth=10, solid_capstyle="butt")    
+    prev = b
+    
+for p in borders:
+    ax.axhline(p + 0.5, c='k', linewidth=1)
[…]

### 2015-02-21 experiments/mouseretina/sparkpipelinecv.py

@@ -1,4 +1,21 @@
 """
+   79  SPARK_DRIVER_MEMORY=32g ~/spark/bin/spark-submit --conf spark.exutorEnv.PYTHONPATH=`pwd` --conf spark.executor.memory=6g --conf spark.task.cpus=4  --conf spark.kryoserializer.buffer.mb=512 --conf spark.akka.frameSize=1000 --py-files=/data/netmotifs.egg,../../code/cvpipelineutil.py sparkpipelinecv.py
+   82  ls sparkdatacv/*.meta
+   84  SPARK_DRIVER_MEMORY=32g ~/spark/bin/spark-submit --conf spark.exutorEnv.PYTHONPATH=`pwd` --conf spark.executor.memory=6g --conf spark.task.cpus=4  --conf spark.kryoserializer.buffer.mb=512 --conf spark.akka.frameSize=1000 --py-files=/data/netmotifs.egg,../../code/cvpipelineutil.py sparkpipelinecv.py
+   85  ls ./sparkdatacv/celegans.2r.ldp.*cv_nfold_10-fixed_20_200-anneal_slow_800.samples.organized/aggstats.pickle
+   86  ls ./sparkdatacv/celegans.2r.ldp.03*cv_nfold_10-fixed_20_200-anneal_slow_800.samples.organized/aggstats.pickle
+   87  rm ./sparkdatacv/celegans.2r.ldp.03*cv_nfold_10-fixed_20_200-anneal_slow_800.samples.organized/aggstats.pickle
+   88  SPARK_DRIVER_MEMORY=32g ~/spark/bin/spark-submit --conf spark.exutorEnv.PYTHONPATH=`pwd` --conf spark.executor.memory=6g --conf spark.task.cpus=4  --conf spark.kryoserializer.buffer.mb=512 --conf spark.akka.frameSize=1000 --py-files=/data/netmotifs.egg,../../code/cvpipelineutil.py sparkpipelinecv.py
+   89  rm ./sparkdatacv/celegans.2r.ldp.*cv_nfold_10-fixed_20_200-anneal_slow_800.samples.organized/aggstats.pickle
[…]

### 2015-02-21 code/linkplots.py
updates to figures and plots

@@ -56,7 +56,9 @@ import util
 # fig.savefig("tgt.bestsample.connections.pdf")
 
 def draw_links(ax, mats, color_func, X_SCALE=6.0, 
-               width_func = None):
+               width_func = None, directed=True, selflinks=False):
+    """
+    """
     if width_func is None:
         width_func = lambda x: x + 1.0
 
@@ -70,12 +72,16 @@ def draw_links(ax, mats, color_func, X_SCALE=6.0,
         m = mats[i]
         for ci in range(m.shape[0]):
             for cj in range(m.shape[1]):
-                if ci <= cj   and ci != cj:
-                    p = m[ci, cj]
-                    points.append([[i*X_SCALE, YOFFSET - ci],
-                                   [(i+1)*X_SCALE, YOFFSET - cj]])
-                    colors.append(color_func(p))
-                    widths.append(width_func(p))
+                if not directed and ci >= cj:
+                    continue
+                if not selflinks and ci == cj:
+                    continue
+
+                p = m[ci, cj]
+                points.append([[i*X_SCALE, YOFFSET - ci],
+                               [(i+1)*X_SCALE, YOFFSET - cj]])
+                colors.append(color_func(p))
+                widths.append(width_func(p))
         lc = LineCollection(points, colors=colors, linewidths=np.array(widths), zorder=1)
         ax.add_collection(lc)
 
@@ -89,12 +95,13 @@ def draw_circles(ax, mats, color_list, X_SCALE=6.0,  CIRCLE_SIZE=0.4):
         ax.add_collection(p)
 
 def plot_helper(ax, mats, node_colors, link_color_func, plot_colorbar = True, 
-                X_SCALE=6.0, link_width_func=None, x_pad=4): 
+                X_SCALE=6.0, link_width_func=None, x_pad=4, selflinks=False, 
+                directed=False): 
 
     YOFFSET = mats.shape[1]
 
     draw_links(ax, mats, link_color_func, X_SCALE=X_SCALE, 
-                width_func=link_width_func)
+                width_func=link_width_func, selflinks=selflinks, directed=directed)
     draw_circles(ax, mats, node_colors, X_SCALE=X_SCALE)
     ax.set_yticks([])
     ax.set_xticks(np.arange(mats.shape[0] + 1) * X_SCALE)

### 2015-02-21 docs/fig1.py

@@ -39,12 +39,13 @@ SIDE_N = 6
 
 
 class_conn = {(0, 1) : (0.7, 0.8), 
-              (1, 2) : (2.0, 0.6), 
+              (1, 2) : (2.2, 0.6), 
               #(3, 2) : (0.7, 0.9), 
-              (2, 1) : (0.7, 0.8), 
-              (0, 3) : (1.7, 0.7), 
+              (2, 1) : (0.2, 0.8), 
+              (0, 3) : (1.4, 0.7), 
               (3, 4) : (2.4, 0.8), 
-              (4, 0) : (1.0, 0.3)}
+              (4, 0) : (1.0, 0.3), 
+              }
 
 
 nodes_with_class, connectivity = irm.data.generate.c_class_neighbors(SIDE_N, 
@@ -167,7 +168,7 @@ f.savefig("source.f1.somadepth.raw.pdf")
 CLASS_N = len(np.unique(c_class))
 
 from brewer2mpl import qualitative
-bmap = qualitative.Dark1[CLASS_N]
+bmap = qualitative.Set1[CLASS_N]
 print bmap
 group_colors = bmap.mpl_colors # (np.linspace(0, 1, CLASS_N))
 
@@ -229,7 +230,7 @@ for c1 in range(CLASS_N):
         c = (c1, c2)
         if c in class_conn:
             mu = class_conn[c][0]
-            lamb = class_conn[c][0]/4
+            lamb = class_conn[c][0]/12
         else:
             mu = 0.0001
             lamb = 0.0001
@@ -244,7 +245,7 @@ f3.savefig("source.f1.latent.pdf")
 
 
 
-dists = np.array([0.1, 1.0, 2.0])
+dists = np.array([0.5, 1.0, 2.0])
 
 mats = np.zeros((len(dists), CLASS_N, CLASS_N))
 
@@ -275,10 +276,10 @@ def color_func(x):
     return np.array([0.0, 0.0, 0.0, x**3])
 
 def width_func(x):
-    return 3.0
+    return 7.0
 linkplots.plot_helper(ax, mats, group_colors, color_func, link_width_func = width_func, 
-                      X_SCALE=3.0, plot_colorbar=False, x_pad =2 )
-ax.set_xticklabels([""] + ["%2.1f um"% x for x in  dists], fontsize=20)
+                      X_SCALE=3.0, plot_colorbar=False, x_pad =2, directed=True)
[…]

### 2015-02-21 experiments/celegans/preprocess.py

@@ -1,0 +1,165 @@
+import numpy as np
+import cPickle as pickle
+from matplotlib import pylab
+from xlrd import open_workbook
+import pandas
+import os
+from ruffus import * 
+from matplotlib import colors
+
+DATA_DIR = "../../../data/celegans/conn2"
+
+@files([os.path.join(DATA_DIR, x) for x in ["NeuronConnect.xls", "NeuronType.xls", '../manualmetadata.xlsx']], "data.pickle")
+def read_data((neuron_connect_file, neuron_type_file, 
+               manual_metadata_file), output_file):
+
+
+    neuron_connect =  open_workbook(neuron_connect_file)
+    connections = {}
+    for s in neuron_connect.sheets():
+        for row_i in range(1, s.nrows):
+            neuron_1 = s.cell(row_i, 0).value
+            neuron_2 = s.cell(row_i, 1).value
+            typ = s.cell(row_i, 2).value
+            nbr = int(s.cell(row_i, 3).value)
+
+            c = neuron_1, neuron_2
+            if c not in connections:
+                connections[c] = []
+            connections[c].append((typ, nbr))
+
+    neurons = {}
+    neuron_connect =  open_workbook(neuron_type_file)
+    for s in neuron_connect.sheets():
+
+        for row_i in range(1, s.nrows):
+            neuron = str(s.cell(row_i, 0).value)
+            soma_pos = float(s.cell(row_i, 1).value)
+            neurons[neuron] = {'soma_pos' : soma_pos}
+
+    # now open the excel workbook of our metadata
+    metadata_df = pandas.io.excel.read_excel(manual_metadata_file, 
+                                             'properties', 
+                                             index_col=0)
+    assert len(metadata_df) == len(neurons)
+    ndf = pandas.Series([d['soma_pos'] for d in neurons.values()], 
+                        neurons.keys())
+    metadata_df['soma_pos'] = ndf
+    n = metadata_df
+
+    # find classes
+    classes = {}
+    orig_names = set(n.index)
+    orig_names_base = set(orig_names)
+    for cur_n in orig_names_base:
+        #cur_n = orig_names.pop()
+        # does its name end in two digits? That's your class
+
+        if cur_n not in orig_names:
+            continue
+        if cur_n[-2:].isdigit():
+            base = cur_n[:-2] 
+            tgts = [s for s in orig_names if s.startswith(base) and s[-2:].isdigit()]
+            classes[base] = tgts
+            orig_names -= set(tgts)
+
[…]

### 2015-02-23 docs/paper.tex
added cluster comparison, cleaned up plots

@@
 \usepackage[backend=biber,style=nature]{biblatex}
-\usepackage{fixme}
+\usepackage[margin=true,inline=true,index=true, status=draft]{fixme}
 \usepackage{grffile}
@@
 
+\listoffixmes
+
 \section*{Introduction}
@@
   \centering 
-  \centerline{\includegraphics[width=183mm]{fig1b.ai}}
-  \caption{Deriving circuitry and types from connectomics data a. As
-    input we take the connectivity between cells (b), the distance
-    between them (c), the depth of the cell bodies (d), and the depth
-    profile of the synapses (e). f. Our algorithm discovers hidden
-    cell types in this connectivity data by assuming all cells of a
-    type share a distance-dependent connectivity profile, similar
+  \centerline{\includegraphics[width=110mm]{fig1b.ai}}
+  \caption{Deriving circuitry and cell types from connectomics data
+    a. As input we take the connectivity between cells (b), the
+    distance between them (c), the depth of the cell bodies (d), and
+    the depth profile of the synapses (e). f. Our algorithm discovers
+    hidden cell types in this connectivity data by assuming all cells
+    of a type share a distance-dependent connectivity profile, similar
     depth, and a similar synaptic density profile, with cells of other
@@
     connection rate to account for errors in data, but the probability
-    is very low.  We can plot this these types (i) to show the
-    relationship and spatial connections, and how probable various
-    types are to connect to one another. j.) The connectivity between
-    our discovered types changes as a function of distance. }
+    is very low.  i.) shows that we also recover the expected
+    laminarity of types and the depth-specific synaptic
+    connectivity. j.) We then plot how the connectivity between these
+    types changes as a function of distance between the cell bodies to
+    better understand short-range and long-range connectivity
+    patterns. }
+
 \label{fig:overview}
@@
 
-\subsection{Metrics}
+\subsection*{Metrics}
 To evaluate the quality of the model fit, we need to use information that quantifies aspects of the data for which we have ground truth information.
@@
 
-\subsection{Validation with simulated data where ground truth is known}
+\subsection*{Validation with simulated data where ground truth is known}
 To validate our model, we performed a series of simulations to test if
@@
 
-\subsection{Model Mismatch}
+\subsection*{Model Mismatch}
 
@@
 
-\subsection{Sensitivity to Edge Effects}
+\subsection*{Sensitivity to Edge Effects}
 
@@
   \centering 
-  % generated from experiments/edgeeffects/edgeeffects.ipynb
   \centerline{\includegraphics[width=3.5in]{edgeeffects.pdf}}
-  \caption{edge effects}
-\label{fig:synthetic}
+  \caption{We generate two sets of synethetic data, one with
+    spatially-dependent connectivity and one without. We measure the
+    variance in the connectivity-distance plot for randomly selected
+    regions of each dataset, ranging from single-cells to the entire
+    volume. We see that while selecting too small of a region can 
+    destroy the appearance of distance-dependent connectivity, 
+    it does not create it in non-spatial data}
+\label{fig:edgeeffects}
 \end{figure}
 
-\fxwarning{cut down on the number of ticks, this will be a small figure, homogenize fonts}
-
 Connectomic efforts so far have reconstructed only small sections of
-tissue at a time. Consequently, many connections to cells outside that
-tissue volume will be lost. We are concerned that this selective
-elimination of connectivity along the boundary might give the
-appearance of distance-dependent connectivity when there is none. We
-thus performed simulations to check if edge effects could destroy
-spatial structure and if edge effects could introduce artificial,
-spurious spatial structure. We measure the degree to which the model
-infers distance dependent connectivity for two synthetic datasets --
-one with distance-dependence and one without. We find that, if there
-is no distance-dependence, edge effects do not artificially introduce
+tissue. Consequently, many connections to cells outside that tissue
+volume will be lost. We are concerned that this selective elimination
+of connectivity along the boundary might give the appearance of
+distance-dependent connectivity when there is none. We thus performed
+simulations to check if edge effects could destroy spatial structure
+and if edge effects could introduce artificial, spurious spatial
+structure. We measure the degree to which distance-dependent effects
+can arise from selecting regions that are smaller than the ``scale''
+of connectivity (fig ~\ref{fig:edgeeffects}). We do this by generating
+two collections of synthetic datasets -- one with distance-dependent
+connectivity and one without.  We then in each dataset randomly
+examine contiguous circular regions with area varying from zero to the
+entire volume, and empirically calculate the spatial variance in
+type-dependent connectivity.  We find that, if there is no
+distance-dependence, edge effects do not artificially introduce
 distance dependence. However, if the section we are examining is too
@@
   \centering 
-  \centerline{\includegraphics{mouseretina.ai}}
-  \caption{Discovering cell classes in the mouse retina connectome.
+  \centerline{\includegraphics[width=160mm]{mouseretina2.ai}}
+  \caption{Discovering cell classes in the mouse retina connectome. Here
+we show the maxium a posteriori (MAP) estimate for the types in the mouse
+retina data. 
     a.) Input connectivity data for 950 cells for which soma positions
-    were known. b. clustered connectivity matrix. c.) connectivity
-    diagram showing our clusters, as well as the cell depth and
-    anatomist-labeled cell type of each cell. Our model shows moderate
-    agreement with the cell types traditionally identified by
-    anatomist laminar-specific connectivity patterns. 
+    were known. b. clustered connectivity matrix, each arbitrary color
+    corresponds to a single type, and will be used to identify that type in the remainder
+    of the plot. c.)  The spatial distribution of our cell types
+    -- each cell type tesselates space. Colors correspond to those in b.)
     d.) Connectivity between our clusters as a function
     of distance -- the cluster consisting primarily of retinal
-    ganglion cells (lower left-center) exhibits the expected near and
-    far connectivity. e.)  The spatial distribution of our cell types
-    -- each cell type tesselates space.}
+    ganglion cells (brown node on graph) exhibits the expected near and
+    far connectivity.}
 \label{fig:mouseretina}
@@
   \centerline{\includegraphics{comparison.ai}}
-  \caption{\textbf{Visualizing type inference uncertainty.} Our fully Bayesian model gives a confidence estimate (posterior probability) that any two given cells are of the same type.
-In b.) we visualize that cell-cell coassignment matrix, showing the probability that cell i is of the same type as cell j on a range from 0.0 to 1.0.
-The block structure shows subsets of cells which are believed to all belong to the same type.
-For comparison, a.) shows the anatomist-defined type for each cell, grouped broadly into the coarse types identified in the previous figure.
-\textbf{Link vs cluster accuracy} c.
-A comparison of the
-    predictive accuracy (area under the curve) for hand-labeled
-    anatomical data, versus inclusion of additional sources of
-    information, as well as the clustering accuracy.
-Note that our model sacrifices very little predictive accuracy for additional clustering accuracy.
-By comparison, conventional methods fail at one or both.
-d.
-) The spatial extent (in depth and area) of the
-    types identified by humans and our various algorithmic
-    approaches.}
-
+  \caption{\textbf{Visualizing type inference uncertainty.} Our fully
+    Bayesian model gives a confidence estimate (posterior probability)
+    that any two given cells are of the same type. In a.) we visualize
+    that cell-cell coassignment matrix, showing the probability that
+    cell i is of the same type as cell j on a range from 0.0 to
+    1.0. The block structure shows subsets of cells which are believed
+    to all belong to the same type. For comparison, b.) shows the
+    anatomist-defined type for each cell, grouped broadly into the
+    coarse types identified in the previous figure. c.) for each cell
+    in the matrix, c.) shows the soma depth as well as the depth
+    histogram for all putative synaptic contacts. \textbf{Link vs
+      cluster accuracy} d.) The posterior distribution of ROC curves
+    from 10-fold cross validation when predicting connectivity, as
+    well as e.) the area under the curve and f.) the type agreements
+    with known neuroanatomist types. g. A comparison of the predictive
+    accuracy (area under the curve) for hand-labeled anatomical data,
+    versus inclusion of additional sources of information, as well as
+    the clustering accuracy. Note that our model sacrifices very
+    little predictive accuracy for additional clustering accuracy. By
+    comparison, conventional methods fail at one or both.  d. ) The
+    spatial extent (in depth and area) of the types identified by
+    humans and our various algorithmic approaches.}
 \label{fig:mouseretina_compare}
@@
   \centering 
-  \centerline{\includegraphics[width=4.7in]{celegans2.ai}}
+  \centerline{\includegraphics[width=7.0in]{celegans2.ai}}
   \caption{Discovering connectivity and type in
-    \textit{C. elegans}. Discovered types a.) The cell/cell
-    coassignment matrix identifies the posterior probability of any
-    two cells in \textit{c. elegans} being of the same type. The large
-    visible clusters roughly correspond to a more-coarse version of
-    types identified by neuroanatomists seen in b.) In particular,
-    robust subpopulations are found corresponding to divisions of
-    moter, interneuron, and sensory neurons. c.) The location of the
-    cell soma, revealing how spatial localizatinion correlates with
-    type. d.) The ROC curves for held-out link probability for both
-    the electrical synapses (gap junctions) and chemical synapses in
-    c. elegans. e.) The posterior distribution of the area under the
-    ROC curve for the curves in d. f.) Measurements of the
-    agreement of our identified cell types compared to neuroanatomists. 
-    See methods for an explanation of metrics. }
+    \textit{C. elegans}. a.) posterior distribution on cell
+    connectivity as a function of discovered type, similar to
+    figure~\ref{mouseretina}. In b.) we plot neuroanatomist-derived
+    types along with their labels. Our model shows high probability of
+    motor neurons, sensory neurons, and various interneuron classes
+    being of the same type. Soma positions along the body axis are
+    plotted in c.) where we see that we cluster spatially-distributed
+    motor neurons together, whereas head sensory neurons are more
+    likely to be grouped together as well.  d.) The ROC curves for
+    held-out link probability for both the electrical synapses (gap
+    junctions) and chemical synapses in c. elegans. e.) The posterior
+    distribution of the area under the ROC curve for the curves in
+    d. f.) Measurements of the agreement of our identified cell types
+    compared to neuroanatomists. The high compleneness but low
+    homogeneity (and corresponding low ARI) reflects our model's
+    tendency to group multiple types into a single type.}
   \label{fig:celegans}
@@
 determined the underlying cell types explained by connectivity and
-distance (fig~\ref{fig:celegans}a). A superficial inspection of the results shows
-clustering into groups consisting roughly homogeneously of motor
-neurons, sensory neurons, and interneurons. Closer examination reveals
-agreement with the classifications originally outlined by White in
-1986.
+distance (fig~\ref{fig:celegans}a). A superficial inspection of the
+results shows clustering into groups consisting roughly homogeneously
+of motor neurons, sensory neurons, and interneurons. Closer
+examination reveals agreement with the classifications originally
+outlined by White in 1986.
 
@@
 
-\fxwarning{Add chip back in}
+\subsection*{Types and connectivity in artificial structures}
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=4in]{mos6502.ai}}
+  \caption{Discovering connectivity and type the 6502 microprocessor.
+    a.) is the micrograph of the original microprocessor, with the
+    region containing the registers under study highlighted. b.) Our
+    graph consists of the interconnections of MOS field-effect
+    transistors with three terminals, Gate, C1, and C2. Reconstruction
+    technique did not permit resolution of C1 and C2 into source and
+    drain. c.) The spatial distribution of the transistors in each
+    cluster show a clear pattern d.) The clusters and connectivity
+    versus distance for connections between Gate and C1, Gate and C2,
+    and C1 and C2 terminals on a transistor. Red and teal types have a
+    terminal pulled down to ground and mostly function as
+    inverters. Purple class are clocked, stateful transistors, orange
+    control the ALU and yellow control the special data bus (SDB).}
+  \label{fig:mos6502}
+\end{figure}
+
+To show the applicability of our method to other connectome-style
+datasets, we obtained the spatial location and interconnectivity of
+the transistors in a classic microprocessor, the MOS Technology 6502
+(used in the Apple II) \autocite{James2010}. Computer architects use
+common patterns of transistors when designing circuits, with each
+transistor having a ``type'' in the circuit. We identified a region of
+the processor with complex but known structure containing the primary
+8-bit registers X, Y, and S (fig~\ref{fig:mos6502}).
+
+Our algorithm identifies areas of spatial homogeneity that mirror the
+known structure in the underlying architectural circuit, segmenting
+transistor types recognizable to computer architects. Using the
+original schematics, we see that one identified type contains the
+``clocked'' transistors, which retain digital state. Two other types
+contain transistors with pins C1 or C2 connected to ground, mostly
+serving as inverters.  An additional identified type controls the
+behavior of the three registers of interest (X, Y, and S) with respect
+to the SB data bus, either allowing them to latch or drive data from
+the bus. The repeat patterns of spatial connectivity are visible in
+figure~\ref{fig:mos6502}c, showing the man-made horizontal and
+vertical layout of the same types of transistors.
+
+\subsection*{Microprocessor}
+We extracted the connection graph for the transistors in the MOS6502
+\autocite{visual6502source}. Each transistor has three terminals (gate,
+source, drain), but the methods of the original dataset were unable to
+consistently resolve which of the C1 and C2 terminals were source and
+drain, leading to ambiguity in our encoding. We identified a region
+consisting of three registers X, Y, and S via visual
+inspection and focused our efforts there. We created a total of six
+connectivity matrices by examining possible terminal pairings. One
+graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
+$e_j$ are connected via pins $g$ and $c_1$.
 
@@
 \section*{Measuring clustering similarity}
-
+\fxwarning{Add an explanation of ARI, etc.} 
 The adjusted rand index (ARI) is a measure of the similarity of two

### 2015-02-27 docs/paper.tex
spell check

@@
 \usepackage{fullpage}
-%\usepackage{algorithm}
-%\usepackage{algorithmic}
 \usepackage{amsmath}
-%\usepackage{subcaption}
-%\usepackage{multicol}
 \usepackage{csquotes}
@@
 \renewcommand{\vec}[1]{\mathbf{#1}}
-
-%\renewcommand{\includegraphics}[2][]{\fbox{}}
 
@@
   patterns in connectomics data. It combines the information
-  traditionally used by biologists, including connectivity, cell body
-  location and the spatial distribution of synapses, in a
-  principled and probabilistically-coherent manner. We show that the
+  traditionally used by biologists in a principled and probabilistically-coherent manner, including  including connectivity, cell body
+  location and the spatial distribution of synapses. We show that the
   approach recovers known neuron types in the retina and enables
@@
 understanding neural circuit structure.  The impact of this kind of
-high-throughput transition has been seen before -- rise of sequencing
+high-throughput transition has been seen before -- the rise of sequencing
 techniques necessitated the development of novel computational methods
-to understand genomic structure, ushering in  bioinformatics
-as an independent discipline \autocite{Koboldt2013}.
-\fxwarning{cite background article, cite the paper from Yuste compare supervised unsupervised on morphology, cite seungs new article}
+to understand genomic structure, ushering in bioinformatics as an
+independent discipline \autocite{Koboldt2013}.
 
@@
 morphological, but they also differ in the way they connect to one
-another. In fact, the idea of well defined, type-dependent local
-connectivity patterns (microcircuits) has been prominent in many
-areas, from sensory (e.g. retina, \autocite{Masland2001} to processing
-(e.g. neocortex \autocite{Mountcastle1997}) to movement (e.g. spinal
-cord) \autocite{Grillner2005}. These sorts of repeated computing
-patterns are a common feature of computing systems, even arising in
-human-made computing circuits. It remains an important challenge to
-develop algorithms to use anatomical data, e.g. connectomics, to
-automatically back out underlying microcircuitry.
+another \autocite{Seung2014}. In fact, the idea of well defined,
+type-dependent local connectivity patterns (microcircuits) has been
+prominent in many areas, from sensory (e.g. retina,
+\autocite{Masland2001} to processing (e.g. neocortex
+\autocite{Mountcastle1997}) to movement (e.g. spinal cord)
+\autocite{Grillner2005}. These sorts of repeated computing patterns
+are a common feature of computing systems, even arising in human-made
+computing circuits. It remains an important challenge to develop
+algorithms to use connectivity-based anatomical data (connectomics), to
+automatically back out underlying microcircuitry.  already begun,
+focusing on
 
@@
 When it comes to neuroscience and other computing systems, we expect
-patterns of connectivity much more complex than traditional motifs \fxwarning{Add citation of Science “motifs” paper},
-exhibiting a strong spatial dependence arising from the complex
-genetic, chemical, and activity-based neural development processes.
-
-To address these challenges, here we describe a Bayesian
-nonparametric model that can discover circuit structure automatically
-from connectomics data: the cell types, their spatial patterns of
+patterns of connectivity much more complex than traditional motifs
+\fxwarning{Add citation of Science “motifs” paper}, exhibiting a
+strong spatial dependence arising from the complex genetic, chemical,
+and activity-based neural development processes.
+
+To address these challenges, here we describe a Bayesian non-parametric
+model that can discover circuit structure automatically from
+connectomics data: the cell types, their spatial patterns of
 interconnection, and the locations of somata and synapses. We show
 that by incorporating this additional information, our model both
-accurately predicts the connection as well as agrees
-with human neuroanatomists as to the identification of cell types.
+accurately predicts the connection as well as agrees with human
+neuroanatomists as to the identification of cell types. We take as
+inspiration previous work on identifying cell types automatically from
+morphology \autocite{Guerra2011} and electrophysiology
+\autocite{Druckmann2013}.
 
 We primarily focus on the recently-released mouse retina connectome
-\autocite{Helmstaedter2013}, but additionally examine the C. elegans
-connectome \autocite{White1986}. 
-Comparing the cell
-types discovered by the algorithms with those obtained manually by
-human anatomists reveals a high degree of agreement. We thus present a
-scalable probabilistic approach to infer microcircuitry from
-connectomics data available today and in the future.
+\autocite{Helmstaedter2013}, but additionally examine the \textit{C. elegans}
+connectome \autocite{White1986}.  Comparing the cell types discovered
+by the algorithms with those obtained manually by human anatomists
+reveals a high degree of agreement. We thus present a scalable
+probabilistic approach to infer microcircuitry from connectomics data
+available today and in the future.
 
@@
 We build a structured probabilistic model which begins with the
-generic notion of a cell being a member of a single, unobserved type
+generic notion of a cell being a member of a single type
 -- and these types affect soma depth, distribution of synapses, as
@@
 example, retinal ganglion cells may synapse on nearby, but not far
-away, amacrine cells, with ganglion cells being superficial and having
-a broad distribution of synapses.
+away, amacrine cells, with bipolar cells clearly tesselating space and
+synapsing on both.
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics{synthetic.ai}}
+  \caption{Correct recovery of true numbers of hidden types in
+    synthetic data when incorporating spatial information. a.) The
+    infinite stochastic block model (which only uses connectivity
+    information) over-estimates the number of classes as it fails to
+    take distance into account, whereas our modeling of the
+    combination of distance and connectivity finds close to the true
+    number of classes. b.) ARI, a measure of ``correct clustering'',
+    for different true class counts, between our model and a
+    traditional iSBM. c.) For synthetic data, type discovery based
+    only on connectivity are tiny spatially-localized groups -- in
+    contrast, our model recovers the true spatial extent of the
+    underlying types.  }
+\label{fig:synthetic}
+\end{figure}
 
@@
 
+We start with a model for connectivity, the infinite stochastic block
+model (iSBM)\autocite{Kemp2006a,Xu2006}, which has been shown to
+meaningfully cluster connection graphs while learning the number of
+hidden groups, or types. We extend this approach by adding distance
+dependence to model salient aspects of microcircuitry via logistic and
+exponential distance-link functions.  We form a unimodial model of cell
+body depth and a multimodal synapse density profile model (see
+Methods for mathematical details).
+
 As an illustrative example, consider a network with only three cell
-types, labeled A, B, and C. Assume the network has three types of
-cells, they are uniformly distributed in space, and that the
-probability of connection between any two cells, $c_i$ and $c_j$,
-depends only on their type and their distance, according to a logistic
-(sigmoidal) function. Let A connect only to nearby B and C cells, but B
-connect to any C regardless of distance.
+types, labeled A, B, and C. Assume these cells are uniformly
+distributed in space, and that the probability of connection between
+any two cells, $c_i$ and $c_j$, depends only on their type and their
+distance, according to a logistic (sigmoidal) function. Let A connect
+only to nearby B and C cells, but B connect to any C regardless of
+distance. This is the prior intuition our model is designed to capture.
 
@@
 $f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
-latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
+latent type, $c_i=m$ and $c_j=n$, via a parameter $\eta_{mn}$, as well
 as a set of global hyper parameters $\theta$, such that the link
@@
 
-We then jointly infer the maximum a posteriori (MAP) estimate of the
+We then jointly infer the posterior distribution of the
 class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix
@@
 
+In our subsequent analysis makes use of both the full posterior
+distribution over these parameters as well as the most probable, or
+maximum \textit{a posteriori (MAP), estimate.
+
 For the retina data, we then extend the model with the additional
 features indicated. Cell soma depth is modeled as a
-cell-type-dependent Gaussian with latent (unknown) per-type mean and
-variance. Similarly, each cell has some number $N_i$ of synapses, 
-each of which is drawn from a cell-type-specific density profile
-with up to three modes.
-
-Inference is performed in three steps via composable transition 
-kernels -- one for structural, one for per-type parameters, and
-one kernel for global parameters and hyperparameters. Details
-of data preprocessing, inference parameters, and runtime can
-be found in the Supplementary Methods section.
-
-We start with a model for connectivity, the infinite stochastic block
-model (iSBM)\autocite{Kemp2006a,Xu2006}, which has been shown to
-meaningfully cluster connection graphs while learning the number of
-hidden groups, or types. We extend this approach by adding distance
-dependence to model salient aspects of microcircuitry via logistic and
-exponential distance-link functions.  We additionally model cell body
-depth unimodally and synapse density profile multimodally (see
-Methods for mathematical details).
-
-\fxwarning{Do we talk about cross-validation early on? What about
-  posterior uncertainty?}
+cell-type-dependent Gaussian distribution with latent (unknown)
+per-type mean and variance. Similarly, each cell has some number $N_i$
+of synapses, each of which is drawn from a cell-type-specific density
+profile with up to three modes.
+
+Inference is performed via Markov-chain Monte Carlo (MCMC) via three
+composable transition kernels -- one for structural, one for per-type
+parameters, and one kernel for global parameters and
+hyperparameters. Details of data preprocessing, inference parameters,
+and runtime can be found in the Supplementary Methods section.
 
 \subsection*{Metrics}
-To evaluate the quality of the model fit, we need to use information that quantifies aspects of the data for which we have ground truth information.
-We focus on two aspects of performance.
-First, if the model works well, then the probability that a pair of neurons is of the same type should be high if the neurons actually are of the same type.
-Second, the model should assign a high probability of connection between two cells if they have a connection in the underlying data.
-We term these two factors clustering accuracy and link prediction accuracy.
+To evaluate the quality of the model fit, we need to use information
+that quantifies aspects of the data for which we have ground truth
+information. We focus on two aspects of performance. First, if the
+model works well, then the probability that a pair of neurons is of
+the same type should be high if the neurons actually are of the same
+type. Second, the model should assign a high probability of connection
+between two cells if they have a connection in the underlying data. We
+term these two factors clustering accuracy and link prediction
+accuracy.
 
 To assess the accuracy of a clustering compared to that determined by
-neuroanatomists, we employ the adjusted Rand index (ARI).
-ARI is an established metric which measures the similarity of two clusterings -- two identical clusterings will have an ARI of 1.0.
-There are no perfect clustering comparison metrics as all suffer from the need to balance
-``overclustering'' (calling a single type two different types) with
-``merging'' (calling two distinct types a single type).  See supplemental for a discussion of
-ARI's shortcomings.
-Despite its shortcomings we use the ARI as a simple measure of clustering accuracy as it quantitatively captures the difference between the types discovered and the types believed by neuroanatomists.
-
-To assess the accuracy of the model for connections, we use link prediction
-accuracy. If our model accurately captures the true structure of
-the data, it should be good at predicting if a link exists.
-We thus train the model on the data with a subset of the links
-marked as unobserved and thus compute our predictive accuracy. We perform 10-way
-cross validation on a given dataset, learn the resulting model, and
-use that model to predict the missing synapses. Each potential
-link between cells is assigned a probability, and we compute
-the area under the curve (AUC) for the resulting ROC curve.
-An AUC of 1.0 means that we perfectly predict the presence and absence of the missing synapses.
-We use link prediction accuracy to quantify how good the model is at discovering the underlying connectivity.
-\fxwarning{Add additional ARI}
+neuroanatomists, we employ three metrics -- clustering homogeneity,
+clustering completness, and the adjusted Rand Index. All metrics equal
+1.0 when two clusterings completely agree. Homogeneity reflects the
+degree to which a found cluster or type contains only a single true
+type. Completeness measures how much of a true type is contained
+within a single identified type -- a completeness of 1.0 means no true
+type is split into multiple subtypes. ARI is a metric that reflects both measures (See
+supplemental for more information).
+
+To assess the accuracy of the model for connections, we use link
+prediction accuracy. If our model accurately captures the true
+structure of the data, it should be good at predicting if a link
+exists.  We thus train the model on the data with a subset of the
+links marked as unobserved and thus compute our predictive
+accuracy. We perform 10-way cross validation on a given dataset \autocite{Guerra2011}, learn
+the resulting model, and use that model to predict the missing
+synapses. Each potential link between cells is assigned a probability,
+and we compute the area under the curve (AUC) for the resulting ROC
+curve. An AUC of 1.0 means that we perfectly predict the presence and
+absence of the missing synapses. We use link prediction accuracy to
+quantify how good the model is at discovering the underlying
+connectivity.
 
 \section*{Results}
-We will first establish that our algorithm works properly and try to understand its properties using simulated data.
-Subsequently we will analyze in detail a dataset on the retina.
-Lastly we will briefly discuss the analysis of data from the worm \textit{c.elegans} and from an ancient microprocessor.
+We will first establish that our algorithm works properly and try to
+understand its properties using simulated data. Subsequently we will
+analyze in detail a dataset on the retina. Lastly we will briefly
+discuss the analysis of data from the worm \textit{c.elegans} and from
+an ancient microprocessor.
 
@@
 (fig~\ref{fig:synthetic}b), and the spatial extent of each type
-(fig~\ref{fig:synthetic}c).  For comparison, we show the results from instead using the infinite
-stochastic block model (fig~\ref{fig:synthetic}a, b, c, black line) which assumes that only cell type matters, and thus finds
-small neighborhoods of connected nodes (instead of global connectivity
-patterns).
-This contrast shows that while the regular block model can not correctly deal with distance dependent connectivity, our model can.
-Our model converges relatively quickly (see supplemental) to an estimate of
-the most probable values for the cell types, which is enabled by using
-a combination of simulated annealing and parallelized Markov chain
-Monte Carlo (see methods for details). Thus our model at least is promising to be applied to biological datasets.
+(fig~\ref{fig:synthetic}c).  For comparison, we show the results from
+instead using the infinite stochastic block model
+(fig~\ref{fig:synthetic}a, b, c, black line) which assumes that only
+cell type matters, and thus finds small neighborhoods of connected
+nodes (instead of global connectivity patterns). This contrast shows
+that while the regular block model can not correctly deal with
+distance dependent connectivity, our model can. Our model converges
+relatively quickly (see supplemental) to an estimate of the most
+probable values for the cell types, which is enabled by using a
+combination of simulated annealing and parallelized Markov chain Monte
+Carlo (see methods for details). Thus our model at least is promising
+to be applied to biological datasets.
 
 \subsection*{Model Mismatch}
-
-\fxwarning{Add xlabel, ylabel, make line width of axes a bit bigger, larger number font, 
-all label fonts should be the same size }
-
-To understand how our model would perform on real datasets that do not
-perfectly adhere to its generative model we next want to analyze its
-performance in such cases. To understand the properties of our model, we
-attempt connectivity inference on four sets of synthetic data. This helps us understand what our model would do if the data does not obey our assumptions.
-
-We thus generate ten sets of synthetic data from each of four existing models. 
-The distance-dependent stochastic block model assumes type depends on distance, 
-the traditional stochastic block model has no notion of distance, the mixed membership block model assumes type is combinatorial, and the latent position cluster model assumes that type is clustered-but-continuous.
-
-If the data is sampled from our model, inference according to our model, unsurprisingly, is good by all measures.
-It correctly estimates the number of cell types; it is good at predicting connectivity (high AUC); it agrees with human classification (Rand index); it discovers all types, and leads to homogenous estimates (Fig.
-\ref{fig:othermodel} first row).
-If the data comes from a block model without distance dependence, we see that it still does pretty well on all meaningful measures (Fig.
-\ref{fig:othermodel} second row).
-It basically estimates correctly that there is no distance dependence.
-For the mixed membership model (Fig.
-\ref{fig:othermodel} third row), the model grossly overestimates the number of types, by basically allocating a type for each combination of memberships.
-Otherwise it still performs relatively well.
-Lastly, for the latent position clustering model (Fig.
-\ref{fig:othermodel} fourth row), the model does poorly.
-If type is continuous instead of discrete then our model is basically trying to cover a continuous set with a discrete scenario leading to rather poor performance.
-However, as we do expect cell types to be real in biology we might expect our model to do well with real data.
+We next analyze how our model performs in cases where the data are
+generated with assumptions different from ours.  To understand the
+properties of our model, we attempt connectivity inference on four
+sets of synthetic data. This helps us understand what our model would
+do if the data does not obey our assumptions.
+
+We thus generate ten sets of synthetic data from each of four existing
+models.  The distance-dependent stochastic block model assumes type
+depends on distance, the traditional stochastic block model has no
+notion of distance, the mixed membership block model assumes type is
+combinatorial, and the latent position cluster model assumes that type
+is clustered-but-continuous.
+
+If the data is sampled from our model, inference according to our
+model, unsurprisingly, is good by all measures. It correctly estimates
+the number of cell types; it is good at predicting connectivity (high
+AUC); it agrees with human classification (Rand index); it discovers
+all types, and leads to homogeneous estimates
+(Fig. \ref{fig:othermodel} first row). If the data comes from a block
+model without distance dependence, we see that it still does well on
+all meaningful measures (Fig. \ref{fig:othermodel} second row). This
+is unsurprising, as our model learns the distance dependence, even its
+absence. For the mixed membership model (Fig. \ref{fig:othermodel}
+third row), the model grossly overestimates the number of types, by
+basically allocating a type for each combination of
+memberships. Otherwise it still performs relatively well. Lastly, for
+the latent position clustering model (Fig. \ref{fig:othermodel} fourth
+row), the model does poorly. If type is continuous instead of discrete
+then our model is basically trying to cover a continuous set with a
+discrete scenario leading to rather poor performance. However, as we
+do expect cell types to have a discrete biological basis we might
+expect our model to do well with real data.
 
@@
 Connectomic efforts so far have reconstructed only small sections of
-tissue. Consequently, many connections to cells outside that tissue
+neural tissue. Consequently, many connections to cells outside that tissue
 volume will be lost. We are concerned that this selective elimination
@@
 
-\fxwarning{Add in a line on this figure to show where the existing data is by looking at old-school tracer reconstruction in the retina. CITE }
-
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics{synthetic.ai}}
-  \caption{Correct recovery of true numbers of hidden types in
-    synthetic data when incorporating spatial information. a.) The
-    infinite stochastic block model (which only uses connectivity
-    information) over-estimates the number of classes as it fails to
-    take distance into account, whereas our modeling of the
-    combination of distance and connectivity finds close to the true
-    number of classes. b.) ARI, a measure of ``correct clustering'',
-    for different true class counts, between our model and a
-    traditional iSBM. c.) For synthetic data, type discovery based
-    only on connectivity are tiny spatially-localized groups -- in
-    contrast, our model recovers the true spatial extent of the
-    underlying types.  }
-\label{fig:synthetic}
-\end{figure}
-
 \subsubsection*{Learning types and circuitry in the retina}
@@
   \caption{Discovering cell classes in the mouse retina connectome. Here
-we show the maxium a posteriori (MAP) estimate for the types in the mouse
+we show the maximum a posteriori (MAP) estimate for the types in the mouse
 retina data. 
@@
     of the plot. c.)  The spatial distribution of our cell types
-    -- each cell type tesselates space. Colors correspond to those in b.)
+    -- each cell type tessellates space. Colors correspond to those in b.)
     d.) Connectivity between our clusters as a function
@@
 
-The algorithm took less than 2 hours to perform inference, dividing
+The algorithm took less than 4 hours to perform inference, dividing
 neurons into a set of cell types (fig~\ref{fig:mouseretina}c, each
@@
 
-As a first measure we compare link prediction accuracy across the
-methods (Fig.\ref{fig:mouseretina_compare} B AUC, red).  We find that
+For our model we show the receiver opererating characteristic curve
+(ROC curve) (fig~\ref{fig:mouseretina_compare}d) whcih shows how the
+true and false positive rates trade off. We plot the posterior
+distribution of the area under this curve in
+figure~\ref{fig:mouseretina_compare}e. We then plot the 
+posterior distribution for cluster agreement metrics -- completeness,
+homogeneity, and ari (figure~\ref{fig:mouseretina_compare}f). We see
+that our model tends to over-cluster -- some of the finest-granularity of
+neuroanatomist types are grouped as a single type by our model.
+
+We compare link prediction accuracy across the
+methods, including our own (Fig.\ref{fig:mouseretina_compare}g AUC, red).  We find that
 given the dataset many techniques allow for good link-predictive
 accuracy. All the methods allow decent link prediction with an AUC in
-the .9 range. However, our algorithm clearly outperforms the simple
+the 0.9 range. However, our algorithm clearly outperforms the simple
 statistical models that only use connectivity.
@@
 As a second measure we compare link prediction accuracy across the
-methods (Fig.\ref{fig:mouseretina_compare} B ARI, blue). We find that our
+methods (Fig.\ref{fig:mouseretina_compare}g ARI, blue). We find that our
 algorithm far outperforms the controls. We also find that when it is
@@
 within a layer and between layers (Fig.\ref{fig:mouseretina_compare}
-C). We see that, in the absence of distance information, mere
+h). We see that, in the absence of distance information, mere
 connectivity information results in types which only span a small
@@
 
-Having shown our model to work on the repeating tessellated, laminar structure of the mammalian retina, we then apply our model to a structurally very different connectome -- the whole body of a small roundworm:
-\textit{Caenorhabditis elegans} is a model system in
+Having shown our model to work on the repeating tessellated, laminar
+structure of the mammalian retina, we then apply our model to a
+structurally very different connectome -- the whole body of a small
+roundworm: \textit{Caenorhabditis elegans} is a model system in
 developmental neuroscience\autocite{White1986}, with the location and
@@
 only the motor neurons in C. elegans exhibit regular distribution in
-space -- along the body axis. Most interneurons are
-concentrated in various ganglia that project throughout the entire
-animal, and the sensory neurons are primarily located in a small
-number of anterior ganglia.
-\textit{C.
-elegans} also differs from the retina in that the measured connectome is actually two separate graphs -- one of directed chemical synapses and another of undirected electrical synapses.
-As this is a very different connectome, it allows an interesting generalization test: how well will our model work on such a distinct data set.
-
-\fxwarning{Fix linewidth too small, align d.e.f. callouts, change multiple fonts. Boxes same size. Need y labels (number of samples in bin) }
+space -- along the body axis. Most interneurons are concentrated in
+various ganglia that project throughout the entire animal, and the
+sensory neurons are primarily located in a small number of anterior
+ganglia. \textit{C. elegans} also differs from the retina in that the
+measured connectome is actually two separate graphs -- one of directed
+chemical synapses and another of undirected electrical synapses. As
+this is a very different connectome, it allows an interesting
+generalization test: how well will our model work on such a distinct
+data set.
 
@@
     connectivity as a function of discovered type, similar to
-    figure~\ref{mouseretina}. In b.) we plot neuroanatomist-derived
-    types along with their labels. Our model shows high probability of
-    motor neurons, sensory neurons, and various interneuron classes
-    being of the same type. Soma positions along the body axis are
-    plotted in c.) where we see that we cluster spatially-distributed
-    motor neurons together, whereas head sensory neurons are more
-    likely to be grouped together as well.  d.) The ROC curves for
-    held-out link probability for both the electrical synapses (gap
-    junctions) and chemical synapses in c. elegans. e.) The posterior
-    distribution of the area under the ROC curve for the curves in
-    d. f.) Measurements of the agreement of our identified cell types
-    compared to neuroanatomists. The high compleneness but low
-    homogeneity (and corresponding low ARI) reflects our model's
-    tendency to group multiple types into a single type.}
+    figure~\ref{fig:mouseretina_compare}. In b.) we plot
+    neuroanatomist-derived types along with their labels. Our model
+    shows high probability of motor neurons, sensory neurons, and
+    various interneuron classes being of the same type. Soma positions
+    along the body axis are plotted in c.) where we see that we
+    cluster spatially-distributed motor neurons together, whereas head
+    sensory neurons are more likely to be grouped together as well.
+    d.) The ROC curves for held-out link probability for both the
+    electrical synapses (gap junctions) and chemical synapses in
+    c. elegans. e.) The posterior distribution of the area under the
+    ROC curve for the curves in d. f.) Measurements of the agreement
+    of our identified cell types compared to neuroanatomists. The high
+    compleneness but low homogeneity (and corresponding low ARI)
+    reflects our model's tendency to group multiple types into a
+    single type.}
   \label{fig:celegans}
@@
   \centering 
-  \centerline{\includegraphics[width=4in]{mos6502.ai}}
+  \centerline{\includegraphics[width=6in]{mos6502.ai}}
   \caption{Discovering connectivity and type the 6502 microprocessor.
@@
     versus distance for connections between Gate and C1, Gate and C2,
-    and C1 and C2 terminals on a transistor. Red and teal types have a
+    and C1 and C2 terminals on a transistor. Purple and yellow types have a
     terminal pulled down to ground and mostly function as
-    inverters. Purple class are clocked, stateful transistors, orange
-    control the ALU and yellow control the special data bus (SDB).}
+    inverters. The blue types are clocked, stateful transistors, green
+    control the ALU and orange control the special data bus (SDB).}
   \label{fig:mos6502}
@@
 vertical layout of the same types of transistors.
-
-\subsection*{Microprocessor}
-We extracted the connection graph for the transistors in the MOS6502
-\autocite{visual6502source}. Each transistor has three terminals (gate,
-source, drain), but the methods of the original dataset were unable to
-consistently resolve which of the C1 and C2 terminals were source and
-drain, leading to ambiguity in our encoding. We identified a region
-consisting of three registers X, Y, and S via visual
-inspection and focused our efforts there. We created a total of six
-connectivity matrices by examining possible terminal pairings. One
-graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
-$e_j$ are connected via pins $g$ and $c_1$.
 
@@
 
+\subsection*{Microprocessor}
+We extracted the connection graph for the transistors in the MOS6502
+\autocite{visual6502source}. Each transistor has three terminals (gate,
+source, drain), but the methods of the original dataset were unable to
+consistently resolve which of the C1 and C2 terminals were source and
+drain, leading to ambiguity in our encoding. We identified a region
+consisting of three registers X, Y, and S via visual
+inspection and focused our efforts there. We created a total of six
+connectivity matrices by examining possible terminal pairings. One
+graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
+$e_j$ are connected via pins $g$ and $c_1$.
+
 % \begin{addendum}
@@
 \section*{Measuring clustering similarity}
-\fxwarning{Add an explanation of ARI, etc.} 
-The adjusted rand index (ARI) is a measure of the similarity of two
-different clusterings \autocite{Hubert1985} -- two identical clusters
-have an ARI of 1.0 while progressively more dissimilar clusters have
-lower ARIs, becoming negative as the clustering gets anti-correlated.
+We compare discovered types to known types via cluster comparison
+metrics: cluster homogeneity, cluster completeness, and the adjusted
+Rand index. Homogeneity measures how many true types are in a given
+found type. If every cell type is split into two types, each subtype
+is still completely homogeneous. Completeness measures how many
+members of a given true type are split across found types.
+
+The adjusted rand index (ARI) takes into account both effects
+\autocite{Hubert1985} -- two identical clusterings have an ARI of 1.0
+while progressively more dissimilar clusters have lower ARIs, becoming
+slightly negative as the clustering gets anti-correlated.
+
+Figure~\ref{fig:s:comparison} shows the result of taking 20 different clusters
+and moving data points between them according to the following operations
+\begin{itemize}
+\item \textbf{distribute}: take a class and distribute its elements
+  uniformly among the remaining types.
+\item \textbf{merge} : take a type and merge it into another existing
+  type
+\item \textbf{split} : take a type and split it into two distinct
+  types.
+\end{itemize}
+
+We can see the impact on ARI, completeness, and homogeneity as we
+perform these operations on more of the original 20 types. In all
+cases, ``distribution'' of one type among the others is detremental to
+the metric. Splitting impacts completeness but not homogeneity, and
+merging impacts homogeneity but not completness.
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=180mm]{cluster_metric_comparison.pdf}}
+  \caption{Type agreement evaluation metrics as a function of splitting types, merging types, and randomly distributing cells between types}
+  \label{fig:s:comparison}
+\end{figure}
 

### 2015-03-02 docs/arxiv.tex
resubmitted

@@
 \usepackage[backend=biber,style=nature]{biblatex}
-\usepackage{todonotes}
+\usepackage[margin=true,inline=true,index=true]{fixme}
 \usepackage{grffile}
@@
   patterns in connectomics data. It combines the information
-  traditionally used by biologists, including connectivity, cell body
+  traditionally used by biologists in a principled and probabilistically-coherent manner, including  including connectivity, cell body
   location and the spatial distribution of synapses. We show that the
@@
   can reveal interesting structure in the nervous system of
-  c. elegans, and automatically discovers the structure of a
-  microprocessor.  Our approach extracts structural meaning from
+  C. elegans and a historic microprocessor.
+  Our approach extracts structural meaning from
   connectomics, enabling new approaches of automatically deriving
@@
 \end{abstract}
+
+\listoffixmes
 
@@
 understanding neural circuit structure.  The impact of this kind of
-high-throughput transition has been seen before -- rise of sequencing
+high-throughput transition has been seen before -- the rise of sequencing
 techniques necessitated the development of novel computational methods
-to understand genomic structure, ushering in  bioinformatics
-as an independent discipline \autocite{Koboldt2013}.
+to understand genomic structure, ushering in bioinformatics as an
+independent discipline \autocite{Koboldt2013}.
 
@@
 morphological, but they also differ in the way they connect to one
-another. In fact, the idea of well defined, type-dependent local
-connectivity patterns (microcircuits) has been prominent in many
-areas, from sensory (e.g. retina, \autocite{Masland2001} to processing
+another \autocite{Seung2014}. In fact, the idea of well defined,
+type-dependent local connectivity patterns (microcircuits) has a long
+history \autocite{Passingham2002}, and is prominent in many areas,
+from sensory (e.g. retina, \autocite{Masland2001} to processing
 (e.g. neocortex \autocite{Mountcastle1997}) to movement (e.g. spinal
@@
 human-made computing circuits. It remains an important challenge to
-develop algorithms to use anatomical data, e.g. connectomics, to
-automatically back out underlying microcircuitry.
+develop algorithms to use connectivity-based anatomical data
+(connectomics), to automatically back out underlying microcircuitry.
+already begun, focusing on
 
@@
 When it comes to neuroscience and other computing systems, we expect
-patterns of connectivity much more complex than traditional motifs,
+patterns of connectivity much more complex than traditional motifs
 exhibiting a strong spatial dependence arising from the complex
@@
 
-To address these challenges, here we describe a Bayesian
-nonparametric model that can discover circuit structure automatically
-from connectomics data: the cell types, their spatial patterns of
+To address these challenges, here we describe a Bayesian non-parametric
+model that can discover circuit structure automatically from
+connectomics data: the cell types, their spatial patterns of
 interconnection, and the locations of somata and synapses. We show
 that by incorporating this additional information, our model both
-accurately predicts the connection as well as agrees
-with human neuroanatomists as to the identification of cell types.
+accurately predicts the connection as well as agrees with human
+neuroanatomists as to the identification of cell types. We take as
+inspiration previous work on identifying cell types automatically from
+morphology \autocite{Guerra2011} and electrophysiology
+\autocite{Druckmann2013}.
 
 We primarily focus on the recently-released mouse retina connectome
-\autocite{Helmstaedter2013}, but additionally examine the c. elegans
-connectome \autocite{White1986}, and then ``connectome'' of a
-classical microprocessor \autocite{James2010}. Comparing the cell
-types discovered by the algorithms with those obtained manually by
-human anatomists reveals a high degree of agreement. We present a
-scalable approach to infer microcircuitry from connectomics data.
+\autocite{Helmstaedter2013}, but additionally examine the \textit{C. elegans}
+connectome \autocite{White1986}.  Comparing the cell types discovered
+by the algorithms with those obtained manually by human anatomists
+reveals a high degree of agreement. We thus present a scalable
+probabilistic approach to infer microcircuitry from connectomics data
+available today and in the future.
 
@@
   \centering 
-  \centerline{\includegraphics[width=183mm]{fig1b.ai}}
-  \caption{An overview of our method. a. Input data is a connectivity
-    matrix indicating whether cell $i$ (row) synapses on cell $j$
-    (column), a distance matrix between cell soma, b. the depth of the
-    cell soma, and the c. depth profile of the cells synaptic
-    connectivity. d. Our algorithm discovers hidden cell types in this
-    connectivity data by assuming all cells of a type share a
-    distance-dependent connectivity profile, similar depth, and a
-    similar synpatic density profile, with cells of other types.  This
-    results in a clustering of the cells by those hidden
-    types. d. shows the cell connectivity matrix with cells of the
-    same type grouped together. e.) shows the learned probability of
+  \centerline{\includegraphics[width=110mm]{fig1b.ai}}
+  \caption{Deriving circuitry and cell types from connectomics data
+    a. As input we take the connectivity between cells (b), the
+    distance between them (c), the depth of the cell bodies (d), and
+    the depth profile of the synapses (e). f. Our algorithm discovers
+    hidden cell types in this connectivity data by assuming all cells
+    of a type share a distance-dependent connectivity profile, similar
+    depth, and a similar synaptic density profile, with cells of other
+    types.  This results in a clustering of the cells by those hidden
+    types. f.) shows the cell connectivity matrix with cells of the
+    same type grouped together. g.) shows the learned probability of
     connection between our different types at various distances -- in
     this case, the cells are likely to connect when they are
-    close. f. shows a connectivity profile with virtually no
-    probability of connectivity. We can plot this these types g.) to
-    show the relationship and spatial connections, and how probable
-    various types are to connect to one another. h. The connectivity
-    between our discovered types changes as a function of distance. }
+    close. h. shows the probability of connection between two cell
+    types that very rarely connect -- there's a background ``base''
+    connection rate to account for errors in data, but the probability
+    is very low.  i.) shows that we also recover the expected
+    laminarity of types and the depth-specific synaptic
+    connectivity. j.) We then plot how the connectivity between these
+    types changes as a function of distance between the cell bodies to
+    better understand short-range and long-range connectivity
+    patterns. }
 
@@
 
-\section{Results}
-
-We start with the generic assumption that cell types exist -- a cell
-is a member of a single, unobserved type -- and these types affect
-soma depth, distribution of synapses, as well as a cell type and
-distance dependent connection probability. From this assumption we
-develop a generative Bayesian model that estimates the underlying cell
-types and how they connect.  We start with a model for connectivity,
-the infinite stochastic block model (iSBM)\autocite{Kemp2006a,Xu2006},
-which has been shown to meaningfully cluster connection graphs while
-learning the number of hidden groups, or types. We extend this
-approach by adding distance dependence to model salient aspects of
-microcircuitry. We additionally model cell body depth as a mixture of
-Gaussians and each cell’s synapse density profile as mixture of
-mixtures of Gaussians.
-
-For example, retinal ganglion cells may synapse on nearby, but not far
-away, amacrine cells, with ganglion cells being superficial and having
-a broad distribution of synapses.  This procedure automatically learns
-the number of cell types, which cells belong to which types, along
-with the distance and type dependent connectivity probabilities (see
-fig~\ref{fig:overview}). We apply our algorithm to datasets from mouse
-retina, c-elegans and a historical microprocessor. Anatomists classify
-cells based on many features, giving us a meaningful baseline to
-compare against.
-
-Before we can use a probabilistic model to analyze real data we first
-need to assess if the model works properly. We thus simulate data for
-which we know the correct structure and comparing the estimated
-structure based on the algorithm (see methods) with the one we used
-for simulation. We find that the model does a good job of recovering
-the correct number of cell types, (fig~\ref{fig:synthetic}a), the the
-cell identities (fig~\ref{fig:synthetic}b), and the spatial extent of
-each type (fig~\ref{fig:synthetic}c).  Indeed, it does far better than
-the existing infinite stochastic block model that assumes that there
-is no distance dependence to the connectivity. The model converges
-relatively quickly to an estimate of the most probable values for the
-cell types, which is enabled by using a combination of simulated
-annealing and parallelized Markov chain Monte Carlo (see methods for
-details). It appears that the model behaves the way it should and
-allows application to relatively large datasets.
+\section*{Methods}
+
+We build a structured probabilistic model which begins with the
+generic notion of a cell being a member of a single type -- and these
+types affect soma depth, distribution of synapses, as well as a cell
+type and distance dependent connection probability. For example,
+retinal ganglion cells may synapse on nearby, but not far away,
+amacrine cells, with bipolar cells clearly tesselating space and
+synapsing on both. In machine learning parlance, our method is
+“unsupervised” -- it seeks to discover structure in data and make
+predictions in the absence of training data. Rather than taking in
+examples of types annotated by human neuroanatomists, we instead start
+with the weakest possible assumption in an attempt to algorithmically
+discover this structure. We contrast this with the supervised
+approaches taken in \autocite{Guerra2011}, where there is high
+confidence in the (morphologically-defined) types and then a
+supervised classifier is built, as our goal here is explicitly
+discovery of types.
 
@@
 
-\subsubsection{Mouse Retina}
+From these assumptions (priors) we develop a generative Bayesian model
+that estimates the underlying cell types and how they connect. We take
+as input (fig ~\ref{fig:overview}a) the connectivity matrix of cells
+(fig ~\ref{fig:overview}b) , a matrix of the distance between cells
+(fig ~\ref{fig:overview}c), the per-cell soma depth (fig
+~\ref{fig:overview}d) and the depth profile of the cell's synapses
+(fig ~\ref{fig:overview}e). We perform joint probabilistic inference
+to automatically learn the number of cell types, which cells belong to
+which type, their type-specific connectivity, and how connections
+between types vary with distance. We also simultaneously learn the
+soma depth associated with each type and the typical synaptic density
+profile (fig ~\ref{fig:overview}f-h.).
+
+We start with a model for connectivity, the infinite stochastic block
+model (iSBM)\autocite{Kemp2006a,Xu2006}, which has been shown to
+meaningfully cluster connection graphs while learning the number of
+hidden groups, or types. We extend this approach by adding distance
+dependence to model salient aspects of microcircuitry via logistic and
+exponential distance-link functions.  We form a unimodial model of cell
+body depth and a multimodal synapse density profile model (see
+Methods for mathematical details).
+
+As an illustrative example, consider a network with only three cell
+types, labeled A, B, and C. Assume these cells are uniformly
+distributed in space, and that the probability of connection between
+any two cells, $c_i$ and $c_j$, depends only on their type and their
+distance, according to a logistic (sigmoidal) function. Let A connect
+only to nearby B and C cells, but B connect to any C regardless of
+distance. This is the prior intuition our model is designed to capture.
+
+For the basic link-distance model, we take as input a connectivity
+matrix $R$ defining the connections between cell $e_i$ and $e_j$, as
+well as a distance function $d(e_i, e_j)$ representing a (physical)
+distance between adjacent cells.  See
+the supplemental material for extension to multiple connectivity
+matrices. We assume there exist an unknown number $K$ of latent
+(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
+cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
+of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
+observed connectivity between two cells $R(e_i, e_j)$ then depends
+only on their latent type and their distance through a link function
+$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
+latent type, $c_i=m$ and $c_j=n$, via a parameter $\eta_{mn}$, as well
+as a set of global hyper parameters $\theta$, such that the link
+function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
+
+We then jointly infer the posterior distribution of the
+class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix
+$\eta_{mn}$, and the global model hyperparameters $\theta$ :
+
+\begin{equation}
+  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
+\end{equation}
+
+In our subsequent analysis makes use of both the full posterior
+distribution over these parameters as well as the most probable, or
+maximum \textit{a posteriori} (MAP), estimate.
+
+For the retina data, we then extend the model with the additional
+features indicated. Cell soma depth is modeled as a
+cell-type-dependent Gaussian distribution with latent (unknown)
+per-type mean and variance. Similarly, each cell has some number $N_i$
+of synapses, each of which is drawn from a cell-type-specific density
+profile with up to three modes.
+
+Inference is performed via Markov-chain Monte Carlo (MCMC) via three
+composable transition kernels -- one for structural, one for per-type
+parameters, and one kernel for global parameters and
+hyperparameters. Details of data preprocessing, inference parameters,
+and runtime can be found in the Supplementary Methods section.
+
+\subsection*{Metrics}
+To evaluate the quality of the model fit, we need to use information
+that quantifies aspects of the data for which we have ground truth
+information. We focus on two aspects of performance. First, if the
+model works well, then the probability that a pair of neurons is of
+the same type should be high if the neurons actually are of the same
+type. Second, the model should assign a high probability of connection
+between two cells if they have a connection in the underlying data. We
+term these two factors clustering accuracy and link prediction
+accuracy.
+
+To assess the accuracy of a clustering compared to that determined by
+neuroanatomists, we employ three metrics -- clustering homogeneity,
+clustering completness, and the adjusted Rand Index. All metrics equal
+1.0 when two clusterings completely agree. Homogeneity reflects the
+degree to which a found cluster or type contains only a single true
+type. Completeness measures how much of a true type is contained
+within a single identified type -- a completeness of 1.0 means no true
+type is split into multiple subtypes. ARI is a metric that reflects both measures (See
+supplemental for more information).
+
+To assess the accuracy of the model for connections, we use link
+prediction accuracy. If our model accurately captures the true
+structure of the data, it should be good at predicting if a link
+exists.  We thus train the model on the data with a subset of the
+links marked as unobserved and thus compute our predictive
+accuracy. We perform 10-way cross validation on a given dataset \autocite{Guerra2011}, learn
+the resulting model, and use that model to predict the missing
+synapses. Each potential link between cells is assigned a probability,
+and we compute the area under the curve (AUC) for the resulting ROC
+curve. An AUC of 1.0 means that we perfectly predict the presence and
+absence of the missing synapses. We use link prediction accuracy to
+quantify how good the model is at discovering the underlying
+connectivity.
+
+\section*{Results}
+We will first establish that our algorithm works properly and try to
+understand its properties using simulated data. Subsequently we will
+analyze in detail a dataset on the retina. Lastly we will briefly
+discuss the analysis of data from the worm \textit{c.elegans} and from
+an ancient microprocessor.
+
+\subsection*{Validation with simulated data where ground truth is known}
+To validate our model, we performed a series of simulations to test if
+the model can accurately recover the true underlying network structure
+and cell type identity.  We thus simulate data for which we know the
+correct structure and compare the estimated structure based on the
+algorithm (see methods) with the one we used for simulation. We find
+that the model does a good job of recovering the correct number of
+cell types, (fig~\ref{fig:synthetic}a), the the cell identities
+(fig~\ref{fig:synthetic}b), and the spatial extent of each type
+(fig~\ref{fig:synthetic}c).  For comparison, we show the results from
+instead using the infinite stochastic block model
+(fig~\ref{fig:synthetic}a, b, c, black line) which assumes that only
+cell type matters, and thus finds small neighborhoods of connected
+nodes (instead of global connectivity patterns). This contrast shows
+that while the regular block model can not correctly deal with
+distance dependent connectivity, our model can. Our model converges
+relatively quickly (see supplemental) to an estimate of the most
+probable values for the cell types, which is enabled by using a
+combination of simulated annealing and parallelized Markov chain Monte
+Carlo (see methods for details). Thus our model at least is promising
+to be applied to biological datasets.
+
+\subsection*{Model Mismatch}
+We next analyze how our model performs in cases where the data are
+generated with assumptions different from ours.  To understand the
+properties of our model, we attempt connectivity inference on four
+sets of synthetic data. This helps us understand what our model would
+do if the data does not obey our assumptions.
+
+We thus generate ten sets of synthetic data from each of four existing
+models.  The distance-dependent stochastic block model assumes type
+depends on distance, the traditional stochastic block model has no
+notion of distance, the mixed membership block model assumes type is
+combinatorial, and the latent position cluster model assumes that type
+is clustered-but-continuous.
+
+If the data is sampled from our model, inference according to our
+model, unsurprisingly, is good by all measures. It correctly estimates
+the number of cell types; it is good at predicting connectivity (high
+AUC); it agrees with human classification (Rand index); it discovers
+all types, and leads to homogeneous estimates
+(Fig. \ref{fig:othermodel} first row). If the data comes from a block
+model without distance dependence, we see that it still does well on
+all meaningful measures (Fig. \ref{fig:othermodel} second row). This
+is unsurprising, as our model learns the distance dependence, even its
+absence. For the mixed membership model (Fig. \ref{fig:othermodel}
+third row), the model grossly overestimates the number of types, by
+basically allocating a type for each combination of
+memberships. Otherwise it still performs relatively well. Lastly, for
+the latent position clustering model (Fig. \ref{fig:othermodel} fourth
+row), the model does poorly. If type is continuous instead of discrete
+then our model is basically trying to cover a continuous set with a
+discrete scenario leading to rather poor performance. However, as we
+do expect cell types to have a discrete biological basis we might
+expect our model to do well with real data.
 
@@
   \centering
-  \centerline{\includegraphics{mouseretina.ai}}
-  \caption{Discovering cell classes in the mouse retina connectome.
+
+  \includegraphics[width=6.0in]{synthdifferent.ai}
+  \caption{Model inferences when true generating model differs from
+    our distance-block-model prior. Horizontal columns show results
+    with synthetic data generated according to the latent position
+    cluster model, the mixed membership block model, the non-distance
+    dependent stochastic block model, and the distance dependent
+    stochastic block model. In all cases histograms represent
+    posterior distribution over indicated metric. a.) The number of
+    types found by the model, vertical dash line indicates ``true''
+    type number (not applicable to mixed membership model). b.) The
+    area under the ROC curve, indicating link prediction accuracy. c,
+    d, e.) Clustering metrics quantifying degree of type agreement
+    with known ground truth. }
+
+  \label{fig:othermodel}
+\end{figure}
+
+\subsection*{Sensitivity to Edge Effects}
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=3.5in]{edgeeffects.pdf}}
+  \caption{We generate two sets of synethetic data, one with
+    spatially-dependent connectivity and one without. We measure the
+    variance in the connectivity-distance plot for randomly selected
+    regions of each dataset, ranging from single-cells to the entire
+    volume. We see that while selecting too small of a region can 
+    destroy the appearance of distance-dependent connectivity, 
+    it does not create it in non-spatial data}
+\label{fig:edgeeffects}
+\end{figure}
+
+Connectomic efforts so far have reconstructed only small sections of
+neural tissue. Consequently, many connections to cells outside that tissue
+volume will be lost. We are concerned that this selective elimination
+of connectivity along the boundary might give the appearance of
+distance-dependent connectivity when there is none. We thus performed
+simulations to check if edge effects could destroy spatial structure
+and if edge effects could introduce artificial, spurious spatial
+structure. We measure the degree to which distance-dependent effects
+can arise from selecting regions that are smaller than the ``scale''
+of connectivity (fig ~\ref{fig:edgeeffects}). We do this by generating
+two collections of synthetic datasets -- one with distance-dependent
+connectivity and one without.  We then in each dataset randomly
+examine contiguous circular regions with area varying from zero to the
+entire volume, and empirically calculate the spatial variance in
+type-dependent connectivity.  We find that, if there is no
+distance-dependence, edge effects do not artificially introduce
+distance dependence. However, if the section we are examining is too
+small, our model can miss the distance dependence. Thus with respect
+to distance-dependent connectivity inference, our model errs on the
+side of caution. But we also find that for spatial extend that is
+similar to the currently available datasets the effects of this are
+quite limited.
+
+\subsubsection*{Learning types and circuitry in the retina}
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=160mm]{mouseretina2.ai}}
+  \caption{Discovering cell classes in the mouse retina connectome. Here
+we show the maximum a posteriori (MAP) estimate for the types in the mouse
+retina data. 
     a.) Input connectivity data for 950 cells for which soma positions
-    were known. b. clustered connectivity matrix. c.) connectivity
-    diagram showing our clusters, as well as the cell depth and
-    anatomist-labeled cell type of each cell. Our model shows moderate
-    agreement with the cell types traditionally identified by
-    anatomist laminar-specific connectivity patterns. 
+    were known. b. clustered connectivity matrix, each arbitrary color
+    corresponds to a single type, and will be used to identify that type in the remainder
+    of the plot. c.)  The spatial distribution of our cell types
+    -- each cell type tessellates space. Colors correspond to those in b.)
     d.) Connectivity between our clusters as a function
     of distance -- the cluster consisting primarily of retinal
-    ganglion cells (lower left-center) exhibits the expected near and
-    far connectivity. e.)  The spatial distribution of our cell types
-    -- each cell type tesselates space.}
+    ganglion cells (brown node on graph) exhibits the expected near and
+    far connectivity.}
 \label{fig:mouseretina}
@@
   \centerline{\includegraphics{comparison.ai}}
-  \caption{a. Cell-cell assignment matrix reflecting the posterior
-    distribution of clusterings found by our model. Large contiguous
-    black areas are cells that were considered the same type in the
-    majority of samples. b. The cells and their true types sorted to
-    agree with the coassignment matrix. c. A comparison of the
-    predictive accuracy (Area under the curve) for hand-labeled
-    anatomical data, versus inclusion of additional sources of
-    information. d. ) The spatial extent (in depth and area) of the
-    types identified by humans and our various algorithmic
-    approaches.}
-
+  \caption{\textbf{Visualizing type inference uncertainty.} Our fully
+    Bayesian model gives a confidence estimate (posterior probability)
+    that any two given cells are of the same type. In a.) we visualize
+    that cell-cell coassignment matrix, showing the probability that
+    cell i is of the same type as cell j on a range from 0.0 to
+    1.0. The block structure shows subsets of cells which are believed
+    to all belong to the same type. For comparison, b.) shows the
+    anatomist-defined type for each cell, grouped broadly into the
+    coarse types identified in the previous figure. c.) for each cell
+    in the matrix, c.) shows the soma depth as well as the depth
+    histogram for all putative synaptic contacts. \textbf{Link vs
+      cluster accuracy} d.) The posterior distribution of ROC curves
+    from 10-fold cross validation when predicting connectivity, as
+    well as e.) the area under the curve and f.) the type agreements
+    with known neuroanatomist types. g. A comparison of the predictive
+    accuracy (area under the curve) for hand-labeled anatomical data,
+    versus inclusion of additional sources of information, as well as
+    the clustering accuracy. Note that our model sacrifices very
+    little predictive accuracy for additional clustering accuracy. By
+    comparison, conventional methods fail at one or both.  d. ) The
+    spatial extent (in depth and area) of the types identified by
+    humans and our various algorithmic approaches.}
 \label{fig:mouseretina_compare}
@@
 
-\subsection{C. elegans}
-
-The roundworm Caenorhabditis elegans is a model system in
+\subsection*{Recovering spatial connectivity in multiple graphs simultaneously}
+
+Having shown our model to work on the repeating tessellated, laminar
+structure of the mammalian retina, we then apply our model to a
+structurally very different connectome -- the whole body of a small
+roundworm: \textit{Caenorhabditis elegans} is a model system in
 developmental neuroscience\autocite{White1986}, with the location and
@@
 leading to early measurement of the connectome. Unlike the retina,
-only the motor neurons in c. elegans exhibit regular distribution in
-space -- along the body axis. Most interneurons are
-concentrated in various ganglia that project throughout the entire
-animal, and the sensory neurons are primarily located in a small
-number of anterior ganglia.
+only the motor neurons in C. elegans exhibit regular distribution in
+space -- along the body axis. Most interneurons are concentrated in
+various ganglia that project throughout the entire animal, and the
+sensory neurons are primarily located in a small number of anterior
+ganglia. \textit{C. elegans} also differs from the retina in that the
+measured connectome is actually two separate graphs -- one of directed
+chemical synapses and another of undirected electrical synapses. As
+this is a very different connectome, it allows an interesting
+generalization test: how well will our model work on such a distinct
+data set.
 
@@
   \centering 
-  \centerline{\includegraphics[width=7in]{otherdata.ai}}
-  \caption{Discovering connectivity and type in both
-    \textit{c. elegans} and the MOS 6502 microprocessor.  a.) Initial
-    \textit{c. elegans} cell connectivity matrix, red are
-    (directional) chemical synapses, blue are electrical
-    synapses. Point size is proportional to synapse count between
-    those cells. b.) Clustered adjacency matrix showing identified
-    clusters. c.) Inter-cluster connectivity, showing the position of
-    each cell along the body axis as well as the ``true'' cell type --
-    both coarse (motor, sensory, and interneuron) as well as fine
-    (small text labels). d.) The clusters and connectivity versus
-    distance for connections between Gate and C1, Gate and C2, and C1
-    and C2 terminals on a transistor. Red and teal types have a
-    terminal pulled down to ground and mostly function as
-    inverters. Purple class are clocked, stateful transistors, orange
-    control the ALU and yellow control the special data bus. e.) The
-    spatial distribution of the transistors in each cluster.}
-  \label{fig:otherdata}
+  \centerline{\includegraphics[width=7.0in]{celegans2.ai}}
+  \caption{Discovering connectivity and type in
+    \textit{C. elegans}. a.) posterior distribution on cell
+    connectivity as a function of discovered type, similar to
+    figure~\ref{fig:mouseretina_compare}. In b.) we plot
+    neuroanatomist-derived types along with their labels. Our model
+    shows high probability of motor neurons, sensory neurons, and
+    various interneuron classes being of the same type. Soma positions
+    along the body axis are plotted in c.) where we see that we
+    cluster spatially-distributed motor neurons together, whereas head
+    sensory neurons are more likely to be grouped together as well.
+    d.) The ROC curves for held-out link probability for both the
+    electrical synapses (gap junctions) and chemical synapses in
+    c. elegans. e.) The posterior distribution of the area under the
+    ROC curve for the curves in d. f.) Measurements of the agreement
+    of our identified cell types compared to neuroanatomists. The high
+    compleneness but low homogeneity (and corresponding low ARI)
+    reflects our model's tendency to group multiple types into a
+    single type.}
+  \label{fig:celegans}
 \end{figure}
@@
 Using both the chemical and electrical connectivity (see methods), we
-determined the underlying clusters explained by connectivity and
-distance (fig~\ref{fig:otherdata}a). A superficial inspection of the results shows
-clustering into groups consisting roughly homogeneously of motor
-neurons, sensory neurons, and interneurons. Closer examination reveals
-agreement with the classifications originally outlined by Brenner in
-1986.
-
-We identify cell types that reflect the known motor/non-motor neuron
-classes, even though this system lacks the strong repeat
-microcircuitry our model was designed for.  Motor-neuron types AS, DA,
-and VA, all exclusively postsynaptic, are identified as a common type,
-as are motor-neuron types VD and DD. Traditional types VC, DB, and VB
-also mostly share a cluster. Various head motor neurons, including
-types SMD and RMD, are clustered together. Interneurons with known
-anatomically-distinct connectivity patterns, such as AVA (2 cells),
-are clustered into pure types.The algorithm even correctly places the
-single-cell types DVB and DVC by themselves.
+determined the underlying cell types explained by connectivity and
+distance (fig~\ref{fig:celegans}a). A superficial inspection of the
+results shows clustering into groups consisting roughly homogeneously
+of motor neurons, sensory neurons, and interneurons. Closer
+examination reveals agreement with the classifications originally
+outlined by White in 1986.
 
@@
 
-\subsection{Microprocessor}
+\subsection*{Types and connectivity in artificial structures}
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=6in]{mos6502.ai}}
+  \caption{Discovering connectivity and type the 6502 microprocessor.
+    a.) is the micrograph of the original microprocessor, with the
+    region containing the registers under study highlighted. b.) Our
+    graph consists of the interconnections of MOS field-effect
+    transistors with three terminals, Gate, C1, and C2. Reconstruction
+    technique did not permit resolution of C1 and C2 into source and
+    drain. c.) The spatial distribution of the transistors in each
+    cluster show a clear pattern d.) The clusters and connectivity
+    versus distance for connections between Gate and C1, Gate and C2,
+    and C1 and C2 terminals on a transistor. Purple and yellow types have a
+    terminal pulled down to ground and mostly function as
+    inverters. The blue types are clocked, stateful transistors, green
+    control the ALU and orange control the special data bus (SDB).}
+  \label{fig:mos6502}
+\end{figure}
 
@@
 the bus. The repeat patterns of spatial connectivity are visible in
-figure~\ref{fig:otherdata}a, showing the man-made horizontal and
+figure~\ref{fig:mos6502}c, showing the man-made horizontal and
 vertical layout of the same types of transistors.
 
-\section{Discussion}
+\section*{Discussion}
 We have presented a machine learning technique that allows cell types
 and microcircuitry to be discovered from connectomics data.  We have
-shown its applicability to a newly published connectomics dataset from
-the mouse retina. When compared to existing methods, we show how the
-incorporation of all of this data yields results that combine both
-high link-prediction accuracy and high agreement with human
-anatomists.  I also reveals interesting structure when applied to the
-classical connectivity of the worm c. elegans, and the connectivity of
-a historical man-made microprocessor. We have found that combining the
-available data types allows us to discover cell types and
-microcircuitry that were known to exist in the systems based on
+shown its applicability to regularly structured laminar neural
+circuits like the retina, as well as a less structured whole neuronal
+organism (\textit{C. elegans}) and a classical processor. When
+compared to existing methods, we show how the incorporation of all of
+this data yields results that combine both high link-prediction
+accuracy and high agreement with human anatomists. We have found that
+combining the available data types allows us to discover cell types
+and microcircuitry that were known to exist in the systems based on
 decades of previous research and allows good prediction of
@@
 parse the huge datasets. This transition may follow a similar
-transition to the one underway in molecular biology (with gene-finding
+transition to that of molecular biology (with gene-finding
 algorithms) and evolutionary biology with (computational
-phylogenetics).  In fact, it is known that there are significant
-disagreements across human anatomists and computational approaches
-promise to make the debates more data driven.
-
-\subsection{Methods Summary}
+phylogenetics). Ultimately, computational approaches may help resolve the significant
+disagreements across human anatomists.
+
+\subsection*{Methods Summary}
 
@@
 \newpage
-\section{Methods }
-
-\subsection{Probabilistic Model}
+\section*{Supplementary Methods }
+
+\subsection*{Probabilistic Model}
 
@@
 
-\subsection{Model Comparison}
+\subsection*{Model Comparison}
 We compare our model with a standard network clustering model, the
-latent-position clutsering model. This model assumes each cell belongs
+latent-position clustering model. This model assumes each cell belongs
 to one of K clusters, and each cluster is associated with a
@@
 \newpage
-\section{Supplemental Material}
-
-\subsection{Other Likelihoods}
+\section*{Supplemental Material}
+
+\subsection*{Other Likelihoods}
 \label{supp:otherlikelihoods}
@@
 
-\section{Measuring clustering similarity}
-
-The adjusted rand index (ARI) is a measure of the similarity of two
-different clusterings \autocite{Hubert1985} -- two identical clusters
-have an ARI of 1.0 while progressively more dissimilar clusters have
-lower ARIs, becoming negative as the clustering gets anti-correlated.
+\section*{Measuring clustering similarity}
+We compare discovered types to known types via cluster comparison
+metrics: cluster homogeneity, cluster completeness, and the adjusted
+Rand index. Homogeneity measures how many true types are in a given
+found type. If every cell type is split into two types, each subtype
+is still completely homogeneous. Completeness measures how many
+members of a given true type are split across found types.
+
+The adjusted rand index (ARI) takes into account both effects
+\autocite{Hubert1985} -- two identical clusterings have an ARI of 1.0
+while progressively more dissimilar clusters have lower ARIs, becoming
+slightly negative as the clustering gets anti-correlated.
+
+Figure~\ref{fig:s:comparison} shows the result of taking 20 different clusters
+and moving data points between them according to the following operations
+\begin{itemize}
+\item \textbf{distribute}: take a class and distribute its elements
+  uniformly among the remaining types.
+\item \textbf{merge} : take a type and merge it into another existing
+  type
+\item \textbf{split} : take a type and split it into two distinct
+  types.
+\end{itemize}
+
+We can see the impact on ARI, completeness, and homogeneity as we
+perform these operations on more of the original 20 types. In all
+cases, ``distribution'' of one type among the others is detremental to
+the metric. Splitting impacts completeness but not homogeneity, and
+merging impacts homogeneity but not completness.
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=180mm]{cluster_metric_comparison.pdf}}
+  \caption{Type agreement evaluation metrics as a function of splitting types, merging types, and randomly distributing cells between types}
+  \label{fig:s:comparison}
+\end{figure}
 

### 2015-04-23 docs/arxiv.tex
updates before push

@@
 connectivity.
+
+\subsection*{Methods Summary}
+
+For the basic link-distance model, we take as input a connectivity
+matrix $R$ defining the connections between cell $e_i$ and $e_j$, as
+well as a distance function $d(e_i, e_j)$ representing a (physical)
+distance between adjacent cells.  See
+the supplemental material for extension to multiple connectivity
+matrices. We assume there exist an unknown number $K$ of latent
+(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
+cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
+of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
+observed connectivity between two cells $R(e_i, e_j)$ then depends
+only on their latent type and their distance through a link function
+$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
+latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
+as a set of global hyper parameters $\theta$, such that the link
+function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
+
+We then jointly infer the maximum a posteriori (MAP) estimate of the
+class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix
+$\eta_{mn}$, and the global model hyperparameters $\theta$ :
+
+\begin{equation}
+  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
+\end{equation}
+
+For the retina data, we then extend the model with the additional
+features indicated. Cell soma depth is modeled as a
+cell-type-dependent Gaussian with latent (unknown) per-type mean and
+variance. Similarly, each cell has some number $N_i$ of synapses, 
+each of which is drawn from a cell-type-specific density profile
+with up to three modes.
+
+Inference is performed in three steps via composable transition 
+kernels -- one for structural, one for per-type parameters, and
+one kernel for global parameters and hyperparameters. Details
+of data preprocessing, inference parameters, and runtime can
+be found in the Supplementary Methods section.
+
+\subsection*{Probabilistic Model}
+
+Our model is a extension of the iSBM
+\autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
+inspired by attempts to extend these models with arbitrary
+discriminative functions\autocite{Murphy2012}.
+
+We take as input a connectivity matrix $R$ defining the connections
+between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
+e_j)$ representing a (physical) distance between adjacent cells. See
+the supplemental material for extension to multiple connectivity
+matrices. We assume there exist an unknown number $K$ of latent
+(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
+cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
+of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
+observed connectivity between two cells $R(e_i, e_j)$ then depends
+only on their latent type and their distance through a link function
+$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
+latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
+as a set of global hyper parameters $\theta$, such that the link
+function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
+
+We then jointly infer the maximum a posteriori (MAP) estimate of the
+class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
+the global model hyperparameters $\theta$ :
+
+\begin{equation}
+  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
+\end{equation}
+
+% Methods section 3000 words
+% methods summary does not appear in the online text
+% methods section can have no figures or tables [grrr]
+
+We describe the spatial ``Logistic-distance Bernoulli''  function here,
+and others in the supplemental material.
+
+% fixme class component, NOT class
+The ``logistic-distance Bernoulli'' spatial model assumes that, if cell
+$e_i$ is of type $m$ and cell $e_j$ is of type $n$, then $\eta_{mn}
+= (\mu_{mn}, \lambda_{mn})$, and the probability that two cells $e_i$
+and $e_j$ are connected is given by
+\begin{eqnarray}
+p^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
+p &= & p^* \cdot (p_{max} - p_{min}) + p_{min}
+\end{eqnarray}
+where $p_{max}$ and $p_{min}$ are global per-graph parameters.
+
+We place an exponential priors on the latent parameters:
+\begin{eqnarray}
+ \mu_{mn} \sim \exp(\mu | \mu^{hp}) \\
+\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})
+\end{eqnarray}
+
+using  $\lambda^{hp}$ and $\mu^{hp}$ as global per-graph hyperparameters.
+
+We use a Dirichlet-process prior on class assignments, which allows
+the number of classs to be determined automatically. In brief, for $N$
+total cells, the probability of a cell belonging to a class is
+proportional to the number of datapoints already in that class, $N_k$,
+such that $p(c_i = k) \propto \frac{m_k}{N + \alpha}$ and the
+probability of the cell belonging to a new class $k'$ is $p(c_i = k')
+\propto \frac{\alpha}{N + \alpha}$. $\alpha$ is the global
+concentration parameter -- larger values of $\alpha$ make the model
+more likely to propose new classes. We grid the parameter $\alpha$ and
+allow the best value to be learned from the data.
+
+Where we model cell depth, we assume that 
+each cell type has a typical depth, and thus a Gaussian distribution
+of $s_i$. We assume $s_i \sim N(\mu^{(s)}_k,
+\sigma^{2(s)}_k)$, where the $(s)$ superscript indicates
+these model parameters are associated with the soma-depth 
+portion of our model.  We use a conjugate prior for $(\mu^{(s)}_k, \sigma^{2(s)}_k)$ with
+$\mu^{(s)}_k \sim N(\mu^{(s)}_{hp}, \sigma^{2(s)}_k/ \kappa^{(s)}_{hp})$ and $\sigma^{2(s)}_k \sim
+\chi^{-1}(\sigma^{2(s)}_{hp}, \nu^{(s)}_{hp}$. The use of conjugacy simplifies inference while allowing for each cell-type to have its own depth mean and distribution.
+
+Where we model synapse depth profile, we assume that each cell
+type has a characteristic depth distribution of synaptic contact points, and thus
+a mixture of Gaussians distribution over cell $i$’s $N_i$ contact points, $\vec{g^i}$.
+We do this by assuming the $g^i_j$
+are drawn from an $M=3$-component mixture of Gaussians. Thus associated with each cell type $k$ is a vector
+of $M$ Gaussian means $(\mu^g_{k,1}, \cdots, \mu^g_{k, M})$, and a
+mixture vector $\pi_k$.
+This representation can thus model depth distributions of contact points that have up to three modes, an assumption that is well matched in the bulk of anatomical studies of cell-type dependent connectivity.
+
+\subsection*{Inference} 
+We perform posterior inference via Markov-chain Monte Carlo (MCMC),
+annealing on the global likelihood during the traditional burn-in
+phase. MCMC transition kernels for different parts of the state space
+can be chained together to construct a kernel whose ergodic
+distribution is the target ergodic distribution over the entire state space.
+
+Our first transition kernel (``structural'') performs gibbs sampling 
+of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
+The lack of conjugacy in our likelihood model makes an explicit 
+evaluation of the conditional assignment probabilities impossible, 
+motivating us to use an auxiliary variable method \autocite{Neal2000}
+in which a collection of ephemeral classs are explicitly represented
+for the duration of the Gibbs scan.
+
+We then employ a transition kernel to update the per-component
+parameter values $\eta_{mn}$. Conditioned on the assignment vector
+$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
+individual $\eta_{mn}$ are independent. We slice sample \autocite{Neal2003}
+each component's parameters, choosing the slice width as a function
+of the global hyperparameter range.
+
+The global hyper-parameters, both $\alpha$ and $\theta$, are allowed
+to take on a discrete set of possible values. As $\theta$ is often a
+tuple of possible values, we explore the cartesian product of all
+possible values. We then Gibbs sample (our final transition kernel),
+which is always possible in a small, finite, discrete state space.
+
+We chain these three kernels together, and then globally anneal on the
+likelihood from a temperature of $T=64$ down to $T=1$ over $900$
+iterations unless otherwise indicated, and then run the chain for
+another $100$ iterations. We then generate at least $20$ samples, each
+taken from the end of a single Markov chain initialized from different
+random initial points in the state space. For visualization we pick the
+chain with the highest log likelihood, but for all numerical
+comparisons (including link probability and cluster accuracy) we use this full
+collection of samples from the posterior distribution to estimate the
+resulting statistics.
+
+\subsection*{Link Prediction}
+To compute link-prediction accuracy we compute the probability of a
+link between two cells using each model, trained via 10-fold cross
+validation.  We use a full collection of posterior samples when computing
+the link probability, and then compute the area under the ROC curve
+for each.
+
+We compare our model with a standard network clustering model, the
+latent-position clustering model. This model assumes each cell belongs
+to one of K clusters, and each cluster is associated with a
+$d-$dimensional Gaussian distribution. The probability of a link is
+then a function of the distance between the data points in this
+continuous-space. We use \autocite{Salter-Townshend2013} a variational
+implementation provided in R, parametrically varying the number of
+latent dimensions and the number of requested groups.  While this
+model provides reasonable link predictive accuracy, the clusterings
+dramatically disagree with those from human anatomists.
+
+\subsection*{Parameters}
+
+Hierarchical generative models can be sensitive to hyperparameter
+settings, thus for most hyperparamters we perform inference. In cases
+where we cannot we run separate collections of markov chains at
+separate settings and show the results across all pooled
+parameters. For the case of the mouse retina data, we consisder
+maximum link probability $p_{max} \in \{0.95, 0.9, 0.7\}$, variance
+scales for the synapse density profile of $\sigma^2 \in \{0.01, 0.1,
+1.0\}$ (of normalized depth), and $K \in \{2, 3\}$ possible synapse
+density profile mixture components. For the connectivity-distance-only
+model we actually perform inference over both $p_{max}$ and $p_{min}$.
+
+\subsubsection*{Mixing of our Markov chains} 
+Evaluating whether or not approximate inference methods, such as MCMC,
+produce samples which are valid approximations of the posterior
+distribution is an ongoing area of research in the computational
+statistics community. We use a rough proxy here -- synthetic
+likelihood evaluation.  For synthetic datasets of sizes comparable to
+our real data size, do we recover known ground truth information after
+running our markov chains for the appropriate amount of time?
+
+Figures \ref{fig:mixing:ari} and \ref{fig:mixing:scorevstime} shows
+the cluster accuracy (ARI) to ground truth and the total log score as a function of
+runtime.  We see dramatic changes in log score initially as we vary
+the temperature, stabilizing as runtime progresses, for each
+chain. Then we see the characteristic jumps between nearby modes
+towards the end of the run, in both log score and ARI.  Importantly,
+regardless of whether our model over- or under-estimates the exact
+posterior variance about the network, we find points in the latent
+variable space that are both predictive \textit{and} parsimonious,
+largely agreeing with the human anatomists and predicting existing
+connections.
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=60mm]{mixing.ari.pdf}}
+  \caption{Adjusted rand index for synthetic data as a function of run iteration. }
+\label{fig:mixing:ari}
+\end{figure}
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=130mm]{mixing.scorevstime.pdf}}
+  \caption{Total model score (log score) vs iteration}
+\label{fig:mixing:scorevstime}
+\end{figure}
+
+\subsection*{Dataset details}
+\subsubsection* {Mouse Retina}
+\label{supp:mouseretina}
+Dense serial electron microscopy of a $114\mu m \times 80 \mu m $ area
+in the mouse retina by \autocite{Helmstaedter2013} yielded a listing
+of places where neurons come into contact. There were over 1000 cells
+originally, and selected the $950$ for which the location of the soma
+could be reconstructed from the provided cell plots (soma locations
+were not provided by the study's authors in machine-readable
+form). Ultimately this left a matrix between the total synapse-like
+contact area between all pairs of 950 cells. Area was thresholded at
+$0.1\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
+matrix that served as input to our algorithm. We measured the distance
+between cells using the reconstructed soma centers, and used the
+Logistic-Distance spatial relation. Hyperprior griddings are shown in
+supplemental section \ref{supp:hyperpriors}.
+
+\subsubsection*{C. elegans}
+
+We obtained the connectome of c. elegans from data published
+previously \autocite{Varshney2011}, and isolated the 279 nonpharyngeal
+neurons, with a total of 6393 chemical synapses and 890 gap junctions
+originally cleaned up in \autocite{Chen2006}. A cell's position was
+its distance along the anterior-posterior axis normalized between zero
+and one. We used both networks, the chemical network as a directed
+graph and the electrical network as undirected graph. We use the
+synapse counts with the logistic-distance poisson likelihood, scaling
+the counts by 4.0 to compensate for the Poisson's overdispersion.
+
+\subsubsection*{Microprocessor}
+We extracted the connection graph for the transistors in the MOS6502
+\autocite{visual6502source}. Each transistor has three terminals (gate,
+source, drain), but the methods of the original dataset were unable to
+consistently resolve which of the C1 and C2 terminals were source and
+drain, leading to ambiguity in our encoding. We identified a region
+consisting of three registers X, Y, and S via visual
+inspection and focused our efforts there. We created a total of six
+connectivity matrices by examining possible terminal pairings. One
+graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
+$e_j$ are connected via pins $g$ and $c_1$.
+
+% \begin{addendum}
+%  \item Put acknowledgements here.
+%  \item[Competing Interests] The authors declare that they have no
+% competing financial interests.
+%  \item[Correspondence] Correspondence and requests for materials
+% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
+% \end{addendum}
+
+\subsection*{Other Likelihoods}
+\label{supp:otherlikelihoods}
+
+We reparameterized the Logistic-Distance Bernoulli likelihood to
+better capture the microprocessor data structure. We are explicitly
+setting the maximum probability $p$ of the logistic function on a
+per-component basis, drawing from a global $p \sim \operatorname{Beta}(\alpha_{hp},
+\beta_{hp})$. Then $\lambda$ is set for each component as a global
+hyperparameter, $\lambda$.
+
+The ``logstic-distance Poisson'' spatial model is used to explicitly
+mode the count of synapses, $c$, between two neurons. The probability
+of c synapses between two neurons is distributed $c \sim
+\textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a
+scaled logistic function (the logistic function has range $[0,
+1]$. For each component $\eta_{mn}$ we learn both the threshold
+$\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$
+and $n$ are likely to have on average $20$ synases if they are closer
+than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$ "
+
+Thus the probability of $R(e_i, e_j) = c$ synapses between two cells $e_i$ and $e_j$ is given by:
+\begin{eqnarray}
+r^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda}}\\
+r & = & r^* \cdot (r_{mn} - r_{min}) + r_{min} \\
+R(e_i, e_j) \sim \textrm{Poisson}(c | r)
+\end{eqnarray}
+
+where $\lambda$ and $r_{min}$ are per-graph parameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $r_{mn} \sim \exp(r_{mn} | r_{scale}^{hp})$.
+
+\subsection*{Source code and data}
+
+All source code and materials for running experiments can be
+obtained from the project website, at \\
+
+\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/netmotifs/}
+
+and the content of this paper along with scripts to run
+experiments and generate all figures can be found at
+
+\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/connect-disco-paper/}
+
+All preprocessed data has been made publicaly available as well.
+
+Please contact the author for pre-publication access.
+
+\subsection*{Extension to multiple graphs}
+\label{supp:multigraph}
+The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by extending the likelihood to include the product of the likelihoods of the individual graphs.
+
+\begin{equation}
+  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}, \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha)
+\end{equation}
+
+\FloatBarrier
+\subsection*{Hyperprior grids and hyperprior inference}
+\label{supp:hyperpriors}
+
+For the mouse retina Logistic-Distance Bernoulli model, we gridded
+$\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0
+and 80.
+
+For the c. elegans data with the Logistic Distance poisson model, we
+gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points
+bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20
+$\log_{10}$-spaced points between 2.0 and 20.0. We globally set
+$rate_{min}=0.01$.
+
+For the microprocessor with the Logistic Distance fixed lambda
+Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced
+points between 10 and 500 and set $\lambda=\mu_{hp}/10$. $p_{min} \in
+\{0.001, 0.01, 0.02\}$ and both $p_\alpha$ and $p_\beta \in \{0.1,
+1.0, 2.0\}$.
+
+\subsection*{Measuring clustering similarity}
+We compare discovered types to known types via cluster comparison
+metrics: cluster homogeneity, cluster completeness, and the adjusted
+Rand index. Homogeneity measures how many true types are in a given
+found type. If every cell type is split into two types, each subtype
+is still completely homogeneous. Completeness measures how many
+members of a given true type are split across found types.
+
+The adjusted rand index (ARI) takes into account both effects
+\autocite{Hubert1985} -- two identical clusterings have an ARI of 1.0
+while progressively more dissimilar clusters have lower ARIs, becoming
+slightly negative as the clustering gets anti-correlated.
+
+Figure~\ref{fig:s:comparison} shows the result of taking 20 different clusters
+and moving data points between them according to the following operations
+\begin{itemize}
+\item \textbf{distribute}: take a class and distribute its elements
+  uniformly among the remaining types.
+\item \textbf{merge} : take a type and merge it into another existing
+  type
+\item \textbf{split} : take a type and split it into two distinct
+  types.
+\end{itemize}
+
+We can see the impact on ARI, completeness, and homogeneity as we
+perform these operations on more of the original 20 types. In all
+cases, ``distribution'' of one type among the others is detremental to
+the metric. Splitting impacts completeness but not homogeneity, and
+merging impacts homogeneity but not completness.
+
+\begin{figure}
+  \centering 
+  \centerline{\includegraphics[width=180mm]{cluster_metric_comparison.pdf}}
+  \caption{Type agreement evaluation metrics as a function of splitting types, merging types, and randomly distributing cells between types}
+  \label{fig:s:comparison}
+\end{figure}
 
@@
 
-\subsection*{Methods Summary}
-
-For the basic link-distance model, we take as input a connectivity
-matrix $R$ defining the connections between cell $e_i$ and $e_j$, as
-well as a distance function $d(e_i, e_j)$ representing a (physical)
-distance between adjacent cells.  See
-the supplemental material for extension to multiple connectivity
-matrices. We assume there exist an unknown number $K$ of latent
-(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
-cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
-of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
-observed connectivity between two cells $R(e_i, e_j)$ then depends
-only on their latent type and their distance through a link function
-$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
-latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
-as a set of global hyper parameters $\theta$, such that the link
-function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
-
-We then jointly infer the maximum a posteriori (MAP) estimate of the
-class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix
-$\eta_{mn}$, and the global model hyperparameters $\theta$ :
-
-\begin{equation}
-  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
-\end{equation}
-
-For the retina data, we then extend the model with the additional
-features indicated. Cell soma depth is modeled as a
-cell-type-dependent Gaussian with latent (unknown) per-type mean and
-variance. Similarly, each cell has some number $N_i$ of synapses, 
-each of which is drawn from a cell-type-specific density profile
-with up to three modes.
-
-Inference is performed in three steps via composable transition 
-kernels -- one for structural, one for per-type parameters, and
-one kernel for global parameters and hyperparameters. Details
-of data preprocessing, inference parameters, and runtime can
-be found in the Supplementary Methods section.
-
 \printbibliography
+\newpage
 
@@
 
-\newpage
-\section*{Supplementary Methods }
-
-\subsection*{Probabilistic Model}
-
-Our model is a extension of the iSBM
-\autocite{Kemp2006a,Xu2006} to incorporate spatial relations between entities,
-inspired by attempts to extend these models with arbitrary
-discriminative functions\autocite{Murphy2012}.
-
-We take as input a connectivity matrix $R$ defining the connections
-between cell $e_i$ and $e_j$, as well as a distance function $d(e_i,
-e_j)$ representing a (physical) distance between adjacent cells. See
-the supplemental material for extension to multiple connectivity
-matrices. We assume there exist an unknown number $K$ of latent
-(unobserved) cell types, $k \in \{1, 2, 3, \dots, K\}$, and that each
-cell $e_i$ belongs to a single cell type. We indicate a cell $e_i$ is
-of type $k$ using the assignment vector $\vec(c)$, so $c_i = k$. The
-observed connectivity between two cells $R(e_i, e_j)$ then depends
-only on their latent type and their distance through a link function
-$f(\cdot, d(e_i, e_j))$. We assume $f$ is parameterized based on the
-latent type, $c_i=m$ and $c_j=n$, via a parameter$\eta_{mn}$, as well
-as a set of global hyper parameters $\theta$, such that the link
-function is $f(d(e_i, e_j) | \eta_{mn}, \theta)$.
-
-We then jointly infer the maximum a posteriori (MAP) estimate of the
-class assignment vector $\vec(c) = \{c_i\}$, the parameter matrix $\eta_{mn}$, and
-the global model hyperparameters $\theta$ :
-
-\begin{equation}
-  p(\vec{c}, \eta, \theta | R ) \propto \prod_{i, j} p(R(e_i, e_j) | f(d(e_i, e_j) | \eta_{c_ic_j}), \theta) \prod_{m, n} p(\eta_{mn} | \theta)  p(\theta) p(\vec{c} | \alpha) p(\alpha) p(\theta)
-\end{equation}
-
-% Methods section 3000 words
-% methods summary does not appear in the online text
-% methods section can have no figures or tables [grrr]
-
-We describe the spatial ``Logistic-distance Bernoulli''  function here,
-and others in the supplemental material.
-
-% fixme class component, NOT class
-The ``logistic-distance Bernoulli'' spatial model assumes that, if cell
-$e_i$ is of type $m$ and cell $e_j$ is of type $n$, then $\eta_{mn}
-= (\mu_{mn}, \lambda_{mn})$, and the probability that two cells $e_i$
-and $e_j$ are connected is given by
-\begin{eqnarray}
-p^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda_{mn}}}\\
-p &= & p^* \cdot (p_{max} - p_{min}) + p_{min}
-\end{eqnarray}
-where $p_{max}$ and $p_{min}$ are global per-graph parameters.
-
-We place an exponential priors on the latent parameters:
-\begin{eqnarray}
- \mu_{mn} \sim \exp(\mu | \mu^{hp}) \\
-\lambda_{mn} \sim \exp(\lambda | \lambda^{hp})
-\end{eqnarray}
-
-using  $\lambda^{hp}$ and $\mu^{hp}$ as global per-graph hyperparameters.
-
-We use a Dirichlet-process prior on class assignments, which allows
-the number of classs to be determined automatically. In brief, for $N$
-total cells, the probability of a cell belonging to a class is
-proportional to the number of datapoints already in that class, $N_k$,
-such that $p(c_i = k) \propto \frac{m_k}{N + \alpha}$ and the
-probability of the cell belonging to a new class $k'$ is $p(c_i = k')
-\propto \frac{\alpha}{N + \alpha}$. $\alpha$ is the global
-concentration parameter -- larger values of $\alpha$ make the model
-more likely to propose new classes. We grid the parameter $\alpha$ and
-allow the best value to be learned from the data.
-
-Where we model cell depth, we assume that 
-each cell type has a typical depth, and thus a Gaussian distribution
-of $s_i$. We assume $s_i \sim N(\mu^{(s)}_k,
-\sigma^{2(s)}_k)$, where the $(s)$ superscript indicates
-these model parameters are associated with the soma-depth 
-portion of our model.  We use a conjugate prior for $(\mu^{(s)}_k, \sigma^{2(s)}_k)$ with
-$\mu^{(s)}_k \sim N(\mu^{(s)}_{hp}, \sigma^{2(s)}_k/ \kappa^{(s)}_{hp})$ and $\sigma^{2(s)}_k \sim
-\chi^{-1}(\sigma^{2(s)}_{hp}, \nu^{(s)}_{hp}$. The use of conjugacy simplifies inference while allowing for each cell-type to have its own depth mean and distribution.
-
-Where we model synapse depth profile, we assume that each cell
-type has a characteristic depth distribution of synaptic contact points, and thus
-a mixture of Gaussians distribution over cell $i$’s $N_i$ contact points, $\vec{g^i}$.
-We do this by assuming the $g^i_j$
-are drawn from an $M=3$-component mixture of Gaussians. Thus associated with each cell type $k$ is a vector
-of $M$ Gaussian means $(\mu^g_{k,1}, \cdots, \mu^g_{k, M})$, and a
-mixture vector $\pi_k$.
-This representation can thus model depth distributions of contact points that have up to three modes, an assumption that is well matched in the bulk of anatomical studies of cell-type dependent connectivity.
-
-\subsection*{Inference} 
-We perform posterior inference via Markov-chain Monte Carlo (MCMC),
-annealing on the global likelihood during the traditional burn-in
-phase. MCMC transition kernels for different parts of the state space
-can be chained together to construct a kernel whose ergodic
-distribution is the target ergodic distribution over the entire state space.
-
-Our first transition kernel (``structural'') performs gibbs sampling 
-of the assignment vector $p(\vec{c} | \eta, \theta, \alpha)$. 
-The lack of conjugacy in our likelihood model makes an explicit 
-evaluation of the conditional assignment probabilities impossible, 
-motivating us to use an auxiliary variable method \autocite{Neal2000}
-in which a collection of ephemeral classs are explicitly represented
-for the duration of the Gibbs scan.
-
-We then employ a transition kernel to update the per-component
-parameter values $\eta_{mn}$. Conditioned on the assignment vector
-$\vec{c}$ and the model hyperparameters $\theta, \alpha$ the 
-individual $\eta_{mn}$ are independent. We slice sample \autocite{Neal2003}
-each component's parameters, choosing the slice width as a function
-of the global hyperparameter range.
-
-The global hyper-parameters, both $\alpha$ and $\theta$, are allowed
-to take on a discrete set of possible values. As $\theta$ is often a
-tuple of possible values, we explore the cartesian product of all
-possible values. We then Gibbs sample (our final transition kernel),
-which is always possible in a small, finite, discrete state space.
-
-We chain these three kernels together, and then globally anneal on the
-likelihood from a temperature of $T=64$ down to $T=1$ over $900$
-iterations unless otherwise indicated, and then run the chain for
-another $100$ iterations. We then generate at least $20$ samples, each
-taken from the end of a single Markov chain initialized from different
-random initial points in the state space. For visualization we pick the
-chain with the highest log likelihood, but for all numerical
-comparisons (including link probability and cluster accuracy) we use this full
-collection of samples from the posterior distribution to estimate the
-resulting statistics.
-
-\subsection*{Link Prediction}
-As a proxy for link-prediction accuracy we compute the probability of a link
-between two cells using each model, trained fully on the data. While this method
-is potentially prone to overfitting, the overfitting will be shared across models
-and in fact will preferentially bias in favor of competing models
-which over-cluster the data. We use a full collection of posterior samples when
-computing the link probability, and then compute the area under the ROC curve for
-each.
-
-\subsection*{Model Comparison}
-We compare our model with a standard network clustering model, the
-latent-position clustering model. This model assumes each cell belongs
-to one of K clusters, and each cluster is associated with a
-$d-$dimensional Gaussian distribution. The probability of a link is
-then a function of the distance between the data points in this
-continuous-space. We use \autocite{Salter-Townshend2013} a variational
-implementation provided in R, parametrically varying the number of
-latent dimensions and the number of requested groups.  While this
-model provides reasonable link predictive accuracy, the clusterings
-dramatically disagree with those from human anatomists.
-
-\subsection*{Parameters}
-
-Hierarchical generative models can be sensitive to hyperparameter
-settings, thus for most hyperparamters we perform inference. In cases
-where we cannot we run separate collections of markov chains at
-separate settings and show the results across all pooled
-parameters. For the case of the mouse retina data, we consisder
-maximum link probability $p_{max} \in \{0.95, 0.9, 0.7\}$, variance
-scales for the synapse density profile of $\sigma^2 \in \{0.01, 0.1,
-1.0\}$ (of normalized depth), and $K \in \{2, 3\}$ possible synapse
-density profile mixture components. For the connectivity-distance-only
-model we actually perform inference over both $p_{max}$ and $p_{min}$.
-
-\subsubsection*{Mixing of our Markov chains} 
-Evaluating whether or not approximate inference methods, such as MCMC,
-produce samples which are valid approximations of the posterior
-distribution is an ongoing area of research in the computational
-statistics community. We use a rough proxy here -- synthetic
-likelihood evaluation.  For synthetic datasets of sizes comparable to
-our real data size, do we recover known ground truth information after
-running our markov chains for the appropriate amount of time?
-
-Figures \ref{fig:mixing:ari} and \ref{fig:mixing:scorevstime} shows
-the cluster accuracy (ARI) to ground truth and the total log score as a function of
-runtime.  We see dramatic changes in log score initially as we vary
-the temperature, stabilizing as runtime progresses, for each
-chain. Then we see the characteristic jumps between nearby modes
-towards the end of the run, in both log score and ARI.  Importantly,
-regardless of whether our model over- or under-estimates the exact
-posterior variance about the network, we find points in the latent
-variable space that are both predictive \textit{and} parsimonious,
-largely agreeing with the human anatomists and predicting existing
-connections.
-
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics[width=60mm]{mixing.ari.pdf}}
-  \caption{Adjusted rand index for synthetic data as a function of run iteration. }
-\label{fig:mixing:ari}
-\end{figure}
-
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics[width=130mm]{mixing.scorevstime.pdf}}
-  \caption{Total model score (log score) vs iteration}
-\label{fig:mixing:scorevstime}
-\end{figure}
-
-\subsection* {Mouse Retina}
-\label{supp:mouseretina}
-Dense serial electron microscopy of a $114\mu m \times 80 \mu m $ area
-in the mouse retina by \autocite{Helmstaedter2013} yielded a listing
-of places where neurons come into contact. There were over 1000 cells
-originally, and selected the $950$ for which the location of the soma
-could be reconstructed from the provided cell plots (soma locations
-were not provided by the study's authors in machine-readable
-form). Ultimately this left a matrix between the total synapse-like
-contact area between all pairs of 950 cells. Area was thresholded at
-$0.1\mu m$, determined by hand, to yield a 950 $\times$ 950 entry
-matrix that served as input to our algorithm. We measured the distance
-between cells using the reconstructed soma centers, and used the
-Logistic-Distance spatial relation. Hyperprior griddings are shown in
-supplemental section \ref{supp:hyperpriors}.
-
-\subsection*{C. elegans}
-
-We obtained the connectome of c. elegans from data published
-previously \autocite{Varshney2011}, and isolated the 279 nonpharyngeal
-neurons, with a total of 6393 chemical synapses and 890 gap junctions
-originally cleaned up in \autocite{Chen2006}. A cell's position was
-its distance along the anterior-posterior axis normalized between zero
-and one. We used both networks, the chemical network as a directed
-graph and the electrical network as undirected graph. We use the
-synapse counts with the logistic-distance poisson likelihood, scaling
-the counts by 4.0 to compensate for the Poisson's overdispersion.
-
-\subsection*{Microprocessor}
-We extracted the connection graph for the transistors in the MOS6502
-\autocite{visual6502source}. Each transistor has three terminals (gate,
-source, drain), but the methods of the original dataset were unable to
-consistently resolve which of the C1 and C2 terminals were source and
-drain, leading to ambiguity in our encoding. We identified a region
-consisting of three registers X, Y, and S via visual
-inspection and focused our efforts there. We created a total of six
-connectivity matrices by examining possible terminal pairings. One
-graph, for example, $R^{gc_1}(e_i, e_j)=1$ if transistor $e_j$ and
-$e_j$ are connected via pins $g$ and $c_1$.
-
-% \begin{addendum}
-%  \item Put acknowledgements here.
-%  \item[Competing Interests] The authors declare that they have no
-% competing financial interests.
-%  \item[Correspondence] Correspondence and requests for materials
-% should be addressed to A.B.C.~(email: myaddress@nowhere.edu).
-% \end{addendum}
-\newpage
-\section*{Supplemental Material}
-
-\subsection*{Other Likelihoods}
-\label{supp:otherlikelihoods}
-
-We reparameterized the Logistic-Distance Bernoulli likelihood to
-better capture the microprocessor data structure. We are explicitly
-setting the maximum probability $p$ of the logistic function on a
-per-component basis, drawing from a global $p \sim \operatorname{Beta}(\alpha_{hp},
-\beta_{hp})$. Then $\lambda$ is set for each component as a global
-hyperparameter, $\lambda$.
-
-The ``logstic-distance Poisson'' spatial model is used to explicitly
-mode the count of synapses, $c$, between two neurons. The probability
-of c synapses between two neurons is distributed $c \sim
-\textrm{Poisson}(c | r)$, where $r$ (the ``rate'') is generated by a
-scaled logistic function (the logistic function has range $[0,
-1]$. For each component $\eta_{mn}$ we learn both the threshold
-$\mu_{mn}$ and the rate scaling factor $r_{mn}$ Thus if for cells $m$
-and $n$ are likely to have on average $20$ synases if they are closer
-than $5 \mu m$, then $\mu_{mn} = 5$ and $r_{mn} = 20$ "
-
-Thus the probability of $R(e_i, e_j) = c$ synapses between two cells $e_i$ and $e_j$ is given by:
-\begin{eqnarray}
-r^* &=& \frac{1.0}{1 + \exp \frac{d(e_i, e_j) - \mu_{mn}}{\lambda}}\\
-r & = & r^* \cdot (r_{mn} - r_{min}) + r_{min} \\
-R(e_i, e_j) \sim \textrm{Poisson}(c | r)
-\end{eqnarray}
-
-where $\lambda$ and $r_{min}$ are per-graph parameters. Per-component parameters $\mu_{mn} \sim \exp(\mu | \mu^{hp})$ and $r_{mn} \sim \exp(r_{mn} | r_{scale}^{hp})$.
-
-\subsection*{Source code and data}
-
-All source code and materials for running experiments can be
-obtained from the project website, at \\
-
-\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/netmotifs/}
-
-and the content of this paper along with scripts to run
-experiments and generate all figures can be found at
-
-\href{http://https://github.com/ericmjonas/netmotifs}{https://github.com/ericmjonas/connect-disco-paper/}
-
-All preprocessed data has been made publicaly available as well.
-
-Please contact the author for pre-publication access.
-
-\subsection*{Extension to multiple graphs}
-\label{supp:multigraph}
-The model can handle multiple graphs $R^q$ simultaneously with a shared clustering by extending the likelihood to include the product of the likelihoods of the individual graphs.
-
-\begin{equation}
-  p(\vec{c}, \{\eta^q\}, \{\theta^q\} | \{R^q\} ) \propto \prod_q \Bigg(\prod_{i, j} p(R^q(e_i, e_j) | f(d(e_i, e_j) | \eta^q_{c_ic_j}, \theta^q) \prod_{m, n} p(\eta^q_{mn} | \theta^q)  p(\theta^q) \Bigg) p(\vec{c} | \alpha) p(\alpha)
-\end{equation}
-
-\FloatBarrier
-\subsection*{Hyperprior grids and hyperprior inference}
-\label{supp:hyperpriors}
-
-For the mouse retina Logistic-Distance Bernoulli model, we gridded
-$\mu^{hp}$ and $\lambda^{hp}$ into 40 $\log_{10}$-spaced points 1.0
-and 80.
-
-For the c. elegans data with the Logistic Distance poisson model, we
-gridded $\mu_{hp}$ and $\lambda$ into 20 $\log_{10}$-spaced points
-bween 0.2 and 2.0, and the $ratescale^{hp}$ parameter into 20
-$\log_{10}$-spaced points between 2.0 and 20.0. We globally set
-$rate_{min}=0.01$.
-
-For the microprocessor with the Logistic Distance fixed lambda
-Bernoulli likelihood, we gridded $mu_{hp}$ into 50 $\log_{10}$-spaced
-points between 10 and 500 and set $\lambda=\mu_{hp}/10$. $p_{min} \in
-\{0.001, 0.01, 0.02\}$ and both $p_\alpha$ and $p_\beta \in \{0.1,
-1.0, 2.0\}$.
-
-\section*{Measuring clustering similarity}
-We compare discovered types to known types via cluster comparison
-metrics: cluster homogeneity, cluster completeness, and the adjusted
-Rand index. Homogeneity measures how many true types are in a given
-found type. If every cell type is split into two types, each subtype
-is still completely homogeneous. Completeness measures how many
-members of a given true type are split across found types.
-
-The adjusted rand index (ARI) takes into account both effects
-\autocite{Hubert1985} -- two identical clusterings have an ARI of 1.0
-while progressively more dissimilar clusters have lower ARIs, becoming
-slightly negative as the clustering gets anti-correlated.
-
-Figure~\ref{fig:s:comparison} shows the result of taking 20 different clusters
-and moving data points between them according to the following operations
-\begin{itemize}
-\item \textbf{distribute}: take a class and distribute its elements
-  uniformly among the remaining types.
-\item \textbf{merge} : take a type and merge it into another existing
-  type
-\item \textbf{split} : take a type and split it into two distinct
-  types.
-\end{itemize}
-
-We can see the impact on ARI, completeness, and homogeneity as we
-perform these operations on more of the original 20 types. In all
-cases, ``distribution'' of one type among the others is detremental to
-the metric. Splitting impacts completeness but not homogeneity, and
-merging impacts homogeneity but not completness.
-
-\begin{figure}
-  \centering 
-  \centerline{\includegraphics[width=180mm]{cluster_metric_comparison.pdf}}
-  \caption{Type agreement evaluation metrics as a function of splitting types, merging types, and randomly distributing cells between types}
-  \label{fig:s:comparison}
-\end{figure}
-
 \end{document}

### 2015-04-23 site/pelicanconf.py
added github pages site generated by pelican

@@ -1,0 +1,39 @@
+#!/usr/bin/env python
+# -*- coding: utf-8 -*- #
+from __future__ import unicode_literals
+
+AUTHOR = u'Eric Jonas (jonas@eecs.berkeley.edu)'
+SITENAME = u'Automatic Discovery of Cell Types'
+SITEURL = ''
+
+THEME = "/Users/jonas/projects/pelican/themes/gum"
+
+# gum 
+
+PATH = 'content'
+
+TIMEZONE = 'Europe/Paris'
+
+DEFAULT_LANG = u'en'
+
+# Feed generation is usually not desired when developing
+FEED_ALL_ATOM = None
+CATEGORY_FEED_ATOM = None
+TRANSLATION_FEED_ATOM = None
+AUTHOR_FEED_ATOM = None
+AUTHOR_FEED_RSS = None
+
+# Blogroll
+LINKS = (('Pelican', 'http://getpelican.com/'),
+         ('Python.org', 'http://python.org/'),
+         ('Jinja2', 'http://jinja.pocoo.org/'),
+         ('You can modify those links in your config file', '#'),)
+
+# Social widget
+SOCIAL = (('You can add links in your config file', '#'),
+          ('Another social link', '#'),)
+
+DEFAULT_PAGINATION = False
+
+# Uncomment following line if you want document-relative URLs when developing
+#RELATIVE_URLS = True
